### 2024-03-13 HickernellFredSpecialSession.tex
Updates from Overleaf

\documentclass[12pt,a4paper,figuresright]{book}

\usepackage{amsmath,amssymb}
\usepackage{tabularx,graphicx,url,xcolor,rotating,multicol,epsfig,colortbl}

\setlength{\textheight}{25.2cm}
\setlength{\textwidth}{16.5cm} %\setlength{\textwidth}{18.2cm}
\setlength{\voffset}{-1.6cm}
\setlength{\hoffset}{-0.3cm} %\setlength{\hoffset}{-1.2cm}
\setlength{\evensidemargin}{-0.3cm} 
\setlength{\oddsidemargin}{0.3cm}
\setlength{\parindent}{0cm} 
\setlength{\parskip}{0.3cm}

% -- adding a talk
\newenvironment{talk}[6]% [1] talk title
                         % [2] speaker name, [3] affiliations, [4] email,
                         % [5] coauthors, [6] special session
                         % [7] time slot
                         % [8] talk id, [9] session id or photo
 {%\needspace{6\baselineskip}%
  \vskip 0pt\nopagebreak%
%   \colorbox{gray!20!white}{\makebox[0.99\textwidth][r]{}}\nopagebreak%
%   \ifthenelse{\equal{#9}{photo}}{%
%                     \\\\\colorbox{gray!20!white}{\makebox{\includegraphics[width=3cm]{#8}}}\nopagebreak}{}%
 \vskip 0pt\nopagebreak%
%  \label{#8}%
  \textbf{#1}\vspace{3mm}\\\nopagebreak%
  \textit{#2}\\\nopagebreak%
  #3\\\nopagebreak%
  \url{#4}\vspace{3mm}\\\nopagebreak%
  \ifthenelse{\equal{#5}{}}{}{Coauthor(s): #5\vspace{3mm}\\\nopagebreak}%
  \ifthenelse{\equal{#6}{}}{}{Special session: #6\quad \vspace{3mm}\\\nopagebreak}%
 }
 {\vspace{1cm}\nopagebreak}%

\pagestyle{empty}

% ------------------------------------------------------------------------
% Document begins here
% ------------------------------------------------------------------------
\begin{document}

\begin{talk}
  {Low Discrepancy Sequences for Kernel Density Estimation}% [1] talk title
  {Fred J. Hickernell}% [2] speaker name
  {Department of Applied Mathematics and Center for Interdisciplinary Scientific Computation, Illinois Institute of Technology}% [3] affiliations
  {hickernell@iit.edu}% [4] email
  {Names of coauthors go here, no affiliations of coauthors please, all affiliations will be included in an appendix of 
  the program book}% [5] coauthors
  {}% [6] special session. Leave this field empty for contributed talks. 
				% Insert the title of the special session if you were invited to give a talk in a special session.

Low discrepancy (LD) sequences, such as lattice, digital, and Halton sequences, can be used to estimate multivariate integrals with greater computational efficiency than independent and identically distributed (IID) sequences.  The use of LD sequences for density estimation has been less studied, but see \cite{AbdEtal21a,LEcuyer2022b} for some recent work.

We

\medskip

If you would like to include references, please do so by creating a simple list numbered by [1], [2], [3], \ldots. See example below.
Please do not use the \texttt{bibliography} environment or \texttt{bibtex} files.
APA reference style is recommended.
\begin{enumerate}
	\item[{[1]}] Niederreiter, Harald (1992). {\it Random number generation and quasi-Monte Carlo methods}. Society for Industrial and Applied Mathematics (SIAM).
	\item[{[2]}] L’Ecuyer, Pierre, \& Christiane Lemieux. (2002). Recent advances in randomized quasi-Monte Carlo methods. Modeling uncertainty: An examination of stochastic theory, methods, and applications, 419-474.
\end{enumerate}

Equations may be used if they are referenced. Please note that the equation numbers may be different (but will be cross-referenced correctly) in the final program book.
\end{talk}

\bibliographystyle{apa}
\bibliography{FJH23,FJHown23}

\end{document}

### 2024-03-13 HickernellFredSpecialSession.tex
Fred 2023 M2 Mac Updates
Updates from Overleaf

@@
   {hickernell@iit.edu}% [4] email
-  {Names of coauthors go here, no affiliations of coauthors please, all affiliations will be included in an appendix of 
-  the program book}% [5] coauthors
-  {}% [6] special session. Leave this field empty for contributed talks. 
+  {}% [5] coauthors
+  {Universality in QMC and related algorithms}% [6] special session. Leave this field empty for contributed talks. 
 				% Insert the title of the special session if you were invited to give a talk in a special session.
@@
 Low discrepancy (LD) sequences, such as lattice, digital, and Halton sequences, can be used to estimate multivariate integrals with greater computational efficiency than independent and identically distributed (IID) sequences.
-The use of LD sequences for density estimation has been less studied, but see \cite{AbdEtal21a,LEcuyer2022b} for some recent work.
+The use of LD sequences for density estimation has been less studied, but see %\cite{AbdEtal21a,LEcuyer2022b,LEcPuc20a}
+[1--3] for some recent work.
 
@@
 \begin{enumerate}
-	\item[{[1]}] Niederreiter, Harald (1992). {\it Random number generation and quasi-Monte Carlo methods}. Society for Industrial and Applied Mathematics (SIAM).
-	\item[{[2]}] L’Ecuyer, Pierre, \& Christiane Lemieux.
-(2002).
-Recent advances in randomized quasi-Monte Carlo methods.
-Modeling uncertainty: An examination of stochastic theory, methods, and applications, 419-474.
+	\item[{[1]}]
+{Ben Abdellah}, A., L'Ecuyer, P., Owen, A.~B., and Puchhammer, F. (2021).
+\newblock Density estimation by randomized quasi-{M}onte {C}arlo.
+\newblock {\em SIAM/ASA J.\ Uncertain.\ Quantif.}, 9(280-301).
+
+\item[{[2]}]
+L'Ecuyer, P. and Puchhammer, F. (2022).
+\newblock Density estimation by Monte Carlo and quasi-Monte Carlo.
+\newblock In Keller, A., editor, {\em {M}onte {C}arlo and Quasi-{M}onte {C}arlo Methods: {MCQMC}, {O}xford, England, {A}ugust 2020}, Springer Proceedings in Mathematics and Statistics. Springer, Cham.
+
+\item[{[3]}]
+L'Ecuyer, P., Puchhammer, F., and {Ben Abdellah}, A. (2022).
+\newblock Monte carlo and quasi{\textendash}{M}onte {C}arlo density estimation via conditioning.
+\newblock {\em INFORMS J.\ Comput.}, 34(3):1729--1748.
 \end{enumerate}
@@
 
-\bibliographystyle{apa}
+\bibliographystyle{apalike}
 \bibliography{FJH23,FJHown23}

### 2024-03-13 HickernellFredSpecialSession.tex
Updates from Overleaf

@@
 Low discrepancy (LD) sequences, such as lattice, digital, and Halton sequences, can be used to estimate multivariate integrals with greater computational efficiency than independent and identically distributed (IID) sequences.
-The use of LD sequences for density estimation has been less studied, but see %\cite{AbdEtal21a,LEcuyer2022b,LEcPuc20a}
-[1--3] for some recent work.
+The use of LD sequences for density estimation has been less studied in the past, but studied recently %\cite{AbdEtal21a,LEcuyer2022b,LEcPuc20a}
+[1--3].
 
-We
+This talk explores further the use of LD sequences for kernel density estimation, but from a deterministic perspective.
+Given a random variable $Y = f(\boldsymbol{X})$, where $\boldsymbol{X} \sim \mathcal{U}[0,1]^d$, with a probability density $\varrho$, and kernel density estimate, $\varrho$ as
 
@@
 
-If you would like to include references, please do so by creating a simple list numbered by [1], [2], [3], \ldots. See example below.
-Please do not use the \texttt{bibliography} environment or \texttt{bibtex} files.
-APA reference style is recommended.
 \begin{enumerate}
@@
 L'Ecuyer, P., Puchhammer, F., and {Ben Abdellah}, A. (2022).
-\newblock Monte carlo and quasi{\textendash}{M}onte {C}arlo density estimation via conditioning.
+\newblock Monte Carlo and quasi{\textendash}{M}onte {C}arlo density estimation via conditioning.
 \newblock {\em INFORMS J.\ Comput.}, 34(3):1729--1748.
@@
 
-Equations may be used if they are referenced. Please note that the equation numbers may be different (but will be cross-referenced correctly) in the final program book.
 \end{talk}
 
-\bibliographystyle{apalike}
-\bibliography{FJH23,FJHown23}
+%\bibliographystyle{apalike}
+%\bibliography{FJH23,FJHown23}
 

### 2024-03-13 HickernellFredSpecialSession.tex
Updates from Overleaf

@@
 
-Low discrepancy (LD) sequences, such as lattice, digital, and Halton sequences, can be used to estimate multivariate integrals with greater computational efficiency than independent and identically distributed (IID) sequences.
-The use of LD sequences for density estimation has been less studied in the past, but studied recently %\cite{AbdEtal21a,LEcuyer2022b,LEcPuc20a}
+Low discrepancy (LD) sequences, such as lattice, digital, and Halton sequences, are well known to estimate multivariate integrals with greater computational efficiency than independent and identically distributed (IID) sequences.
+The use of LD sequences for density estimation has been studied recently %\cite{AbdEtal21a,LEcuyer2022b,LEcPuc20a}
 [1--3].

### 2024-06-09 HickernellFredTutorial.tex
Updates from Overleaf

--- 2024-03-13 HickernellFredSpecialSession.tex
@@
 \begin{talk}
-  {Low Discrepancy Sequences for Kernel Density Estimation}% [1] talk title
+  {Quasi-Monte Carlo Methods:  What, Why, and How?}% [1] talk title
   {Fred J. Hickernell}% [2] speaker name
@@
   {}% [5] coauthors
-  {Universality in QMC and related algorithms}% [6] special session. Leave this field empty for contributed talks. 
+  {}% [6] special session. Leave this field empty for contributed talks. 
 				% Insert the title of the special session if you were invited to give a talk in a special session.
 
-Low discrepancy (LD) sequences, such as lattice, digital, and Halton sequences, are well known to estimate multivariate integrals with greater computational efficiency than independent and identically distributed (IID) sequences.
-The use of LD sequences for density estimation has been studied recently %\cite{AbdEtal21a,LEcuyer2022b,LEcPuc20a}
-[1--3].
+Many problems in  quantitative finance, uncertainty quantification, and other areas can be formulated as computing $\mu := \mathbb{E}(Y)$, where instances of $Y:=f(\boldsymbol{X})$ are generated by numerical simulation.
+The population mean, $\mu$, can be approximated by the sample mean, $\hat{\mu}_n := n^{-1} \sum_{i=1}^n f(\boldsymbol{X}_i)$.
+Computing $\mu$ is equivalent to computing a $d$-dimensional integral.
 
-This talk explores further the use of LD sequences for kernel density estimation, but from a deterministic perspective.
-Given a random variable $Y = f(\boldsymbol{X})$, where $\boldsymbol{X} \sim \mathcal{U}[0,1]^d$, with a probability density $\varrho$, and kernel density estimate, $\varrho$ as
+Quasi-Monte Carlo methods replace independent and identically distributed  sequences of random vectors, $\{\boldsymbol{X}_1, \boldsymbol{X}_2, \ldots \}$, by low discrepancy sequences.
+This accelerates the convergence of $\hat{\mu}_n$ to $\mu$ as $n \to \infty$.
 
-\medskip
-
-\begin{enumerate}
-	\item[{[1]}]
-{Ben Abdellah}, A., L'Ecuyer, P., Owen, A.~B., and Puchhammer, F. (2021).
-\newblock Density estimation by randomized quasi-{M}onte {C}arlo.
-\newblock {\em SIAM/ASA J.\ Uncertain.\ Quantif.}, 9(280-301).
-
-\item[{[2]}]
-L'Ecuyer, P. and Puchhammer, F. (2022).
-\newblock Density estimation by Monte Carlo and quasi-Monte Carlo.
-\newblock In Keller, A., editor, {\em {M}onte {C}arlo and Quasi-{M}onte {C}arlo Methods: {MCQMC}, {O}xford, England, {A}ugust 2020}, Springer Proceedings in Mathematics and Statistics. Springer, Cham.
-
-\item[{[3]}]
-L'Ecuyer, P., Puchhammer, F., and {Ben Abdellah}, A. (2022).
-\newblock Monte Carlo and quasi{\textendash}{M}onte {C}arlo density estimation via conditioning.
-\newblock {\em INFORMS J.\ Comput.}, 34(3):1729--1748.
-\end{enumerate}
+This tutorial describes  low discrepancy sequences  and their quality measures.
+We demonstrate the performance gains possible with quasi-Monte Carlo methods.
+Moreover, we describe how to formulate problems to realize the most increase in performance using quasi-Monte Carlo methods.
+We also briefly describe the use of quasi-Monte Carlo methods for problems beyond computing the mean.
 
@@
 
-%\bibliographystyle{apalike}
-%\bibliography{FJH23,FJHown23}
+%\bibliographystyle{apalike} \bibliography{FJH23,FJHown23}
 

### 2024-09-20 authorTemplateMCQMC2024.tex
Fred M3 Mac Updates

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a proceedings volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% to be used for MCQMC 2024
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{svproc}
%
% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%

% to typeset URLs, URIs, and DOIs
\usepackage{url}
\def\UrlFont{\rmfamily}

%%%%%%%added CL
\usepackage{hyperref}
\usepackage{type1cm}        % activate if the above 3 fonts are
% not available on your system
%
\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool
                             % when including figure files
%%% if you are including figures and images, you are encouraged to put them in sub-directory(ies)
%%% whose path(s) can be provided using the following command
%%% \graphicspath{{<path 1>}{<path 2>}}
%%% for example if your sub-directory is called MyFigures then you would use the command
%%% \graphicspath{{MyFigures/}}

\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom

\usepackage{newtxtext}       %
\usepackage[varvw]{newtxmath}       % selects Times Roman as basic font

% see the list of further useful packages
% in the Reference Guide

%\makeindex             % used for the subject index
% please use the style svind.ist with
% your makeindex program

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Additional packages
\usepackage{bm}% for bold math symbols
\usepackage{siunitx}% for table alignments

% Additional environment
% \usepackage{amsmath}
% \newtheorem{algorithm}{\upshape Algorithm}[1][]

\newcounter{algorithm}%[section]
\newenvironment{algorithm}[1][]{\refstepcounter{algorithm}\par\medskip
  \noindent\textbf{Algorithm~\thealgorithm\mbox{ }#1}}{\medskip}

\input{macros}
%%%% end added CL

\begin{document}

\mainmatter              % start of a contribution
%
\title{Hamiltonian Mechanics unter besonderer Ber\"ucksichtigung der
h\"ohereren Lehranstalten}
%
\titlerunning{Hamiltonian Mechanics}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Ivar Ekeland\inst{1} \and Roger Temam\inst{2}
Jeffrey Dean \and David Grove \and Craig Chambers \and Kim~B.~Bruce \and
Elsa Bertino}
%
\authorrunning{Ivar Ekeland et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Ivar Ekeland, Roger Temam, Jeffrey Dean, David Grove,
Craig Chambers, Kim B. Bruce, and Elisa Bertino}
%
\institute{Princeton University, Princeton NJ 08544, USA,\\
\email{I.Ekeland@princeton.edu},\\ WWW home page:
\texttt{http://users/\homedir iekeland/web/welcome.html}
\and
Universit\'{e} de Paris-Sud,
Laboratoire d'Analyse Num\'{e}rique, B\^{a}timent 425,\\
F-91405 Orsay Cedex, France}

\maketitle              % typeset the title of the contribution

\begin{abstract}
The abstract should summarize the contents of the paper
using at least 70 and at most 150 words. It will be set in 9-point
font size and be inset 1.0 cm from the right and left margins.
There will be two blank lines before and after the Abstract. \dots
% We would like to encourage you to list your keywords within
% the abstract section using the \keywords{...} command.
\keywords{computational geometry, graph theory, Hamilton cycles}
\end{abstract}
%
\section{Fixed-Period Problems: The Sublinear Case}
%
With this chapter, the preliminaries are over, and we begin the search
for periodic solutions to Hamiltonian systems. All this will be done in
the convex case; that is, we shall study the boundary-value problem
\begin{eqnarray*}
  \dot{x}&=&JH' (t,x)\\
  x(0) &=& x(T)
\end{eqnarray*}
with $H(t,\cdot)$ a convex function of $x$, going to $+\infty$ when
$\left\|x\right\| \to \infty$. We also want to test various things such as whether $\setN$ and $\var(\bsbeta)$ or $\bsu$ are correctly working.

%
\subsection{Autonomous Systems}
%
In this section, we will consider the case when the Hamiltonian $H(x)$
is autonomous. For the sake of simplicity, we shall also assume that it
is $C^{1}$.

We shall first consider the question of nontriviality, within the
general framework of
$\left(A_{\infty},B_{\infty}\right)$-subquadratic Hamiltonians. In
the second subsection, we shall look into the special case when $H$ is
$\left(0,b_{\infty}\right)$-subquadratic,
and we shall try to derive additional information.
%
\subsubsection{The General Case: Nontriviality.}
%
We assume that $H$ is
$\left(A_{\infty},B_{\infty}\right)$-sub\-qua\-dra\-tic at infinity,
for some constant symmetric matrices $A_{\infty}$ and $B_{\infty}$,
with $B_{\infty}-A_{\infty}$ positive definite. Set:
\begin{eqnarray}
\gamma :&=&{\rm smallest\ eigenvalue\ of}\ \ B_{\infty} - A_{\infty} \\
  \lambda : &=& {\rm largest\ negative\ eigenvalue\ of}\ \
  J \frac{d}{dt} +A_{\infty}\ .
\end{eqnarray}

Theorem~\ref{ghou:pre} tells us that if $\lambda +\gamma < 0$, the
boundary-value problem:
\begin{equation}
\begin{array}{rcl}
  \dot{x}&=&JH' (x)\\
  x(0)&=&x (T)
\end{array}
\end{equation}
has at least one solution
$\overline{x}$, which is found by minimizing the dual
action functional:
\begin{equation}
  \psi (u) = \int_{o}^{T} \left[\frac{1}{2}
  \left(\Lambda_{o}^{-1} u,u\right) + N^{\ast} (-u)\right] dt
\end{equation}
on the range of $\Lambda$, which is a subspace $R (\Lambda)_{L}^{2}$
with finite codimension. Here
\begin{equation}
  N(x) := H(x) - \frac{1}{2} \left(A_{\infty} x,x\right)
\end{equation}
is a convex function, and
\begin{equation}
  N(x) \le \frac{1}{2}
  \left(\left(B_{\infty} - A_{\infty}\right) x,x\right)
  + c\ \ \ \forall x\ .
\end{equation}

%
\begin{proposition}
Assume $H'(0)=0$ and $ H(0)=0$. Set:
\begin{equation}
  \delta := \liminf_{x\to 0} 2 N (x) \left\|x\right\|^{-2}\ .
  \label{eq:one}
\end{equation}

If $\gamma < - \lambda < \delta$,
the solution $\overline{u}$ is non-zero:
\begin{equation}
  \overline{x} (t) \ne 0\ \ \ \forall t\ .
\end{equation}
\end{proposition}
%
\begin{proof}
Condition (\ref{eq:one}) means that, for every
$\delta ' > \delta$, there is some $\varepsilon > 0$ such that
\begin{equation}
  \left\|x\right\| \le \varepsilon \Rightarrow N (x) \le
  \frac{\delta '}{2} \left\|x\right\|^{2}\ .
\end{equation}

It is an exercise in convex analysis, into which we shall not go, to
show that this implies that there is an $\eta > 0$ such that
\begin{equation}
  f\left\|x\right\| \le \eta
  \Rightarrow N^{\ast} (y) \le \frac{1}{2\delta '}
  \left\|y\right\|^{2}\ .
  \label{eq:two}
\end{equation}

\begin{figure}
\vspace{2.5cm}
%%%You can un-comment the \includegraphics{} line below and comment the one above
%%%if you have copied in your directory the file figure.eps
%%%from Springer zip file. As mentioned in the preamble, in your manuscript you are encouraged to put
%%%figures and images in a sub-directory which can be provided in the preamble via the 
%%%command \graphicspath{{<path 1>}{<path 2>}}
%\includegraphics{figure.eps}
\caption{This is the caption of the figure displaying a white eagle and
a white horse on a snow field}
\end{figure}

Since $u_{1}$ is a smooth function, we will have
$\left\|hu_{1}\right\|_\infty \le \eta$
for $h$ small enough, and inequality (\ref{eq:two}) will hold,
yielding thereby:
\begin{equation}
  \psi (hu_{1}) \le \frac{h^{2}}{2}
  \frac{1}{\lambda} \left\|u_{1} \right\|_{2}^{2} + \frac{h^{2}}{2}
  \frac{1}{\delta '} \left\|u_{1}\right\|^{2}\ .
\end{equation}

If we choose $\delta '$ close enough to $\delta$, the quantity
$\left(\frac{1}{\lambda} + \frac{1}{\delta '}\right)$
will be negative, and we end up with
\begin{equation}
  \psi (hu_{1}) < 0\ \ \ \ \ {\rm for}\ \ h\ne 0\ \ {\rm small}\ .
\end{equation}

On the other hand, we check directly that $\psi (0) = 0$. This shows
that 0 cannot be a minimizer of $\psi$, not even a local one.
So $\overline{u} \ne 0$ and
$\overline{u} \ne \Lambda_{o}^{-1} (0) = 0$. \qed
\end{proof}
%
\begin{corollary}
Assume $H$ is $C^{2}$ and
$\left(a_{\infty},b_{\infty}\right)$-subquadratic at infinity. Let
$\xi_{1},\allowbreak\dots,\allowbreak\xi_{N}$  be the
equilibria, that is, the solutions of $H' (\xi ) = 0$.
Denote by $\omega_{k}$
the smallest eigenvalue of $H'' \left(\xi_{k}\right)$, and set:
\begin{equation}
  \omega : = {\rm Min\,} \left\{\omega_{1},\dots,\omega_{k}\right\}\ .
\end{equation}
If:
\begin{equation}
  \frac{T}{2\pi} b_{\infty} <
  - E \left[- \frac{T}{2\pi}a_{\infty}\right] <
  \frac{T}{2\pi}\omega
  \label{eq:three}
\end{equation}
then minimization of $\psi$ yields a non-constant $T$-periodic solution
$\overline{x}$.
\end{corollary}
%

We recall once more that by the integer part $E [\alpha ]$ of
$\alpha \in \bbbr$, we mean the $a\in \bbbz$
such that $a< \alpha \le a+1$. For instance,
if we take $a_{\infty} = 0$, Corollary 2 tells
us that $\overline{x}$ exists and is
non-constant provided that:

\begin{equation}
  \frac{T}{2\pi} b_{\infty} < 1 < \frac{T}{2\pi}
\end{equation}
or
\begin{equation}
  T\in \left(\frac{2\pi}{\omega},\frac{2\pi}{b_{\infty}}\right)\ .
  \label{eq:four}
\end{equation}

%
\begin{proof}
The spectrum of $\Lambda$ is $\frac{2\pi}{T} \bbbz +a_{\infty}$. The
largest negative eigenvalue $\lambda$ is given by
$\frac{2\pi}{T}k_{o} +a_{\infty}$,
where
\begin{equation}
  \frac{2\pi}{T}k_{o} + a_{\infty} < 0
  \le \frac{2\pi}{T} (k_{o} +1) + a_{\infty}\ .
\end{equation}
Hence:
\begin{equation}
  k_{o} = E \left[- \frac{T}{2\pi} a_{\infty}\right] \ .
\end{equation}

The condition $\gamma < -\lambda < \delta$ now becomes:
\begin{equation}
  b_{\infty} - a_{\infty} <
  - \frac{2\pi}{T} k_{o} -a_{\infty} < \omega -a_{\infty}
\end{equation}
which is precisely condition (\ref{eq:three}).\qed
\end{proof}
%

\begin{lemma}
Assume that $H$ is $C^{2}$ on $\bbbr^{2n} \setminus \{ 0\}$ and
that $H'' (x)$ is non-de\-gen\-er\-ate for any $x\ne 0$. Then any local
minimizer $\widetilde{x}$ of $\psi$ has minimal period $T$.
\end{lemma}
%
\begin{proof}
We know that $\widetilde{x}$, or
$\widetilde{x} + \xi$ for some constant $\xi
\in \bbbr^{2n}$, is a $T$-periodic solution of the Hamiltonian system:
\begin{equation}
  \dot{x} = JH' (x)\ .
\end{equation}

There is no loss of generality in taking $\xi = 0$. So
$\psi (x) \ge \psi (\widetilde{x} )$
for all $\widetilde{x}$ in some neighbourhood of $x$ in
$W^{1,2} \left(\bbbr / T\bbbz ; \bbbr^{2n}\right)$.

But this index is precisely the index
$i_{T} (\widetilde{x} )$ of the $T$-periodic
solution $\widetilde{x}$ over the interval
$(0,T)$, as defined in Sect.~2.6. So
\begin{equation}
  i_{T} (\widetilde{x} ) = 0\ .
  \label{eq:five}
\end{equation}

Now if $\widetilde{x}$ has a lower period, $T/k$ say,
we would have, by Corollary 31:
\begin{equation}
  i_{T} (\widetilde{x} ) =
  i_{kT/k}(\widetilde{x} ) \ge
  ki_{T/k} (\widetilde{x} ) + k-1 \ge k-1 \ge 1\ .
\end{equation}

This would contradict (\ref{eq:five}), and thus cannot happen.\qed
\end{proof}
%
\paragraph{Notes and Comments.}
The results in this section are a
refined version of \cite{smit:wat};
the minimality result of Proposition
14 was the first of its kind.

To understand the nontriviality conditions, such as the one in formula
(\ref{eq:four}), one may think of a one-parameter family
$x_{T}$, $T\in \left(2\pi\omega^{-1}, 2\pi b_{\infty}^{-1}\right)$
of periodic solutions, $x_{T} (0) = x_{T} (T)$,
with $x_{T}$ going away to infinity when $T\to 2\pi \omega^{-1}$,
which is the period of the linearized system at 0.

\begin{table}
\caption{This is the example table taken out of {\it The
\TeX{}book,} p.\,246}
\begin{center}
\begin{tabular}{r@{\quad}rl}
\hline
\multicolumn{1}{l}{\rule{0pt}{12pt}
                   Year}&\multicolumn{2}{l}{World population}\\[2pt]
\hline\rule{0pt}{12pt}
8000 B.C.  &     5,000,000& \\
  50 A.D.  &   200,000,000& \\
1650 A.D.  &   500,000,000& \\
1945 A.D.  & 2,300,000,000& \\
1980 A.D.  & 4,400,000,000& \\[2pt]
\hline
\end{tabular}
\end{center}
\end{table}
%
\begin{theorem} [Ghoussoub-Preiss]\label{ghou:pre}
Assume $H(t,x)$ is
$(0,\varepsilon )$-subquadratic at
infinity for all $\varepsilon > 0$, and $T$-periodic in $t$
\begin{equation}
  H (t,\cdot )\ \ \ \ \ {\rm is\ convex}\ \ \forall t
\end{equation}
\begin{equation}
  H (\cdot ,x)\ \ \ \ \ {\rm is}\ \ T{\rm -periodic}\ \ \forall x
\end{equation}
\begin{equation}
  H (t,x)\ge n\left(\left\|x\right\|\right)\ \ \ \ \
  {\rm with}\ \ n (s)s^{-1}\to \infty\ \ {\rm as}\ \ s\to \infty
\end{equation}
\begin{equation}
  \forall \varepsilon > 0\ ,\ \ \ \exists c\ :\
  H(t,x) \le \frac{\varepsilon}{2}\left\|x\right\|^{2} + c\ .
\end{equation}

Assume also that $H$ is $C^{2}$, and $H'' (t,x)$ is positive definite
everywhere. Then there is a sequence $x_{k}$, $k\in \bbbn$, of
$kT$-periodic solutions of the system
\begin{equation}
  \dot{x} = JH' (t,x)
\end{equation}
such that, for every $k\in \bbbn$, there is some $p_{o}\in\bbbn$ with:
\begin{equation}
  p\ge p_{o}\Rightarrow x_{pk} \ne x_{k}\ .
\end{equation}
\qed
\end{theorem}
%
\begin{example} [{{\rm External forcing}}]
Consider the system:
\begin{equation}
  \dot{x} = JH' (x) + f(t)
\end{equation}
where the Hamiltonian $H$ is
$\left(0,b_{\infty}\right)$-subquadratic, and the
forcing term is a distribution on the circle:
\begin{equation}
  f = \frac{d}{dt} F + f_{o}\ \ \ \ \
  {\rm with}\ \ F\in L^{2} \left(\bbbr / T\bbbz; \bbbr^{2n}\right)\ ,
\end{equation}
where $f_{o} : = T^{-1}\int_{o}^{T} f (t) dt$. For instance,
\begin{equation}
  f (t) = \sum_{k\in \bbbn} \delta_{k} \xi\ ,
\end{equation}
where $\delta_{k}$ is the Dirac mass at $t= k$ and
$\xi \in \bbbr^{2n}$ is a
constant, fits the prescription. This means that the system
$\dot{x} = JH' (x)$ is being excited by a
series of identical shocks at interval $T$.
\end{example}
%
\begin{definition}
Let $A_{\infty} (t)$ and $B_{\infty} (t)$ be symmetric
operators in $\bbbr^{2n}$, depending continuously on
$t\in [0,T]$, such that
$A_{\infty} (t) \le B_{\infty} (t)$ for all $t$.

A Borelian function
$H: [0,T]\times \bbbr^{2n} \to \bbbr$
is called
$\left(A_{\infty} ,B_{\infty}\right)$-{\it subquadratic at infinity}
if there exists a function $N(t,x)$ such that:
\begin{equation}
  H (t,x) = \frac{1}{2} \left(A_{\infty} (t) x,x\right) + N(t,x)
\end{equation}
\begin{equation}
  \forall t\ ,\ \ \ N(t,x)\ \ \ \ \
  {\rm is\ convex\ with\  respect\  to}\ \ x
\end{equation}
\begin{equation}
  N(t,x) \ge n\left(\left\|x\right\|\right)\ \ \ \ \
  {\rm with}\ \ n(s)s^{-1}\to +\infty\ \ {\rm as}\ \ s\to +\infty
\end{equation}
\begin{equation}
  \exists c\in \bbbr\ :\ \ \ H (t,x) \le
  \frac{1}{2} \left(B_{\infty} (t) x,x\right) + c\ \ \ \forall x\ .
\end{equation}

If $A_{\infty} (t) = a_{\infty} I$ and
$B_{\infty} (t) = b_{\infty} I$, with
$a_{\infty} \le b_{\infty} \in \bbbr$,
we shall say that $H$ is
$\left(a_{\infty},b_{\infty}\right)$-subquadratic
at infinity. As an example, the function
$\left\|x\right\|^{\alpha}$, with
$1\le \alpha < 2$, is $(0,\varepsilon )$-subquadratic at infinity
for every $\varepsilon > 0$. Similarly, the Hamiltonian
\begin{equation}
H (t,x) = \frac{1}{2} k \left\|k\right\|^{2} +\left\|x\right\|^{\alpha}
\end{equation}
is $(k,k+\varepsilon )$-subquadratic for every $\varepsilon > 0$.
Note that, if $k<0$, it is not convex.
\end{definition}
%

\paragraph{Notes and Comments.}
The first results on subharmonics were
obtained by Rabinowitz in \cite{fo:kes:nic:tue}, who showed the existence of
infinitely many subharmonics both in the subquadratic and superquadratic
case, with suitable growth conditions on $H'$. Again the duality
approach enabled Clarke and Ekeland in \cite{may:ehr:stein} to treat the
same problem in the convex-subquadratic case, with growth conditions on
$H$ only.

Recently, Michalek and Tarantello (see \cite{fost:kes} and \cite{czaj:fitz})
have obtained lower bound on the number of subharmonics of period $kT$,
based on symmetry considerations and on pinching estimates, as in
Sect.~5.2 of this article.

%%% To ensure the bibliography has the correct style please run bibtex with the spmpsci style
%%% which is included in the Springer zip file
%%% Here we assume refs.bib would be the name of the bib file containing bibliographic info
%%% You can then copy the .bbl produced, as given in the example below
%%%
%\bibliographystyle{spmpsci}
%\bibliography{refs}
%
% ---- Bibliography ----
%
\begin{thebibliography}{6}
%

\bibitem {smit:wat}
Smith, T.F., Waterman, M.S.: Identification of common molecular subsequences.
J. Mol. Biol. 147, 195?197 (1981). \url{doi:10.1016/0022-2836(81)90087-5}

\bibitem {may:ehr:stein}
May, P., Ehrlich, H.-C., Steinke, T.: ZIB structure prediction pipeline:
composing a complex biological workflow through web services.
In: Nagel, W.E., Walter, W.V., Lehner, W. (eds.) Euro-Par 2006.
LNCS, vol. 4128, pp. 1148?1158. Springer, Heidelberg (2006).
\url{doi:10.1007/11823285_121}

\bibitem {fost:kes}
Foster, I., Kesselman, C.: The Grid: Blueprint for a New Computing Infrastructure.
Morgan Kaufmann, San Francisco (1999)

\bibitem {czaj:fitz}
Czajkowski, K., Fitzgerald, S., Foster, I., Kesselman, C.: Grid information services
for distributed resource sharing. In: 10th IEEE International Symposium
on High Performance Distributed Computing, pp. 181?184. IEEE Press, New York (2001).
\url{doi: 10.1109/HPDC.2001.945188}

\bibitem {fo:kes:nic:tue}
Foster, I., Kesselman, C., Nick, J., Tuecke, S.: The physiology of the grid: an open grid services architecture for distributed systems integration. Technical report, Global Grid
Forum (2002)

\bibitem {onlyurl}
National Center for Biotechnology Information. \url{http://www.ncbi.nlm.nih.gov}

\end{thebibliography}
\end{document}

### 2024-09-20 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

--- 2024-09-20 authorTemplateMCQMC2024.tex
@@
 %
-\title{Hamiltonian Mechanics unter besonderer Ber\"ucksichtigung der
-h\"ohereren Lehranstalten}
+\title{Quasi-Monte Carlo Methods:  What, Why, and How?}
 %
@@
 %
-\author{Ivar Ekeland\inst{1} \and Roger Temam\inst{2}
-Jeffrey Dean \and David Grove \and Craig Chambers \and Kim~B.~Bruce \and
-Elsa Bertino}
+\author{Fred J. Hickernell\inst{1} \and Aleksei G. Sorokin\inst{2}}
 %
-\authorrunning{Ivar Ekeland et al.} % abbreviated author list (for running head)
+\authorrunning{Hickernell and Sorokin} % abbreviated author list (for running head)
 %
 %%%% list of authors for the TOC (use if author list has to be modified)
-\tocauthor{Ivar Ekeland, Roger Temam, Jeffrey Dean, David Grove,
-Craig Chambers, Kim B. Bruce, and Elisa Bertino}
+\tocauthor{Fred J. Hickernell, Aleksei G. Sorokin}
 %
-\institute{Princeton University, Princeton NJ 08544, USA,\\
-\email{I.Ekeland@princeton.edu},\\ WWW home page:
-\texttt{http://users/\homedir iekeland/web/welcome.html}
+\institute{Departmet of Applied Mathematics, Illinois Institute of Technology, Chicago, IL, 60616, USA\\
+\email{hickernell@iit.edu}, \ 
+\texttt{http://www.iit.edu/~hickernell}
 \and
-Universit\'{e} de Paris-Sud,
-Laboratoire d'Analyse Num\'{e}rique, B\^{a}timent 425,\\
-F-91405 Orsay Cedex, France}
+Departmet of Applied Mathematics, Illinois Institute of Technology, Chicago, IL, 60616, USA\\
+\email{asorokin@hawk.iit.edu}}
 
@@
 \begin{abstract}
-The abstract should summarize the contents of the paper
-using at least 70 and at most 150 words. It will be set in 9-point
-font size and be inset 1.0 cm from the right and left margins.
-There will be two blank lines before and after the Abstract. \dots
-% We would like to encourage you to list your keywords within
-% the abstract section using the \keywords{...} command.
-\keywords{computational geometry, graph theory, Hamilton cycles}
+Many problems in  quantitative finance, uncertainty quantification, and other areas can be formulated as computing $\mu := \mathbb{E}(Y)$, where instances of $Y:=f(\boldsymbol{X})$ are generated by numerical simulation.
+The population mean, $\mu$, can be approximated by the sample mean, $\hat{\mu}_n := n^{-1} \sum_{i=1}^n f(\boldsymbol{X}_i)$.
+Computing $\mu$ is equivalent to computing a $d$-dimensional integral.
+
+Quasi-Monte Carlo methods replace independent and identically distributed  sequences of random vectors, $\{\boldsymbol{X}_1, \boldsymbol{X}_2, \ldots \}$, by low discrepancy sequences.
+This accelerates the convergence of $\hat{\mu}_n$ to $\mu$ as $n \to \infty$.
+
+This tutorial describes  low discrepancy sequences  and their quality measures.
+We demonstrate the performance gains possible with quasi-Monte Carlo methods.
+Moreover, we describe how to formulate problems to realize the most increase in performance using quasi-Monte Carlo methods.
+We also briefly describe the use of quasi-Monte Carlo methods for problems beyond computing the mean.
+
+\keywords{}
 \end{abstract}

### 2024-10-21 HickernellMCQMC2024_QMCTutorial.tex
M3 Mac updates

@@
 %
-\section{Fixed-Period Problems: The Sublinear Case}
+\section{Introduction} \label{sec:intro}
 %
-With this chapter, the preliminaries are over, and we begin the search
-for periodic solutions to Hamiltonian systems. All this will be done in
-the convex case; that is, we shall study the boundary-value problem
-\begin{eqnarray*}
-  \dot{x}&=&JH' (t,x)\\
-  x(0) &=& x(T)
-\end{eqnarray*}
-with $H(t,\cdot)$ a convex function of $x$, going to $+\infty$ when
-$\left\|x\right\| \to \infty$. We also want to test various things such as whether $\setN$ and $\var(\bsbeta)$ or $\bsu$ are correctly working.
+%
+\section{Where in Practice} \label{sec:practice}
+\section{Low Discrepancy Sequence Constructions} \label{sec:construct}
+\section{Discrepancy} \label{sec:discrepancy}
+\section{Stopping Criteria} \label{sec:stop}
+\section{Reformulating Our Problem} \label{sec:reformulate}
+\section{Ongoing Research} \label{sec:onging}
+\section{Conclusion} \label{sec:conclusion}
 
-%
-\subsection{Autonomous Systems}
+\end{document}
 %

### 2024-10-23 HickernellMCQMC2024_QMCTutorial.tex
M3 Mac updates

@@
 
+%%Fred's temporary macros
+
+%%End of Fred's temporary macros
+
 \begin{document}
@@
 %
-\titlerunning{Hamiltonian Mechanics}  % abbreviated title (for running head)
+\titlerunning{Quasi-Monte Carlo Methods}  % abbreviated title (for running head)
 %                                     also used for the TOC unless
@@
 \section{Introduction} \label{sec:intro}
-%
+There are may settings where key quantities are unknown, such as future financial market forces, the porosity of rock containing oil or gas, or particle interactions leading to observations in a high energy physics experiment.
+Such situations are often modeled using random variables or stochastic processes.
+Computations are performed by generating a multidude of scenarios informed by the assumed probability distribution.
+One may then estimate the mean outcome, quantiles, and distributions.
+This is the essence of the Monte Carlo method.
+
+In mathematical terms, the outcome is $Y := f(\boldsymbol{X})$, where $\boldsymbol{X}$ is a vector random variable driving the outcome, and $f$ is a function whose values can be computed by an algorithm, whose complexity may be such that it can be considered a ``black box''.
+The scenarios are the $\bsX_0, \bsX_1, \ldots$ that give rise to $Y_0 = f(\bsX_0), Y_1 = f(\bsX_1), \ldots$.
+The $Y_i$ form the basis for approximating $\bbE(Y)$ and other quantities of interest.
+
+How one generates scenarios
 %

### 2024-10-24 HickernellMCQMC2024_QMCTutorial.tex
M2 Mac updates

--- 2024-10-21 HickernellMCQMC2024_QMCTutorial.tex
@@
 \section{Introduction} \label{sec:intro}
-%
+There are may settings where key underlying quantities that affect the outcome are unknown, e.g.,
+\begin{itemize}
+	\item Future market forces, which affect financial risk,
+	\item The porosity of rock, which affect the optimal parameters for extracting oil or gas, or
+	\item Elemetary particle interactions in a high energy physics experiment, which affect what is observed.
+\end{itemize}
+In such situations the unknown inputs are often modeled using random variables or stochastic processes.
+Computations are performed by generating a multidude of possible scenarios informed by the assumed probability distribution of the input.
+These are used to estimate the mean, quantiles, and/or probability distribution of the outcome.
+This is the essence of the Monte Carlo method.
+
+In mathematical terms, the outcome is $Y := f(\boldsymbol{X})$, where $\boldsymbol{X}$ is a vector random variable of inputs determining the outcome.
+The values of the function $f$ can be computed by an algorithm, whose complexity may be such that $f$ can be considered a ``black box''.
+The scenarios are $\bsX_0, \bsX_1, \ldots$, which give rise to $Y_0 = f(\bsX_0), Y_1 = f(\bsX_1), \ldots$.
+The $Y_i$ form the basis for approximating $\bbE(Y)$ and other quantities of interest.
+
+How one generates scenarios
 %

### 2024-10-24 HickernellMCQMC2024_QMCTutorial.tex
M2 Mac updates

@@
 	\item Future market forces, which affect financial risk,
-	\item The porosity of rock, which affect the optimal parameters for extracting oil or gas, or
+	\item The porosity of rock, which affect the extraction of oil or gas, or
 	\item Elemetary particle interactions in a high energy physics experiment, which affect what is observed.
@@
 The values of the function $f$ can be computed by an algorithm, whose complexity may be such that $f$ can be considered a ``black box''.
-The scenarios are $\bsX_0, \bsX_1, \ldots$, which give rise to $Y_0 = f(\bsX_0), Y_1 = f(\bsX_1), \ldots$.
+The scenarios are $\bsX_0, \bsX_1, \ldots$, which give rise to outcomes $Y_0 = f(\bsX_0), Y_1 = f(\bsX_1), \ldots$.
 The $Y_i$ form the basis for approximating $\bbE(Y)$ and other quantities of interest.

### 2024-10-25 HickernellMCQMC2024_QMCTutorial.tex
txs auto checkin
M2 Mac updates

@@
 
-How one generates scenarios
+Simple or independent and identically distributed (IID)  Monte Carlo chooses $\bsX_0, \bsX_1, \ldots$ to be IID.
+Quasi-Monte Carlo (qMC) methods choose $\bsX_0, \bsX_1, \ldots$ \emph{not} IID, but to have an empirical distribution that is close to the actual probability distribution of $\bsX$.
+The difference between these two distributions is called a \emph{discrepancy}, and the sequences used in qMC are called low discrepancy sequences.
+
+This tutorial describes how
+
 %

### 2025-01-07 HickernellMCQMC2024_QMCTutorial.tex
M2 Mac updates

--- 2024-10-24 HickernellMCQMC2024_QMCTutorial.tex
@@
 %%Fred's temporary macros
+\newcommand{\figpath}{Figures}
 
@@
 These are used to estimate the mean, quantiles, and/or probability distribution of the outcome.
-This is the essence of the Monte Carlo method.
+This is the essence of the Monte Carlo (MC) method.
 
@@
 
-How one generates scenarios
+Simple or independent and identically distributed (IID)  Monte Carlo chooses the sequence $\bsX_0, \bsX_1, \ldots$ to be IID.
+Quasi-Monte Carlo (qMC) methods choose $\bsX_0, \bsX_1, \ldots$ \emph{not} IID, but with an empirical distribution that approximates the true probability distribution of $\bsX$.
+The difference between these two distributions is called a \emph{discrepancy}, and the sequences used in qMC are called low discrepancy sequences.
+
+This tutorial describes what qMC is, why we would want to use qMC, and how qMC can be implemented well.
+The next session provides an illustration of qMC in practice.
+This is followed by a description of how low discrepancy sequences are constructed and various measures of quality.
+We then explain how randomization can improve qMC, how to decide what sample size, $n$, is sufficient, and how to rewrite the problem of interest in a qMC-friendly way.
+
 %
-\section{Where in Practice} \label{sec:practice}
+\section{An Illustration of Quasi-Monte Carlo} \label{sec:practice}
+
+To illustrate the benefits of qMC, consider the example from Keister \cite[(13)]{Kei96}, motivated by computational physics:
+
+\begin{equation}\label{eq:keisterA}
+	\mu := \int_{\mathbb{R}^d} \cos(\lVert \bst \rVert) \exp(-\lVert \bst \rVert^2) \, \rmd \bst,
+\end{equation}
+where $\lVert \bst \rVert := \sqrt{t_1^2 + \cdots + t_d^2}$.  This integral may be evaluated numerically by re-writing it in spherical coordinates as
+\begin{equation}\label{eq:keisterExact}
+	\mu = \frac{2 \pi^{d/2}}{\Gamma(d/2)}\int_{0}^{\infty} \cos(r) \exp(-r^2) \, r^{d-1} \rmd r,
+\end{equation}
+where $2 \pi^{d/2}/\Gamma(d/2)$ is the surface area of the sphere in $d$ dimensions, and $\Gamma$ is the Gamma function.
+Having a non-trivial test case with a value that can be easily calculated allows us to compute the numerical errors of various cubature schemes.
+
+The aim is approximate $\mu$ by a sample mean,
+\begin{equation} \label{eq:sample_mean}
+	\hat{\mu}_n := \frac 1n \sum_{i=1}^n f(\boldsymbol{X}_i)
+\end{equation}
+but to do so requires some further preparation.
+This integral above, $\mu$, may be thought of as the expectation of $Y := g(\bsT) := \pi^{d/2} \cos(\lVert \bsT \rVert)$, where $T_1, \ldots, T_d$ are IID Gaussian random variables with zero mean and variance $1/2$, i.e., $\bsT \sim \calN(\bszero,\mathsf{I}/2)$:
+\begin{equation}\label{eq:keisterB}
+	\mu = \int_{\mathbb{R}^d} \underbrace{\pi^{d/2} \cos(\lVert \bst \rVert)}_{=:g(\bst)} \cdot \underbrace{\frac{\exp(-\lVert \bst \rVert^2)}{\pi^{d/2}}}_{\text{density of } \calN(\bszero,\mathsf{I}/2)} \, \rmd \bst = \bbE(Y) = \bbE[g(\bsT)].
+\end{equation}
+
+Since the low discrepancy sequences that drive qMC are defined on the unit cube, $[0,1]^d$, we perform a variable transformation $\bst = \bigl (\Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d) \bigr )/\sqrt{2}$, where is $\Phi$ is the cumulative distribution function of the standard Gaussian random variable.
+This reimagines the integral $I$ as the expectation of a function, $f$, of a standard uniform random variable:
+\begin{multline}\label{eq:keisterB}
+	\mu = \int_{[0,1]^d} \underbrace{\pi^{d/2} \cos\Bigl (\sqrt{2} \bigl \lVert \bigl( \Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d)\bigr) \bigr\rVert \Bigr)}_{=:f(\bsx)} \, \rmd \bsx \\
+	= \bbE[Y]
+	= \bbE[f(\bsX)], \qquad \bsX \sim \calU[0,1]^d.
+\end{multline}
+
+Consider the specific case of $d=6$ for which $\mu \approx -2.327303729298$.
+We approximate $\mu$ by the sample mean defined in \eqref{eq:sample_mean} for three different kinds of nodes, $\bsX_0, \bsX_1, \ldots, \bsX_{n-1}$, for various sample sizes, $n$, and plot the relative errors, $\lvert (\mu - 	\hat{\mu}_n)/\mu\rvert$ in Figure \ref{fig:keister-err}.
+The three kind of nodes are
+\begin{enumerate}
+	\renewcommand{\labelenumi}{\roman{enumi}.}
+	\item Cartesian grids, $\{1/(2m), \ldots, (2m-1)/(2m) \}^6$, for $m = 2, 3, \ldots$ (blue),
+	\item IID simple random  with arbitrary $n$ (orange), and
+	\item Randomized low discrepancy (Sobol') sequences with $n = 1, 2, \ldots, 2^m, \ldots $ (green).
+\end{enumerate}
+
+\begin{figure}
+	\centering
+	\includegraphics[width =10cm]{\figpath/n_is_16777216_d_is_6_n_rep_is_50_KeisterErrors.pdf}
+	\caption{The relative error of approximationg $\mu$ defined in \eqref{eq:keisterB} by the sample mean, $\hat{\mu}_n$ defined in \eqref{eq:sample_mean} for various choices of nodes, $\bsX_0, \bsX_1, \ldots$.
+Grids have the largest error and low discrepancy (qMC) nodes have the smallest error.
+\label{fig:keister-err}}
+\end{figure}
+
+Note the following from this example:
+\begin{itemize}
+	\item This integral is not particularly easy.  Even the best choice of nodes requires at least $n=100$ to get a relative error below $10\%$.
+
+	\item Although grids may be attractive for low dimensional problems, e.g., $d = 1$, $2$, or $3$, they do poorly for this modest dimension, $d=6$.
+For modest $d$, the possible sample sizes, $n$, are quite sparse, as is shown in Figure \ref{fig:keister-err}.
+Choosing the nodes to lie on a grid corresponds to a tensor product midpoint cubature rule, which would normally be expected to have an error of $\mathcal{O}(n^{-2/d})$  for general $d$.
+The error decay of $\mathcal{O}(n^{-1/5})$ rather than $\mathcal{O}(n^{-2/6})$ for this example may be due to the lack of smoothness of $f$ near the boundaries of the unit cube.
+	\item Simple MC (orange) are a substantial improvement over grid nodes.
+Their root mean squared error is known to decay as $\mathcal{O}(n^{-1/2})$, as is observed here.
+As a reminder, for IID nodes the root mean squared error
+	\begin{equation}\label{eq:IIDerror}
+		\sqrt{\mathbb{E}[(\mu - \hat{\mu})^2]} = \sqrt{\var(\hat{\mu})} = \sqrt{\frac{\var(f(\bsX))}{n}} = \frac{\mathrm{Std}(f(\bsX))}{n^{1/2}},
+	\end{equation*}
+	where $\var$ denotes the variance and $\mathrm{Std}$ the standard deviation.  For simple MC
+\end{itemize}
+
 \section{Low Discrepancy Sequence Constructions} \label{sec:construct}
 \section{Discrepancy} \label{sec:discrepancy}
+\section{Randomization} \label{sec:random}
 \section{Stopping Criteria} \label{sec:stop}
@@
 \section{Conclusion} \label{sec:conclusion}
+
+\bibliographystyle{spmpsci}
+\bibliography{FJH23,FJHown23}
 

### 2025-01-08 HickernellMCQMC2024_QMCTutorial.tex
M2 Mac updates

@@
 
-	\item Although grids may be attractive for low dimensional problems, e.g., $d = 1$, $2$, or $3$, they do poorly for this modest dimension, $d=6$.
-For modest $d$, the possible sample sizes, $n$, are quite sparse, as is shown in Figure \ref{fig:keister-err}.
+	\item \emph{Grid nodes.} Although grids may be attractive for low dimensional problems, e.g., $d = 1$, $2$, or $3$, they do poorly for this modest dimension, $d=6$.
+For this modest $d$, the possible sample sizes, $n$, are quite sparse, as shown in Figure \ref{fig:keister-err}.
+
 	Choosing the nodes to lie on a grid corresponds to a tensor product midpoint cubature rule, which would normally be expected to have an error of $\mathcal{O}(n^{-2/d})$  for general $d$.
 The error decay of $\mathcal{O}(n^{-1/5})$ rather than $\mathcal{O}(n^{-2/6})$ for this example may be due to the lack of smoothness of $f$ near the boundaries of the unit cube.
-	\item Simple MC (orange) are a substantial improvement over grid nodes.
-Their root mean squared error is known to decay as $\mathcal{O}(n^{-1/2})$, as is observed here.
-As a reminder, for IID nodes the root mean squared error
+
+	\item \emph{IID nodes.}  Simple MC (orange) is a substantial improvement over grid nodes. As a reminder, for IID nodes the root mean squared error is
 	\begin{equation}\label{eq:IIDerror}
 		\sqrt{\mathbb{E}[(\mu - \hat{\mu})^2]} = \sqrt{\var(\hat{\mu})} = \sqrt{\frac{\var(f(\bsX))}{n}} = \frac{\mathrm{Std}(f(\bsX))}{n^{1/2}},
-	\end{equation*}
-	where $\var$ denotes the variance and $\mathrm{Std}$ the standard deviation.  For simple MC
+	\end{equation}
+	where $\var$ denotes the variance and $\mathrm{Std}$ the standard deviation.
+This $\mathcal{O}(n^{-1/2})$ decay is observed in Figure \ref{fig:keister-err}.
+For simple MC the sample size, $n$, can be any positive integer without affecting the rate of decay of the error.
+
+	\item \emph{Low discrepancy nodes.}  For qMC methods the error decays nearly like $\mathcal{O}(n^{-1})$, which for this example correponds to a reduction in error of several orders of magnitude for large enough $n$.
+Typically, low discrepancy sequences have preferred sample sizes.
+For this Sobol' sequence, $n$ an positive integer power of $2$ tends to give better accuracy than arbitrary $n$.
+
 \end{itemize}

### 2025-01-08 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

@@
 This reimagines the integral $I$ as the expectation of a function, $f$, of a standard uniform random variable:
-\begin{multline}\label{eq:keisterB}
+\begin{multline}\label{eq:keisterC}
 	\mu = \int_{[0,1]^d} \underbrace{\pi^{d/2} \cos\Bigl (\sqrt{2} \bigl \lVert \bigl( \Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d)\bigr) \bigr\rVert \Bigr)}_{=:f(\bsx)} \, \rmd \bsx \\
@@
 	\includegraphics[width =10cm]{\figpath/n_is_16777216_d_is_6_n_rep_is_50_KeisterErrors.pdf}
-	\caption{The relative error of approximationg $\mu$ defined in \eqref{eq:keisterB} by the sample mean, $\hat{\mu}_n$ defined in \eqref{eq:sample_mean} for various choices of nodes, $\bsX_0, \bsX_1, \ldots$.
+	\caption{The relative error of approximationg $\mu$ defined in \eqref{eq:keisterC} by the sample mean, $\hat{\mu}_n$ defined in \eqref{eq:sample_mean} for various choices of nodes, $\bsX_0, \bsX_1, \ldots$.
 Grids have the largest error and low discrepancy (qMC) nodes have the smallest error.
@@
 Typically, low discrepancy sequences have preferred sample sizes.
-For this Sobol' sequence, $n$ an positive integer power of $2$ tends to give better accuracy than arbitrary $n$.
+For this Sobol' sequence, $n$ a positive integer power of $2$ tends to give better accuracy than arbitrary $n$.
 

### 2025-01-08 HickernellMCQMC2024_QMCTutorial.tex
M2 Mac updates

@@
 
-	\item \emph{Low discrepancy nodes.}  For qMC methods the error decays nearly like $\mathcal{O}(n^{-1})$, which for this example correponds to a reduction in error of several orders of magnitude for large enough $n$.
+	\item \emph{Low discrepancy nodes.}  For qMC methods the error decays nearly like $\mathcal{O}(n^{-1})$, which for this example correponds to a reduction in error of several orders of magnitude compared to simple MC for large enough $n$.
 Typically, low discrepancy sequences have preferred sample sizes.
-For this Sobol' sequence, $n$ a positive integer power of $2$ tends to give better accuracy than arbitrary $n$.
+For this Sobol' sequence, $n$ equal to a positive integer power of $2$ tends to give better accuracy than arbitrary $n$.
 

### 2025-01-08 HickernellMCQMC2024_QMCTutorial.tex
M2 Mac updates

@@
 \begin{abstract}
-Many problems in  quantitative finance, uncertainty quantification, and other areas can be formulated as computing $\mu := \mathbb{E}(Y)$, where instances of $Y:=f(\boldsymbol{X})$ are generated by numerical simulation.
-The population mean, $\mu$, can be approximated by the sample mean, $\hat{\mu}_n := n^{-1} \sum_{i=1}^n f(\boldsymbol{X}_i)$.
+Many problems in  quantitative finance, uncertainty quantification, and other areas have answers of the form $\mu := \mathbb{E}(Y)$, where instances of $Y:=f(\boldsymbol{X})$ may be generated by numerical simulation.
+The population mean, $\mu$, can be approximated by the sample mean, $\hat{\mu}_n := n^{-1} \sum_{i=0}^{n-1} f(\boldsymbol{X}_i)$ for a well chosen sequence of nodes, $\{\bsX_0, \bsX_1, \ldots\}$.
 Computing $\mu$ is equivalent to computing a $d$-dimensional integral.
 
-Quasi-Monte Carlo methods replace independent and identically distributed  sequences of random vectors, $\{\boldsymbol{X}_1, \boldsymbol{X}_2, \ldots \}$, by low discrepancy sequences.
+Quasi-Monte Carlo methods replace independent and identically distributed  sequences of random vector nodes, $\{\boldsymbol{X}_i \}$, by low discrepancy sequences.
 This accelerates the convergence of $\hat{\mu}_n$ to $\mu$ as $n \to \infty$.
@@
 Moreover, we describe how to formulate problems to realize the most increase in performance using quasi-Monte Carlo methods.
-We also briefly describe the use of quasi-Monte Carlo methods for problems beyond computing the mean.
+We also briefly describe the use of quasi-Monte Carlo methods for problems beyond computing the mean, $\mu$.
 
@@
 	\item The porosity of rock, which affect the extraction of oil or gas, or
-	\item Elemetary particle interactions in a high energy physics experiment, which affect what is observed.
+	\item Elemetary particle interactions in a high energy physics experiment, which affect what is observed by detectors.
 \end{itemize}
-In such situations the unknown inputs are often modeled using random variables or stochastic processes.
+In such situations, the unknown inputs are often modeled using random variables or stochastic processes.
 Computations are performed by generating a multidude of possible scenarios informed by the assumed probability distribution of the input.
@@
 In mathematical terms, the outcome is $Y := f(\boldsymbol{X})$, where $\boldsymbol{X}$ is a vector random variable of inputs determining the outcome.
-The values of the function $f$ can be computed by an algorithm, whose complexity may be such that $f$ can be considered a ``black box''.
+The values of $f$ can be computed by an algorithm, whose complexity may be such that $f$ can be considered a ``black box''.
 The scenarios are $\bsX_0, \bsX_1, \ldots$, which give rise to outcomes $Y_0 = f(\bsX_0), Y_1 = f(\bsX_1), \ldots$.
 The $Y_i$ form the basis for approximating $\bbE(Y)$ and other quantities of interest.
+These scenarios are also commonly called instances of $\bsX$ or nodes.
+We use these terms interchangeably.
 
-Simple or independent and identically distributed (IID)  Monte Carlo chooses the sequence $\bsX_0, \bsX_1, \ldots$ to be IID.
-Quasi-Monte Carlo (qMC) methods choose $\bsX_0, \bsX_1, \ldots$ \emph{not} IID, but with an empirical distribution that approximates the true probability distribution of $\bsX$.
-The difference between these two distributions is called a \emph{discrepancy}, and the sequences used in qMC are called low discrepancy sequences.
+Simple or independent and identically distributed (IID)  Monte Carlo chooses the sequence of nodes $\{\bsX_0, \bsX_1, \ldots\}$ to be IID.
+Quasi-Monte Carlo (qMC) methods choose the nodes to be \emph{not} IID, but with an empirical distribution that approximates well the true probability distribution of $\bsX$.
+The difference between these two distributions is called a \emph{discrepancy}, and the node sequences used in qMC are called low discrepancy sequences.
 
 This tutorial describes what qMC is, why we would want to use qMC, and how qMC can be implemented well.
-The next session provides an illustration of qMC in practice.
-This is followed by a description of how low discrepancy sequences are constructed and various measures of quality.
+The next session provides an illustrative example of the advantages of qMC.
+This is followed by a description of how low discrepancy sequences are constructed and various measures of discrepancy.
 We then explain how randomization can improve qMC, how to decide what sample size, $n$, is sufficient, and how to rewrite the problem of interest in a qMC-friendly way.
@@
 where $2 \pi^{d/2}/\Gamma(d/2)$ is the surface area of the sphere in $d$ dimensions, and $\Gamma$ is the Gamma function.
-Having a non-trivial test case with a value that can be easily calculated allows us to compute the numerical errors of various cubature schemes.
+This non-trivial test case with a true value that can be easily calculated allows us to compute the numerical errors of various cubature schemes.
 
-The aim is approximate $\mu$ by a sample mean,
+The aim is to approximate $\mu$ by a sample mean,
 \begin{equation} \label{eq:sample_mean}
-	\hat{\mu}_n := \frac 1n \sum_{i=1}^n f(\boldsymbol{X}_i)
+	\hat{\mu}_n := \frac 1n \sum_{i=1}^n f(\boldsymbol{X}_i),
 \end{equation}
 but to do so requires some further preparation.
-This integral above, $\mu$, may be thought of as the expectation of $Y := g(\bsT) := \pi^{d/2} \cos(\lVert \bsT \rVert)$, where $T_1, \ldots, T_d$ are IID Gaussian random variables with zero mean and variance $1/2$, i.e., $\bsT \sim \calN(\bszero,\mathsf{I}/2)$:
+This integral above, $\mu$, may be thought of as the expectation of $Y := g(\bsT) := \pi^{d/2} \cos(\lVert \bsT \rVert)$, where $T_1, \ldots, T_d$ are IID Gaussian random variables with zero mean and variance $1/2$, i.e., $\bsT :=(T_1, \ldots, T_d) \sim \calN(\bszero,\mathsf{I}/2)$:
 \begin{equation}\label{eq:keisterB}
@@
 
-Since the low discrepancy sequences that drive qMC are defined on the unit cube, $[0,1]^d$, we perform a variable transformation $\bst = \bigl (\Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d) \bigr )/\sqrt{2}$, where is $\Phi$ is the cumulative distribution function of the standard Gaussian random variable.
-This reimagines the integral $I$ as the expectation of a function, $f$, of a standard uniform random variable:
+Since the low discrepancy sequences that underly qMC are defined to approximate the standard uniform distribution, we perform a variable transformation $\bst = \bigl (\Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d) \bigr )/\sqrt{2}$, where is $\Phi$ is the cumulative distribution function of the standard Gaussian random variable.
+This reimagines the integral $\mu$ as the expectation of a function, $f$, of a standard uniform random variable:
 \begin{multline}\label{eq:keisterC}
-	\mu = \int_{[0,1]^d} \underbrace{\pi^{d/2} \cos\Bigl (\sqrt{2} \bigl \lVert \bigl( \Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d)\bigr) \bigr\rVert \Bigr)}_{=:f(\bsx)} \, \rmd \bsx \\
+	\mu = \int_{[0,1]^d} \underbrace{\pi^{d/2} \cos\Bigl (\bigl \lVert \bigl( \Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d)\bigr) \bigr\rVert/\sqrt{2}  \Bigr)}_{=:f(\bsx)} \, \rmd \bsx \\
 	= \bbE[Y]
@@
 Consider the specific case of $d=6$ for which $\mu \approx -2.327303729298$.
-We approximate $\mu$ by the sample mean defined in \eqref{eq:sample_mean} for three different kinds of nodes, $\bsX_0, \bsX_1, \ldots, \bsX_{n-1}$, for various sample sizes, $n$, and plot the relative errors, $\lvert (\mu - 	\hat{\mu}_n)/\mu\rvert$ in Figure \ref{fig:keister-err}.
+We approximate $\mu$ by the sample mean defined in \eqref{eq:sample_mean} for three different kinds of nodes, $\{\bsX_i\}_{i=0}^{n-1}$, for various sample sizes, $n$, and plot the relative errors, $\lvert (\mu - 	\hat{\mu}_n)/\mu\rvert$ in Figure \ref{fig:keister-err}.
 The three kind of nodes are
@@
 	\item \emph{Grid nodes.} Although grids may be attractive for low dimensional problems, e.g., $d = 1$, $2$, or $3$, they do poorly for this modest dimension, $d=6$.
-For this modest $d$, the possible sample sizes, $n$, are quite sparse, as shown in Figure \ref{fig:keister-err}.
+Figure \ref{} shows a grid with $n = 2$, $3$, and $6$For this modest $d$, the possible sample sizes, $n$, are quite sparse, as shown in Figure \ref{fig:keister-err}.
 

### 2025-01-08 HickernellMCQMC2024_QMCTutorial.tex
M2 Mac updates

@@
 Many problems in  quantitative finance, uncertainty quantification, and other areas have answers of the form $\mu := \mathbb{E}(Y)$, where instances of $Y:=f(\boldsymbol{X})$ may be generated by numerical simulation.
-The population mean, $\mu$, can be approximated by the sample mean, $\hat{\mu}_n := n^{-1} \sum_{i=0}^{n-1} f(\boldsymbol{X}_i)$ for a well chosen sequence of nodes, $\{\bsX_0, \bsX_1, \ldots\}$.
+The population mean, $\mu$, can be approximated by the sample mean, $\hat{\mu}_n := n^{-1} \sum_{i=0}^{n-1} f(\bsx_i)$ for a well chosen sequence of nodes, $\{\bsx_0, \bsx_1, \ldots\}$.
 Computing $\mu$ is equivalent to computing a $d$-dimensional integral.
@@
 
-In mathematical terms, the outcome is $Y := f(\boldsymbol{X})$, where $\boldsymbol{X}$ is a vector random variable of inputs determining the outcome.
+In mathematical terms, the outcome is $Y := f(\bsX)$, where $\bsX$ is a vector random variable of inputs determining the outcome.
 The values of $f$ can be computed by an algorithm, whose complexity may be such that $f$ can be considered a ``black box''.
-The scenarios are $\bsX_0, \bsX_1, \ldots$, which give rise to outcomes $Y_0 = f(\bsX_0), Y_1 = f(\bsX_1), \ldots$.
-The $Y_i$ form the basis for approximating $\bbE(Y)$ and other quantities of interest.
+The scenarios are $\bsx_0, \bsx_1, \ldots$, which give rise to outcomes $y_0 = f(\bsx_0), y_1 = f(\bsx_1), \ldots$.
+The $y_i$ form the basis for approximating $\bbE(Y)$ and other quantities of interest.
 These scenarios are also commonly called instances of $\bsX$ or nodes.
@@
 
-Simple or independent and identically distributed (IID)  Monte Carlo chooses the sequence of nodes $\{\bsX_0, \bsX_1, \ldots\}$ to be IID.
+Simple or independent and identically distributed (IID)  Monte Carlo chooses the sequence of nodes $\{\bsx_0, \bsx_1, \ldots\}$ to be IID.
 Quasi-Monte Carlo (qMC) methods choose the nodes to be \emph{not} IID, but with an empirical distribution that approximates well the true probability distribution of $\bsX$.
-The difference between these two distributions is called a \emph{discrepancy}, and the node sequences used in qMC are called low discrepancy sequences.
+The difference between these two distributions is called a \emph{discrepancy}, and the node sequences used in qMC are called low discrepancy (LD) sequences.
 
@@
 The next session provides an illustrative example of the advantages of qMC.
-This is followed by a description of how low discrepancy sequences are constructed and various measures of discrepancy.
+This is followed by a description of how LD sequences are constructed and various measures of discrepancy.
 We then explain how randomization can improve qMC, how to decide what sample size, $n$, is sufficient, and how to rewrite the problem of interest in a qMC-friendly way.
@@
 \begin{equation} \label{eq:sample_mean}
-	\hat{\mu}_n := \frac 1n \sum_{i=1}^n f(\boldsymbol{X}_i),
+	\hat{\mu}_n := \frac 1n \sum_{i=1}^n f(\bsx_i),
 \end{equation}
-but to do so requires some further preparation.
-This integral above, $\mu$, may be thought of as the expectation of $Y := g(\bsT) := \pi^{d/2} \cos(\lVert \bsT \rVert)$, where $T_1, \ldots, T_d$ are IID Gaussian random variables with zero mean and variance $1/2$, i.e., $\bsT :=(T_1, \ldots, T_d) \sim \calN(\bszero,\mathsf{I}/2)$:
+but to do so requires some further preparation.  More on this preparation process is discussed in Section \ref{sec:reformulate}.
+
+This integral, $\mu$, may be thought of as the expectation of $Y := g(\bsT) := \pi^{d/2} \cos(\lVert \bsT \rVert)$, where $T_1, \ldots, T_d$ are IID Gaussian random variables with zero mean and variance $1/2$, i.e., $\bsT :=(T_1, \ldots, T_d) \sim \calN(\bszero,\mathsf{I}/2)$:
 \begin{equation}\label{eq:keisterB}
@@
 
-Since the low discrepancy sequences that underly qMC are defined to approximate the standard uniform distribution, we perform a variable transformation $\bst = \bigl (\Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d) \bigr )/\sqrt{2}$, where is $\Phi$ is the cumulative distribution function of the standard Gaussian random variable.
+Since the LD sequences that underly qMC are defined to approximate the standard uniform distribution, we perform a variable transformation $\bst = \bigl (\Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d) \bigr )/\sqrt{2}$, where is $\Phi$ is the cumulative distribution function of the standard Gaussian random variable.
 This reimagines the integral $\mu$ as the expectation of a function, $f$, of a standard uniform random variable:
@@
 Consider the specific case of $d=6$ for which $\mu \approx -2.327303729298$.
-We approximate $\mu$ by the sample mean defined in \eqref{eq:sample_mean} for three different kinds of nodes, $\{\bsX_i\}_{i=0}^{n-1}$, for various sample sizes, $n$, and plot the relative errors, $\lvert (\mu - 	\hat{\mu}_n)/\mu\rvert$ in Figure \ref{fig:keister-err}.
+We approximate $\mu$ by the sample mean defined in \eqref{eq:sample_mean} for three different kinds of nodes, $\{\bsx_i\}_{i=0}^{n-1}$, for various sample sizes, $n$, and plot the relative errors, $\lvert (\mu - 	\hat{\mu}_n)/\mu\rvert$ in Figure \ref{fig:keister-err}.
 The three kind of nodes are
@@
 	\item IID simple random  with arbitrary $n$ (orange), and
-	\item Randomized low discrepancy (Sobol') sequences with $n = 1, 2, \ldots, 2^m, \ldots $ (green).
+	\item Randomized LD (Sobol') sequences with $n = 1, 2, \ldots, 2^m, \ldots $ (green).
 \end{enumerate}
@@
 	\includegraphics[width =10cm]{\figpath/n_is_16777216_d_is_6_n_rep_is_50_KeisterErrors.pdf}
-	\caption{The relative error of approximationg $\mu$ defined in \eqref{eq:keisterC} by the sample mean, $\hat{\mu}_n$ defined in \eqref{eq:sample_mean} for various choices of nodes, $\bsX_0, \bsX_1, \ldots$.
-Grids have the largest error and low discrepancy (qMC) nodes have the smallest error.
+	\caption{The relative error of approximationg $\mu$ defined in \eqref{eq:keisterC} by the sample mean, $\hat{\mu}_n$ defined in \eqref{eq:sample_mean} for various choices of nodes, $\bsx_0, \bsx_1, \ldots$.
+Grids have the largest error and LD nodes have the smallest error.
 \label{fig:keister-err}}
@@
 \begin{itemize}
-	\item This integral is not particularly easy.  Even the best choice of nodes requires at least $n=100$ to get a relative error below $10\%$.
+	\item This integral is not particularly easy to evaluate numerically.  Even the best choice of nodes requires at least $n=100$ to get a relative error below $10\%$.
 
 	\item \emph{Grid nodes.} Although grids may be attractive for low dimensional problems, e.g., $d = 1$, $2$, or $3$, they do poorly for this modest dimension, $d=6$.
-Figure \ref{} shows a grid with $n = 2$, $3$, and $6$For this modest $d$, the possible sample sizes, $n$, are quite sparse, as shown in Figure \ref{fig:keister-err}.
+Figure \ref{fig:grid} compares a $64$ point grid with $d = 2$ and $6$.
+For $d = 6$, the possible sample sizes, $n$, are quite sparse, as shown in Figure \ref{fig:keister-err}.
 
@@
 	\begin{equation}\label{eq:IIDerror}
-		\sqrt{\mathbb{E}[(\mu - \hat{\mu})^2]} = \sqrt{\var(\hat{\mu})} = \sqrt{\frac{\var(f(\bsX))}{n}} = \frac{\mathrm{Std}(f(\bsX))}{n^{1/2}},
+		\sqrt{\mathbb{E}[(\mu - \hat{\mu})^2]} = \sqrt{\var(\hat{\mu})} = \sqrt{\frac{\var(f(\bsx))}{n}} = \frac{\mathrm{Std}(f(\bsx))}{n^{1/2}},
 	\end{equation}
@@
 
-	\item \emph{Low discrepancy nodes.}  For qMC methods the error decays nearly like $\mathcal{O}(n^{-1})$, which for this example correponds to a reduction in error of several orders of magnitude compared to simple MC for large enough $n$.
-Typically, low discrepancy sequences have preferred sample sizes.
-For this Sobol' sequence, $n$ equal to a positive integer power of $2$ tends to give better accuracy than arbitrary $n$.
+	Whereas grid points collapse on top of one another when viewed in low dimensional projections, are IID points may be seen when viewed in any lower dimensional projection, as seen in Figure \ref{fig:iid}.
+The disadvantage of IID points is that they form clusters and leave gaps.
+This is because the position of any one node is independent of the position of the others.
+
+	\item \emph{LD nodes.}  For qMC methods the error decays nearly like $\mathcal{O}(n^{-1})$, which for this example correponds to a reduction in error of several orders of magnitude compared to simple MC for large enough $n$.
+Typically, LD sequences have preferred sample sizes.
+For this Sobol' sequence, $n$ is chosen to be positive integer powers of $2$, which tends to give better accuracy than arbitrary $n$.
+
+	Figure \ref{fig:ld} shows typical two-dimensional projections of a $64$ LD node set.
+Visually, these nodes fill the unit cube better than IID nodes.
+A quantitative measure of this, the discrepancy, is defined in Section \ref{sec:discrepancy}.
+Constructions of LD node set are explained in the next section.
 
@@
 
-\section{Low Discrepancy Sequence Constructions} \label{sec:construct}
+\begin{figure}
+	\centering
+	\includegraphics[width = 0.48\textwidth]{\figpath/64gridpts_d2.pdf}\quad
+	%\includegraphics[width = 5cm]{\figpath/64gridpts_d3.pdf}\quad
+	\includegraphics[width =  0.48\textwidth]{\figpath/64gridpts_d6.pdf}
+	\caption{Although a two-dimensional grid covers the unit square rather well, a $d$-dimensional grid for modest $d$ does \emph{not} cover the unit cube well.
+For example, one can only see four distinct points in a two dimensional projection of a six-dimensional grid with $64$ points.
+\label{fig:grid}}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width =\textwidth]{\figpath/64iidpts_d6.pdf}
+	\caption{IID points cover the unit cube better than grid points, although one does observe clusters and gaps.
+In any  projection there is  a similar looking distribution of all $64$ points \label{fig:iid}}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width =\textwidth]{\figpath/64sobolpts_d6.pdf}
+	\caption{LD points cover the unit cube better even better than IID points or grids.  In any  projection there is  a similar looking distribution of all $64$ points. \label{fig:ld}}
+\end{figure}
+
+\section{LD Sequence Constructions} \label{sec:construct}
+This section introduces some of the most popular LD sequences.  What sets these apart from IID sequences is that the nodes are deliberately chosen and highly correlated to one another.
+
+\subsection{Lattice Sequences} \label{sec:lattice}
+The simplest LD construction is the family of good lattice points \cite{DicEtal22a,SloJoe94}.  As a finite node set, they are defined as follows:
+\begin{equation}
+	\bsx_i = i\bsh/n \pmod \bsone, \qquad i = 0, \ldots, n-1
+\end{equation}
+where $\bsh$ is a well chosen $d$-dimensional integer vector.
+Note that the set $\{\bsx_0, \ldots, \bsx_{n-1}\}$ is closed under addition modulo $\bsone$ so that it forms a group.
+Figure \ref{fig:latticeconstruct} illustrates this construction for $n = 16$ and $\bsh = (1,11)$.
+
+\begin{figure}
+	\centering
+	\includegraphics[width = \textwidth]{\figpath/16_lattice_construct_d2.pdf}
+	\caption{The construction of a good lattice point set     \label{fig:latticeconstruct}}
+\end{figure}
+
+One disadvantage of this construction is that it is not readily extensible.
+For the example in Figure \ref{fig:latticeconstruct} the first $8$ points do not fill the unit square.
+However, the set of points defined by even $i$ do a reasonable job.
+This suggests a method for defining lattice sequences that was proposed independently by \cite{Mai81a} and \cite{HicEtal00}.
+
+Let's start with the one-dimensional extensible lattice that is defined by the van der Corput sequence, $\phi_b(0), \phi_b(1), \ldots$.
+This involves reflecting the digits of the integer $i$ in base $b$ about the decimal point.
+For example, $i = 6 = 110_{2}$ give $x_6 = {}_20.011 = 3/8$, or in general
+\begin{equation}
+	i = i_0 + i_1b + i_2 b^2 + \cdots \implies x_i = \phi_b(i) := i_0 b^{-1} + i_1 b^{-2} + i_2 b^{-3} + \cdots \in [0,1).
+\end{equation}
+
 \section{Discrepancy} \label{sec:discrepancy}
@@
 \end{document}
-%
-In this section, we will consider the case when the Hamiltonian $H(x)$
-is autonomous. For the sake of simplicity, we shall also assume that it
-is $C^{1}$.
-
-We shall first consider the question of nontriviality, within the
-general framework of
-$\left(A_{\infty},B_{\infty}\right)$-subquadratic Hamiltonians. In
-the second subsection, we shall look into the special case when $H$ is
-$\left(0,b_{\infty}\right)$-subquadratic,
-and we shall try to derive additional information.
-%
-\subsubsection{The General Case: Nontriviality.}
-%
-We assume that $H$ is
-$\left(A_{\infty},B_{\infty}\right)$-sub\-qua\-dra\-tic at infinity,
-for some constant symmetric matrices $A_{\infty}$ and $B_{\infty}$,
-with $B_{\infty}-A_{\infty}$ positive definite. Set:
-\begin{eqnarray}
-\gamma :&=&{\rm smallest\ eigenvalue\ of}\ \ B_{\infty} - A_{\infty} \\
-  \lambda : &=& {\rm largest\ negative\ eigenvalue\ of}\ \
-  J \frac{d}{dt} +A_{\infty}\ .
-\end{eqnarray}
-
-Theorem~\ref{ghou:pre} tells us that if $\lambda +\gamma < 0$, the
-boundary-value problem:
-\begin{equation}
-\begin{array}{rcl}
-  \dot{x}&=&JH' (x)\\
-  x(0)&=&x (T)
-\end{array}
-\end{equation}
-has at least one solution
-$\overline{x}$, which is found by minimizing the dual
-action functional:
-\begin{equation}
-  \psi (u) = \int_{o}^{T} \left[\frac{1}{2}
-  \left(\Lambda_{o}^{-1} u,u\right) + N^{\ast} (-u)\right] dt
-\end{equation}
-on the range of $\Lambda$, which is a subspace $R (\Lambda)_{L}^{2}$
-with finite codimension. Here
-\begin{equation}
-  N(x) := H(x) - \frac{1}{2} \left(A_{\infty} x,x\right)
-\end{equation}
-is a convex function, and
-\begin{equation}
-  N(x) \le \frac{1}{2}
-  \left(\left(B_{\infty} - A_{\infty}\right) x,x\right)
-  + c\ \ \ \forall x\ .
-\end{equation}
-
-%
-\begin{proposition}
-Assume $H'(0)=0$ and $ H(0)=0$. Set:
-\begin{equation}
-  \delta := \liminf_{x\to 0} 2 N (x) \left\|x\right\|^{-2}\ .
-  \label{eq:one}
-\end{equation}
-
-If $\gamma < - \lambda < \delta$,
-the solution $\overline{u}$ is non-zero:
-\begin{equation}
-  \overline{x} (t) \ne 0\ \ \ \forall t\ .
-\end{equation}
-\end{proposition}
-%
-\begin{proof}
-Condition (\ref{eq:one}) means that, for every
-$\delta ' > \delta$, there is some $\varepsilon > 0$ such that
-\begin{equation}
-  \left\|x\right\| \le \varepsilon \Rightarrow N (x) \le
-  \frac{\delta '}{2} \left\|x\right\|^{2}\ .
-\end{equation}
-
-It is an exercise in convex analysis, into which we shall not go, to
-show that this implies that there is an $\eta > 0$ such that
-\begin{equation}
-  f\left\|x\right\| \le \eta
-  \Rightarrow N^{\ast} (y) \le \frac{1}{2\delta '}
-  \left\|y\right\|^{2}\ .
-  \label{eq:two}
-\end{equation}
-
-\begin{figure}
-\vspace{2.5cm}
-%%%You can un-comment the \includegraphics{} line below and comment the one above
-%%%if you have copied in your directory the file figure.eps
-%%%from Springer zip file. As mentioned in the preamble, in your manuscript you are encouraged to put
-%%%figures and images in a sub-directory which can be provided in the preamble via the
-%%%command \graphicspath{{<path 1>}{<path 2>}}
-%\includegraphics{figure.eps}
-\caption{This is the caption of the figure displaying a white eagle and
-a white horse on a snow field}
-\end{figure}
-
-Since $u_{1}$ is a smooth function, we will have
-$\left\|hu_{1}\right\|_\infty \le \eta$
-for $h$ small enough, and inequality (\ref{eq:two}) will hold,
-yielding thereby:
-\begin{equation}
-  \psi (hu_{1}) \le \frac{h^{2}}{2}
-  \frac{1}{\lambda} \left\|u_{1} \right\|_{2}^{2} + \frac{h^{2}}{2}
-  \frac{1}{\delta '} \left\|u_{1}\right\|^{2}\ .
-\end{equation}
-
-If we choose $\delta '$ close enough to $\delta$, the quantity
-$\left(\frac{1}{\lambda} + \frac{1}{\delta '}\right)$
-will be negative, and we end up with
-\begin{equation}
-  \psi (hu_{1}) < 0\ \ \ \ \ {\rm for}\ \ h\ne 0\ \ {\rm small}\ .
-\end{equation}
-
-On the other hand, we check directly that $\psi (0) = 0$. This shows
-that 0 cannot be a minimizer of $\psi$, not even a local one.
-So $\overline{u} \ne 0$ and
-$\overline{u} \ne \Lambda_{o}^{-1} (0) = 0$. \qed
-\end{proof}
-%
-\begin{corollary}
-Assume $H$ is $C^{2}$ and
-$\left(a_{\infty},b_{\infty}\right)$-subquadratic at infinity. Let
-$\xi_{1},\allowbreak\dots,\allowbreak\xi_{N}$  be the
-equilibria, that is, the solutions of $H' (\xi ) = 0$.
-Denote by $\omega_{k}$
-the smallest eigenvalue of $H'' \left(\xi_{k}\right)$, and set:
-\begin{equation}
-  \omega : = {\rm Min\,} \left\{\omega_{1},\dots,\omega_{k}\right\}\ .
-\end{equation}
-If:
-\begin{equation}
-  \frac{T}{2\pi} b_{\infty} <
-  - E \left[- \frac{T}{2\pi}a_{\infty}\right] <
-  \frac{T}{2\pi}\omega
-  \label{eq:three}
-\end{equation}
-then minimization of $\psi$ yields a non-constant $T$-periodic solution
-$\overline{x}$.
-\end{corollary}
-%
-
-We recall once more that by the integer part $E [\alpha ]$ of
-$\alpha \in \bbbr$, we mean the $a\in \bbbz$
-such that $a< \alpha \le a+1$. For instance,
-if we take $a_{\infty} = 0$, Corollary 2 tells
-us that $\overline{x}$ exists and is
-non-constant provided that:
-
-\begin{equation}
-  \frac{T}{2\pi} b_{\infty} < 1 < \frac{T}{2\pi}
-\end{equation}
-or
-\begin{equation}
-  T\in \left(\frac{2\pi}{\omega},\frac{2\pi}{b_{\infty}}\right)\ .
-  \label{eq:four}
-\end{equation}
-
-%
-\begin{proof}
-The spectrum of $\Lambda$ is $\frac{2\pi}{T} \bbbz +a_{\infty}$. The
-largest negative eigenvalue $\lambda$ is given by
-$\frac{2\pi}{T}k_{o} +a_{\infty}$,
-where
-\begin{equation}
-  \frac{2\pi}{T}k_{o} + a_{\infty} < 0
-  \le \frac{2\pi}{T} (k_{o} +1) + a_{\infty}\ .
-\end{equation}
-Hence:
-\begin{equation}
-  k_{o} = E \left[- \frac{T}{2\pi} a_{\infty}\right] \ .
-\end{equation}
-
-The condition $\gamma < -\lambda < \delta$ now becomes:
-\begin{equation}
-  b_{\infty} - a_{\infty} <
-  - \frac{2\pi}{T} k_{o} -a_{\infty} < \omega -a_{\infty}
-\end{equation}
-which is precisely condition (\ref{eq:three}).\qed
-\end{proof}
-%
-
-\begin{lemma}
-Assume that $H$ is $C^{2}$ on $\bbbr^{2n} \setminus \{ 0\}$ and
-that $H'' (x)$ is non-de\-gen\-er\-ate for any $x\ne 0$. Then any local
-minimizer $\widetilde{x}$ of $\psi$ has minimal period $T$.
-\end{lemma}
-%
-\begin{proof}
-We know that $\widetilde{x}$, or
-$\widetilde{x} + \xi$ for some constant $\xi
-\in \bbbr^{2n}$, is a $T$-periodic solution of the Hamiltonian system:
-\begin{equation}
-  \dot{x} = JH' (x)\ .
-\end{equation}
-
-There is no loss of generality in taking $\xi = 0$. So
-$\psi (x) \ge \psi (\widetilde{x} )$
-for all $\widetilde{x}$ in some neighbourhood of $x$ in
-$W^{1,2} \left(\bbbr / T\bbbz ; \bbbr^{2n}\right)$.
-
-But this index is precisely the index
-$i_{T} (\widetilde{x} )$ of the $T$-periodic
-solution $\widetilde{x}$ over the interval
-$(0,T)$, as defined in Sect.~2.6. So
-\begin{equation}
-  i_{T} (\widetilde{x} ) = 0\ .
-  \label{eq:five}
-\end{equation}
-
-Now if $\widetilde{x}$ has a lower period, $T/k$ say,
-we would have, by Corollary 31:
-\begin{equation}
-  i_{T} (\widetilde{x} ) =
-  i_{kT/k}(\widetilde{x} ) \ge
-  ki_{T/k} (\widetilde{x} ) + k-1 \ge k-1 \ge 1\ .
-\end{equation}
-
-This would contradict (\ref{eq:five}), and thus cannot happen.\qed
-\end{proof}
-%
-\paragraph{Notes and Comments.}
-The results in this section are a
-refined version of \cite{smit:wat};
-the minimality result of Proposition
-14 was the first of its kind.
-
-To understand the nontriviality conditions, such as the one in formula
-(\ref{eq:four}), one may think of a one-parameter family
-$x_{T}$, $T\in \left(2\pi\omega^{-1}, 2\pi b_{\infty}^{-1}\right)$
-of periodic solutions, $x_{T} (0) = x_{T} (T)$,
-with $x_{T}$ going away to infinity when $T\to 2\pi \omega^{-1}$,
-which is the period of the linearized system at 0.
-
-\begin{table}
-\caption{This is the example table taken out of {\it The
-\TeX{}book,} p.\,246}
-\begin{center}
-\begin{tabular}{r@{\quad}rl}
-\hline
-\multicolumn{1}{l}{\rule{0pt}{12pt}
-                   Year}&\multicolumn{2}{l}{World population}\\[2pt]
-\hline\rule{0pt}{12pt}
-8000 B.C.  &     5,000,000& \\
-  50 A.D.  &   200,000,000& \\
-1650 A.D.  &   500,000,000& \\
-1945 A.D.  & 2,300,000,000& \\
-1980 A.D.  & 4,400,000,000& \\[2pt]
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-%
-\begin{theorem} [Ghoussoub-Preiss]\label{ghou:pre}
-Assume $H(t,x)$ is
-$(0,\varepsilon )$-subquadratic at
-infinity for all $\varepsilon > 0$, and $T$-periodic in $t$
-\begin{equation}
-  H (t,\cdot )\ \ \ \ \ {\rm is\ convex}\ \ \forall t
-\end{equation}
-\begin{equation}
-  H (\cdot ,x)\ \ \ \ \ {\rm is}\ \ T{\rm -periodic}\ \ \forall x
-\end{equation}
-\begin{equation}
-  H (t,x)\ge n\left(\left\|x\right\|\right)\ \ \ \ \
-  {\rm with}\ \ n (s)s^{-1}\to \infty\ \ {\rm as}\ \ s\to \infty
-\end{equation}
-\begin{equation}
-  \forall \varepsilon > 0\ ,\ \ \ \exists c\ :\
-  H(t,x) \le \frac{\varepsilon}{2}\left\|x\right\|^{2} + c\ .
-\end{equation}
-
-Assume also that $H$ is $C^{2}$, and $H'' (t,x)$ is positive definite
-everywhere. Then there is a sequence $x_{k}$, $k\in \bbbn$, of
-$kT$-periodic solutions of the system
-\begin{equation}
-  \dot{x} = JH' (t,x)
-\end{equation}
-such that, for every $k\in \bbbn$, there is some $p_{o}\in\bbbn$ with:
-\begin{equation}
-  p\ge p_{o}\Rightarrow x_{pk} \ne x_{k}\ .
-\end{equation}
-\qed
-\end{theorem}
-%
-\begin{example} [{{\rm External forcing}}]
-Consider the system:
-\begin{equation}
-  \dot{x} = JH' (x) + f(t)
-\end{equation}
-where the Hamiltonian $H$ is
-$\left(0,b_{\infty}\right)$-subquadratic, and the
-forcing term is a distribution on the circle:
-\begin{equation}
-  f = \frac{d}{dt} F + f_{o}\ \ \ \ \
-  {\rm with}\ \ F\in L^{2} \left(\bbbr / T\bbbz; \bbbr^{2n}\right)\ ,
-\end{equation}
-where $f_{o} : = T^{-1}\int_{o}^{T} f (t) dt$. For instance,
-\begin{equation}
-  f (t) = \sum_{k\in \bbbn} \delta_{k} \xi\ ,
-\end{equation}
-where $\delta_{k}$ is the Dirac mass at $t= k$ and
-$\xi \in \bbbr^{2n}$ is a
-constant, fits the prescription. This means that the system
-$\dot{x} = JH' (x)$ is being excited by a
-series of identical shocks at interval $T$.
-\end{example}
-%
-\begin{definition}
-Let $A_{\infty} (t)$ and $B_{\infty} (t)$ be symmetric
-operators in $\bbbr^{2n}$, depending continuously on
-$t\in [0,T]$, such that
-$A_{\infty} (t) \le B_{\infty} (t)$ for all $t$.
-
-A Borelian function
-$H: [0,T]\times \bbbr^{2n} \to \bbbr$
-is called
-$\left(A_{\infty} ,B_{\infty}\right)$-{\it subquadratic at infinity}
-if there exists a function $N(t,x)$ such that:
-\begin{equation}
-  H (t,x) = \frac{1}{2} \left(A_{\infty} (t) x,x\right) + N(t,x)
-\end{equation}
-\begin{equation}
-  \forall t\ ,\ \ \ N(t,x)\ \ \ \ \
-  {\rm is\ convex\ with\  respect\  to}\ \ x
-\end{equation}
-\begin{equation}
-  N(t,x) \ge n\left(\left\|x\right\|\right)\ \ \ \ \
-  {\rm with}\ \ n(s)s^{-1}\to +\infty\ \ {\rm as}\ \ s\to +\infty
-\end{equation}
-\begin{equation}
-  \exists c\in \bbbr\ :\ \ \ H (t,x) \le
-  \frac{1}{2} \left(B_{\infty} (t) x,x\right) + c\ \ \ \forall x\ .
-\end{equation}
-
-If $A_{\infty} (t) = a_{\infty} I$ and
-$B_{\infty} (t) = b_{\infty} I$, with
-$a_{\infty} \le b_{\infty} \in \bbbr$,
-we shall say that $H$ is
-$\left(a_{\infty},b_{\infty}\right)$-subquadratic
-at infinity. As an example, the function
-$\left\|x\right\|^{\alpha}$, with
-$1\le \alpha < 2$, is $(0,\varepsilon )$-subquadratic at infinity
-for every $\varepsilon > 0$. Similarly, the Hamiltonian
-\begin{equation}
-H (t,x) = \frac{1}{2} k \left\|k\right\|^{2} +\left\|x\right\|^{\alpha}
-\end{equation}
-is $(k,k+\varepsilon )$-subquadratic for every $\varepsilon > 0$.
-Note that, if $k<0$, it is not convex.
-\end{definition}
-%
-
-\paragraph{Notes and Comments.}
-The first results on subharmonics were
-obtained by Rabinowitz in \cite{fo:kes:nic:tue}, who showed the existence of
-infinitely many subharmonics both in the subquadratic and superquadratic
-case, with suitable growth conditions on $H'$. Again the duality
-approach enabled Clarke and Ekeland in \cite{may:ehr:stein} to treat the
-same problem in the convex-subquadratic case, with growth conditions on
-$H$ only.
-
-Recently, Michalek and Tarantello (see \cite{fost:kes} and \cite{czaj:fitz})
-have obtained lower bound on the number of subharmonics of period $kT$,
-based on symmetry considerations and on pinching estimates, as in
-Sect.~5.2 of this article.
-
-%%% To ensure the bibliography has the correct style please run bibtex with the spmpsci style
-%%% which is included in the Springer zip file
-%%% Here we assume refs.bib would be the name of the bib file containing bibliographic info
-%%% You can then copy the .bbl produced, as given in the example below
-%%%
-%\bibliographystyle{spmpsci}
-%\bibliography{refs}
-%
-% ---- Bibliography ----
-%
-\begin{thebibliography}{6}
-%
-
-\bibitem {smit:wat}
-Smith, T.F., Waterman, M.S.: Identification of common molecular subsequences.
-J. Mol. Biol. 147, 195?197 (1981). \url{doi:10.1016/0022-2836(81)90087-5}
-
-\bibitem {may:ehr:stein}
-May, P., Ehrlich, H.-C., Steinke, T.: ZIB structure prediction pipeline:
-composing a complex biological workflow through web services.
-In: Nagel, W.E., Walter, W.V., Lehner, W. (eds.) Euro-Par 2006.
-LNCS, vol. 4128, pp. 1148?1158. Springer, Heidelberg (2006).
-\url{doi:10.1007/11823285_121}
-
-\bibitem {fost:kes}
-Foster, I., Kesselman, C.: The Grid: Blueprint for a New Computing Infrastructure.
-Morgan Kaufmann, San Francisco (1999)
-
-\bibitem {czaj:fitz}
-Czajkowski, K., Fitzgerald, S., Foster, I., Kesselman, C.: Grid information services
-for distributed resource sharing. In: 10th IEEE International Symposium
-on High Performance Distributed Computing, pp. 181?184. IEEE Press, New York (2001).
-\url{doi: 10.1109/HPDC.2001.945188}
-
-\bibitem {fo:kes:nic:tue}
-Foster, I., Kesselman, C., Nick, J., Tuecke, S.: The physiology of the grid: an open grid services architecture for distributed systems integration. Technical report, Global Grid
-Forum (2002)
-
-\bibitem {onlyurl}
-National Center for Biotechnology Information. \url{http://www.ncbi.nlm.nih.gov}
-
-\end{thebibliography}
-\end{document}

### 2025-01-08 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

@@
 	\includegraphics[width =10cm]{\figpath/n_is_16777216_d_is_6_n_rep_is_50_KeisterErrors.pdf}
-	\caption{The relative error of approximationg $\mu$ defined in \eqref{eq:keisterC} by the sample mean, $\hat{\mu}_n$ defined in \eqref{eq:sample_mean} for various choices of nodes, $\bsx_0, \bsx_1, \ldots$.
+	\caption{The relative error of approximating $\mu$ defined in \eqref{eq:keisterC} by the sample mean, $\hat{\mu}_n$ defined in \eqref{eq:sample_mean} for various choices of nodes, $\bsx_0, \bsx_1, \ldots$.
 Grids have the largest error and LD nodes have the smallest error.

### 2025-01-09 HickernellMCQMC2024_QMCTutorial.tex
M2 Mac updates

@@
 The simplest LD construction is the family of good lattice points \cite{DicEtal22a,SloJoe94}.  As a finite node set, they are defined as follows:
-\begin{equation}
-	\bsx_i = i\bsh/n \pmod \bsone, \qquad i = 0, \ldots, n-1
+\begin{equation} \label{eq:latticepts}
+	\bsx_i = i\bsh/n \pmod \bsone, \qquad i = 0, \ldots, n-1,
 \end{equation}
 where $\bsh$ is a well chosen $d$-dimensional integer vector.
-Note that the set $\{\bsx_0, \ldots, \bsx_{n-1}\}$ is closed under addition modulo $\bsone$ so that it forms a group.
 Figure \ref{fig:latticeconstruct} illustrates this construction for $n = 16$ and $\bsh = (1,11)$.
+Note that the set $\{\bsx_0, \ldots, \bsx_{n-1}\}$ defined in \eqref{eq:latticepts} is closed under addition modulo $\bsone$ so that it forms a group.
 
@@
 	\includegraphics[width = \textwidth]{\figpath/16_lattice_construct_d2.pdf}
-	\caption{The construction of a good lattice point set     \label{fig:latticeconstruct}}
+	\caption{The construction of a good lattice point set in two dimensions with $16$ nodes is obtained by moving $\bsh/n$ beyond the present point and wrapping around the boundaries of the square until one returns to the origin.
+\label{fig:latticeconstruct}}
 \end{figure}
@@
 
-Let's start with the one-dimensional extensible lattice that is defined by the van der Corput sequence, $\phi_b(0), \phi_b(1), \ldots$.
+Let's start with the one-dimensional extensible lattice that is defined by the van der Corput sequence, $\{\phi_b(0), \phi_b(1), \ldots\}$.
 This involves reflecting the digits of the integer $i$ in base $b$ about the decimal point.
-For example, $i = 6 = 110_{2}$ give $x_6 = {}_20.011 = 3/8$, or in general
+For example, $i = 6 = 110_{2}$ gives $\phi_2(6) = {}_20.011 = 3/8$, or in general,
 \begin{equation}
-	i = i_0 + i_1b + i_2 b^2 + \cdots \implies x_i = \phi_b(i) := i_0 b^{-1} + i_1 b^{-2} + i_2 b^{-3} + \cdots \in [0,1).
+	i = i_0 + i_1b + i_2 b^2 + \cdots \implies \phi_b(i) := i_0 b^{-1} + i_1 b^{-2} + i_2 b^{-3} + \cdots \in [0,1).
 \end{equation}
+The first $n = b^m$ nodes in the van der Corput sequence correspond to the evenly spaced nodes $\{0, b^{-m}, \ldots, 1 - b^{-m} \}$---albeit in a different order---for all positive integers $m$.
+
+To construct an extensible lattice, we replace $i/n$ in \eqref{eq:latticepts} by $\phi_b(i)$ to get
+\begin{equation} \label{eq:extensiblelattice}
+	\bsx_i =\phi_b( i)\bsh \pmod \bsone, \qquad i = 0, 1 , \ldots.
+\end{equation}
+This reordering of points from the original construction allows us to preserve the lattice structure for the first $b^m$ points for any non-negative integer $m$.
+That is, $\{\bsx_0, \ldots, \bsx_{2^m-1}\}$ is a closed under addition modulo $\bsone$.
+
+Figure \ref{fig:extensiblelatticeconstruct} shows the first $4$, $8$, and $16$ points of the earlier example in Figure \ref{fig:latticeconstruct}.
+The blue nodes are a copy of the nodes in the plot to the left, and the the orange nodes correspond to a shifted copy of the blue nodes modulo $\bsone$.
+For the left plot, the shift is $\phi_2(2)\bsh = (1,11)/4$ or equivalently $(0.25,0.75)$.
+For the middle plot, the shift is $\phi_2(4)\bsh = (1,11)/8$ or equivalently $(0.125,0.375)$.
+For the right plot, the shift is $\phi_2(8)\bsh = (1,11)/16$  or equivalently $(0.0625,0.6875)$.
+
+For the example in Figure \ref{fig:extensiblelatticeconstruct}, increasing $n$ beyond $16$ leads to repetition of the orginal points because the generator $\bsh$ has integers less than $16$.
+Obtaining a truly extensible lattice sequence that  the
+
+\begin{figure}
+	\centering
+	\includegraphics[width = \textwidth]{\figpath/16_extensible_lattice_construct_d2.pdf}
+	\caption{An extensible lattice corresponding to  Figure \ref{fig:latticeconstruct}  with the nodes reordered using the van der Corput sequence in base $2$.
+For each plot the blue nodes are a copy of the nodes to the left and the orange nodes are a shifted copy (modulo $\bsone$) of the blue nodes.
+\label{fig:extensiblelatticeconstruct}}
+\end{figure}
 

### 2025-01-09 HickernellMCQMC2024_QMCTutorial.tex
M2 Mac updates

@@
 \newcommand{\figpath}{Figures}
+\usepackage{xspace}
 
@@
 For example, $i = 6 = 110_{2}$ gives $\phi_2(6) = {}_20.011 = 3/8$, or in general,
-\begin{equation}
-	i = i_0 + i_1b + i_2 b^2 + \cdots \implies \phi_b(i) := i_0 b^{-1} + i_1 b^{-2} + i_2 b^{-3} + \cdots \in [0,1).
+\begin{equation} \label{eq:vdc}
+	\phi_b(i) := i_0 b^{-1} + i_1 b^{-2} + i_2 b^{-3} + \cdots \in [0,1) \quad \text{for }
+		i = i_0 + i_1b + i_2 b^2 + \cdots .
 \end{equation}
-The first $n = b^m$ nodes in the van der Corput sequence correspond to the evenly spaced nodes $\{0, b^{-m}, \ldots, 1 - b^{-m} \}$---albeit in a different order---for all positive integers $m$.
+For all non-negative integers, $m$, the first $n = b^m$ nodes in the van der Corput sequence correspond to the evenly spaced nodes $\{0, b^{-m}, \ldots, 1 - b^{-m} \}$---albeit in a different order.
 
@@
 
-For the example in Figure \ref{fig:extensiblelatticeconstruct}, increasing $n$ beyond $16$ leads to repetition of the orginal points because the generator $\bsh$ has integers less than $16$.
-Obtaining a truly extensible lattice sequence that  the
-
 \begin{figure}
@@
 \end{figure}
+
+For the example in Figure \ref{fig:extensiblelatticeconstruct}, increasing $n$ beyond $16$ repeats the orginal points because the generating vector $\bsh$ has integers less than $16$.
+Obtaining a truly extensible lattice sequence requires that $\bsh$ be a vector of generalized integers---essentially integers with infinite numbers of nonzero digits---as explained in \cite{HicNie03a}, where the existence of good generating vectors is also proved.
+In practice, one searches computationally for a $\bsh$ that produces good set of lattice nodes of size $n = b^m$ for a range of $m$ \cite{HicEtal00}.
+
+\subsection{Digital Sequences} \label{sec:digital}
+
+Another family of LD sequences that also builds from the van der Corput sequence of \eqref{eq:vdc} is digital sequences \cite{DicPil10a,Nie92}.
+For simplicity, we restrict ourselves to $b = 2$ and let $\oplus$ denote digitwise addition, e.g., $3/8 \oplus 3/4 = {}_20.011 \oplus {}_20.110 = {}_20.101 = 5/8$.
+A digital sequence is defined as
+\begin{equation} \label{eq:vdc}
+	\bsx_i := i_0 \bsc_1 \oplus i_1 \bsc_2 \oplus i_2 \bsc_3 \oplus \cdots \in [0,1)^d \quad \text{for }
+	i = i_0 + i_12 + i_2 2^2 + \cdots,
+\end{equation}
+where $\bsc_1, \bsc_2, \ldots$ are carefully chosen.
+If $c_{ij\ell}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ coordinate of $\bsc_i$, then in the literature one may see the generators of digital sequences stored as generating matrices, $\mathsf{C}_j = (c_{ij\ell})_{\ell,i = 1}^{M,N}$ for $j = 1, \ldots, d$, where $M$ is th maximum number of bits in the expression for $\bsx_i$, say $52$, and $2^N$ is the maximum number of nodes that is intended to be generated.
+
+Digital sequence generators can be found via number theory or numerical search.  The
 

### 2025-01-10 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf
M2 Mac updates

--- 2025-01-09 HickernellMCQMC2024_QMCTutorial.tex
@@
 \begin{abstract}
-Many problems in  quantitative finance, uncertainty quantification, and other areas have answers of the form $\mu := \mathbb{E}(Y)$, where instances of $Y:=f(\boldsymbol{X})$ may be generated by numerical simulation.
+Many problems in  quantitative finance, uncertainty quantification, and other disiplines have answers of the form $\mu := \mathbb{E}(Y)$, where instances of $Y:=f(\boldsymbol{X})$ may be generated by numerical simulation.
 The population mean, $\mu$, can be approximated by the sample mean, $\hat{\mu}_n := n^{-1} \sum_{i=0}^{n-1} f(\bsx_i)$ for a well chosen sequence of nodes, $\{\bsx_0, \bsx_1, \ldots\}$.
-Computing $\mu$ is equivalent to computing a $d$-dimensional integral.
+Computing $\mu$ is equivalent to computing a $d$-dimensional integral, $\int f(\bsx) \varrho(\bsx) \, \mathrm{d} \bsx$, where $\varrho$ is the probability density for $\bsX$.
 
-Quasi-Monte Carlo methods replace independent and identically distributed  sequences of random vector nodes, $\{\boldsymbol{X}_i \}$, by low discrepancy sequences.
+Quasi-Monte Carlo methods replace independent and identically distributed  sequences of random vector nodes, $\{\bsx_i \}$, by low discrepancy sequences.
 This accelerates the convergence of $\hat{\mu}_n$ to $\mu$ as $n \to \infty$.
@@
 	\item Future market forces, which affect financial risk,
-	\item The porosity of rock, which affect the extraction of oil or gas, or
+	\item The porosity of rock, which affects the extraction of oil or gas, or
 	\item Elemetary particle interactions in a high energy physics experiment, which affect what is observed by detectors.
 \end{itemize}
-In such situations, the unknown inputs are often modeled using random variables or stochastic processes.
-Computations are performed by generating a multidude of possible scenarios informed by the assumed probability distribution of the input.
+In such situations, the unknown inputs are often modeled using random vector variables or stochastic processes.
+Computations are performed by generating a multidude of possible outcomes informed by the assumed probability distribution of the input.
 These are used to estimate the mean, quantiles, and/or probability distribution of the outcome.
@@
 
-In mathematical terms, the outcome is $Y := f(\bsX)$, where $\bsX$ is a vector random variable of inputs determining the outcome.
-The values of $f$ can be computed by an algorithm, whose complexity may be such that $f$ can be considered a ``black box''.
-The scenarios are $\bsx_0, \bsx_1, \ldots$, which give rise to outcomes $y_0 = f(\bsx_0), y_1 = f(\bsx_1), \ldots$.
+In mathematical terms, the random outcome is $Y := f(\bsX)$, where $\bsX$ is a vector random input.
+Given a particular input value, $\bsx$, the corresponding outcome  $y = f(\bsx)$ can be computed by an algorithm, whose complexity may be such that $f$ can be considered a ``black box''.
+The computer experiment selects a sequence of input values or \emph{nodes} $\bsx_0, \bsx_1, \ldots$, which give rise to the observed outcomes $y_0 = f(\bsx_0), y_1 = f(\bsx_1), \ldots$.
 The $y_i$ form the basis for approximating $\bbE(Y)$ and other quantities of interest.
-These scenarios are also commonly called instances of $\bsX$ or nodes.
-We use these terms interchangeably.
 
-Simple or independent and identically distributed (IID)  Monte Carlo chooses the sequence of nodes $\{\bsx_0, \bsx_1, \ldots\}$ to be IID.
-Quasi-Monte Carlo (qMC) methods choose the nodes to be \emph{not} IID, but with an empirical distribution that approximates well the true probability distribution of $\bsX$.
+Simple or independent and identically distributed (IID)  Monte Carlo chooses the nodes $\{\bsx_0, \bsx_1, \ldots\}$ to be IID.
+Quasi-Monte Carlo (qMC) methods choose the nodes \emph{not} IID, but with an empirical distribution that approximates well the true probability distribution of $\bsX$.
 The difference between these two distributions is called a \emph{discrepancy}, and the node sequences used in qMC are called low discrepancy (LD) sequences.
@@
 
-The aim is to approximate $\mu$ by a sample mean,
+However, for the purpose of this illustration, we work with $\mu$ in its original form, \eqref{eq:keisterA} and approximate it by a sample mean,
 \begin{equation} \label{eq:sample_mean}
-	\hat{\mu}_n := \frac 1n \sum_{i=1}^n f(\bsx_i),
+	\hat{\mu}_n := \frac 1n \sum_{i=1}^n f(\bsx_i).
 \end{equation}
-but to do so requires some further preparation.  More on this preparation process is discussed in Section \ref{sec:reformulate}.
+This does require some further preparation, which is discussed in greater generality in Section \ref{sec:reformulate}.
 
-This integral, $\mu$, may be thought of as the expectation of $Y := g(\bsT) := \pi^{d/2} \cos(\lVert \bsT \rVert)$, where $T_1, \ldots, T_d$ are IID Gaussian random variables with zero mean and variance $1/2$, i.e., $\bsT :=(T_1, \ldots, T_d) \sim \calN(\bszero,\mathsf{I}/2)$:
+This $\mu$, may be thought of as the expectation of $Y := g(\bsT) := \pi^{d/2} \cos(\lVert \bsT \rVert)$, where $T_1, \ldots, T_d$ are IID Gaussian (normal) random variables with zero mean and variance $1/2$, i.e., $\bsT :=(T_1, \ldots, T_d) \sim \calN(\bszero,\mathsf{I}/2)$:
 \begin{equation}\label{eq:keisterB}
@@
 \end{equation}
-
-Since the LD sequences that underly qMC are defined to approximate the standard uniform distribution, we perform a variable transformation $\bst = \bigl (\Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d) \bigr )/\sqrt{2}$, where is $\Phi$ is the cumulative distribution function of the standard Gaussian random variable.
+Nearly all LD sequences, which underly qMC, are defined to approximate the standard uniform distribution, $\calU[0,1]^d$.
+Thus, we perform a variable transformation $\bst = \bigl (\Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d) \bigr )/\sqrt{2}$, where is $\Phi$ is the cumulative distribution function of the standard Gaussian random variable.
 This reimagines the integral $\mu$ as the expectation of a function, $f$, of a standard uniform random variable:
@@
 
+Now we can apply IID MC, qMC, and other computational methods to approximate $\mu$.
 Consider the specific case of $d=6$ for which $\mu \approx -2.327303729298$.
@@
 	\renewcommand{\labelenumi}{\roman{enumi}.}
-	\item Cartesian grids, $\{1/(2m), \ldots, (2m-1)/(2m) \}^6$, for $m = 2, 3, \ldots$ (blue),
-	\item IID simple random  with arbitrary $n$ (orange), and
+	\item Cartesian grids, $\{1/(2m), \ldots, (2m-1)/(2m) \}^6$, for $m = 2, 3, \ldots$ and $n = m^6$ (blue),
+	\item IID sequences with arbitrary $n$ (orange), and
 	\item Randomized LD (Sobol') sequences with $n = 1, 2, \ldots, 2^m, \ldots $ (green).
@@
 	\includegraphics[width =10cm]{\figpath/n_is_16777216_d_is_6_n_rep_is_50_KeisterErrors.pdf}
-	\caption{The relative error of approximating $\mu$ defined in \eqref{eq:keisterC} by the sample mean, $\hat{\mu}_n$ defined in \eqref{eq:sample_mean} for various choices of nodes, $\bsx_0, \bsx_1, \ldots$.
+	\caption{The relative error of approximating $\mu$ defined in \eqref{eq:keisterC} by the sample mean, $\hat{\mu}_n$, defined in \eqref{eq:sample_mean} for various choices of nodes, $\bsx_0, \bsx_1, \ldots$.
 Grids have the largest error and LD nodes have the smallest error.
@@
 	\begin{equation}\label{eq:IIDerror}
-		\sqrt{\mathbb{E}[(\mu - \hat{\mu})^2]} = \sqrt{\var(\hat{\mu})} = \sqrt{\frac{\var(f(\bsx))}{n}} = \frac{\mathrm{Std}(f(\bsx))}{n^{1/2}},
+		\sqrt{\mathbb{E}[(\mu - \hat{\mu})^2]} = \sqrt{\var(\hat{\mu})} = \sqrt{\frac{\var(f(\bsX))}{n}} = \frac{\mathrm{Std}(f(\bsX))}{n^{1/2}},
 	\end{equation}
@@
 
-	Whereas grid points collapse on top of one another when viewed in low dimensional projections, are IID points may be seen when viewed in any lower dimensional projection, as seen in Figure \ref{fig:iid}.
+	Whereas grid points collapse on top of one another when viewed in low dimensional projections (Figure \ref{fig:grid}), all IID points may be seen when viewed in any lower dimensional projection, as seen in Figure \ref{fig:iid}.
 The disadvantage of IID points is that they form clusters and leave gaps.
@@
 Typically, LD sequences have preferred sample sizes.
-For this Sobol' sequence, $n$ is chosen to be positive integer powers of $2$, which tends to give better accuracy than arbitrary $n$.
+For this Sobol' sequence, $n$ is chosen to be non-negative integer powers of $2$, which tends to give better accuracy than arbitrary $n$.
 
@@
 A quantitative measure of this, the discrepancy, is defined in Section \ref{sec:discrepancy}.
-Constructions of LD node set are explained in the next section.
+Constructions of LD node sequences are explained in the next section.
 
@@
 \section{LD Sequence Constructions} \label{sec:construct}
-This section introduces some of the most popular LD sequences.  What sets these apart from IID sequences is that the nodes are deliberately chosen and highly correlated to one another.
+This section introduces some of the most popular LD sequences.
+What sets these apart from IID sequences is that the nodes are deliberately chosen and highly correlated to one another, whether they be deterministic or randomized.
 
 \subsection{Lattice Sequences} \label{sec:lattice}
-The simplest LD construction is the family of good lattice points \cite{DicEtal22a,SloJoe94}.  As a finite node set, they are defined as follows:
+One of the simplest LD constructions is the family of good lattice points \cite{DicEtal22a,SloJoe94}.  As a finite node set, they are defined as follows:
 \begin{equation} \label{eq:latticepts}
@@
 One disadvantage of this construction is that it is not readily extensible.
-For the example in Figure \ref{fig:latticeconstruct} the first $8$ points do not fill the unit square.
+For the example in Figure \ref{fig:latticeconstruct} the first $8$ points do not fill the unit square well.
 However, the set of points defined by even $i$ do a reasonable job.
-This suggests a method for defining lattice sequences that was proposed independently by \cite{Mai81a} and \cite{HicEtal00}.
+This suggests a method for defining extensible lattice sequences that was proposed independently by \cite{Mai81a} and \cite{HicEtal00}.
 
@@
 This involves reflecting the digits of the integer $i$ in base $b$ about the decimal point.
-For example, $i = 6 = 110_{2}$ gives $\phi_2(6) = {}_20.011 = 3/8$, or in general,
-\begin{equation}
-	i = i_0 + i_1b + i_2 b^2 + \cdots \implies \phi_b(i) := i_0 b^{-1} + i_1 b^{-2} + i_2 b^{-3} + \cdots \in [0,1).
-\end{equation}
-The first $n = b^m$ nodes in the van der Corput sequence correspond to the evenly spaced nodes $\{0, b^{-m}, \ldots, 1 - b^{-m} \}$---albeit in a different order---for all positive integers $m$.
+For example, $\phi_2(6) = \phi_2(110_{2}) = {}_20.011 = 3/8$.
+In general,
+\begin{multline} \label{eq:vdc}
+	\phi_b(_0 + i_1b + i_2 b^2 + \cdots ) := i_0 b^{-1} + i_1 b^{-2} + i_2 b^{-3} + \cdots \in [0,1)
+	\\
+	 \text{where } i_0, i_1, \ldots \in \{0,\ldots b-1\}.
+\end{multline}
+For all non-negative integers, $m$, the first $n = b^m$ nodes in the van der Corput sequence correspond to the evenly spaced nodes $\{0, b^{-m}, \ldots, 1 - b^{-m} \}$---albeit in a different order.
 
@@
 
+For the example in Figure \ref{fig:extensiblelatticeconstruct}, increasing $n$ beyond $16$ repeats the orginal points because the generating vector $\bsh$ contains integers all less than $16$.
+Obtaining a truly extensible lattice sequence requires that $\bsh$ be a vector of generalized integers---essentially integers with infinite numbers of nonzero digits---as explained in \cite{HicNie03a}, where the existence of good generating vectors is also proved.
+In practice, one searches computationally for a $\bsh$ that produces good set of lattice nodes of size $n = b^m$ for a range of $m$ \cite{HicEtal00}.
+
+The construction of great generating vectors for lattices has attracted a great deal of interest.
+The component-by-component constructions pioneered by Frances Kuo, Josef Dick, Ian Sloan, Dirk Nuyens and others are surveyed in \cite[Chapter ?]{DicEtal22a}.
+LatNet builder \cite{LEcEtal22a,LatNet} is a software package by Pierre L'Ecuyer and collaborators, which constructs generators for lattices and digital sequences.
+
+\subsection{Digital Sequences} \label{sec:digital}
+
+Another family of LD sequences also built upon the van der Corput sequence \eqref{eq:vdc} is digital sequences \cite{DicPil10a,Nie92}.
+For simplicity, we restrict ourselves to $b = 2$ and let $\oplus$ denote binary digitwise addition modulo $2$ (also known as digitwise exclusive or), e.g., $3/8 \oplus 3/4 = {}_20.011 \oplus {}_20.110 = {}_20.101 = 5/8$.
+A digital sequence is defined as
+\begin{equation} \label{eq:digital}
+	\bsx_i := i_0 \bsx_1 \oplus i_1 \bsx_2 \oplus i_2 \bsx_4 \oplus \cdots \in [0,1)^d \quad \text{for }
+	i = i_0 + i_12 + i_2 2^2 + \cdots,
+\end{equation}
+where $\bsx_1, \bsx_2, \ldots$ are carefully chosen.  The node set $\{ \bsx_0, \ldots, \bsx_{2^m-1}\}$  for non-negative integer $m$ is called a digital net and is closed under $\oplus$.
+
+If $x_{ij\ell}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ coordinate of $\bsx_i$, then in the literature one may see the generators of digital sequences stored as generating matrices, $\mathsf{C}_j = (x_{ij\ell})_{\ell,i = 1}^{M,N}$ for $j = 1, \ldots, d$, where $M$ is th maximum number of bits in the expression for $\bsx_i$, say $52$, and $2^N$ is the maximum number of nodes that is intended to be generated.
+In this form
+\begin{equation} \label{eq:digitalB}
+	\overrightarrow{x}_{ij} = \mathsf{C}_j \overrightarrow{i} \quad \pmod{2}
+\end{equation}
+where $\overrightarrow{x}_{ij}$ denotes the binary digits of $x_{ij}$ and $\overrightarrow{i}$ denotes the binary digits of $i$.
+
+Imagine a box in $[0,1)^d$ of the form
+\begin{multline} \label{eq:element_box}
+	[a_1 b^{-k_1}, (a_1 + 1) 2^{-k_1}) \times \cdots \times [a_d b^{-k_d}, (a_d + 1) 2^{-k_d}), \\
+	a_j \in \{0, \ldots, 2^{k_j} - 1\}, \ k_j \in \{0,1, \ldots \}, \qquad j =1, \ldots d.
+\end{multline}
+This box has volume $2^{-(k_1 + \cdots +k_d)} = 2^{-\lVert\bsk\rVert_1}$.
+For a digital net with $2^m$ points and $m \ge \lVert \bsk \rVert_1$, a "fair share" of nodes would be for this box would be $2^{m- \lVert \bsk \rVert_1}$.
+This will occur if
+\begin{multline} \label{eq:tcond}
+	\text{the first $k_1$ rows of the first $m$ columns of $\mathsf{C}_1$ plus} \\
+	\text{the first $k_2$ rows of the first $m$ columns of $\mathsf{C}_2$ plus} \\
+	 \cdots \\
+	\text{the first $k_d$ rows of the first $m$ columns of $\mathsf{C}_d$} \\
+	\text{are linearly independent over the finite field with the elements $\{0,1\}$. }
+\end{multline}
+This above consdition is necessary and sufficient regardless of the choice of $\bsa = (a_1, \ldots, a_d)$.
+The $t$-value of a digital net with $2^m$ nodes is defined as the smallest $t$ for which condition \eqref{eq:tcond} holds for all $\bsk$ with $\lVert \bsk \rVert_1 \le m-t$.
+This digital net is then called a $(t,m,d)$-net.
+A digital sequence for which this condition holds for all non-negative $m$ is called a  $(t,d)$-sequence.
+
+To illustrate this idea of the $t$-value, consider the following digital net with $2^3$ (eight) nodes, i.e., $m = 3$:
+\begin{multline} \label{eq:smallnet}
+	\{(0, 0,   0),
+	(0.5,   0.5,   0.5  ),
+	(0.25,  0.75,  0.75 ),
+	(0.75,  0.25,  0.25 ),
+	(0.125, 0.625, 0.375),\\
+	(0.625, 0.125, 0.875),
+	(0.375, 0.375, 0.625),
+	(0.875, 0.875, 0.125)\}.
+\end{multline}
+Figure \ref{fig:elementinterval} whows two dimensional projections of these nodes and the two-dimensional boxes (rectangles) of the form \eqref{eq:element_box} with $\lVert\bsk\rVert_1 = 3$.
+For most of the boxes there is a fare share of one node, but the box $[0,1/4) \times [0,1) \times [0,1/2)$ contains no nodes and the adjacent box, $[1/4,1/2) \times [0,1) \times [0,1/2)$ contains no nodes.
+Thus, the $t$-value cannot be $0$.
+However, the $t$-value is $1$ because all boxes of the form  \eqref{eq:element_box} with $\lVert\bsk\rVert_1 = 2$, i.e., a volume of $2^{-2} = 1/4$ contain a fair share of $2^{m- \lVert \bsk \rVert_1} = 2$ points.
+The node set \eqref{eq:smallnet} is a $(1,3,3)$-net.
+
+This conclusion can also be reached by looking at the first three rows and columns of the generating matrices for this digital net, which are
+\begin{equation*}
+	\mathsf{C}_1 = \begin{pmatrix}
+		1 & 0 & 0 & \cdots \\ 0 & 1 & 0 & \cdots \\ 0 & 0 & 1 & \cdots \\ \vdots & \vdots & \vdots & \ddots
+	\end{pmatrix}, \quad
+	\mathsf{C}_2 = \begin{pmatrix}
+		1 & 1 & 1 & \cdots \\ 0 & 1 & 0 & \cdots \\ 0 & 0 & 1 & \cdots \\ \vdots & \vdots & \vdots & \ddots
+	\end{pmatrix}, \quad
+	\mathsf{C}_3 = \begin{pmatrix}
+	1 & 1 & 0 & \cdots \\ 0 & 1 & 1 & \cdots \\ 0 & 0 & 1 & \cdots \\ \vdots & \vdots & \vdots & \ddots
+\end{pmatrix}.
+\end{equation*}
+For $k_1=2$, $k_2 = 0$, and $k_3 = 1$ condition \eqref{eq:tcond} is not satisfied, and so $t$ cannot be $0$ for the node set  \eqref{eq:smallnet}.
+However, condition \eqref{eq:tcond} is satisfied for all $\bsk$ with $\lVert \bsk \rVert = 2$, again confirming that we have a $(1,3,3)$-net.
+
+If we consider only the first two coordinates of the node set defined in \eqref{eq:smallnet}, then  condition \eqref{eq:tcond} is satisfied for all $\bsk$ with $\lVert \bsk \rVert = 3$.
+The first two coordinates of \eqref{eq:smallnet} are a $(0,3,2)$-net
+
+\begin{figure}
+	\centering
+	\includegraphics[width = \textwidth]{\figpath/8_sobol_sequence_elementary_intervals_d3_coord_(1,2)}
+	\includegraphics[width = \textwidth]{\figpath/8_sobol_sequence_elementary_intervals_d3_coord_(1,3)}
+	\caption{Each box in the two dimensional projections of the node set \eqref{eq:smallnet} plotted above show one point in each box except for the second row, third plot from the left.
+Thus, this node set is not a $(0,3,3)$-net.
+It is however a $(1,3,3)$-net.
+\label{fig:elementinterval}}
+\end{figure}
+
+Digital sequence generators can be found via number theory or numerical search.  The
+
+(t,m,d)
+
+\section{Randomization} \label{sec:random}
+randomizations
+random generators
+
 \section{Discrepancy} \label{sec:discrepancy}
-\section{Randomization} \label{sec:random}
+\subsection{Discrepancies Defined by Kernels} \label{sec:kerdisc}
+\subsection{Geometrically Motivated Discrepancies} \label{sec:geodisc}
+\subsection{Coordinate weights} \label{sec:coordwts}
+\subsection{Discrepancies for Lattices and Nets} \label{sec:kerdisc}
+
 \section{Stopping Criteria} \label{sec:stop}
+Art
+
+Tony
+
+Jags
+
 \section{Reformulating Our Problem} \label{sec:reformulate}

### 2025-01-10 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

@@
 
-For the example in Figure \ref{fig:extensiblelatticeconstruct}, increasing $n$ beyond $16$ repeats the orginal points because the generating vector $\bsh$ contains integers all less than $16$.
+For the example in Figure \ref{fig:extensiblelatticeconstruct}, increasing $n$ beyond $16$ repeats the original points because the generating vector $\bsh$ contains integers all less than $16$.
 Obtaining a truly extensible lattice sequence requires that $\bsh$ be a vector of generalized integers---essentially integers with infinite numbers of nonzero digits---as explained in \cite{HicNie03a}, where the existence of good generating vectors is also proved.

### 2025-01-10 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

@@
 %
-\author{Fred J. Hickernell\inst{1} \and Aleksei G. Sorokin\inst{2}}
+\author{Fred J. Hickernell\inst{1} \and Nathan Kirk\inst{2} \and Aleksei G. Sorokin\inst{3}}
 %
-\authorrunning{Hickernell and Sorokin} % abbreviated author list (for running head)
+\authorrunning{Hickernell, Kirk, and Sorokin} % abbreviated author list (for running head)
 %
 %%%% list of authors for the TOC (use if author list has to be modified)
-\tocauthor{Fred J. Hickernell, Aleksei G. Sorokin}
+\tocauthor{Fred J. Hickernell, Nathan Kirk, Aleksei G. Sorokin}
 %
-\institute{Departmet of Applied Mathematics, Illinois Institute of Technology, Chicago, IL, 60616, USA\\
+\institute{
+Department of Applied Mathematics, Illinois Institute of Technology, \\ Chicago, IL, 60616, USA\\
 \email{hickernell@iit.edu}, \
@@
 \and
-Departmet of Applied Mathematics, Illinois Institute of Technology, Chicago, IL, 60616, USA\\
-\email{asorokin@hawk.iit.edu}}
+\email{nkirk@iit.edu}
+\and
+\email{asorokin@hawk.iit.edu}
+}
 

### 2025-01-12 HickernellMCQMC2024_QMCTutorial.tex
M2 Mac updates

@@
 Given a particular input value, $\bsx$, the corresponding outcome  $y = f(\bsx)$ can be computed by an algorithm, whose complexity may be such that $f$ can be considered a ``black box''.
-The computer experiment selects a sequence of input values or \emph{nodes} $\bsx_0, \bsx_1, \ldots$, which give rise to the observed outcomes $y_0 = f(\bsx_0), y_1 = f(\bsx_1), \ldots$.
-The $y_i$ form the basis for approximating $\bbE(Y)$ and other quantities of interest.
+The user selects a sequence of input values, also known as data sites or \emph{nodes}, $\bsx_0, \bsx_1, \ldots$, which give rise to the observed outcomes $y_0 = f(\bsx_0), y_1 = f(\bsx_1), \ldots$.
+The $y_i$ form the basis for approximating $\bbE(Y)$, quartiles of $Y$, the probability density of $Y$, and other quantities of interest.
 
-Simple or independent and identically distributed (IID)  Monte Carlo chooses the nodes $\{\bsx_0, \bsx_1, \ldots\}$ to be IID.
+Simple  Monte Carlo chooses the nodes $\{\bsx_0, \bsx_1, \ldots\}$ to be independent and identically distributed (IID).
 Quasi-Monte Carlo (qMC) methods choose the nodes \emph{not} IID, but with an empirical distribution that approximates well the true probability distribution of $\bsX$.
@@
 
-To illustrate the benefits of qMC, consider the example from Keister \cite[(13)]{Kei96}, motivated by computational physics:
+We illustrate the benefits of qMC with an example from Keister \cite[(13)]{Kei96}, motivated by computational physics:
 
@@
 Nearly all LD sequences, which underly qMC, are defined to approximate the standard uniform distribution, $\calU[0,1]^d$.
-Thus, we perform a variable transformation $\bst = \bigl (\Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d) \bigr )/\sqrt{2}$, where is $\Phi$ is the cumulative distribution function of the standard Gaussian random variable.
+Thus, we perform a variable transformation $bst = \bigl (\Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d) \bigr )/\sqrt{2}$,
+where is $\Phi$ is the cumulative distribution function of the standard Gaussian random variable.
 This reimagines the integral $\mu$ as the expectation of a function, $f$, of a standard uniform random variable:
@@
 
-Now we can apply IID MC, qMC, and other computational methods to approximate $\mu$.
+Now we can apply IID MC, qMC, and other computational methods to approximate $\mu$ by the sample mean
+\begin{equation} \label{eq:sample_mean_Keister}
+	\hat{\mu}_n = \frac 1n \sum_{i=1}^n pi^{d/2} \cos\Bigl (\bigl \lVert \bigl( \Phi^{-1}(x_{i1}), \ldots, \Phi^{-1}(x_{id})\bigr) \bigr\rVert/\sqrt{2}  \Bigr).
+\end{equation}
 Consider the specific case of $d=6$ for which $\mu \approx -2.327303729298$.
@@
 
-If $x_{ij\ell}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ coordinate of $\bsx_i$, then in the literature one may see the generators of digital sequences stored as generating matrices, $\mathsf{C}_j = (x_{ij\ell})_{\ell,i = 1}^{M,N}$ for $j = 1, \ldots, d$, where $M$ is th maximum number of bits in the expression for $\bsx_i$, say $52$, and $2^N$ is the maximum number of nodes that is intended to be generated.
-In this form
+If $x_{ij\ell}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ coordinate of $\bsx_i$, then the literature often defineds generating matrices, $\mathsf{C}_j = (x_{ij\ell})_{\ell,i = 1}^{M,N}$ for $j = 1, \ldots, d$, where $M$ is th maximum number of bits in the expression for $\bsx_i$, say $52$, and $2^N$ is the maximum number of nodes that is intended to be generated.
+The digital sequence can then be defined equivalently as
 \begin{equation} \label{eq:digitalB}
-	\overrightarrow{x}_{ij} = \mathsf{C}_j \overrightarrow{i} \quad \pmod{2}
+	\begin{pmatrix} x_{ij1} \\ x_{ij2} \\ \vdots \end{pmatrix}
+	= \mathsf{C}_j \begin{pmatrix} i_0 \\ i_1 \\ \vdots \end{pmatrix} \quad \pmod{2}, \qquad j = 1,\ldots, d, \ i = 0, 1, \ldots.
 \end{equation}
-where $\overrightarrow{x}_{ij}$ denotes the binary digits of $x_{ij}$ and $\overrightarrow{i}$ denotes the binary digits of $i$.
 
-Imagine a box in $[0,1)^d$ of the form
+To understand how these points might be evenly distributed over $[0,1)^d$, imagine a box of the form
 \begin{multline} \label{eq:element_box}
@@
 This box has volume $2^{-(k_1 + \cdots +k_d)} = 2^{-\lVert\bsk\rVert_1}$.
-For a digital net with $2^m$ points and $m \ge \lVert \bsk \rVert_1$, a "fair share" of nodes would be for this box would be $2^{m- \lVert \bsk \rVert_1}$.
-This will occur if
+For a digital net with $2^m$ points and $m \ge \lVert \bsk \rVert_1$, a ``fair share" of nodes for this box would be $2^{m- \lVert \bsk \rVert_1}$.
+This will occur if and only if
 \begin{multline} \label{eq:tcond}
@@
 	\text{the first $k_d$ rows of the first $m$ columns of $\mathsf{C}_d$} \\
-	\text{are linearly independent over the finite field with the elements $\{0,1\}$. }
+	\text{are linearly independent over the finite field with the elements $\{0,1\}$, }
 \end{multline}
-This above consdition is necessary and sufficient regardless of the choice of $\bsa = (a_1, \ldots, a_d)$.
+regardless of the choice of $\bsa = (a_1, \ldots, a_d)$.
 The $t$-value of a digital net with $2^m$ nodes is defined as the smallest $t$ for which condition \eqref{eq:tcond} holds for all $\bsk$ with $\lVert \bsk \rVert_1 \le m-t$.
-This digital net is then called a $(t,m,d)$-net.
-A digital sequence for which this condition holds for all non-negative $m$ is called a  $(t,d)$-sequence.
+Equivalently, this $t$ is the smallest value for which every box of the form \eqref{eq:element_box} with volume $2^{\lVert \bsk \rVert_1 - m}$ contains its fair share of $2^t$ nodes.
+Such a digital net is then called a $(t,m,d)$-net.
+An infinite sequence of the form \eqref{eq:digitalB} for which this condition holds for all non-negative $m$ is called a  $(t,d)$-sequence.
 
-To illustrate this idea of the $t$-value, consider the following digital net with $2^3$ (eight) nodes, i.e., $m = 3$:
+To illustrate the $t$-value, consider the following digital net with $2^3$ (eight) nodes, i.e., $m = 3$:
 \begin{multline} \label{eq:smallnet}
@@
 \end{multline}
-Figure \ref{fig:elementinterval} whows two dimensional projections of these nodes and the two-dimensional boxes (rectangles) of the form \eqref{eq:element_box} with $\lVert\bsk\rVert_1 = 3$.
-For most of the boxes there is a fare share of one node, but the box $[0,1/4) \times [0,1) \times [0,1/2)$ contains no nodes and the adjacent box, $[1/4,1/2) \times [0,1) \times [0,1/2)$ contains no nodes.
+Figure \ref{fig:elementinterval} shows several two dimensional projections of these nodes and the two-dimensional boxes (rectangles) of the form \eqref{eq:element_box} with $\lVert\bsk\rVert_1 = 3$.
+Most of the boxes contain their fair share of one node, but the box $[0,1/4) \times [0,1) \times [0,1/2)$ contains two nodes and the adjacent box, $[1/4,1/2) \times [0,1) \times [0,1/2)$ contains no nodes.
 Thus, the $t$-value cannot be $0$.
@@
 For $k_1=2$, $k_2 = 0$, and $k_3 = 1$ condition \eqref{eq:tcond} is not satisfied, and so $t$ cannot be $0$ for the node set  \eqref{eq:smallnet}.
-However, condition \eqref{eq:tcond} is satisfied for all $\bsk$ with $\lVert \bsk \rVert = 2$, again confirming that we have a $(1,3,3)$-net.
+However, condition \eqref{eq:tcond} is satisfied for all $\bsk$ with $\lVert \bsk \rVert_1 = 2$, again confirming that we have a $(1,3,3)$-net.
 
-If we consider only the first two coordinates of the node set defined in \eqref{eq:smallnet}, then  condition \eqref{eq:tcond} is satisfied for all $\bsk$ with $\lVert \bsk \rVert = 3$.
+If we consider only the first two coordinates of the node set defined in \eqref{eq:smallnet}, then  condition \eqref{eq:tcond} is satisfied for all $\bsk$ with $\lVert \bsk \rVert_1 = 3$.
 The first two coordinates of \eqref{eq:smallnet} are a $(0,3,2)$-net
@@
 
-Digital sequence generators can be found via number theory or numerical search.  The
+Digital sequence generators can be found via number theory or numerical search \cite{}.  The earliest instance is due to Sobol' \cite{}.
 
-(t,m,d)
+\subsection{Halton sequences} \label{sec:Halton}  While lattice sequences and digital sequences have preferred sample sizes, $n$, Halton sequences have no preferred sample size.
+The Halton sequence is defined in terms of the van der Corput sequence
+\begin{equation}\label{eq:Halton}
+	\bsx_i = \bigl( \phi_{b_1}(i), \ldots, \phi_{b_d}(i) \bigr), \qquad i = 0, 1, \ldots,
+\end{equation}
+where $b_1, \ldots, b_d$ is a choice of $d$ distinct prime numbers.  Often they are chosen as the first $d$ prime numbers.  Figure \ref{fig:Halton} shows two dimensional projections of Halton points
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{\figpath/64Halton_pts_d6.pdf}
+	\caption{Halton points as defined in \eqref{eq:Halton} have low discrepancy but no preferred sample size. \label{fig:Halton}}
+\end{figure}
 
 \section{Randomization} \label{sec:random}
+The LD constructions described in the previous section have so far been \emph{deterministic}. The sample mean, $\hat{\mu}_n$, does not change each time it is computed like it would for IID nodes.
+
+While determinism has some advantages randomization has advantages as well.
+\begin{itemize}
+	\item Randomization done right removes bias in the estimator.
+	\item Replications of random estimators can facilitate error estimates for the sample mean.
+	\item Often the application of interest requires transforming the LD nodes to mimic other distributions.
+This is the case in the Keister example of Section \ref{sec:intro}, as seen in \eqref{eq:sample_mean_Keister}.
+There the nodes, $\{\bsx_i\}_{i=0}^{n-1}$, are transformed to mimic a Gaussian distribution.
+
+	The LD sequences defined in the previous section start with $\bsx_0 = \bszero$, as can be seen in \eqref{eq:latticepts}, \eqref{eq:digital}, \eqref{eq:Halton}, and Figures \ref{fig:latticeconstruct}--\ref{fig:Halton}.
+The node $\bszero$ becomes infinite under a transformation to mimic a Gaussian distribution, which can trigger runtime errors.
+Randomization eliminates nodes on the boundary of the unit cube.
+\end{itemize}
+The key to good randomization is to preserve the LD quality of the node sequence.
+
+shifts
+scrambles
+
 randomizations

### 2025-01-13 HickernellMCQMC2024_QMCTutorial.tex
M2 Mac updates

@@
 \begin{multline} \label{eq:vdc}
-	\phi_b(_0 + i_1b + i_2 b^2 + \cdots ) := i_0 b^{-1} + i_1 b^{-2} + i_2 b^{-3} + \cdots \in [0,1)
+	\phi_b(i_0 + i_1b + i_2 b^2 + \cdots ) := i_0 b^{-1} + i_1 b^{-2} + i_2 b^{-3} + \cdots \in [0,1)
 	\\
@@
 
-shifts
-scrambles
+\subsection{Shifts} \label{sec:shifts}
+The simplest randomization is to shift the node sequence by $\bsDelta  \sim \calU[0,1)^d$.  For lattice  sequences the shift is applied modulo $\bsone$ so that shifted extensible lattice turns
+\eqref{eq:extensiblelattice} into
+\begin{equation} \label{eq:sh_extensiblelattice}
+	\bsx^{\bsDelta\text{-lat}}_i := \bsx_i + \bsDelta \pmod \bsone =\phi_b( i)\bsh + \bsDelta \pmod \bsone, \qquad i = 0, 1 , \ldots.
+\end{equation}
+Since the first $2^m$ elements of the lattice sequence form a group under modulo $\bsone$ addition, the coresponding shifted node set is a coset.
+Figure \ref{fig:shift_lat} illustrates three shifts of the lattice plotted in Figure \ref{fig:latticeconstruct}.
 
-randomizations
-random generators
+\begin{figure}
+	\centering
+	\includegraphics[width = \textwidth]{\figpath/16_shifted_lattice_construct_d2.pdf}
+	\caption{The original unshifted lattice (left) and three different shifts modulo $\bsone$ of the lattice.  Note that the structure of the lattice is maintained. \label{fig:shift_lat}}
+\end{figure}
+
+For digital sequences the shift should be applied using digitwise addition.  A digital shift of the digital sequence defined in  \eqref{eq:digital} then becomes
+\begin{equation} \label{eq:sh_digital}
+	\bsx_i^{\bsDelta\text{-dig}} := \bsx_i \oplus \bsDelta = i_0 \bsx_1 \oplus i_1 \bsx_2 \oplus i_2 \bsx_4 \oplus \cdots \oplus \bsDelta \in [0,1)^d \quad \text{for }
+	i = i_0 + i_12 + i_2 2^2 + \cdots.
+\end{equation}
+Three digital shifts of the same net are given in Figure \ref{fig:shift_net}.  A digital shift does not alter the $t$-value of a $(t,m,d)$-net.
+
+\begin{figure}
+	\centering
+	\includegraphics[width = \textwidth]{\figpath/n_8_n_sh_3_shifted_sobol_sequence_d2.pdf}
+	\caption{Three digital shifts of a digital $(0,3,2)$-net with nodes denoted by three different colors.
+There is one node in each smaller box of the form \eqref{eq:element_box} for $\lVert \bsk \rVert = 3$, for each digitally shifted net.
+\label{fig:shift_net}}
+\end{figure}
+
+Let $\{\bsx_i^{\bsDelta}\}_{i=0}^{n-1}$ denote a random shift of any deterministic original node set, $\{\bsx_i \}_{i=0}^{n-1}$ as described above.
+That is, $\bsx_i^{\bsDelta} : = \bsx_i + \bsDelta \pmod \bsone$ or  $\bsx_i^{\bsDelta} : = \bsx_i \oplus \bsDelta$ and $\bsDelta \sim \calU[0,1)^d$.
+This then implies that each $\bsx_i^{\bsDelta} \sim \calU[0,1)^d$ and thus $\hat{\mu}_n$ is an unbiased estimator of $\mu$, which was mentioned at the beginning of this section.
+
+\subsection{Digital Scrambles} \label{sec:scrambles}
+For digital nets one can randomize even further by a scrambling that preserves the $t$-value.
+Owen \cite{Owe95} proposed a scrambling of nets called nested uniform scrambling.
+A simpler version called linear matrix scrambling was proposed by Matou\v{s}ek \cite{Mat98} and is described here.
+Starting from the formulation of digital sequences involving generator matrices in \ref{eq:digitalB}, we multiply each generator matrix on the left by a lower triangular matrix with ones along the diagonals and elements below the diagonal that are randomly chosen to be $0$ or $1$ with equal probability:
+\begin{subequations} \label{eq:scr_lattice}
+\begin{gather} \label{eq:digitalscr}
+	\begin{pmatrix} x^{\text{scr}}_{ij1} \\ x^{\text{scr}}_{ij2} \\ \vdots \end{pmatrix}
+	= \mathsf{L_j} \mathsf{C}_j \begin{pmatrix} i_0 \\ i_1 \\ \vdots \end{pmatrix} + \begin{pmatrix} \Delta_{1j} \\ \Delta_{2j} \\ \vdots \end{pmatrix} \quad \pmod{2}, \qquad j = 1,\ldots, d, \ i = 0, 1, \ldots, \\
+	\mathsf{L_j} :=
+	\begin{pmatrix}
+		1 & 0 & 0 & 0 & \cdots \\
+		l_{21} & 1 & 0 & 0 & \cdots \\
+		l_{31} & l_{32} & 1 & 0 & \cdots \\
+		\vdots & \vdots & \vdots & \ddots & \ddots
+	\end{pmatrix}, \qquad l_{\ell j} \overset {\text{IID}}{\sim} \calU\{0,1\}
+\end{gather}
+\end{subequations}
+A random digital shift, $\bsDelta$ is also added, where $\Delta_{\ell j}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ of the shift.
+
+\begin{figure}
+	\centering
+	\includegraphics[width = \textwidth]{\figpath/n_8_n_scr_3_scrambled_sobol_sequence_d2.pdf}
+	\caption{Three linear scrambles of a digital $(0,3,2)$-net with nodes denoted by three different colors.
+There is one node in each smaller box of the form \eqref{eq:element_box} for $\lVert \bsk \rVert = 3$, for each digitally shifted net.
+\label{fig:scrambled_net}}
+\end{figure}
+
+[make a scramble of an image]
+
+\subsection{Other Randomizations for Halton}
+
+\subsection{Randomized Generators}
 
 \section{Discrepancy} \label{sec:discrepancy}
+So far, we have relied on eye tests and the Keister example to show how LD sequences are better than IID.
+This section introduces the theory that shows way minimizing discrepancy leads to better approximations to $\mu$.
+
 \subsection{Discrepancies Defined by Kernels} \label{sec:kerdisc}
+Let $K: [0,1)^d \times [0,1)^2 \to \mathbb{R}$ be a function satisfying the following two conditions:
+\begin{subequations}
+	\begin{align}
+		\text{Symmetry:} \quad & K(\bst,\bsx) = K(\bsx,\bst) \qquad \forall \bst,\bsx \in [0,1)^d \\
+		\nonumber
+		\text{Postitive definiteness:} \quad & \sum_{i,j = 0}^{n-1} c_i K(\bsx_i,\bsx_j) c_j > 0 \qquad  \forall n \in \mathbb{N}, \ \bsc \ne 0, \\
+		& \qquad \qquad  \ \text{distinct } \bsx_0, \ldots, \bsx_{n-1} \in [0,1)^d.
+	\end{align}
+\end{subequations}
+Then it is known that this $K$ is the reproducing kernel for a Hilbert space, $\calH_K$ with associated inner product $\langle \cdot, \cdot \rangle_{\calH_K}$, such that
+\begin{subequations}
+	\begin{align}
+	\text{Belonging:} \quad & K(\cdot,\bsx) \in \calH_K \qquad \forall \bsx \in [0,1)^d \\
+	\text{Reproducing:} \quad & f(\bsx) = \langle K(\cdot,\bsx),f \rangle \qquad  \forall \bsx \in [0,1)^d, \ f \in \calH_K.
+\end{align}
+\end{subequations}
+
+Having a reproducing kernel Hilbert space with a known kernel $K$ allows us to derive a rigorous error bound for $\mu - \hat{\mu}_n$.
+First note that for any bounded, linear functional, $L$ on the Hilbert space $\calH_K$, there is some $\eta_L \in \calH_K$ such that
+\begin{equation*}
+L(f) = \langle \eta_L , f \rangle_{\calH_K} \qquad \forall f \in \calH_K.
+\end{equation*}
+This is guaranteed by the Riesz Representation Theorm and $\eta_L$ is called the representer of $L$.  Using the reproducing property of $K$, one may evaluate
+\begin{equation*}
+\eta_L(\bsx) = \langle K(\cdot, \bsx), \eta_L \rangle_{\calH_K} = L\bigl( K(\cdot,\bsx) \bigr).
+\end{equation*}
+From an explict expression for $\eta_L(\bsx)$ one may then calculate the squared norm of $L$ as
+\begin{equation*}
+	\lVert \eta_L \rVert_{\calH_K}^2 = \langle\eta_L, \eta_L \rangle_{\calH_K} = L(\eta_L) = L^{\cdot\cdot} \Bigl(L^{\cdot}\bigl( K(\cdot,\cdot \cdot) \bigr) \Bigr).
+\end{equation*}
+
+If $K$ has reasonable properties, we may use the argument above to derive our error bound.
+First we write the dependence of $\mu - \hat{\mu}_n$ on  $f$ explicitly as $\mu(f) - \hat{\mu}_n(f)$ and note that the error functional, $\mu(\cdot) - \hat{\mu}_n(\cdot)$, is  linear.
+The Riesz Representation Theorem and the Cauchy Schawarz inequality imply a tight error bound:
+\begin{align}
+	\nonumber
+	\lvert\mu(f) - \hat{\mu}_n(f) \rvert
+	& = \lvert \langle \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)}, f \rangle_{\calH_K} \rvert \\
+	& \le \lVert  \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)} \rVert_{\calH_K} \, \lVert  f \rVert_{\calH_K} \qquad \forall f \in \calH_K
+	\intertext{where the squared norm of the error funtional can be expressed explicitly as}
+	\nonumber
+	\lVert  \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)} \rVert_{\calH_K}^2 &=
+	\bigl(\mu(\cdot\cdot) - \hat{\mu}_n(\cdot\cdot)\bigr) \Bigl(\bigl(\mu(\cdot) - \hat{\mu}_n(\cdot)\bigr)\bigl( K(\cdot,\cdot \cdot) \bigr) \Bigr) \\
+	\nonumber
+	& = \int_{[0,1)^d \times [0,1)^d} K(\bst,\bsx) \, \mathrm{d} \bst \, \mathrm{d} \bsx  -\frac 2n  \sum_{i=0}^{n-1} \int_{[0,1)^d} K(\bst,\bsx_i) \, \mathrm{d} \bst \\
+	& \qquad \qquad + \frac{1}{n^2} \sum_{i,j=0}^{n-1}  K(\bsx_i,\bsx_j).
+\end{align}
+
 \subsection{Geometrically Motivated Discrepancies} \label{sec:geodisc}
 \subsection{Coordinate weights} \label{sec:coordwts}
-\subsection{Discrepancies for Lattices and Nets} \label{sec:kerdisc}
+\subsection{Discrepancies for Lattices and Nets} \label{sec:latticenetdis}
 

### 2025-01-13 HickernellMCQMC2024_QMCTutorial.tex
M2 Mac updates
txs auto checkin

@@
 % Additional environment
-% \usepackage{amsmath}
+ \usepackage{mathtools}
 % \newtheorem{algorithm}{\upshape Algorithm}[1][]
@@
 \begin{abstract}
-Many problems in  quantitative finance, uncertainty quantification, and other disiplines have answers of the form $\mu := \mathbb{E}(Y)$, where instances of $Y:=f(\boldsymbol{X})$ may be generated by numerical simulation.
-The population mean, $\mu$, can be approximated by the sample mean, $\hat{\mu}_n := n^{-1} \sum_{i=0}^{n-1} f(\bsx_i)$ for a well chosen sequence of nodes, $\{\bsx_0, \bsx_1, \ldots\}$.
+Many problems in  quantitative finance, uncertainty quantification, and other disiplines are answered by computing the population mean, $\mu := \mathbb{E}(Y)$, where instances of $Y:=f(\boldsymbol{X})$ may be generated by numerical simulation and $\bsX$ has a simple probability  distribution.
+The population mean can be approximated by the sample mean, $\hat{\mu}_n := n^{-1} \sum_{i=0}^{n-1} f(\bsx_i)$ for a well chosen sequence of nodes, $\{\bsx_0, \bsx_1, \ldots\}$ and a sufficiently large sample size, $n$.
 Computing $\mu$ is equivalent to computing a $d$-dimensional integral, $\int f(\bsx) \varrho(\bsx) \, \mathrm{d} \bsx$, where $\varrho$ is the probability density for $\bsX$.
 
-Quasi-Monte Carlo methods replace independent and identically distributed  sequences of random vector nodes, $\{\bsx_i \}$, by low discrepancy sequences.
+Quasi-Monte Carlo methods replace independent and identically distributed  sequences of random vector nodes, $\{\bsx_i \}_{i = 0}^{\infty}$, by low discrepancy sequences.
 This accelerates the convergence of $\hat{\mu}_n$ to $\mu$ as $n \to \infty$.
@@
 We demonstrate the performance gains possible with quasi-Monte Carlo methods.
-Moreover, we describe how to formulate problems to realize the most increase in performance using quasi-Monte Carlo methods.
+Moreover, we describe how to formulate problems to realize the greatest performance gains using quasi-Monte Carlo methods.
 We also briefly describe the use of quasi-Monte Carlo methods for problems beyond computing the mean, $\mu$.
@@
 Simple  Monte Carlo chooses the nodes $\{\bsx_0, \bsx_1, \ldots\}$ to be independent and identically distributed (IID).
-Quasi-Monte Carlo (qMC) methods choose the nodes \emph{not} IID, but with an empirical distribution that approximates well the true probability distribution of $\bsX$.
+Quasi-Monte Carlo (qMC)  chooses the nodes \emph{not} IID, but with an empirical distribution that approximates well the true probability distribution of $\bsX$.
 The difference between these two distributions is called a \emph{discrepancy}, and the node sequences used in qMC are called low discrepancy (LD) sequences.
@@
 This tutorial describes what qMC is, why we would want to use qMC, and how qMC can be implemented well.
-The next session provides an illustrative example of the advantages of qMC.
-This is followed by a description of how LD sequences are constructed and various measures of discrepancy.
-We then explain how randomization can improve qMC, how to decide what sample size, $n$, is sufficient, and how to rewrite the problem of interest in a qMC-friendly way.
+The next session illustrates by example of the advantages of qMC.
+This is followed by a description of how deterministic LD sequences are constructed and randomizations that preserve their low discrepancy.
+We describe various measures of discrepancy.
+We explain how to decide what sample size, $n$, is sufficient to meet the user's error criterion.
+We then discuss how to rewrite the problem of interest in a qMC-friendly way.
 
@@
 where $2 \pi^{d/2}/\Gamma(d/2)$ is the surface area of the sphere in $d$ dimensions, and $\Gamma$ is the Gamma function.
+The resulting one-dimensional integral is amenable to quadrature methods \cite{}.
 This non-trivial test case with a true value that can be easily calculated allows us to compute the numerical errors of various cubature schemes.
@@
 	& = \lvert \langle \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)}, f \rangle_{\calH_K} \rvert \\
-	& \le \lVert  \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)} \rVert_{\calH_K} \, \lVert  f \rVert_{\calH_K} \qquad \forall f \in \calH_K
+	& \le \lVert  \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)} \rVert_{\calH_K} \, \lVert  f \rVert_{\calH_K} \qquad \forall f \in \calH_K,
 	\intertext{where the squared norm of the error funtional can be expressed explicitly as}
@@
 \section{Reformulating Our Problem} \label{sec:reformulate}
+\subsection{Variable Transformations}
+\subsection{Variation Reduction}
+\subsection{Multilevel Methods}
+
 \section{Ongoing Research} \label{sec:onging}
@@
 
+\section*{Appendix A}
+The integral $I_{\cos}(d) := \int_{0}^{\infty} \cos(r) \exp(-r^2) \, r^{d-1} \rmd r$ and its relative, the integral $I_{\sin}(d) := \int_{0}^{\infty} \sin(r) \exp(-r^2) \, r^{d-1} \rmd r$, may be evaluated iteratively for $d = 1, 2, \ldots$.
+Note that using integration by parts:
+\begin{align*}
+I_{\cos}(d) 
+	&=  \left .   \cos(r) \exp(-r^2) \, \frac{r^{d}}{d}  \right \rvert_0^{\infty} \\
+	& \qquad \qquad - \int_{0}^{\infty} [\cos(r) (-2r )\exp(-r^2) -\sin(r)  \exp(-r^2)] \,  \frac{r^{d}}{d} \rmd r \\
+	&= \frac{2 I_{\cos}(d+2) + I_{\sin}(d+1)}{d} \\
+	I_{\sin}(d) &=  \left .   \sin(r) \exp(-r^2) \, \frac{r^{d}}{d}  \right \rvert_0^{\infty} \\
+	&\qquad \qquad - \int_{0}^{\infty} [\sin(r) (-2r )\exp(-r^2) + \cos(r)  \exp(-r^2)] \,  \frac{r^{d}}{d} \rmd r \\
+&= \frac{2 I_{\sin}(d+2) - I_{\cos}(d+1)}{d} 
+\end{align*}
+This implies that
+\begin{align*}
+	I_{\cos}(d+2) & = \frac{d I_{\cos}(d) - I_{\sin}(d+1)}{2},   \\
+	I_{\sin}(d+2) &=   \frac{d I_{\sin}(d) + I_{\cos}(d+1)}{2},
+\end{align*}
+
 \end{document}

### 2025-01-13 HickernellMCQMC2024_QMCTutorial.tex
Fred M3 Mac Updates
txs auto checkin
Updates from Overleaf

@@
 
-If $x_{ij\ell}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ coordinate of $\bsx_i$, then the literature often defineds generating matrices, $\mathsf{C}_j = (x_{ij\ell})_{\ell,i = 1}^{M,N}$ for $j = 1, \ldots, d$, where $M$ is th maximum number of bits in the expression for $\bsx_i$, say $52$, and $2^N$ is the maximum number of nodes that is intended to be generated.
+If $x_{ij\ell}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ coordinate of $\bsx_i$, then the literature often defines generating matrices, $\mathsf{C}_j = (x_{ij\ell})_{\ell,i = 1}^{M,N}$ for $j = 1, \ldots, d$, where $M$ is th maximum number of bits in the expression for $\bsx_i$, say $52$, and $2^N$ is the maximum number of nodes that is intended to be generated.
 The digital sequence can then be defined equivalently as
@@
 
+Need to finish
+
 \end{document}

### 2025-01-14 HickernellMCQMC2024_QMCTutorial.tex
txs auto checkin

@@
 \end{subequations}
-Then it is known that this $K$ is the reproducing kernel for a Hilbert space, $\calH_K$ with associated inner product $\langle \cdot, \cdot \rangle_{\calH_K}$, such that
+Then it is known \cite{Aro50} that this $K$ is the reproducing kernel for a Hilbert space, $\calH_K$ with associated inner product $\langle \cdot, \cdot \rangle_{\calH_K}$, such that
 \begin{subequations}
@@
 \end{equation*}
-This is guaranteed by the Riesz Representation Theorm and $\eta_L$ is called the representer of $L$.  Using the reproducing property of $K$, one may evaluate
+This is guaranteed by the Riesz Representation Theorm, and $\eta_L$ is called the representer of $L$.
+Using the reproducing property of $K$, one may derove am explict formula for $\eta_L(\bsx)$, namely,
 \begin{equation*}
@@
 \end{equation*}
-From an explict expression for $\eta_L(\bsx)$ one may then calculate the squared norm of $L$ as
+From this expression for $\eta_L(\bsx)$ one may then calculate the squared norm of the linear functional $L$ as the squared norm of its representer:
 \begin{equation*}
@@
 
-If $K$ has reasonable properties, we may use the argument above to derive our error bound.
+If $\mu(\cdot)$ is a bounded linear functional on $\calH_K$ is, we may use the argument above to derive our error bound.
 First we write the dependence of $\mu - \hat{\mu}_n$ on  $f$ explicitly as $\mu(f) - \hat{\mu}_n(f)$ and note that the error functional, $\mu(\cdot) - \hat{\mu}_n(\cdot)$, is  linear.
-The Riesz Representation Theorem and the Cauchy Schawarz inequality imply a tight error bound:
+The Riesz Representation Theorem and the Cauchy-Schawarz inequality imply a tight error bound:
 \begin{align}
@@
 	\lvert\mu(f) - \hat{\mu}_n(f) \rvert
-	& = \lvert \langle \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)}, f \rangle_{\calH_K} \rvert \\
-	& \le \lVert  \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)} \rVert_{\calH_K} \, \lVert  f \rVert_{\calH_K} \qquad \forall f \in \calH_K,
-	\intertext{where the squared norm of the error funtional can be expressed explicitly as}
+	& = \lvert \langle \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)}, f \rangle_{\calH_K} \rvert
+	 \le \lVert  \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)} \rVert_{\calH_K} \, \lVert f \rVert_{\calH_K}\\
+	\lvert\mu(f) - \hat{\mu}_n(f) \rvert
+	& \le \underbrace{\lVert  \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)} \rVert_{\calH_K}}_{\text{discrepancy}\bigl(\{\bsx\}_{i=0}^{n-1},K \bigr)}
+	\, \underbrace{\inf_{c \in \RR}\lVert  f - c \rVert_{\calH_K}}_{\text{variation}(f,K)} \qquad \forall f \in \calH_K,  \label{eq:cuberrbd}
+	\intertext{since $\hat{\mu}_n$ is exact for constants.
+(It is assumed that $\calH$ contains constant functions so that $f \in \calH_K$ implies that $f - c \in \calH_K$.)  The squared norm of the error funtional can be expressed explicitly as}
 	\nonumber
@@
 \end{align}
+
+Error bound \eqref{eq:cuberrbd} has two factors:
+\begin{itemize}
+	\item the discrepancy, which is the norm of the error functional, is a measure of deficiency of the node set, $\{\bsx\}_{i=0}^{n-1}$, and
+	\item the variation is a measure of the roughness of the function defining our random variable whose expectation we wish to compute.
+\end{itemize}
+In general, designers of qMC methods want to construct sets or sequences of nodes with  discrepancy as small as possible, either by increasing $n$, if the computational budget allows, or by better placement of the nodes.
+The constructions described in Section \ref{sec:construct} are ways of ensuring better placement of the nodes, provided that the generators are chosen well.
+Practitioners of qMC want to formulate $\mu$ in a way to make the variation as small as possible.
+This discussed in Section \ref{sec:reformulate}.
+
+Although error bound \eqref{eq:cuberrbd} is elegant, it leaves several questions unanswered:
+\begin{itemize}
+	\item The reproducing kernel, $K$, defined both the discrepancy and the variation.
+This means that differnt choices of $K$ will lead to different error bounds for the same $\hat{\mu}_n$, even though the error is unchanged.
+	\item The Hilbert space
+\end{itemize}
 
@@
 
-Need to finish
+Need to finish and check against VS Code today
 

### 2025-01-14 HickernellMCQMC2024_QMCTutorial.tex
txs auto checkin

@@
 
-Although error bound \eqref{eq:cuberrbd} is elegant, it leaves several questions unanswered:
+Although error bound \eqref{eq:cuberrbd} is elegant, it leaves several matters unresolved:
 \begin{itemize}
-	\item The reproducing kernel, $K$, defined both the discrepancy and the variation.
+	\item The reproducing kernel $K$ defines both the discrepancy and the variation.
 This means that differnt choices of $K$ will lead to different error bounds for the same $\hat{\mu}_n$, even though the error is unchanged.
-	\item The Hilbert space
+	\item The Hilbert space $\calH_K$ contains ones assumptions about $f$, such as smoothness or periodicity.
+Knowing $K$, however, does not automatically lead to an explict formula for the inner product of the associated Hilbert space, $\langle \cdot, \cdot \rangle_{\calH_K}$.
+Conversely, having an explict formula for  $\langle \cdot, \cdot \rangle_{\calH_K}$ does not automatically lead to an explict formula for the reproducing kernel, $K$.
+In both cases educated guesswork is involved.
+	\item Choosing a more restrictive $\calH_K$ may lead to a sharper error bound, however, it is often infeasible in practice to check whether the ``black box'' $f$ lies in $\calH_K$.
+	\item   Moreover, even if one is confident that $f \in \calH_K$, it is usually impractical to compute variation$(f,K)$, as we shall see below.
+Error bound \eqref{eq:cuberrbd} cannot be used as a stopping criterion to determine the $n$ needed to satisfy an error tolerance.
 \end{itemize}
+Nonetheless \eqref{eq:cuberrbd}
 
@@
 \subsection{Discrepancies for Lattices and Nets} \label{sec:latticenetdis}
+
+\subsection{Randomized Error}
 

### 2025-01-15 HickernellMCQMC2024_QMCTutorial.tex
txs auto checkin

@@
 Let $K: [0,1)^d \times [0,1)^2 \to \mathbb{R}$ be a function satisfying the following two conditions:
-\begin{subequations}
+\begin{subequations} \label{eq:Kcond}
 	\begin{align}
+		\label{eq:KcondSym}
 		\text{Symmetry:} \quad & K(\bst,\bsx) = K(\bsx,\bst) \qquad \forall \bst,\bsx \in [0,1)^d \\
@@
 		\text{Postitive definiteness:} \quad & \sum_{i,j = 0}^{n-1} c_i K(\bsx_i,\bsx_j) c_j > 0 \qquad  \forall n \in \mathbb{N}, \ \bsc \ne 0, \\
-		& \qquad \qquad  \ \text{distinct } \bsx_0, \ldots, \bsx_{n-1} \in [0,1)^d.
+		& \qquad \qquad  \ \text{distinct } \bsx_0, \ldots, \bsx_{n-1} \in [0,1)^d. \label{eq:KcondPD}
 	\end{align}
@@
 \end{itemize}
-Nonetheless \eqref{eq:cuberrbd}
+In spite of these drawbacks, \eqref{eq:cuberrbd} is quite useful:
+\begin{itemize}
+	\item There are a family of commonly used reproducing kernels \cite{BerT-A04,Hic97a,Hic98b,Hic99b}.
+Once $K$ has been fixed, the discrepancy may be used as a comparison measure among different LD sequences, and its rate of decay---which sometimes can be computed theoretically---indicates how quickly $\hat{\mu}_n$ approaches $\mu$.
+	\item When an explict formula for the inner product of the associated Hilbert space, $\langle \cdot, \cdot \rangle_{\calH_K}$ can be deduced, it suggests how to $\mu$ to reduce variation$(f)$.
+\end{itemize}
+
+\subsection{The Centered Discrepancy}
+An example of a reproducing kernel defined on $[0,1)^d \times [0,1)^d$ is
+\begin{equation} \label{eq:ctrkernel}
+	K(\bst,\bsx) = \prod_{\ell = 1}^d \left[ 1 + \frac 12 \bigl ( \lvert t_\ell - 1/2 \rvert + \lvert x_\ell - 1/2 \rvert - \lvert t_\ell - x_\ell \rvert \bigr ) \right] .
+\end{equation}
+One can easily see that this $K$ satisfies the symmetry condition \eqref{eq:KcondSym}.  It also is symmetric about the middle of the unit cube, i.e.,
+\begin{equation*}
+	K\bigl((t_1,\ldots, t_{\ell-1}, 1 - t_\ell,  t_{\ell+1}, \ldots, t_d),
+	(x_1,\ldots, x_{\ell-1}, 1 - x_\ell,  x_{\ell+1}, \ldots, x_d)
+	\bigr) =
+	K(\bst,\bsx),
+\end{equation*}
+which is a desirable property for problems where there is no preferred corner.
+
+The inner product for the Hilbert space $\calH_K$ with reproducing kernel $K$ is
+\begin{align}
+	\langle f, g \rangle_{\calH_K} &= f(\boldsymbol{1/2}) g(\boldsymbol{1/2}) \nonumber
+	\\
+	&
+	+ \int_{[0,1)} [D^{\{1\}}f](x_1) \, [D^{\{1\}}g](x_1)\, \mathrm{d} x_1
+	+ \int_{[0,1)} [D^{\{2\}}f](x_2) \, [D^{\{2\}}g](x_2)\, \mathrm{d} x_2 \nonumber \\
+	& \qquad \qquad + \cdots \nonumber\\
+	&
+	+ \int_{[0,1)^2} [D^{{\{1,2\}}}f](x_1,x_2) \, [D^{\{1,2\}}g](x_1,x_2)\, \mathrm{d} x_1 \mathrm{d} x_2 + \cdots \nonumber \\
+	&  \qquad \qquad + \cdots + \int_{[0,1)^d} [D^{\{1,\ldots, d\}}f](\bsx) \, [D^{\{1,\ldots, d\}}g](\bsx) \, \mathrm{d} \bsx \nonumber \\
+	& = \sum_{\bsell \subseteq \{1, \ldots, d\}} \langle D^{\bsell}f, D^{\bsell}g \rangle_2
+\end{align}
+where $\langle \cdot,\cdot \rangle_2$ denotes the $L^2$ inner product,
+\begin{multline*}
+	[D^{\{\ell_1, \ell_2, \ldots, \ell_s\} }f](x_1, x_2, \ldots, x_s) \\
+	: = \frac{\partial^s f(1/2,\ldots, 1/2, x_{\ell_1}, 1/2, \ldots, 1/2, x_{\ell_2}, 1/2, \ldots, 1/2, x_{\ell_s}, 1/2, \ldots, 1/2)}{\partial x_{\ell_1} \partial x_{\ell_2} \cdots \partial x_s},
+\end{multline*}
+and $D^{\emptyset}f$ is understood to be the constant $f(1/2, \ldots, 1/2)$ \cite{Hic97a}.
+
+We show that $K$ defined in \eqref{eq:ctrkernel} is the reproducing kernel for the $\calH_K$ with the above inner product for $d=1$ and assuming $0 \le x \le 1/2$:
+\begin{align*}
+	\langle K(\cdot,x), f \rangle_{\calH_K}
+	& =
+	K(1/2,x) f(1/2) + \int_0^1 \frac{\partial K(t,x)}{\partial t} f'(t) \, \mathrm{d} t \\
+	& = 1 \times f(1/2) +  \int_0^x \frac 12 [-1  - (-1) ] f'(t) \, \mathrm{d} t +
+	\int_x^{1/2} \frac 12 [-1  - 1 ] f'(t) \, \mathrm{d} t \\
+	& \qquad \qquad +
+	\int_{1/2}^{1} \frac 12 [1  - 1 ] f'(t) \, \mathrm{d} t \\
+	& = f(1/2) + 0 - \int_x^{1/2} f'(t) \, \mathrm{d} t + 0 = f(1/2) - [f(1/2) - f(x)] \\
+	& =  f(x).
+\end{align*}
+The extension to general $1/2 < x < 1$ and to $d > 1$ is straightforward but a bit tedious.
+
+Figure
+
+\begin{figure}
+	\centering
+	\caption{A plot of the o
+\end{figure}
 

### 2025-01-16 HickernellMCQMC2024_QMCTutorial.tex
txs auto checkin
M2 Mac updates

@@
 %
+\setcounter{tocdepth}{2}
+\tableofcontents
+
 \section{Introduction} \label{sec:intro}
@@
 
+\subsection{LD Nodes by Optimization}\label{sec:optLD}
+[Nathan Kirk write here, please]
+
 \section{Randomization} \label{sec:random}
@@
 From this expression for $\eta_L(\bsx)$ one may then calculate the squared norm of the linear functional $L$ as the squared norm of its representer:
-\begin{equation*}
+\begin{equation} \label{eq:sqnormL}
 	\lVert \eta_L \rVert_{\calH_K}^2 = \langle\eta_L, \eta_L \rangle_{\calH_K} = L(\eta_L) = L^{\cdot\cdot} \Bigl(L^{\cdot}\bigl( K(\cdot,\cdot \cdot) \bigr) \Bigr).
-\end{equation*}
+\end{equation}
 
@@
 	\lvert\mu(f) - \hat{\mu}_n(f) \rvert
-	& \le \underbrace{\lVert  \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)} \rVert_{\calH_K}}_{\text{discrepancy}\bigl(\{\bsx\}_{i=0}^{n-1},K \bigr)}
-	\, \underbrace{\inf_{c \in \RR}\lVert  f - c \rVert_{\calH_K}}_{\text{variation}(f,K)} \qquad \forall f \in \calH_K,  \label{eq:cuberrbd}
+	& \le \underbrace{\lVert  \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)} \rVert_{\calH_K}}_{=:\text{discrepancy}\bigl(\{\bsx\}_{i=0}^{n-1},K \bigr)}
+	\, \underbrace{\inf_{c \in \RR}\lVert  f - c \rVert_{\calH_K}}_{=:\text{variation}(f,K)} \qquad \forall f \in \calH_K,  \label{eq:cuberrbd}
 	\intertext{since $\hat{\mu}_n$ is exact for constants.
-(It is assumed that $\calH$ contains constant functions so that $f \in \calH_K$ implies that $f - c \in \calH_K$.)  The squared norm of the error funtional can be expressed explicitly as}
+(It is assumed that $\calH$ contains constant functions so that $f \in \calH_K$ implies that $f - c \in \calH_K$.)  The squared norm of the error funtional can be expressed explicitly via \eqref{eq:sqnormL} as}
 	\nonumber
-	\lVert  \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)} \rVert_{\calH_K}^2 &=
+	\text{discrepancy}^2\bigl(\{\bsx\}_{i=0}^{n-1},K \bigr) & = \lVert  \eta_{\mu(\cdot) - \hat{\mu}_n(\cdot)} \rVert_{\calH_K}^2 \\
+	\nonumber
+	& =
 	\bigl(\mu(\cdot\cdot) - \hat{\mu}_n(\cdot\cdot)\bigr) \Bigl(\bigl(\mu(\cdot) - \hat{\mu}_n(\cdot)\bigr)\bigl( K(\cdot,\cdot \cdot) \bigr) \Bigr) \\
@@
 	& = \int_{[0,1)^d \times [0,1)^d} K(\bst,\bsx) \, \mathrm{d} \bst \, \mathrm{d} \bsx  -\frac 2n  \sum_{i=0}^{n-1} \int_{[0,1)^d} K(\bst,\bsx_i) \, \mathrm{d} \bst \\
-	& \qquad \qquad + \frac{1}{n^2} \sum_{i,j=0}^{n-1}  K(\bsx_i,\bsx_j).
+	& \qquad \qquad + \frac{1}{n^2} \sum_{i,j=0}^{n-1}  K(\bsx_i,\bsx_j). \label{eq:sqdisc}
 \end{align}
@@
 \end{align*}
-The extension to general $1/2 < x < 1$ and to $d > 1$ is straightforward but a bit tedious.
+The extensions to general $1/2 < x < 1$ and to $d > 1$ is straightforward but a bit tedious \cite{Hic97a}.
 
-Figure
+Figure \ref{fig:ctrKer} is a color map of the reproducing kernel, $K$, defined in \eqref{eq:ctrkernel} for $d=1$.
+As shown in the figure, the matrix $\mathsf{K} := \bigl( K(x_i,x_j) \bigr)_{i,j=1}^n$ that arises in condition \eqref{eq:KcondPD} is formed by taking $n$ rows and corresponding columns of this plot.
+Any such matrix $\mathsf{K}$ is strictly positive definite, as is suggested by the higher values near the diagonal.
 
@@
 	\centering
-	\caption{A plot of the o
+	\includegraphics[width = \textwidth]{\figpath/centered_discrepancy_kernel.pdf}
+	\caption{A plot of the reproducing kernel for the centered discrepancy defined in \eqref{eq:ctrkernel} for $d=1$.
+Note that the vertical axis is inverted so that the plot is a picture of the ``matrix'' $\bigl( K(t,x) \bigr)_{t,x \in [0,1)]}$.
+Any submatrix is positive definite by property \eqref{eq:KcondPD}.
+\label{fig:ctrKer}}
 \end{figure}
 
-\subsection{Geometrically Motivated Discrepancies} \label{sec:geodisc}
+The squared centered discrepancy is given by substituting the formula for $K$ into \eqref{eq:sqdisc}, which implies
+\begin{align} \label{eq:ctrdisc}
+	\nonumber
+	\MoveEqLeft\text{ctr-discrepancy}^2\bigl(\{ \bsx_i\}_{i=0}^{n-1} \bigr) \\
+	\nonumber
+	&= \left ( \frac{13}{12} \right )^d
+	- \frac 2n \sum_{i=0}^{n-1} \prod_{\ell=1}^d \left [ 1 + \frac 12 \bigl ( \lvert x_{i\ell} - 1/2 \rvert - \lvert x_{i\ell} - 1/2 \rvert^d \bigr )\right ] \\
+	& \qquad \qquad \frac{1}{n^2} \sum_{i,j=0}^{n-1} \prod_{\ell = 1}^d \left[ 1 + \frac 12 \bigl ( \lvert x_{i\ell} - 1/2 \rvert + \lvert x_{j\ell} - 1/2 \rvert - \lvert x_{i\ell} - x_{j\ell} \rvert \bigr ) \right].
+\end{align}
+As is true for most cases, the computational cost of computing the discrepancy is $\Order(dn^2)$ as $d$ and/or $n$ tend to infinity.
+
+Figure \ref{fig:ctrdiscmodestd} displays the discrepancy for a Sobol' LD digital net for several modest values of $d$ and for $n =  1, 2, 4, \ldots$.
+The decay rate of nearly $\calO(n^{-1})$ is observed for small $d$ and can be anticipated for a bit larger $d$.
+Thus, the decay of the discrepancy mirrors the convergenc rate for LD sequences for the Kiester example in Section \ref{sec:intro} and Figure \ref{fig:keister-err}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width = \textwidth]{\figpath/SobolCtrDisc_n_8192_d8.pdf}
+	\caption{The centered discrepancy for several modest values of $d$.
+They show an asymptotic decay close to $\calO(n^{-1})$, but this rate is achieved only at higher $n$ for larger values of $d$ \label{fig:ctrdiscmodestd}}
+\end{figure}
+
+The centered discrepancy also has a geometric interpretation, as is described in \cite{Hic97a}. Other choices of kernels and their geomeetric interpretations are explained there as well.
+
 \subsection{Coordinate weights} \label{sec:coordwts}
@@
 \end{document}
+
+\iffalse
+\subsection{Geometrically Motivated Discrepancies} \label{sec:geodisc}
+
+The centered discrepancy also has a geometric explanation, as is explained in \cite{Hic97a}.
+Let's explain for $d=1$, which means that we are measuring the discrepancy of points on $[0,1)$.
+Consider the line segment between an arbitrary point $x$ and closesr end of the line segment.
+The length of that line segment is $\lvert \ind_{[1/2,1)}(x) - x \rvert$, ,where $\ind$ is the indicator function.
+The fraction of nodes that lie inside this line segment is
+\begin{equation*}
+	\frac 1n \sum_{i=0}^{n-1} \bigl \lvert \ind_{[1/2,1)}(x) - \ind_{[x_i,1)}(x) \bigr \rvert
+\end{equation*}
+A measure of discrepancy is the root mean square integral of the difference between the `fare share value'' and the actual value, i.e.,
+\begin{align*}
+	\text{geo-discrepancy}\bigl( \{x_i\|_{i=0}^{n-1}\}\bigr) & =
+	\int_0^1 \left [ \lvert \ind_{[1/2,1)}(x) - x \rvert - \frac 1n \sum_{i=0}^{n-1} \bigl \lvert \ind_{[1/2,1)}(x) - \ind_{[x_i,1)}(x) \bigr \rvert \right ] ^2
+\end{align*}
+
+\fi

### 2025-01-16 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

@@
 This tutorial describes what qMC is, why we would want to use qMC, and how qMC can be implemented well.
-The next session illustrates by example of the advantages of qMC.
+The next session illustrates by example the advantages of qMC.
 This is followed by a description of how deterministic LD sequences are constructed and randomizations that preserve their low discrepancy.
@@
 
-We illustrate the benefits of qMC with an example from Keister \cite[(13)]{Kei96}, motivated by computational physics:
+We illustrate the benefits of qMC with an example from Keister \cite{Kei96}, motivated by computational physics:
 
@@
 
-This $\mu$, may be thought of as the expectation of $Y := g(\bsT) := \pi^{d/2} \cos(\lVert \bsT \rVert)$, where $T_1, \ldots, T_d$ are IID Gaussian (normal) random variables with zero mean and variance $1/2$, i.e., $\bsT :=(T_1, \ldots, T_d) \sim \calN(\bszero,\mathsf{I}/2)$:
+This $\mu$ may be thought of as the expectation of $Y := g(\bsT) := \pi^{d/2} \cos(\lVert \bsT \rVert)$, where $T_1, \ldots, T_d$ are IID Gaussian (normal) random variables with zero mean and variance $1/2$, i.e., $\bsT :=(T_1, \ldots, T_d) \sim \calN(\bszero,\mathsf{I}/2)$:
 \begin{equation}\label{eq:keisterB}
@@
 \end{equation}
-Nearly all LD sequences, which underly qMC, are defined to approximate the standard uniform distribution, $\calU[0,1]^d$.
-Thus, we perform a variable transformation $bst = \bigl (\Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d) \bigr )/\sqrt{2}$,
+Nearly all LD sequences, which underlie qMC, are defined to approximate the standard uniform distribution, $\calU[0,1]^d$.
+Thus, we perform a variable transformation $\bst = \bigl (\Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d) \bigr )/\sqrt{2}$,
 where is $\Phi$ is the cumulative distribution function of the standard Gaussian random variable.
@@
 \begin{equation} \label{eq:sample_mean_Keister}
-	\hat{\mu}_n = \frac 1n \sum_{i=1}^n pi^{d/2} \cos\Bigl (\bigl \lVert \bigl( \Phi^{-1}(x_{i1}), \ldots, \Phi^{-1}(x_{id})\bigr) \bigr\rVert/\sqrt{2}  \Bigr).
+	\hat{\mu}_n = \frac 1n \sum_{i=1}^n \pi^{d/2} \cos\Bigl (\bigl \lVert \bigl( \Phi^{-1}(x_{i1}), \ldots, \Phi^{-1}(x_{id})\bigr) \bigr\rVert/\sqrt{2}  \Bigr).
 \end{equation}
@@
 
-	Figure \ref{fig:ld} shows typical two-dimensional projections of a $64$ LD node set.
+	Figure \ref{fig:ld} shows typical two-dimensional projections of a size $64$ LD node set.
 Visually, these nodes fill the unit cube better than IID nodes.
@@
 	\includegraphics[width =\textwidth]{\figpath/64sobolpts_d6.pdf}
-	\caption{LD points cover the unit cube better even better than IID points or grids.  In any  projection there is  a similar looking distribution of all $64$ points. \label{fig:ld}}
+	\caption{LD points cover the unit cube even better than IID points or grids.  In any  projection there is  a similar looking distribution of all $64$ points. \label{fig:ld}}
 \end{figure}
@@
 
-If $x_{ij\ell}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ coordinate of $\bsx_i$, then the literature often defines generating matrices, $\mathsf{C}_j = (x_{ij\ell})_{\ell,i = 1}^{M,N}$ for $j = 1, \ldots, d$, where $M$ is th maximum number of bits in the expression for $\bsx_i$, say $52$, and $2^N$ is the maximum number of nodes that is intended to be generated.
+If $x_{ij\ell}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ coordinate of $\bsx_i$, then the literature often defines generating matrices, $\mathsf{C}_j = (c_{ij\ell})_{\ell,i = 1}^{M,N}$ for $j = 1, \ldots, d$, where $M$ is the maximum number of bits in the expression for $\bsx_i$, say $52$, and $2^N$ is the maximum number of nodes that is intended to be generated.
 The digital sequence can then be defined equivalently as
@@
 \subsection{LD Nodes by Optimization}\label{sec:optLD}
-[Nathan Kirk write here, please]
+
+Most quasi-Monte Carlo constructions are studied based on their asymptotic behavior as \(n \to \infty\), with numerous results on asymptotic rates \cite{NIED1978QMCBook,NovWoz10a}.
+However, few methods address finding optimal point configurations for specific \(n\) and \(d\) which can be critical in practical applications where each function evaluation is costly.
+Of course, one possible strategy is to truncate asymptotically optimal sequences to the desired \(n\), however, as discussed above many of the standard LD constructions have a preferred number of points with truncations leading to imbalances in equidistribution.
+We can often do better optimizing directly for the target \(n\) and \(d\).
+
+%When optimizing a sample point distribution, one requires a figure-of-merit to serve as the objective function for the optimization problem.
+
+\vspace{2mm}
+\noindent
+\textbf{Combinatorial Optimization.}
+For several decades, combinatorial optimization techniques have been employed to obtain good generating vectors for rank-1 lattice points.
+Recall that for given \(n\) and \(d\), a lattice rule is fully determined by the generating vector \(\boldsymbol{h} \in \mathbb{Z}_n^d \coloneqq \{0, 1, \ldots, n-1\}^d\).
+While good generating vectors are known for \(d \leq 2\) \cite{DicEtal22a,SloJoe94}, the same does not hold for \(d \geq 3\), often resorting to finding good $\boldsymbol{h}$ via computer search to minimize a chosen figure-of-merit.
+An exhaustive search is infeasible due to the exponential growth of the search space with \(d\).
+The component-by-component (CBC) construction, introduced by Korobov \cite{kor63} and later revisited by Sloan and collaborators \cite{Slo02}, reduces the search space to size \(dn\) by constructing \(\boldsymbol{h}\) one component at a time.
+%Further simplification is possible by restricting the search to Korobov lattices of the form \(\boldsymbol{h} = (1, h, h^2, \ldots, h^{d-1}) \in \mathbb{Z}^d\). 
+For certain choices of function spaces, there exist fast component-by-component construction implementations using fast Fourier transforms (FFTs) as the primary tool.
+We refer to \cite{DicKuo04a,JoeKuo03,KuoJoe02b} and references therein for works on the CBC construction for lattice rules and \cite{NuyCoo06a,NuyCoo06b} for their fast implementations.
+
+\vspace{2mm}
+\noindent
+\textbf{Global Optimization.}
+
+\vspace{2mm}
+\noindent
+\textbf{Other Notable Methods.}
 
@@
 
-While determinism has some advantages randomization has advantages as well.
+While determinism has some advantages, randomization has advantages as well.
 \begin{itemize}
@@
 \subsection{Discrepancies Defined by Kernels} \label{sec:kerdisc}
-Let $K: [0,1)^d \times [0,1)^2 \to \mathbb{R}$ be a function satisfying the following two conditions:
+Let $K: [0,1)^d \times [0,1)^d \to \mathbb{R}$ be a function satisfying the following two conditions:
 \begin{subequations} \label{eq:Kcond}

### 2025-01-16 HickernellMCQMC2024_QMCTutorial.tex
M2 Mac updates

@@
 \end{multline*}
-and $D^{\emptyset}f$ is understood to be the constant $f(1/2, \ldots, 1/2)$ \cite{Hic97a}.
+and $D^{\emptyset}f$ is understood to be the constant $f(1/2, \ldots, 1/2)$ \cite{Hic97a}.  The variation, which corresponds to the non-constant part of the function, is
+\begin{equation*}
+		\text{variation}(f) := \sqrt{\sum_{\emptyset \ne \bsell \subseteq \{1, \ldots, d\}} \lVert D^{\bsell}f\rVert_2^2}.
+\end{equation*}
 
@@
 \end{align*}
-The extensions to general $1/2 < x < 1$ and to $d > 1$ is straightforward but a bit tedious \cite{Hic97a}.
+The extensions to general $1/2 < x < 1$ and to $d > 1$ are straightforward but tedious \cite{Hic97a}.
 
@@
 \end{align}
-As is true for most cases, the computational cost of computing the discrepancy is $\Order(dn^2)$ as $d$ and/or $n$ tend to infinity.
+As is true for most cases, the computational cost of computing the discrepancy is $\mathcal{O}(dn^2)$ as $d$ and/or $n$ tend to infinity.
 
@@
 \subsection{Coordinate weights} \label{sec:coordwts}
+When the dimension, $d$, is increased, the discrepancy increases and the rate of decay with increasing sample size deteriorates, as can be seen in Figure ??.
+This raises the question of whether the LD sequences lose their effectiveness for higher dimensions or whether the theory described in the pervious section does not well describe what may occur in practice.
+
+One thing to note is that the discrepancy of the null set, which is equivalent to the norm of the linear functional $\mu(\cdot)$, is
+\begin{equation}
+		\text{discrepancy}\bigl(\emptyset,K \bigr)  = \sqrt{\int_{[0,1)^d \times [0,1)^d} K(\bst,\bsx) \, \mathrm{d} \bst \, \mathrm{d} \bsx},
+\end{equation}
+which corresponds to $(13/12)^{d/2}$ for the centered discrepancy.
+This means that integration gets exponentially harder as $d$ increases.
+One may scale the discrepancy of a node set by dividing it by discrepancy$(\emptyset,K)$, but this does not recover the desired nearly $\mathcal{O}(n^{-1})$ decay rate.
+
+A better approach is to introduce coordinate weights, $\gamma_1, \ldots, \gamma_d$ into the reproducing kernel.  For the kernel defining the centered discrepancy, this means
+\begin{align} \label{eq:wtctrkernel}
+	K(\bst,\bsx) &= \prod_{\ell = 1}^d \biggl [ 1 + \frac {\gamma_\ell^2}2 \bigl ( \lvert t_\ell - 1/2 \rvert + \lvert x_\ell - 1/2 \rvert - \lvert t_\ell - x_\ell \rvert \bigr ) \biggr] , \\
+	\langle f, g \rangle_{\calH_K}  &= \sum_{\bsell \subseteq \{1, \ldots, d\}} \prod_k \gamma_{\ell_k}^2 \langle D^{\bsell}f, D^{\bsell}g \rangle_2, \\
+	\text{variation}(f)  &= \sqrt{\sum_{\bsell \subseteq \{1, \ldots, d\}} \prod_k \gamma_{\ell_k}^2 \lVert D^{\bsell}f \rVert_2^2}, \\
+	\label{eq:wt-ctrdisc}
+	\nonumber
+	\MoveEqLeft[4]\text{wt-ctr-discrepancy}^2\bigl(\{ \bsx_i\}_{i=0}^{n-1} \bigr) \\
+	\nonumber
+	&= \prod_{\ell=1}^d \left ( 1 + \frac{\gamma_{\ell}^2}{12} \right )^d
+	- \frac 2n \sum_{i=0}^{n-1} \prod_{\ell=1}^d \left [ 1 + \frac {\gamma_{\ell}^2}2 \bigl ( \lvert x_{i\ell} - 1/2 \rvert - \lvert x_{i\ell} - 1/2 \rvert^d \bigr )\right ] \\
+	& \qquad \qquad \frac{1}{n^2} \sum_{i,j=0}^{n-1} \prod_{\ell = 1}^d \left[ 1 + \frac {\gamma_{\ell}^2}2 \bigl ( \lvert x_{i\ell} - 1/2 \rvert + \lvert x_{j\ell} - 1/2 \rvert - \lvert x_{i\ell} - x_{j\ell} \rvert \bigr ) \right].
+\end{align}
+
 \subsection{Discrepancies for Lattices and Nets} \label{sec:latticenetdis}

### 2025-01-16 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

--- 2025-01-16 HickernellMCQMC2024_QMCTutorial.tex
@@
 
-%When optimizing a sample point distribution, one requires a figure-of-merit to serve as the objective function for the optimization problem.
+When optimizing LD nodes, one requires an objective function which in some sense must assesses the quality of the LD nodes themselves, or the expected performance when implemented in a qMC estimate.
+We will refer to such an objective function as a figure-of-merit (FOM) and popular choices include discrepancy or worst-case error metrics (see Section \ref{sec:discrepancy}) and spectral test...
+[come back to this]
 
@@
 \textbf{Combinatorial Optimization.}
-For several decades, combinatorial optimization techniques have been employed to obtain good generating vectors for rank-1 lattice points.
-Recall that for given \(n\) and \(d\), a lattice rule is fully determined by the generating vector \(\boldsymbol{h} \in \mathbb{Z}_n^d \coloneqq \{0, 1, \ldots, n-1\}^d\).
-While good generating vectors are known for \(d \leq 2\) \cite{DicEtal22a,SloJoe94}, the same does not hold for \(d \geq 3\), often resorting to finding good $\boldsymbol{h}$ via computer search to minimize a chosen figure-of-merit.
+For several decades, combinatorial optimization techniques have been employed to obtain good generating parameters for LD constructions.
+For example, in the case of a lattice rule which is fully determined by the generating vector \(\boldsymbol{h} \in \mathbb{Z}_n^d \coloneqq \{0, 1, \ldots, n-1\}^d\) given $n$ and $d$,
+%While good generating vectors are known for \(d \leq 2\) \cite{DicEtal22a,SloJoe94}, the same does not hold for \(d \geq 3\), 
+one often resorts to finding good $\boldsymbol{h}$ via computer search to minimize a chosen figure-of-merit.
 An exhaustive search is infeasible due to the exponential growth of the search space with \(d\).
-The component-by-component (CBC) construction, introduced by Korobov \cite{kor63} and later revisited by Sloan and collaborators \cite{Slo02}, reduces the search space to size \(dn\) by constructing \(\boldsymbol{h}\) one component at a time.
+Thus, the component-by-component (CBC) construction, introduced by Korobov \cite{kor63} and later revisited by Sloan and collaborators \cite{Slo02}, reduces the search space to size \(dn\) by constructing \(\boldsymbol{h}\) one component at a time while keeping the previous components fixed.
+This greedy procedure also applies to finding good generating matrices $\mathsf{C}_j$ for digital sequences.
 %Further simplification is possible by restricting the search to Korobov lattices of the form \(\boldsymbol{h} = (1, h, h^2, \ldots, h^{d-1}) \in \mathbb{Z}^d\). 
 For certain choices of function spaces, there exist fast component-by-component construction implementations using fast Fourier transforms (FFTs) as the primary tool.
-We refer to \cite{DicKuo04a,JoeKuo03,KuoJoe02b} and references therein for works on the CBC construction for lattice rules and \cite{NuyCoo06a,NuyCoo06b} for their fast implementations.
+We refer to \cite{DicKuo04a,JoeKuo03,KuoJoe02b} and references therein for works on the CBC construction for lattice rules and digital sequences, \cite{NuyCoo06a,NuyCoo06b} for their fast implementations and \cite{LatNet} for a QMC software tool called LatNet Builder by L'Ecuyer to generate LD nodes which relies heavily upon the CBC construction method.
 
@@
 \noindent
-\textbf{Global Optimization.}
+\textbf{Global Optimization.} Global optimization of LD nodes is often challenging due to the non-convex nature of the FOM as a function of the nodes.
+Traditional global optimization methods, such as regular gradient descent, struggle in this setting becoming trapped in suboptimal local minima and failing to identify the global optimum.
+One of the first works to globally optimize LD nodes, albeit a heuristic approach, was by Winker and Fang \cite{WinFan97b} employing the threshold accepting optimization algorithm.
+More recently, Rusch, Kirk, Bronstein, Lemieux and Rus \cite{ruschkirk24} developed the first machine-learning assisted optimization tool for LD node optimization in the Message-Passing Monte Carlo (MPMC) method.
+MPMC implements tools for geometric deep learning imposing a computational graph onto an initialized IID node set and learning a transformation mapping
 
@@
 \noindent
-\textbf{Other Notable Methods.}
+\textbf{Other Notable Methods.} Recently, a method called subset selection is formulated to choose from an $n−$element point set, the $k<n$ points which yield the smallest discrepancy.
+An exact selection algorithm was presented in \cite{cle22}, while a swap-based heuristic approach was used in \cite{cle24}.
+Furthermore, a method to generate optimal star-discrepancy point sets for fixed N and d based on a non-linear programming approach was suggested in \cite{}.
 
@@
 \end{subequations}
-A random digital shift, $\bsDelta$ is also added, where $\Delta_{\ell j}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ of the shift.
+A random digital shift, $\bsDelta$ is also added, where $\Delta_{\ell j}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ dimension of the shift.
 
@@
 
+As Halton points can be viewed as digital nets with different bases for each dimension, the randomization techniques used for digital nets can be applied to Halton pointsets.
+Random digital shifts for Halton set $x_{ij\ell} \mapsto (x_{ij\ell} + \Delta_{j\ell}) \mod b_j$ where the shift $\Delta_{j\ell} \in \{0,\dots,b_j-1\}$.
+More generally, a permutation scramble for digital nets sets $x_{ij\ell} \mapsto \Pi_{j\ell}(x_{ij\ell})$ where $\Pi_{j\ell}$ are independent random permutations of $\{0,\dots,b_j-1\}$.
+Random nested uniform scrambling and linear matrix scrambling have also been considered for Halton pointsets \cite{owen2024gain}.
+
 \subsection{Randomized Generators}
+
+Instead of explicitly searching for good generating vectors or matrices, one can instead sample a few random generating vectors or generating matrices and then prune out bad choices \cite{}.
+Random generating vectors $\bsh$ take $h_j \sim \calU\{0,\dots,n-1\}$.
+Random generating matrices for base $2$ digital nets set $c_{ij\ell} \sim \calU\{0,1\}$ for $i<\ell$ while enforcing $c_{iji}=1$ and $c_{ij\ell}=0$ for $i>\ell$.
+One idea is to generate
 
@@
 So far, we have relied on eye tests and the Keister example to show how LD sequences are better than IID.
-This section introduces the theory that shows way minimizing discrepancy leads to better approximations to $\mu$.
+This section introduces the theory that shows why minimizing discrepancy leads to better approximations to $\mu$.
 
@@
 \end{equation*}
-This is guaranteed by the Riesz Representation Theorm, and $\eta_L$ is called the representer of $L$.
-Using the reproducing property of $K$, one may derove am explict formula for $\eta_L(\bsx)$, namely,
+This is guaranteed by the Riesz Representation Theorem, and $\eta_L$ is called the representer of $L$.
+Using the reproducing property of $K$, one may derive an explicit formula for $\eta_L(\bsx)$, namely,
 \begin{equation*}
@@
 
-If $\mu(\cdot)$ is a bounded linear functional on $\calH_K$ is, we may use the argument above to derive our error bound.
+If $\mu(\cdot)$ is a bounded linear functional on $\calH_K$, we may use the argument above to derive our error bound.
 First we write the dependence of $\mu - \hat{\mu}_n$ on  $f$ explicitly as $\mu(f) - \hat{\mu}_n(f)$ and note that the error functional, $\mu(\cdot) - \hat{\mu}_n(\cdot)$, is  linear.
-The Riesz Representation Theorem and the Cauchy-Schawarz inequality imply a tight error bound:
+The Riesz Representation Theorem and the Cauchy-Schwarz inequality imply a tight error bound:
 \begin{align}
@@
 Practitioners of qMC want to formulate $\mu$ in a way to make the variation as small as possible.
-This discussed in Section \ref{sec:reformulate}.
+This is discussed in Section \ref{sec:reformulate}.
 
@@
 	\item The Hilbert space $\calH_K$ contains ones assumptions about $f$, such as smoothness or periodicity.
-Knowing $K$, however, does not automatically lead to an explict formula for the inner product of the associated Hilbert space, $\langle \cdot, \cdot \rangle_{\calH_K}$.
-Conversely, having an explict formula for  $\langle \cdot, \cdot \rangle_{\calH_K}$ does not automatically lead to an explict formula for the reproducing kernel, $K$.
+Knowing $K$, however, does not automatically lead to an explicit formula for the inner product of the associated Hilbert space, $\langle \cdot, \cdot \rangle_{\calH_K}$.
+Conversely, having an explicit formula for  $\langle \cdot, \cdot \rangle_{\calH_K}$ does not automatically lead to an explicit formula for the reproducing kernel, $K$.
 In both cases educated guesswork is involved.
@@
 Once $K$ has been fixed, the discrepancy may be used as a comparison measure among different LD sequences, and its rate of decay---which sometimes can be computed theoretically---indicates how quickly $\hat{\mu}_n$ approaches $\mu$.
-	\item When an explict formula for the inner product of the associated Hilbert space, $\langle \cdot, \cdot \rangle_{\calH_K}$ can be deduced, it suggests how to $\mu$ to reduce variation$(f)$.
+	\item When an explicit formula for the inner product of the associated Hilbert space, $\langle \cdot, \cdot \rangle_{\calH_K}$, can be deduced, it suggests how to formulate $\mu$ to reduce variation$(f)$.
 \end{itemize}
@@
 \section{Stopping Criteria} \label{sec:stop}
+
 Art
@@
 \bibliographystyle{spmpsci}
-\bibliography{FJH23,FJHown23}
+\bibliography{FJH23,FJHown23,main}
 

### 2025-01-17 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

@@
 When optimizing LD nodes, one requires an objective function which in some sense must assesses the quality of the LD nodes themselves, or the expected performance when implemented in a qMC estimate.
-We will refer to such an objective function as a figure-of-merit (FOM) and popular choices include discrepancy or worst-case error metrics (see Section \ref{sec:discrepancy}) and spectral test...
-[come back to this]
+We will refer to such an objective function as a figure-of-merit (FOM) and popular choices include the discrepancy (see Section \ref{sec:discrepancy}) or worst-case error metrics when considering a specific space of functions; see \cite{Hic00a,LatNet}.
 
@@
 %While good generating vectors are known for \(d \leq 2\) \cite{DicEtal22a,SloJoe94}, the same does not hold for \(d \geq 3\), 
-one often resorts to finding good $\boldsymbol{h}$ via computer search to minimize a chosen figure-of-merit.
+one often resorts to finding good $\boldsymbol{h}$ via computer search to minimize a chosen FOM.
 An exhaustive search is infeasible due to the exponential growth of the search space with \(d\).
@@
 \textbf{Global Optimization.} Global optimization of LD nodes is often challenging due to the non-convex nature of the FOM as a function of the nodes.
-Traditional global optimization methods, such as regular gradient descent, struggle in this setting becoming trapped in suboptimal local minima and failing to identify the global optimum.
-One of the first works to globally optimize LD nodes, albeit a heuristic approach, was by Winker and Fang \cite{WinFan97b} employing the threshold accepting optimization algorithm.
+Traditional global optimization methods, such as gradient descent or its variants, struggle in this setting becoming trapped in suboptimal local minima and failing to identify the global optimum configuration of nodes.
+
+One of the first works to globally optimize LD nodes, albeit a heuristic approach, was by Winker and Fang \cite{WinFan97b} employing the threshold accepting optimization algorithm to directly optimize the uniformity of the LD nodes with respect to the discrepancy FOM.
+
 More recently, Rusch, Kirk, Bronstein, Lemieux and Rus \cite{ruschkirk24} developed the first machine-learning assisted optimization tool for LD node optimization in the Message-Passing Monte Carlo (MPMC) method.
-MPMC implements tools for geometric deep learning imposing a computational graph onto an initialized IID node set and learning a transformation mapping
+MPMC implements tools from geometric deep learning imposing a computational graph onto an initialized IID node set to transform the points to LD node set via a learned mapping, again guided by the minimization of a discrepancy-based FOM.
+
+Lastly, we highlight the method proposed by Hinrichs and Oettershagen \cite{hinoet16}, which generates optimal LD nodes for a specific class of functions in two-dimensions.
+To address the non-convexity challenge, the authors decompose the global optimization problem into a large number of smaller, convex sub-problems.
 
@@
 \noindent
-\textbf{Other Notable Methods.} Recently, a method called subset selection is formulated to choose from an $n−$element point set, the $k<n$ points which yield the smallest discrepancy.
-An exact selection algorithm was presented in \cite{cle22}, while a swap-based heuristic approach was used in \cite{cle24}.
-Furthermore, a method to generate optimal star-discrepancy point sets for fixed N and d based on a non-linear programming approach was suggested in \cite{}.
+\textbf{Other Notable Methods.} 
+There are several other attempts to construct optimal LD nodes, many of which are well summarized in the thesis of Cl\'{e}ment \cite{}.
+We highlight the subset selection method proposed in \cite{cle22} to choose from an $n$-element initialization set, the $k<n$ points which minimize a discrepancy FOM.
+A heuristic approach to this problem was later shown in \cite{cle24}.
+%Furthermore, a method to generate optimal star-discrepancy point sets for fixed $n$ and $d$ based on a non-linear programming approach was suggested in \cite{}.
+
+It is commonplace in qMC constructions that equidistribution is required not just in the (possibly high dimensional) $d$-dimensional space, but also for all or specific low-dimensional projections; see Section \ref{sec:coordwts}.
+MatBuilder \cite{paulin2022} is a software tool developed precisely for this task; to generate digital sequence LD nodes when low-dimensional uniformity is critical.
+Uniformity constraints related to digital sequence stratification across boxes of the form \eqref{eq:element_box} are expressed as integer linear programming (ILP) problems and subsequently solved, while optimizing an objective function relating to the ILP reformulation, to determine the next columns in the generator matrices for the digital sequence.
 
@@
 
-\subsection{Discrepancies for Lattices and Nets} \label{sec:latticenetdis}
+Figure \ref{fig:scaledweightedcentered} plots the scaled centered discrepancy with coordinate weights $\bsgamma = (1/\ell)_{\ell =1}^d$,
+\begin{equation*}
+    \text{wt-ctr-discrepancy}\bigl(\{ \bsx_i\}_{i=0}^{n-1} \bigr)/\text{wt-ctr-discrepancy}(\emptyset ),
+\end{equation*}
+against sample size for a variety of $d$.  Now, even for larger $d$ there is a reasonable decay that approximates $\mathcal{O}(n^{-1})$.
 
-\subsection{Randomized Error}
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{\figpath/IIDSobolScaledWtCtrDisc_n_8192_d64.pdf}
+    \caption{The scaled centered discrepancy with coordinate weights for a Sobol' LD sequence for various $d$ and the root mean squared scaled centered discrepancy with coordinate weights for IID nodes.
+LD nodes have a much smaller discrepancy than IID nodes.}
+    \label{fig:scaledweightedcentered}
+\end{figure}
+
+For IID samples the root mean squared discrepancy for a general reproducing kernel, $K$, is 
+\begin{multline}
+    \mathbb{E}[\text{discrepancy}^2\bigl(\{\bsx\}_{i=0}^{n-1},K \bigr)]  \\
+    = \frac 1{\sqrt{n}}  \sqrt{  \int_{[0,1)^d} K(\bsx,\bsx) \, \mathrm{d} \bsx - 
+	 \int_{[0,1)^d \times [0,1)^d} K(\bst,\bsx) \, \mathrm{d} \bst \, \mathrm{d} \bsx}.
+\end{multline}
+This is also plotted in Figure \ref{fig:scaledweightedcentered}. which highlights the superiority of LD nodes.
+
+\subsection{Discrepancies for Lattices and Digital Nets} \label{sec:latticenetdis}
+Many useful constructions of LD sequences are constructed numerically by minimizing the discrepancy, however, as noted earlier, computing the discrepancy requires $\mathcal{O}(dn^2)$ operations, which can be expensive.
+However, there is a way to reduce this computational cost for lattices and digital nets when using friendly kernels.
+
+The mean squared discrepancy of a randomly shifted node set, $\{\bsx_i^{\bsDelta} =  \bsx_i + \bsDelta \bmod \bsone \}_{i=0}^{n-1}$ for an arbitrary kernel takes the form
+\begin{align}
+\nonumber
+\MoveEqLeft \mathbb{E}\bigl[\text{discrepancy}^2 \bigl (\{ \bsx_{i} + \bsDelta \bmod \bsone \}_{i=0}^{n-1} \bigr) \bigr] \\
+\nonumber
+& = \int_{[0,1)^d \times [0,1)^d} K(\bst,\bsx) \, \mathrm{d} \bst \, \mathrm{d} \bsx  -\frac 2n  \sum_{i=0}^{n-1}  \int_{[0,1)^d} \mathbb{E} \bigl [K(\bst,\bsx_i + \bsDelta \bmod \bsone) \bigr]  \, \mathrm{d} \bst \\
+\nonumber
+	& \qquad \qquad + \frac{1}{n^2} \sum_{i,j=0}^{n-1}  \mathbb{E} \bigl [ K(\bsx_i + \bsDelta \bmod \bsone,\bsx_j + \bsDelta \bmod \bsone) \bigr] \\
+\nonumber
+    & = \frac{1}{n^2} \sum_{i,j=0}^{n-1}  \mathbb{E} \bigl [ K(\bsx_i - \bsx_j  + \bsDelta \bmod \bsone, \bsDelta \bmod \bsone) \bigr] - \int_{[0,1)^d \times [0,1)^d} K(\bst,\bsx) \, \mathrm{d} \bst \, \mathrm{d} \bsx ,\\
+    \intertext{where we have applied a variable transformation $\bsDelta \to \bsDelta - \bsx_j$,}
+    & = \frac{1}{n^2} \sum_{i,j=0}^{n-1}  \widetilde{K}(\bsx_i - \bsx_j \bmod \bsone ) - \int_{[0,1)^d} \widetilde{K}(\bsx) \, \mathrm{d} \bsx, \\
+    \intertext{where the filtered version of the reproducing kernel is defined as}
+    \widetilde{K}(\bsx) &: = \int_{[0,1)^d} K(\bsx + \bst \bmod \bsone, \bst) \, \mathrm{d} \bst.
+\end{align}
+
+For arbitrary node sets, this mean squared discrepancy is of similar computational cost as the original, but if the unshifted node set $\{\bsx_i\}_{i=0}^{n-1}$ is a lattice as defined in \eqref{eq:latticepts}, then for any $j$, the set $\{\bsx_i - \bsx_j \bmod \bsone \}_{i=0}^{n-1}$ has the same values.
+Thus the mean squared discrepancy for a randomly shifted lattice can be computed in only $\mathcal{O}(dn)$ operations, assuming that $K(\bsx)$ requires $d$ operations to compute:
+\begin{equation}
+\mathbb{E}\bigl[\text{discrepancy}^2 \bigl (\{ \bsx^{\bsDelta-\text{lat}}_{i} \}_{i=0}^{n-1} \bigr) \bigr] = \frac{1}{n} \sum_{i=0}^{n-1}  \widetilde{K}(\bsx_i) - \int_{[0,1)^d} \widetilde{K}(\bsx) \, \mathrm{d} \bsx.
+\end{equation}
+For the kernel defining the centered discrepancy with coordinate weights, 
+\begin{equation}
+    \widetilde{K}(\bsx) = \int_{[0,1)^d} \prod_{\ell = 1}^d \Bigl [ 1 + \gamma_\ell^2 \{ 1/4 - x_\ell (1-x_\ell ) \} \Bigr], \quad \int_{[0,1)^d} \widetilde{K}(\bsx) \, \mathrm{d} \bsx = \prod_{\ell = 1}^d \biggl [ 1 + \frac{\gamma_\ell^2}{12} \biggr ].
+\end{equation}
+
+\subsection{Randomized Error and Bayesian Error}
+The error analysis leading to the
 
@@
 
+\section*{Appendix B}
+We derive the filtered version of the centered discrepancy with coordinate weights:
+\begin{align*}
+    \widetilde{K}(\bsx)
+    &: = \int_{[0,1)^d} K(\bsx + \bst \bmod \bsone, \bst) \, \mathrm{d} \bst \\
+    & = \int_{[0,1)^d} \prod_{\ell = 1}^d \biggl [ 1 + \frac {\gamma_\ell^2}2 \bigl ( \lvert (x_\ell + t_\ell) \bmod 1 - 1/2 \rvert + \lvert t_\ell - 1/2 \rvert   \\
+    & \qquad \qquad  - \lvert (x_\ell + t_\ell) \bmod 1 - t_\ell \rvert \bigr ) \biggr] \, \mathrm{d} \bst \\
+    & = \prod_{\ell = 1}^d \int_0^1 \biggl [ 1 + \frac {\gamma_\ell^2}2 \bigl ( \lvert (x + t) \bmod 1 - 1/2 \rvert + \lvert t - 1/2 \rvert - \lvert (x + t) \bmod 1 - t \rvert \bigr ) \biggr] \, \mathrm{d} t
+\end{align*}
+Now
+\begin{align*}
+   \int_0^1   \lvert (x + t) \bmod 1 - 1/2 \rvert  \, \mathrm{d} t & =  \int_0^1   \lvert t - 1/2 \rvert  \, \mathrm{d} t \\
+   & = \int_0^{1/2} (1/2-t) \, \mathrm{d} t + \int_{1/2}^1 (t - 1/2) \, \mathrm{d} t \\
+   & = \left. - \frac{(1/2-t)^2}{2} \right \rvert_0^{1/2} + \left . - \frac{(t-1/2)^2}{2} \right \rvert _{1/2}^1 = \frac 14, \\
+   \int_0^1   \lvert (x + t) \bmod 1 - t \rvert  \, \mathrm{d} t & =
+   \int_0^{1-x} \lvert x+t - t \rvert \, \mathrm{d} t + \int_{1-x}^1 \lvert x+t -1 - t \rvert \, \mathrm{d} t \\
+   & = x(1-x) + x(1-x) = 2x(1-x).
+\end{align*}
+Thus, 
+\begin{equation}
+    \widetilde{K}(\bsx) = \int_{[0,1)^d} \prod_{\ell = 1}^d \Bigl [ 1 + \gamma_\ell^2 \{ 1/4 - x_\ell (1-x_\ell ) \} \Bigr]
+\end{equation}
+Moreover,
+\begin{align*}
+    \int_{[0,1)^d} \widetilde{K}(\bsx) \, \mathrm{d} \bsx & =
+    \prod_{\ell = 1}^d \Bigl [ 1 + \gamma_\ell^2  \bigl\{ 1/4 - \int_0^1 x (1-x ) \, \mathrm{d}x \bigr\} \Bigr] \\
+    & = \prod_{\ell = 1}^d \Bigl [ 1 + \gamma_\ell^2 \bigl \{ 1/4 - \left . [x^2/2 - x^3 / 3] \right \rvert_0^1  \bigr \} \Bigr ]    \\
+    & = \prod_{\ell = 1}^d \Bigl [ 1 + \frac{\gamma_\ell^2}{12} \Bigr ]
+\end{align*}
+
 \end{document}

### 2025-01-17 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

@@
 \subsection{Randomized Error and Bayesian Error}
-The error analysis leading to the
+The error analysis leading to error bound \eqref{eq:cuberrbd} is a worst-case error analysis, meaning that the error in approximating $\mu(f)$ by $\hat{\mu}_n(f)$ is definitely smaller than the discrepancy times the variation, provided that $f$ lies inside the Hilbert space with reproducing kernel $K$.
+There are other error measures that we mention briefly here.
+More detail is given in \cite{HicEtal17a}.
+
+For sample means where the nodes, $\{\bsx_i\}_{i=0}^{n-1}$ are randomized, one can define a measure of their quality based on the root mean squared error of $\hat{\mu}_n(f)$:
+\begin{align}
+    \text{RMSE}(\hat{\mu}_n(f)) & : = \sqrt{\mathbb{E}[(\mu(f) - \hat{\mu}_n(f))^2]} \\
+    &\le \underbrace{ \sup_{\text{variation}(f,K) \le 1} \sqrt{\mathbb{E}[(\mu(f) - \hat{\mu}_n(f))^2]}}_{\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)} \, \text{variation}(f,K).
+\end{align}
+This quality measure $\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ is typically more optimistic than the root mean squared of  $\text{discrepancy}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ because the latter is finding a worst case function for each random instance of nodes, while the former is finding one worst function.
+
+Unfortunately, $\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ is normally not as easy to compute as the discrepancy.
+If $\hat{\mu}_n$ is unbiased, then $\sqrt{\mathbb{E}[(\mu(f) - \hat{\mu}_n(f))^2]} = \text{std}(\hat{\mu}_n(f))$.
+For IID nodes we know that this is $\text{std}(f(\bsX))/\sqrt{n}$.
+
+For randomly shifted lattices
+
+Bayesian
 

### 2025-01-18 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

@@
 \usepackage{xspace}
+\renewcommand{\textfraction}{0.1}  % Minimum fraction of text
+\renewcommand{\floatpagefraction}{0.9}  % Maximum fraction of a page for floats
 
@@
 In mathematical terms, the random outcome is $Y := f(\bsX)$, where $\bsX$ is a vector random input.
-Given a particular input value, $\bsx$, the corresponding outcome  $y = f(\bsx)$ can be computed by an algorithm, whose complexity may be such that $f$ can be considered a ``black box''.
-The user selects a sequence of input values, also known as data sites or \emph{nodes}, $\bsx_0, \bsx_1, \ldots$, which give rise to the observed outcomes $y_0 = f(\bsx_0), y_1 = f(\bsx_1), \ldots$.
-The $y_i$ form the basis for approximating $\bbE(Y)$, quartiles of $Y$, the probability density of $Y$, and other quantities of interest.
+Given an input, $\bsx$, the corresponding outcome  $y = f(\bsx)$ can be computed by an algorithm, whose complexity may might make $f$ a \emph{black box}, or at least a gray box.
+The user selects a sequence of input values, also known as data sites or \emph{nodes}, $\bsx_0, \bsx_1, \ldots$, which result in observed outcomes $y_0 = f(\bsx_0), y_1 = f(\bsx_1), \ldots$.
+The $y_i$ may be used to approximate $\bbE(Y)$, quartiles of $Y$, the probability density of $Y$, and other quantities of interest.
 
-Simple  Monte Carlo chooses the nodes $\{\bsx_0, \bsx_1, \ldots\}$ to be independent and identically distributed (IID).
-Quasi-Monte Carlo (qMC)  chooses the nodes \emph{not} IID, but with an empirical distribution that approximates well the true probability distribution of $\bsX$.
-The difference between these two distributions is called a \emph{discrepancy}, and the node sequences used in qMC are called low discrepancy (LD) sequences.
+Simple  Monte Carlo chooses the nodes, $\{\bsx_0, \bsx_1, \ldots\}$, to be independent and identically distributed (IID).
+Quasi-Monte Carlo (qMC)  chooses the nodes differently, so that their empirical distribution approximates well the true probability distribution of $\bsX$.
+The difference between  the true and empirical distributions is called a \emph{discrepancy}, and the node sequences used in qMC are called low discrepancy (LD) sequences.
 
@@
 \begin{equation}\label{eq:keisterA}
-	\mu := \int_{\mathbb{R}^d} \cos(\lVert \bst \rVert) \exp(-\lVert \bst \rVert^2) \, \rmd \bst,
+	\mu := \int_{\mathbb{R}^d} \cos(\lVert \bst \rVert_2) \exp(-\lVert \bst \rVert_2^2) \, \rmd \bst,
 \end{equation}
-where $\lVert \bst \rVert := \sqrt{t_1^2 + \cdots + t_d^2}$.  This integral may be evaluated numerically by re-writing it in spherical coordinates as
+where $\lVert \bst \rVert_2 := \sqrt{t_1^2 + \cdots + t_d^2}$.  This integral may be evaluated numerically by re-writing it in spherical coordinates as
 \begin{equation}\label{eq:keisterExact}
@@
 where $2 \pi^{d/2}/\Gamma(d/2)$ is the surface area of the sphere in $d$ dimensions, and $\Gamma$ is the Gamma function.
-The resulting one-dimensional integral is amenable to quadrature methods \cite{}.
+The resulting one-dimensional integral is amenable to quadrature methods.
 This non-trivial test case with a true value that can be easily calculated allows us to compute the numerical errors of various cubature schemes.
 
-However, for the purpose of this illustration, we work with $\mu$ in its original form, \eqref{eq:keisterA} and approximate it by a sample mean,
+For the purpose of this illustration, we work with $\mu$ in its original form, \eqref{eq:keisterA} and approximate it by a sample mean,
 \begin{equation} \label{eq:sample_mean}
@@
 \end{equation}
-This does require some further preparation, which is discussed in greater generality in Section \ref{sec:reformulate}.
+This does require some preparation, which is discussed in generality in Section \ref{sec:reformulate}.
 
-This $\mu$ may be thought of as the expectation of $Y := g(\bsT) := \pi^{d/2} \cos(\lVert \bsT \rVert)$, where $T_1, \ldots, T_d$ are IID Gaussian (normal) random variables with zero mean and variance $1/2$, i.e., $\bsT :=(T_1, \ldots, T_d) \sim \calN(\bszero,\mathsf{I}/2)$:
+This $\mu$ is the expectation of $Y := g(\bsT) := \pi^{d/2} \cos(\lVert \bsT \rVert)$, where $T_1, \ldots, T_d$ are IID Gaussian (normal) random variables with zero mean and variance $1/2$, i.e., $\bsT :=(T_1, \ldots, T_d) \sim \calN(\bszero,\mathsf{I}/2)$:
 \begin{equation}\label{eq:keisterB}
@@
 \end{equation}
-Nearly all LD sequences, which underlie qMC, are defined to approximate the standard uniform distribution, $\calU[0,1]^d$.
+Virtually all LD sequences, which underlie qMC, are defined to approximate the standard uniform distribution, $\calU[0,1]^d$.
 Thus, we perform a variable transformation $\bst = \bigl (\Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d) \bigr )/\sqrt{2}$,
@@
 \begin{equation} \label{eq:sample_mean_Keister}
-	\hat{\mu}_n = \frac 1n \sum_{i=1}^n \pi^{d/2} \cos\Bigl (\bigl \lVert \bigl( \Phi^{-1}(x_{i1}), \ldots, \Phi^{-1}(x_{id})\bigr) \bigr\rVert/\sqrt{2}  \Bigr).
+	\hat{\mu}_n = \frac 1n \sum_{i=0}^{n-1} f (\bsx) = \frac 1n \sum_{i=1}^n \pi^{d/2} \cos\Bigl (\bigl \lVert \bigl( \Phi^{-1}(x_{i1}), \ldots, \Phi^{-1}(x_{id})\bigr) \bigr\rVert/\sqrt{2}  \Bigr).
 \end{equation}
 Consider the specific case of $d=6$ for which $\mu \approx -2.327303729298$.
-We approximate $\mu$ by the sample mean defined in \eqref{eq:sample_mean} for three different kinds of nodes, $\{\bsx_i\}_{i=0}^{n-1}$, for various sample sizes, $n$, and plot the relative errors, $\lvert (\mu - 	\hat{\mu}_n)/\mu\rvert$ in Figure \ref{fig:keister-err}.
-The three kind of nodes are
+We approximate $\mu$ by $\hat{\mu}_n$ for different, $\{\bsx_i\}_{i=0}^{n-1}$, for various sample sizes, $n$, and plot the relative errors, $\lvert (\mu - 	\hat{\mu}_n)/\mu\rvert$ in Figure \ref{fig:keister-err}.
+The three kinds of nodes are
 \begin{enumerate}
 	\renewcommand{\labelenumi}{\roman{enumi}.}
-	\item Cartesian grids, $\{1/(2m), \ldots, (2m-1)/(2m) \}^6$, for $m = 2, 3, \ldots$ and $n = m^6$ (blue),
-	\item IID sequences with arbitrary $n$ (orange), and
-	\item Randomized LD (Sobol') sequences with $n = 1, 2, \ldots, 2^m, \ldots $ (green).
+	\item Cartesian grids, $\{1/(2m), \ldots, (2m-1)/(2m) \}^6$, for $m = 2, 3, \ldots$ and $n = m^6$ (blue dots),
+	\item IID sequences with arbitrary $n$ (orange diamonds), and
+	\item Randomized LD (Sobol') sequences with $n = 1, 2, \ldots, 2^m, \ldots $ (green squares).
 \end{enumerate}
+There are $50$ replications for each of the latter two cases with the average errors being plotted as trend lines.
 
@@
 	\centering
-	\includegraphics[width =10cm]{\figpath/n_is_16777216_d_is_6_n_rep_is_50_KeisterErrors.pdf}
+	\includegraphics[width =0.7\textwidth]{\figpath/n_is_16777216_d_is_6_n_rep_is_50_KeisterErrors.pdf}
 	\caption{The relative error of approximating $\mu$ defined in \eqref{eq:keisterC} by the sample mean, $\hat{\mu}_n$, defined in \eqref{eq:sample_mean} for various choices of nodes, $\bsx_0, \bsx_1, \ldots$.
 Grids have the largest error and LD nodes have the smallest error.
+The numerical computations behind this figure and the others in the is chapter may be reproduced using the Jupyter notebook \url{??}.
 \label{fig:keister-err}}
@@
 
-	\item \emph{Grid nodes.} Although grids may be attractive for low dimensional problems, e.g., $d = 1$, $2$, or $3$, they do poorly for this modest dimension, $d=6$.
+	\item \emph{Grids.} Although they may be attractive for low dimensional problems, e.g., $d = 1$, $2$, or $3$, grids do poorly for this modest dimension, $d=6$.
 Figure \ref{fig:grid} compares a $64$ point grid with $d = 2$ and $6$.
@@
 
-	\item \emph{IID nodes.}  Simple MC (orange) is a substantial improvement over grid nodes. As a reminder, for IID nodes the root mean squared error is
+	\item \emph{IID.}  Simple MC (orange diamonds) is a substantial improvement over grid nodes. As a reminder, for IID nodes the root mean squared error is
 	\begin{equation}\label{eq:IIDerror}
@@
 
-	\item \emph{LD nodes.}  For qMC methods the error decays nearly like $\mathcal{O}(n^{-1})$, which for this example correponds to a reduction in error of several orders of magnitude compared to simple MC for large enough $n$.
+	\item \emph{LD.}  For qMC methods the error decays nearly like $\mathcal{O}(n^{-1})$, which for this example correponds to a reduction in error of several orders of magnitude compared to simple MC for large enough $n$.
 Typically, LD sequences have preferred sample sizes.
@@
 	\centering
-	\includegraphics[width = 0.48\textwidth]{\figpath/64gridpts_d2.pdf}\quad
-	%\includegraphics[width = 5cm]{\figpath/64gridpts_d3.pdf}\quad
-	\includegraphics[width =  0.48\textwidth]{\figpath/64gridpts_d6.pdf}
+	\includegraphics[width = 0.3\textwidth]{\figpath/64gridpts_d2.pdf}\qquad
+	\includegraphics[width =  0.3\textwidth]{\figpath/64gridpts_d6.pdf}
 	\caption{Although a two-dimensional grid covers the unit square rather well, a $d$-dimensional grid for modest $d$ does \emph{not} cover the unit cube well.
@@
 This section introduces some of the most popular LD sequences.
-What sets these apart from IID sequences is that the nodes are deliberately chosen and highly correlated to one another, whether they be deterministic or randomized.
+What sets these apart from IID sequences is that the nodes are deliberately chosen and highly correlated, whether they be deterministic or randomized.
+\texttt{qmcpy} \cite{QMCPy2020a} is a Python library that contains many of the constructions discussed here, as well as the stopping criteria discussed in Section \ref{sec:stop}.
 
 \subsection{Lattice Sequences} \label{sec:lattice}
-One of the simplest LD constructions is the family of good lattice points \cite{DicEtal22a,SloJoe94}.  As a finite node set, they are defined as follows:
+One of the simplest LD constructions is the family of good lattice points \cite{DicEtal22a,SloJoe94}.  As a finite node set, they are defined as
 \begin{equation} \label{eq:latticepts}
-	\bsx_i = i\bsh/n \pmod \bsone, \qquad i = 0, \ldots, n-1,
+	\bsx_i^{\text{lat}} = i\bsh/n \pmod \bsone, \qquad i = 0, \ldots, n-1,
 \end{equation}
@@
 	\includegraphics[width = \textwidth]{\figpath/16_lattice_construct_d2.pdf}
-	\caption{The construction of a good lattice point set in two dimensions with $16$ nodes is obtained by moving $\bsh/n$ beyond the present point and wrapping around the boundaries of the square until one returns to the origin.
+	\caption{The construction of a good lattice point set in two dimensions with $16$ nodes is obtained by moving $\bsh/n$ beyond the present point and wrapping around the boundaries of the square (hypercube in $d$ dimensions) until one returns to the origin.
 \label{fig:latticeconstruct}}
@@
 
-One disadvantage of this construction is that it is not readily extensible.
-For the example in Figure \ref{fig:latticeconstruct} the first $8$ points do not fill the unit square well.
-However, the set of points defined by even $i$ do a reasonable job.
+One disadvantage of this construction is that it is not extensible.
+For the example in Figure \ref{fig:latticeconstruct} the first $8$ nodes do not fill the unit square well.
+However, the set of nodes defined by even $i$ do a reasonable job.
 This suggests a method for defining extensible lattice sequences that was proposed independently by \cite{Mai81a} and \cite{HicEtal00}.
 
-Let's start with the one-dimensional extensible lattice that is defined by the van der Corput sequence, $\{\phi_b(0), \phi_b(1), \ldots\}$.
+The one-dimensional extensible lattice is defined by the van der Corput sequence, $\{\phi_b(0), \phi_b(1), \ldots\}$.
 This involves reflecting the digits of the integer $i$ in base $b$ about the decimal point.
@@
 \begin{equation} \label{eq:extensiblelattice}
-	\bsx_i =\phi_b( i)\bsh \pmod \bsone, \qquad i = 0, 1 , \ldots.
+	\bsx_i^{\text{lat}} =\phi_b( i)\bsh \pmod \bsone, \qquad i = 0, 1 , \ldots.
 \end{equation}
-This reordering of points from the original construction allows us to preserve the lattice structure for the first $b^m$ points for any non-negative integer $m$.
-That is, $\{\bsx_0, \ldots, \bsx_{2^m-1}\}$ is a closed under addition modulo $\bsone$.
+This reordering of nodes from the original construction allows us to preserve the lattice structure for the first $b^m$ nodes for any non-negative integer $m$.
+That is, $\{\bsx_0, \ldots, \bsx_{b^m-1}\}$ is a closed under addition modulo $\bsone$.
 
-Figure \ref{fig:extensiblelatticeconstruct} shows the first $4$, $8$, and $16$ points of the earlier example in Figure \ref{fig:latticeconstruct}.
+Figure \ref{fig:extensiblelatticeconstruct} shows the first $4$, $8$, and $16$ nodes of an extensible lattice in base $2$ with the same generator as in Figure \ref{fig:latticeconstruct}.
 The blue nodes are a copy of the nodes in the plot to the left, and the the orange nodes correspond to a shifted copy of the blue nodes modulo $\bsone$.
@@
 	\centering
-	\includegraphics[width = \textwidth]{\figpath/16_extensible_lattice_construct_d2.pdf}
+	\includegraphics[width = 0.7\textwidth]{\figpath/16_extensible_lattice_construct_d2.pdf}
 	\caption{An extensible lattice corresponding to  Figure \ref{fig:latticeconstruct}  with the nodes reordered using the van der Corput sequence in base $2$.
@@
 
-For the example in Figure \ref{fig:extensiblelatticeconstruct}, increasing $n$ beyond $16$ repeats the original points because the generating vector $\bsh$ contains integers all less than $16$.
+For the example in Figure \ref{fig:extensiblelatticeconstruct}, increasing $n$ beyond $16$ repeats the original nodes because the generating vector, $\bsh$, contains integers all less than $16$.
 Obtaining a truly extensible lattice sequence requires that $\bsh$ be a vector of generalized integers---essentially integers with infinite numbers of nonzero digits---as explained in \cite{HicNie03a}, where the existence of good generating vectors is also proved.
@@
 
-Another family of LD sequences also built upon the van der Corput sequence \eqref{eq:vdc} is digital sequences \cite{DicPil10a,Nie92}.
+Another family of LD sequences built upon the van der Corput sequence \eqref{eq:vdc} is digital sequences \cite{DicPil10a,Nie92}.
 For simplicity, we restrict ourselves to $b = 2$ and let $\oplus$ denote binary digitwise addition modulo $2$ (also known as digitwise exclusive or), e.g., $3/8 \oplus 3/4 = {}_20.011 \oplus {}_20.110 = {}_20.101 = 5/8$.
@@
 \begin{equation} \label{eq:digital}
-	\bsx_i := i_0 \bsx_1 \oplus i_1 \bsx_2 \oplus i_2 \bsx_4 \oplus \cdots \in [0,1)^d \quad \text{for }
+	\bsx_i^{\text{dig}} := i_0 \bsx^{\text{dig}} _1 \oplus i_1 \bsx^{\text{dig}} _2 \oplus i_2 \bsx^{\text{dig}} _4 \oplus \cdots \in [0,1)^d \quad \text{for }
 	i = i_0 + i_12 + i_2 2^2 + \cdots,
 \end{equation}
-where $\bsx_1, \bsx_2, \ldots$ are carefully chosen.  The node set $\{ \bsx_0, \ldots, \bsx_{2^m-1}\}$  for non-negative integer $m$ is called a digital net and is closed under $\oplus$.
+where $\bsx_1, \bsx_2, \ldots$ are carefully chosen.
+(This is the van  der Corput sequence in base $2$ for $d=1$ and $x_\ell = 2^{-\ell}$.) The node set $\{ \bsx_0, \ldots, \bsx_{2^m-1}\}$  for non-negative integer $m$ is called a digital net and is closed under $\oplus$.
 
@@
 \begin{equation} \label{eq:digitalB}
-	\begin{pmatrix} x_{ij1} \\ x_{ij2} \\ \vdots \end{pmatrix}
+\renewcommand{\arraystretch}{1.5}
+	\begin{pmatrix} x^{\text{dig}}_{ij1} \\ x^{\text{dig}}_{ij2} \\ \vdots \end{pmatrix}
 	= \mathsf{C}_j \begin{pmatrix} i_0 \\ i_1 \\ \vdots \end{pmatrix} \quad \pmod{2}, \qquad j = 1,\ldots, d, \ i = 0, 1, \ldots.
@@
 
-To understand how these points might be evenly distributed over $[0,1)^d$, imagine a box of the form
+To understand how these nodes might be evenly distributed over $[0,1)^d$, imagine a box of the form
 \begin{multline} \label{eq:element_box}
 	[a_1 b^{-k_1}, (a_1 + 1) 2^{-k_1}) \times \cdots \times [a_d b^{-k_d}, (a_d + 1) 2^{-k_d}), \\
-	a_j \in \{0, \ldots, 2^{k_j} - 1\}, \ k_j \in \{0,1, \ldots \}, \qquad j =1, \ldots d.
+	a_j \in \{0, \ldots, 2^{k_j} - 1\}, \ k_j \in \{0,1, \ldots \}, \qquad j =1, \ldots d, \ \bsk \in \NN_0.
 \end{multline}
 This box has volume $2^{-(k_1 + \cdots +k_d)} = 2^{-\lVert\bsk\rVert_1}$.
-For a digital net with $2^m$ points and $m \ge \lVert \bsk \rVert_1$, a ``fair share" of nodes for this box would be $2^{m- \lVert \bsk \rVert_1}$.
+For a digital net with $2^m$ nodes and $\lVert \bsk \rVert_1 \le m$, a ``fair share" of nodes for this box would be $2^{m- \lVert \bsk \rVert_1}$.
 This will occur if and only if
@@
 Thus, the $t$-value cannot be $0$.
-However, the $t$-value is $1$ because all boxes of the form  \eqref{eq:element_box} with $\lVert\bsk\rVert_1 = 2$, i.e., a volume of $2^{-2} = 1/4$ contain a fair share of $2^{m- \lVert \bsk \rVert_1} = 2$ points.
+However, the $t$-value is $1$ because all boxes of the form  \eqref{eq:element_box} with $\lVert\bsk\rVert_1 = 2$, i.e., a volume of $2^{-2} = 1/4$ contain a fair share of $2^{m- \lVert \bsk \rVert_1} = 2$ nodes.
 The node set \eqref{eq:smallnet} is a $(1,3,3)$-net.
@@
 	\includegraphics[width = \textwidth]{\figpath/8_sobol_sequence_elementary_intervals_d3_coord_(1,3)}
-	\caption{Each box in the two dimensional projections of the node set \eqref{eq:smallnet} plotted above show one point in each box except for the second row, third plot from the left.
+	\caption{Each box in the two dimensional projections of the node set \eqref{eq:smallnet} plotted above show one node in each box except for the second row, third plot from the left.
 Thus, this node set is not a $(0,3,3)$-net.
@@
 \begin{equation}\label{eq:Halton}
-	\bsx_i = \bigl( \phi_{b_1}(i), \ldots, \phi_{b_d}(i) \bigr), \qquad i = 0, 1, \ldots,
+	\bsx^{\text{Hal}}_i = \bigl( \phi_{b_1}(i), \ldots, \phi_{b_d}(i) \bigr), \qquad i = 0, 1, \ldots,
 \end{equation}
-where $b_1, \ldots, b_d$ is a choice of $d$ distinct prime numbers.  Often they are chosen as the first $d$ prime numbers.  Figure \ref{fig:Halton} shows two dimensional projections of Halton points
+where $b_1, \ldots, b_d$ is a choice of $d$ distinct prime numbers.  Often they are chosen as the first $d$ prime numbers.  Figure \ref{fig:Halton} shows two dimensional projections of Halton nodes
 
@@
 	\includegraphics[width=\textwidth]{\figpath/64Halton_pts_d6.pdf}
-	\caption{Halton points as defined in \eqref{eq:Halton} have low discrepancy but no preferred sample size. \label{fig:Halton}}
+	\caption{Halton nodes as defined in \eqref{eq:Halton} have low discrepancy but no preferred sample size. \label{fig:Halton}}
 \end{figure}
@@
 Most quasi-Monte Carlo constructions are studied based on their asymptotic behavior as \(n \to \infty\), with numerous results on asymptotic rates \cite{NIED1978QMCBook,NovWoz10a}.
-However, few methods address finding optimal point configurations for specific \(n\) and \(d\) which can be critical in practical applications where each function evaluation is costly.
-Of course, one possible strategy is to truncate asymptotically optimal sequences to the desired \(n\), however, as discussed above many of the standard LD constructions have a preferred number of points with truncations leading to imbalances in equidistribution.
+However, few methods address finding optimal node configurations for specific \(n\) and \(d\) which can be critical in practical applications where each function evaluation is costly.
+Of course, one possible strategy is to truncate asymptotically optimal sequences to the desired \(n\), however, as discussed above many of the standard LD constructions have a preferred number of nodes with truncations leading to imbalances in equidistribution.
 We can often do better optimizing directly for the target \(n\) and \(d\).
 
-When optimizing LD nodes, one requires an objective function which in some sense must assesses the quality of the LD nodes themselves, or the expected performance when implemented in a qMC estimate.
-We will refer to such an objective function as a figure-of-merit (FOM) and popular choices include the discrepancy (see Section \ref{sec:discrepancy}) or worst-case error metrics when considering a specific space of functions; see \cite{Hic00a,LatNet}.
+When optimizing LD nodes, one requires an objective function which in some sense must assesses the quality of the LD nodes directly, or the expected performance when the node set is implemented in a qMC estimate \eqref{eq:sample_mean}.
+We will refer to such an objective function as a figure-of-merit (FOM) and popular choices include the discrepancy (see Section \ref{sec:discrepancy}) or worst-case error metrics when restricted a specific space of functions; see \cite{Hic00a,LatNet}.
 
@@
 More recently, Rusch, Kirk, Bronstein, Lemieux and Rus \cite{ruschkirk24} developed the first machine-learning assisted optimization tool for LD node optimization in the Message-Passing Monte Carlo (MPMC) method.
-MPMC implements tools from geometric deep learning imposing a computational graph onto an initialized IID node set to transform the points to LD node set via a learned mapping, again guided by the minimization of a discrepancy-based FOM.
+MPMC implements tools from geometric deep learning imposing a computational graph onto an initialized IID node set to transform the nodes to LD node set via a learned mapping, again guided by the minimization of a discrepancy-based FOM.
 
@@
 There are several other attempts to construct optimal LD nodes, many of which are well summarized in the thesis of Cl\'{e}ment \cite{}.
-We highlight the subset selection method proposed in \cite{cle22} to choose from an $n$-element initialization set, the $k<n$ points which minimize a discrepancy FOM.
+We highlight the subset selection method proposed in \cite{cle22} to choose from an $n$-element initialization set, the $k<n$ nodes which minimize a discrepancy FOM.
 A heuristic approach to this problem was later shown in \cite{cle24}.
-%Furthermore, a method to generate optimal star-discrepancy point sets for fixed $n$ and $d$ based on a non-linear programming approach was suggested in \cite{}.
+%Furthermore, a method to generate optimal star-discrepancy node sets for fixed $n$ and $d$ based on a non-linear programming approach was suggested in \cite{}.
 
@@
 
-While determinism has some advantages, randomization has advantages as well.
+While determinism has some advantages, randomization has substantial advantages as well.
 \begin{itemize}
@@
 \end{itemize}
-The key to good randomization is to preserve the LD quality of the node sequence.
+The key to good randomization is to preserve the LD quality of the node sequence.  In general, we recommend randomization, if available.
 
@@
 \begin{equation} \label{eq:sh_extensiblelattice}
-	\bsx^{\bsDelta\text{-lat}}_i := \bsx_i + \bsDelta \pmod \bsone =\phi_b( i)\bsh + \bsDelta \pmod \bsone, \qquad i = 0, 1 , \ldots.
+	\bsx^{\bsDelta\text{-lat}}_i := \bsx_i^{\text{lat}} + \bsDelta \pmod \bsone =\phi_b( i)\bsh + \bsDelta \pmod \bsone, \qquad i = 0, 1 , \ldots.
 \end{equation}
@@
 For digital sequences the shift should be applied using digitwise addition.  A digital shift of the digital sequence defined in  \eqref{eq:digital} then becomes
-\begin{equation} \label{eq:sh_digital}
-	\bsx_i^{\bsDelta\text{-dig}} := \bsx_i \oplus \bsDelta = i_0 \bsx_1 \oplus i_1 \bsx_2 \oplus i_2 \bsx_4 \oplus \cdots \oplus \bsDelta \in [0,1)^d \quad \text{for }
+\begin{multline} \label{eq:sh_digital}
+	\bsx_i^{\bsDelta\text{-dig}} := \bsx_i^{\text{dig}} \oplus \bsDelta = i_0 \bsx^{\text{dig}}_1 \oplus i_1 \bsx^{\text{dig}}_2 \oplus i_2 \bsx^{\text{dig}}_4 \oplus \cdots \oplus \bsDelta \in [0,1)^d \\
+    \text{for }
 	i = i_0 + i_12 + i_2 2^2 + \cdots.
-\end{equation}
+\end{multline}
 Three digital shifts of the same net are given in Figure \ref{fig:shift_net}.  A digital shift does not alter the $t$-value of a $(t,m,d)$-net.
@@
 
-Let $\{\bsx_i^{\bsDelta}\}_{i=0}^{n-1}$ denote a random shift of any deterministic original node set, $\{\bsx_i \}_{i=0}^{n-1}$ as described above.
+Let $\{\bsx_i^{\bsDelta}\}_{i=0}^{n-1}$ denote a uniform (digital or modulo $\bsone$) random shift of any deterministic original node set, $\{\bsx_i \}_{i=0}^{n-1}$.
 That is, $\bsx_i^{\bsDelta} : = \bsx_i + \bsDelta \pmod \bsone$ or  $\bsx_i^{\bsDelta} : = \bsx_i \oplus \bsDelta$ and $\bsDelta \sim \calU[0,1)^d$.
@@
 
-As Halton points can be viewed as digital nets with different bases for each dimension, the randomization techniques used for digital nets can be applied to Halton pointsets.
+As Halton nodes can be viewed as digital nets with different bases for each dimension, the randomization techniques used for digital nets can be applied to Halton pointsets.
 Random digital shifts for Halton set $x_{ij\ell} \mapsto (x_{ij\ell} + \Delta_{j\ell}) \mod b_j$ where the shift $\Delta_{j\ell} \in \{0,\dots,b_j-1\}$.
@@
 
-The centered discrepancy also has a geometric interpretation, as is described in \cite{Hic97a}. Other choices of kernels and their geomeetric interpretations are explained there as well.
+The centered discrepancy also has a geometric interpretation, as is described in \cite{Hic97a}. Other choices of kernels and their geometric interpretations are explained there as well.
 
@@
 For IID samples the root mean squared discrepancy for a general reproducing kernel, $K$, is 
-\begin{multline}
-    \mathbb{E}[\text{discrepancy}^2\bigl(\{\bsx\}_{i=0}^{n-1},K \bigr)]  \\
+\begin{multline} \label{eq:discIID}
+    \sqrt{\mathbb{E}[\text{discrepancy}^2\bigl(\{\bsx\}_{i=0}^{n-1},K \bigr)] } \\
     = \frac 1{\sqrt{n}}  \sqrt{  \int_{[0,1)^d} K(\bsx,\bsx) \, \mathrm{d} \bsx - 
@@
 
+Under modest smoothness assumptions, the discrepancy decays nearly as fast as $\mathcal{O}(n^{-1})$ for typical LD sequences.
+However, when the reproducing kernel, $K$, is constructed to ensure that functions in $\calH_K$ have mixed partial derivatives of up to order $r$ in each coordinate and are periodic in all lower order derivatives, then the discrepancy of lattices can achieve a nearly $\mathcal{O}(n^{-r})$ decay rate.
+
 \subsection{Randomized Error and Bayesian Error}
@@
 If $\hat{\mu}_n$ is unbiased, then $\sqrt{\mathbb{E}[(\mu(f) - \hat{\mu}_n(f))^2]} = \text{std}(\hat{\mu}_n(f))$.
-For IID nodes we know that this is $\text{std}(f(\bsX))/\sqrt{n}$.
+For IID nodes we know that this is $\text{std}(f(\bsX))/\sqrt{n}$, which has the same order of decay as root mean squared discrepancy of an IID node set in \eqref{eq:discIID}, but a smaller constant.
 
-For randomly shifted lattices
+For randomly shifted lattices $\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ tends to have the same rate of decay.
+However, for randomly scrambled and shifted digital nets, there can be an improvement
 
+The symmetric, positive definite reproducing kernel, $K$, which appears in the worst-case error analysis in Section \ref{sec:kerdisc} may instead take the role of the covariance kernel of a Gaussian process with zero mean \cite{}.
 Bayesian

### 2025-01-20 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

@@
 
-\vspace{2mm}
-\noindent
-\textbf{Combinatorial Optimization.}
 For several decades, combinatorial optimization techniques have been employed to obtain good generating parameters for LD constructions.
@@
 
-\vspace{2mm}
-\noindent
-\textbf{Global Optimization.} Global optimization of LD nodes is often challenging due to the non-convex nature of the FOM as a function of the nodes.
+It is commonplace in qMC constructions that equidistribution is required not just in the (possibly high dimensional) $d$-dimensional space, but also for all or specific low-dimensional projections; see Section \ref{sec:coordwts}.
+MatBuilder \cite{paulin2022} is a software tool developed precisely for this task; to generate digital sequence LD nodes when low-dimensional uniformity is critical.
+Uniformity constraints related to digital sequence stratification across boxes of the form \eqref{eq:element_box} are expressed as integer linear programming (ILP) problems and subsequently solved to optimize an objective function relating to the ILP reformulation.
+The solutions then determine the next columns in the generator matrices for the digital sequence.
+
+As discussed above, the Halton sequence has the benefit of not having a preferred number of points for its construction.
+It does however possess the drawback of exhibiting poor uniformity in the lower dimensional projections when $d$ is moderate to high due to intrinsic number theoretic correlation between the coordinates \cite{kirklem24}.
+One popular strategy to circumvent this poor behavior, dating back to Faure \cite{fau78} and Braaten and Weller \cite{brawell79}, is to randomly or deterministically permute the digits of the base $b_j$ expansions $j = 1,2\ldots,d$ to break the observed correlation between the components.
+To do so, we first define the generalized van der Corput sequence by obtaining a sequence $\Sigma = (\sigma_r)_{r\geq0}$ of permutations of $\{0,\ldots,b-1\}$ where the $i^{th}$ term of the one-dimensional sequence is defined as
+\begin{multline}\label{eq:gen_vdc}
+    \phi_b^{\Sigma}(i_0 + i_1b + i_2 b^2 + \cdots ) := \sigma_0(i_0) b^{-1} + \sigma_1(i_1) b^{-2} + \sigma_2(i_2) b^{-3} + \cdots \in [0,1)
+	\\
+	 \text{where } i_0, \sigma_0(i_0), i_1, \sigma_1(i_1), \ldots \in \{0,\ldots b-1\}.
+\end{multline}
+The generalized Halton sequence is subsequently defined by choosing $d$ sequences of permutations $\Sigma_j = (\sigma_{j,r})_{r\geq0}, j = 1, 2,\ldots,d$ and setting
+$$\bsx_i^{\text{G-Hal}} = \left( \phi^{\Sigma_1}_{b_1}(i), \ldots, \phi^{\Sigma_d}_{b_d}(i)\right), \quad i= 0,1,\ldots,$$
+There exist many attempts to optimize the permutations defining the generalized sequence, most of which involve so-called linear digit scramblings.
+We point the reader towards \cite{} and the excellent survey of the work in this direction (until 2008) by Faure and Lemieux \cite{}.
+
+So far, we have discussed the attempts of researchers to optimize specific classes of LD constructions; lattices and digital sequences, and Halton sequences.
+There also exist several efforts, some very recently, to generate novel LD constructions which don't follow any number theoretic or any other underlying structure, but none-the-less, possess a high level of uniformity.
+These methods are often global optimization methods, which is often challenging due to the non-convex nature of the FOM as a function of the nodes.
 Traditional global optimization methods, such as gradient descent or its variants, struggle in this setting becoming trapped in suboptimal local minima and failing to identify the global optimum configuration of nodes.
-
 One of the first works to globally optimize LD nodes, albeit a heuristic approach, was by Winker and Fang \cite{WinFan97b} employing the threshold accepting optimization algorithm to directly optimize the uniformity of the LD nodes with respect to the discrepancy FOM.
-
 More recently, Rusch, Kirk, Bronstein, Lemieux and Rus \cite{ruschkirk24} developed the first machine-learning assisted optimization tool for LD node optimization in the Message-Passing Monte Carlo (MPMC) method.
 MPMC implements tools from geometric deep learning imposing a computational graph onto an initialized IID node set to transform the nodes to LD node set via a learned mapping, again guided by the minimization of a discrepancy-based FOM.
-
-Lastly, we highlight the method proposed by Hinrichs and Oettershagen \cite{hinoet16}, which generates optimal LD nodes for a specific class of functions in two-dimensions.
-To address the non-convexity challenge, the authors decompose the global optimization problem into a large number of smaller, convex sub-problems.
-
-\vspace{2mm}
-\noindent
-\textbf{Other Notable Methods.} 
+We also highlight the method proposed by Hinrichs and Oettershagen \cite{hinoet16}, which generates optimal LD nodes for a specific class of functions in two-dimensions.
+To address the non-convexity challenge, the authors decompose the global optimization of a specific FOM into a large number of smaller, convex sub-problems which are solved individually.
 There are several other attempts to construct optimal LD nodes, many of which are well summarized in the thesis of Cl\'{e}ment \cite{}.
@@
 A heuristic approach to this problem was later shown in \cite{cle24}.
-%Furthermore, a method to generate optimal star-discrepancy node sets for fixed $n$ and $d$ based on a non-linear programming approach was suggested in \cite{}.
-
-It is commonplace in qMC constructions that equidistribution is required not just in the (possibly high dimensional) $d$-dimensional space, but also for all or specific low-dimensional projections; see Section \ref{sec:coordwts}.
-MatBuilder \cite{paulin2022} is a software tool developed precisely for this task; to generate digital sequence LD nodes when low-dimensional uniformity is critical.
-Uniformity constraints related to digital sequence stratification across boxes of the form \eqref{eq:element_box} are expressed as integer linear programming (ILP) problems and subsequently solved, while optimizing an objective function relating to the ILP reformulation, to determine the next columns in the generator matrices for the digital sequence.
+Furthermore, a method to generate optimal star-discrepancy node sets for fixed $n$ and $d$ based on a non-linear programming approach was suggested in \cite{}.
 
@@
 \subsection{Variation Reduction}
-\subsection{Multilevel Methods}
+\subsection{Multilevel Methods} \label{sec:multi}
+The cost of computing the sample mean defined in \eqref{eq:sample_mean} is typically $\mathcal{O}(dn)$ as the dimension, $d$, and the sample size, $n$, tend to infinity.
+Some applications have $d$ either huge or infinite.
+On example is in quantitative finance, where $d$ denotes the number of times that the asset price is monitored over the life of a financial derivative times the number of assets involved.
+However, there is a way to reduce the cost of approximating $\mu$ by employing multilevel methods \cite{}.
+
+Briefly, let $\{Y_l := f_l(\bsX_{l})\}_{l=1}^L$ be a sequence of approximations to $Y = f(\bsX)$, where $\bsX_l \sim \mathcal{U}[0,1)^{d_l}$.
+Here we assume that $d_1 < \cdots < d_L$, and that the cost of evaluating $f_l(\bsx_{l})$ is $\mathcal{O}(d)$.
+Also, $d_L = d$, $Y_{d_L} = T$, and $f_L = f$.
+Then
+\begin{align*}
+    \mu & := \mathbb{E}(Y) = \mathbb{E}\bigl[f(\bsX)\bigr] 
+    = \mathbb{E}(Y_1) + \mathbb{E}(Y_2 - Y_1) + \cdots + \mathbb{E}(Y_{d_{L-1}} - Y_d) \\
+    & = \int_{[0,1)^{d_1}} f_1(\bsx^{(1)}) \, \mathrm{d} \bsx^{(1)} +  \int_{[0,1)^{d_2}} \bigl[f_2(\bsx^{(2)}) - f_1(\bsx^{(1)})\bigr] \, \mathrm{d} \bsx^{(2)} + \cdots \\
+    & \qquad \qquad + \int_{[0,1)^{d_L}} \bigl[f_L(\bsx^{(L)}) - f_{L-1}(\bsx^{(L-1)}) \bigr] \, \mathrm{d} \bsx^{(L)}
+    \intertext{where}
+    \bsx^{(1)} & = (x_1, \dots, x_{d_1}), \quad \bsx^{(2)} = (x_1, \dots, x_{d_2}), \ldots, \bsx^{(L)} = (x_1, \dots, x_{d_L}).
+\end{align*}
+This series of integrals may be approximated by a series of sample means, namely
+\begin{multline*}
+    \hat{\mu}^{\text{ML}}_{\bsn,\bsd}  := \frac{1}{n_1}  \sum_{i=0}^{n_1-1} f_1(\bsx^{(1)}_i)  +  \frac{1}{n_2}  \sum_{i=0}^{n_2-1}  \bigl[f_2(\bsx^{(2)}_i) - f_1(\bsx^{(1)}_i)\bigr] + \cdots \\
+     + \frac{1}{n_L}  \sum_{i=0}^{n_L-1}  \bigl[f_L(\bsx_i^{(L)}) - f_{L-1}(\bsx_i^{(L-1)}) \bigr].
+\end{multline*}
+The computational cost of this approximation is $\mathcal{O}(d_1n_1 + \cdots + d_Ln_L)$.
+
+A worst case error analysis as described in Section \ref{sec:discrepancy} implies that
+\begin{multline*}
+    \bigl \lvert \mu - \hat{\mu}^{\text{ML}}_{\bsn,\bsd} \bigr \rvert \le 
+    \text{discrepancy}\bigl(\{\bsx_i^{(1)}\}_{i=0}^{n_1-1}, K^{(1)} \bigr) \text{variation}(f^{(1)}, K^{(1)}) \\
+    + \text{discrepancy}\bigl(\{\bsx_i^{(2)}\}_{i=0}^{n_2-1}, K^{(2)} \bigr) \text{variation}(f^{(2)} - f^{(1)}, K^{(2)}) + \cdots \\
+    + \text{discrepancy}\bigl(\{\bsx_i^{(L)}\}_{i=0}^{n_L-1}, K^{(L)} \bigr) \text{variation}(f^{(L)} - f^{(L-1)}, K^{(L)}),
+\end{multline*}
+where $K^{(l)}$ is the reproducing kernel corresponding to the relevant  Hilbert space of functions of $d_l$ variables.
 
@@
 
+\end{document}
 \section*{Appendix A}

### 2025-01-20 FindPDFinLaTeX.py
M2 Mac updates

@@ -1,0 +1,44 @@
+import re
+import os
+
+def find_pdf_files_in_latex(latex_file):
+    """
+    Scans a LaTeX file to find PDF files included in the document.
+
+    Args:
+        latex_file (str): Path to the LaTeX file.
+
+    Returns:
+        list: A list of PDF file names found in the LaTeX file.
+    """
+    pdf_files = []
+
+    # Regular expression to capture \includegraphics{file.pdf} or \includegraphics[options]{file.pdf}
+    graphics_regex = re.compile(r"\\includegraphics(?:\[.*?\])?\{(.*?\.pdf)\}")
+
+    try:
+        with open(latex_file, 'r', encoding='utf-8') as file:
+            for line in file:
+                matches = graphics_regex.findall(line)
+                if matches:
+                    pdf_files.extend(matches)
+    except FileNotFoundError:
+        print(f"Error: File '{latex_file}' not found.")
+    except Exception as e:
+        print(f"An error occurred: {e}")
+
+    return pdf_files
+
+# Example usage
+if __name__ == "__main__":
+    # Specify the path to your LaTeX file
+    latex_file_path = "HickernellMCQMC2024_QMCTutorial.tex"
+
+    # Find and print the PDF files included in the LaTeX file
+    pdf_files = find_pdf_files_in_latex(latex_file_path)
+    if pdf_files:
+        print("PDF files used in the LaTeX file:")
+        for pdf in pdf_files:
+            print(pdf)
+    else:
+        print("No PDF files found in the LaTeX file.")

### 2025-01-20 HickernellMCQMC2024_QMCTutorial.tex
M2 Mac updates
Updates from Overleaf

@@
 
-It is commonplace in qMC constructions that equidistribution is required not just in the (possibly high dimensional) $d$-dimensional space, but also for all or specific low-dimensional projections; see Section \ref{sec:coordwts}.
-MatBuilder \cite{paulin2022} is a software tool developed precisely for this task; to generate digital sequence LD nodes when low-dimensional uniformity is critical.
-Uniformity constraints related to digital sequence stratification across boxes of the form \eqref{eq:element_box} are expressed as integer linear programming (ILP) problems and subsequently solved to optimize an objective function relating to the ILP reformulation.
-The solutions then determine the next columns in the generator matrices for the digital sequence.
+It is commonplace in qMC constructions that equidistribution is required not just in the (possibly high dimensional) $d$-dimensional space, but also for all or specific low-dimensional projections.
+MatBuilder \cite{paulin2022} is a software tool developed precisely for this task; to generate digital sequence LD nodes when low dimensional uniformity is critical.
+Uniformity constraints related to digital sequence stratification across boxes of the form \eqref{eq:element_box} are expressed as integer linear programming (ILP) problems and subsequently solved to optimize an objective function relating to the ILP reformulation ensuring good quality low dimensional projections.
+The solutions determine then build the columns of the generator matrices in a greedy manner for the LD digital sequence.
 
-As discussed above, the Halton sequence has the benefit of not having a preferred number of points for its construction.
+As discussed in the previous section, the Halton sequence has the benefit of not having a preferred number of points for its construction.
 It does however possess the drawback of exhibiting poor uniformity in the lower dimensional projections when $d$ is moderate to high due to intrinsic number theoretic correlation between the coordinates \cite{kirklem24}.
-One popular strategy to circumvent this poor behavior, dating back to Faure \cite{fau78} and Braaten and Weller \cite{brawell79}, is to randomly or deterministically permute the digits of the base $b_j$ expansions $j = 1,2\ldots,d$ to break the observed correlation between the components.
+One popular strategy to circumvent this poor behavior, dating back to Faure \cite{fau78} and Braaten and Weller \cite{brawell79}, is to randomly or deterministically permute the digits of the base $b_j$ expansions $j = 1,2\ldots,d$ to disrupt correlations between the components.
 To do so, we first define the generalized van der Corput sequence by obtaining a sequence $\Sigma = (\sigma_r)_{r\geq0}$ of permutations of $\{0,\ldots,b-1\}$ where the $i^{th}$ term of the one-dimensional sequence is defined as
@@
 $$\bsx_i^{\text{G-Hal}} = \left( \phi^{\Sigma_1}_{b_1}(i), \ldots, \phi^{\Sigma_d}_{b_d}(i)\right), \quad i= 0,1,\ldots,$$
-There exist many attempts to optimize the permutations defining the generalized sequence, most of which involve so-called linear digit scramblings.
-We point the reader towards \cite{} and the excellent survey of the work in this direction (until 2008) by Faure and Lemieux \cite{}.
+There exist many proposed algorithms to optimize the permutations defining the generalized sequence, most of which involve so-called linear digit scramblings and usually involve directly optimizing a discrepancy FOM.
+We point the reader towards \cite{MasChiWar05,tuffin98,Atan04,vancools06} and the excellent survey of the work in this direction (until 2008) by Faure and Lemieux \cite{faulem09}.
 
@@
 Traditional global optimization methods, such as gradient descent or its variants, struggle in this setting becoming trapped in suboptimal local minima and failing to identify the global optimum configuration of nodes.
-One of the first works to globally optimize LD nodes, albeit a heuristic approach, was by Winker and Fang \cite{WinFan97b} employing the threshold accepting optimization algorithm to directly optimize the uniformity of the LD nodes with respect to the discrepancy FOM.
+One of the first works to globally optimize LD nodes, albeit a heuristic approach, was by Winker and Fang \cite{WinFan97b} employing the threshold accepting optimization algorithm to directly optimize the uniformity of the LD nodes with respect to a discrepancy FOM.
 More recently, Rusch, Kirk, Bronstein, Lemieux and Rus \cite{ruschkirk24} developed the first machine-learning assisted optimization tool for LD node optimization in the Message-Passing Monte Carlo (MPMC) method.
@@
 To address the non-convexity challenge, the authors decompose the global optimization of a specific FOM into a large number of smaller, convex sub-problems which are solved individually.
-There are several other attempts to construct optimal LD nodes, many of which are well summarized in the thesis of Cl\'{e}ment \cite{}.
+There are several other recent attempts to construct optimal LD nodes, many of which are well summarized in the thesis of Cl\'{e}ment \cite{clethesis24}.
 We highlight the subset selection method proposed in \cite{cle22} to choose from an $n$-element initialization set, the $k<n$ nodes which minimize a discrepancy FOM.
 A heuristic approach to this problem was later shown in \cite{cle24}.
-Furthermore, a method to generate optimal star-discrepancy node sets for fixed $n$ and $d$ based on a non-linear programming approach was suggested in \cite{}.
+Furthermore, a method to generate optimal star-discrepancy node sets for fixed $n$ and $d$ based on a non-linear programming approach was suggested in \cite{cle24}, however is computationally restricted to very small numbers of points in dimensions two and three only.
 

### 2025-01-20 FindPDFinLaTeX.py
M2 Mac updates

@@ -27,7 +27,8 @@ def find_pdf_files_in_latex(latex_file):
     except Exception as e:
         print(f"An error occurred: {e}")
 
-    return pdf_files
+    #return pdf_files
+    return sorted(pdf_files)
 
 # Example usage
 if __name__ == "__main__":

### 2025-01-20 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

@@
 %%End of Fred's temporary macros
+
+%%Figures used in this document
+\iffalse
+\figpath/16_extensible_lattice_construct_d2.pdf
+\figpath/16_lattice_construct_d2.pdf
+\figpath/16_shifted_lattice_construct_d2.pdf
+\figpath/64Halton_pts_d6.pdf
+\figpath/64gridpts_d2.pdf
+\figpath/64gridpts_d6.pdf
+\figpath/64iidpts_d6.pdf
+\figpath/64sobolpts_d6.pdf
+\figpath/IIDSobolScaledWtCtrDisc_n_8192_d64.pdf
+\figpath/SobolCtrDisc_n_8192_d8.pdf
+\figpath/centered_discrepancy_kernel.pdf
+\figpath/n_8_n_scr_3_scrambled_sobol_sequence_d2.pdf
+\figpath/n_8_n_sh_3_shifted_sobol_sequence_d2.pdf
+\figpath/n_is_16777216_d_is_6_n_rep_is_50_KeisterErrors.pdf
+
+Using pkl files 
+Keister_d_6_n_max_16777216_n_rep_50.pkl
+CtrDisc_dmax_64_nmax_8192.pkl
+\fi
 
@@
 	\centering
-	\includegraphics[width = \textwidth]{\figpath/8_sobol_sequence_elementary_intervals_d3_coord_(1,2)}
-	\includegraphics[width = \textwidth]{\figpath/8_sobol_sequence_elementary_intervals_d3_coord_(1,3)}
+	\includegraphics[width = \textwidth]{\figpath/8_sobol_sequence_elementary_intervals_d3_coord_(1,2).pdf}
+	\includegraphics[width = \textwidth]{\figpath/8_sobol_sequence_elementary_intervals_d3_coord_(1,3).pdf}
 	\caption{Each box in the two dimensional projections of the node set \eqref{eq:smallnet} plotted above show one node in each box except for the second row, third plot from the left.

### 2025-01-21 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

@@
 \figpath/64sobolpts_d6.pdf
+\figpath/8_sobol_sequence_elementary_intervals_d3_coord_(1,2).pdf
+\figpath/8_sobol_sequence_elementary_intervals_d3_coord_(1,3).pdf
 \figpath/IIDSobolScaledWtCtrDisc_n_8192_d64.pdf
@@
 	\centering
-	\includegraphics[width =\textwidth]{\figpath/64sobolpts_d6.pdf}
+	\includegraphics[width = \textwidth]{\figpath/64sobolpts_d6.pdf}
 	\caption{LD points cover the unit cube even better than IID points or grids.  In any  projection there is  a similar looking distribution of all $64$ points. \label{fig:ld}}
@@
 	\centering
-	\includegraphics[width = 0.7\textwidth]{\figpath/16_extensible_lattice_construct_d2.pdf}
+	\includegraphics[width = \textwidth]{\figpath/16_extensible_lattice_construct_d2.pdf}
 	\caption{An extensible lattice corresponding to  Figure \ref{fig:latticeconstruct}  with the nodes reordered using the van der Corput sequence in base $2$.
@@
 This series of integrals may be approximated by a series of sample means, namely
-\begin{multline*}
+\begin{multline} \label{eq:MLmuhat}
     \hat{\mu}^{\text{ML}}_{\bsn,\bsd}  := \frac{1}{n_1}  \sum_{i=0}^{n_1-1} f_1(\bsx^{(1)}_i)  +  \frac{1}{n_2}  \sum_{i=0}^{n_2-1}  \bigl[f_2(\bsx^{(2)}_i) - f_1(\bsx^{(1)}_i)\bigr] + \cdots \\
      + \frac{1}{n_L}  \sum_{i=0}^{n_L-1}  \bigl[f_L(\bsx_i^{(L)}) - f_{L-1}(\bsx_i^{(L-1)}) \bigr].
-\end{multline*}
+\end{multline}
 The computational cost of this approximation is $\mathcal{O}(d_1n_1 + \cdots + d_Ln_L)$.
@@
 A worst case error analysis as described in Section \ref{sec:discrepancy} implies that
-\begin{multline*}
-    \bigl \lvert \mu - \hat{\mu}^{\text{ML}}_{\bsn,\bsd} \bigr \rvert \le
-    \text{discrepancy}\bigl(\{\bsx_i^{(1)}\}_{i=0}^{n_1-1}, K^{(1)} \bigr) \text{variation}(f^{(1)}, K^{(1)}) \\
-    + \text{discrepancy}\bigl(\{\bsx_i^{(2)}\}_{i=0}^{n_2-1}, K^{(2)} \bigr) \text{variation}(f^{(2)} - f^{(1)}, K^{(2)}) + \cdots \\
-    + \text{discrepancy}\bigl(\{\bsx_i^{(L)}\}_{i=0}^{n_L-1}, K^{(L)} \bigr) \text{variation}(f^{(L)} - f^{(L-1)}, K^{(L)}),
-\end{multline*}
-where $K^{(l)}$ is the reproducing kernel corresponding to the relevant  Hilbert space of functions of $d_l$ variables.
-
-\section{Ongoing Research} \label{sec:onging}
+\begin{align}
+    \label{eq:MLerrbd}
+    \bigl \lvert \mu - \hat{\mu}^{\text{ML}}_{\bsn,\bsd} \bigr \rvert &\le
+    V_1/n_1 + V_2/n_2 + \cdots + V_L/n_L,  \\
+    \intertext{where} 
+    \nonumber
+    V_1 &= n_1 \, \text{discrepancy}\bigl(\{\bsx_i^{(1)}\}_{i=0}^{n_2-1}, K^{(1)} \bigr) \, \text{variation}(f^{(1)}, K^{(1)})\\
+    \nonumber
+    V_l & = n_l \,\text{discrepancy}\bigl(\{\bsx_i^{(l)}\}_{i=0}^{n_l-1}, K^{(l)} \bigr) \, \text{variation}(f^{(l)} - f^{(l-1)}, K^{(l)}), \\
+    \nonumber
+    & \qquad \qquad l = 2, \ldots, L,
+\end{align}
+and $K^{(l)}$ is the reproducing kernel corresponding to the relevant  Hilbert space of functions of $d_l$ variables.
+The $V_l$ are defined such that we expect them to have little dependence on $n_l$ since we are using quasi-Monte Carlo nodes.
+
+If $C_1$ is the computational cost of $f^{(1)}(\bsx^{(1)})$ and $C_l$ is the computational cost of $f^{(l)}(\bsx^{(1)})$ of $f_L(\bsx_i^{(l)}) - f_{l-1}(\bsx_i^{(l-1)})$ for $l = 2, \ldots, L$, then the total computational cost of
+\eqref{eq:MLmuhat} is 
+\begin{equation}
+    \text{cost}(\bsn,\bsd) = n_1 C_1 + \cdots + n_L C_L.
+\end{equation}
+Optimizing $\bsn$ to minimize the error bound in \eqref{eq:MLerrbd} for a fixed cost and assuming constant $\bsC$ and $\bsV$, yields
+\begin{align}
+    \bigl \lvert \mu - \hat{\mu}^{\text{ML}}_{\bsn,\bsd} \bigr \rvert &\le \varepsilon \\
+\intertext{for}
+    n_l & = \frac{\sqrt{V_l} \bigl(\sqrt{V_1 C_1} + \ldots + \sqrt{V_LC_L}\bigr)}{\varepsilon \sqrt{C_l} }, \qquad l = 1, \ldots, d,\\
+    \text{cost}(\bsn,\bsd) & = \frac{\bigl(\sqrt{V_1 C_1} + \ldots + \sqrt{V_LC_L}\bigr)^2}{\varepsilon}. \label{eq:MLcost}
+\end{align}
+
+To understand this result, suppose that the cost per function value at each level, $l$, is proportional to the dimension, $d_l$, and that the dimensions increase exponentially, i.e.,  $C_l \le \alpha d_1^l$.
+Moreover, suppose that the $V_l$, which are roughly the variation, decrease exponentially, i.e., $V_l \le \beta r^{-l} $.
+Then the computational cost in \eqref{eq:MLcost} can be bounded above by
+\begin{align*}
+     \text{cost}(\bsn,\bsd) & \le \frac{\alpha \beta \bigl((d_1/r)^{-1/2} + \ldots + (d_1/r)^{-L/2}\bigr)^2}{\varepsilon} \\
+     & = \frac{\alpha \beta \bigl((d_1/r)^{1/2} - (d_1/r)^{(L+1)/2}\bigr)^2}{\varepsilon \bigl( 1 - (d_1/r)^{1/2}\bigr)^2}.
+\end{align*}
+If the $V_l$ decay quickly enough, which means that the corrections $f_l - f_{l-1}$ depending on many variables decay quickly enough, in particular, if $r > d_1$, then 
+\begin{equation}
+         \text{cost}(\bsn,\bsd) \le \frac{\alpha \beta d_1} {r \varepsilon \bigl( 1 - (d_1/r)^{1/2}\bigr)^2} = \mathcal{O}(d_1/\varepsilon).
+\end{equation}
+
+\section{Ongoing Research} \label{sec:ongoing}
 \section{Conclusion} \label{sec:conclusion}

### 2025-01-22 HickernellMCQMC2024_QMCTutorial.tex
aleksei: stopping criterion

@@
 Random generating matrices for base $2$ digital nets set $c_{ij\ell} \sim \calU\{0,1\}$ for $i<\ell$ while enforcing $c_{iji}=1$ and $c_{ij\ell}=0$ for $i>\ell$.
-One idea is to generate
 
@@
 
-Figure \ref{fig:ctrKer} is a color map of the reproducing kernel, $K$, defined in \eqref{eq:ctrkernel} for $d=1$.
+Figure \ref{fig:ctrKer} (left) is a color map of the reproducing kernel, $K$, defined in \eqref{eq:ctrkernel} for $d=1$.
 As shown in the figure, the matrix $\mathsf{K} := \bigl( K(x_i,x_j) \bigr)_{i,j=1}^n$ that arises in condition \eqref{eq:KcondPD} is formed by taking $n$ rows and corresponding columns of this plot.
@@
 	\centering
-	\includegraphics[width = \textwidth]{\figpath/centered_discrepancy_kernel.pdf}
-	\caption{A plot of the reproducing kernel for the centered discrepancy defined in \eqref{eq:ctrkernel} for $d=1$.
+	\includegraphics[width = 0.48\textwidth]{\figpath/centered_discrepancy_kernel.pdf} \quad 
+    \includegraphics[width = 0.48\textwidth]{\figpath/SobolCtrDisc_n_8192_d8.pdf}
+	\caption{Left:  A plot of the reproducing kernel for the centered discrepancy defined in \eqref{eq:ctrkernel} for $d=1$.
 Note that the vertical axis is inverted so that the plot is a picture of the ``matrix'' $\bigl( K(t,x) \bigr)_{t,x \in [0,1)]}$.
 Any submatrix is positive definite by property \eqref{eq:KcondPD}.
-\label{fig:ctrKer}}
+\\
+    Right: The centered discrepancy for several modest values of $d$.
+They show an asymptotic decay close to $\calO(n^{-1})$, but this rate is achieved only at higher $n$ for larger values of $d$ \label{fig:ctrKer}}
 \end{figure}
@@
 
-Figure \ref{fig:ctrdiscmodestd} displays the discrepancy for a Sobol' LD digital net for several modest values of $d$ and for $n =  1, 2, 4, \ldots$.
+Figure  \ref{fig:ctrKer} (right)  displays the discrepancy for a Sobol' LD digital net for several modest values of $d$ and for $n =  1, 2, 4, \ldots$.
 The decay rate of nearly $\calO(n^{-1})$ is observed for small $d$ and can be anticipated for a bit larger $d$.
 Thus, the decay of the discrepancy mirrors the convergenc rate for LD sequences for the Kiester example in Section \ref{sec:intro} and Figure \ref{fig:keister-err}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width = \textwidth]{\figpath/SobolCtrDisc_n_8192_d8.pdf}
-	\caption{The centered discrepancy for several modest values of $d$.
-They show an asymptotic decay close to $\calO(n^{-1})$, but this rate is achieved only at higher $n$ for larger values of $d$ \label{fig:ctrdiscmodestd}}
-\end{figure}
 
@@
 
-\section{Stopping Criteria} \label{sec:stop}
-
-Art
-
-Tony
-
-Jags
+\section{Stopping Criteria \textcolor{blue}{Aleksei}} \label{sec:stop}
+
+Monte Carlo approximates the true mean $\mu = \bbE[f(\bsX)]$, $X \sim \mathcal{U}[0,1]^d$ by the sample mean $\hat{\mu} = n^{-1} \sum_{i=0}^{n-1} f(\bsx_i)$ \textcolor{green}{maybe these formulas are already in the paper.}.
+Stopping criterion determine how large $n$ should be so that the error $\lvert \mu - \hat{\mu} \rvert$ is below an error tolerance $\epsilon$ either certainly or with sufficiently low uncertainty $\alpha$.
+The error tolerance and probability threshold are problem dependent.
+For example, in option pricing one may set the error tolerance to a penny, $\varepsilon=0.01$, and wish to find error bounds which hold with 95\% confidence i.e.
+$\alpha=.05$.
+
+For classic Monte Carlo with IID points, the Central Limit Theorem (CLT) may be used to get an approximate confidence interval $(\mu^-,\mu^+)$ where $\mu^\pm = \hat{\mu} \pm Z^{1-\alpha/2} \sigma/\sqrt{n}$ where $\sigma = \mathrm{Std}[f(X)]$ and $Z^{1-\alpha/2}$ is the $1-\alpha/2$ quantile of the standard Gaussian distribution.
+If
+\begin{equation}
+    n > (2 Z^{1-\alpha/2} \sigma/\varepsilon)^2
+    \label{eq:sc_clt_iid_n}
+\end{equation} then $\mu \in (\mu^-,\mu^+)$ with probability $1-\alpha$.
+A simple IID Monte Carlo stopping criterion uses an initial sample of $n_0$ points to approximate $\sigma^2 \approx (n_0-1)^{-1} \sum_{i=0}^{n_0-1}(f(\bsx_i)-\hat{\mu})^2$, and then plugs this into \eqref{eq:sc_clt_iid_n} to determine $n$.
+As the CLT only holds as $n$ goes to infinity, this is only a approximate stopping criterion.
+For integrands $f$ with bounded kurtosis, a guaranteed version of this two-stage-sampling idea has been developed to account for the finite sample size using Berry-Esseen inequalities \cite{JiaHic16a}.
+
+A straightforward Quasi-Monte Carlo stopping criterion may also be derived from the CLT using a few randomizations of a low-discrepancy pointset.
+Let $\bsx_{ir}$ be the $i^\mathrm{th}$ point in the $r^\mathrm{th}$ randomization of a low discrepancy pointset with $0 \leq i < n$ and $1 \leq r \leq R$.
+For example, we may use $R$ independent shifts of a lattice pointset or independent digital shifts of a digital net.
+Crucially, the randomized pointsets presented in Section \ref{sec:random} are also low discrepancy.
+Let $\hat{\mu}_r = n^{-1} \sum_{i=0}^{n-1} f(\bsx_{ir})$ be approximate means for each randomization and set $\hat{\mu} = R^{-1} \sum_{r=1}^R \hat{\mu}_r$ to be the aggregate approximation.
+In the spirit of CLT, a standard approximate $1-\alpha$ confidence interval $(\mu^-,\mu^+)$ has $\mu^\pm = \hat{\mu} \pm t_{R-1}^{1-\alpha/2} S/\sqrt{R}$ where $S^2 = (R-1)^{-1} \sum_{r=1}^R (\hat{\mu}_r - \hat{\mu})^2$ and $t_{R-1}^{1-\alpha/2}$ is the $1-\alpha/2$ quantile of the Student $t$ distribution with $R-1$ degrees of freedom.
+Here we have used the Student $t$ distribution since $R$ is typically small, for example $R=15$.
+In \cite{l2023confidence}, comprehensive numerical experiments show that this straightforward QMC stopping criterion achieve better coverage than more involved bootstrap $t$ methods for a variety of test functions.
+To use this confidence interval for a stopping criterion, one may iteratively double the sample size $n$ until the interval width is sufficiently low.
+
+One of the drawbacks of this QMC-CLT stopping criterion is that, at least intuitively, independent randomizations of a low discrepancy pointset are less efficient than simply using more low discrepancy points from a single sequence.
+Concretely, we would expect an approximation with $n=2^8$ and $R=2^3$ to perform worse than simply taking $n=2^{11}$ points from a single randomized low discrepancy sequence ($R=1$).
+One idea in this direction is to approximate the decay rate of Fourier coefficients of $f$.
+Bounds of absolute certainty have be derived for cones of functions with sufficient decay \cite{HicEtal17a}.
+As in the QMC CLT stopping criterion, one may iteratively double $n$ until the bounds are sufficiently tight.
+
+One cone tracks the decay of approximate complex-exponential Fourier coefficients \cite{JimHic16a}.
+When paired with a randomly shifted lattice, these approximate coefficients may be computed at only $\calO(n \log n)$ cost using the fast Fourier transform.
+Another cone tracks the decay of approximate Walsh coefficients  \cite{HicJim16a}.
+When paired with a digitally shifted digital net, these approximate coefficients may be computed at only $\calO(n \log n)$ using the fast Walsh-Hadamard transform.
+
+Another idea which avoids multiple randomizations is to treat $f$ as a Gaussian process.
+Conditioned on observations $f(\bsx_i)$, a closed form expression exists for the $1-\alpha$ credible intervals for $\mu$.
+While also applicable when using IID points, using low discrepancy sequences paired with special kernels enables these credible intervals to be computed at only $\calO(n \log n)$ cost \cite{rathinavel2019fast}.
+Again we may iteratively double the sample size $n$ until the credible intervals are sufficiently tight.
+
+Specifically, pairing shifted lattices with shift-invariant kernels yields circulant Gram matrices for which computations can be done quickly using the fast Fourier transform \cite{RatHic19a}.
+Similarly, pairing digitally shifted digital nets with digitally-shift-invariant kernels yields recursive-block Toeplitz Gram matrices for which computations can be done quickly using fast Walsh-Hadamard transform \cite{JagHic22a}.
+
+Extensions of these stopping criteria to functions of multiple expectations was explored in \cite{sorokin2022bounding}.
+Such problems include computing the posterior mean in a Bayesian context or the sensitivity indices for quantifying feature importance to a functions output variability.
+This work also considered sharing samples across approximations and generalized error tolerances such as relative error tolerance.
 
 \section{Reformulating Our Problem} \label{sec:reformulate}
-\subsection{Variable Transformations}
-\subsection{Variation Reduction}
-\subsection{Multilevel Methods} \label{sec:multi}
+\subsection{Variable Transformations \textcolor{blue}{Aleksei}} 
+\subsection{Variation Reduction \textcolor{blue}{Fred}} 
+\subsection{Multilevel Methods \textcolor{blue}{Fred}} \label{sec:multi}
 The cost of computing the sample mean defined in \eqref{eq:sample_mean} is typically $\mathcal{O}(dn)$ as the dimension, $d$, and the sample size, $n$, tend to infinity.
-Some applications have $d$ either huge or infinite.
+Some applications have $d$ that is either huge or infinite.
 On example is in quantitative finance, where $d$ denotes the number of times that the asset price is monitored over the life of a financial derivative times the number of assets involved.
-However, there is a way to reduce the cost of approximating $\mu$ by employing multilevel methods \cite{}.
-
-Briefly, let $\{Y_l := f_l(\bsX_{l})\}_{l=1}^L$ be a sequence of approximations to $Y = f(\bsX)$, where $\bsX_l \sim \mathcal{U}[0,1)^{d_l}$.
-Here we assume that $d_1 < \cdots < d_L$, and that the cost of evaluating $f_l(\bsx_{l})$ is $\mathcal{O}(d)$.
-Also, $d_L = d$, $Y_{d_L} = T$, and $f_L = f$.
+
+However, there is a way to reduce the cost of approximating $\mu$ by employing multilevel methods so that the computational cost may be comparable to solving a low dimensional approximation \cite{Gil15a}.
+We give an introduction to these methods here.
+
+Let $\{Y_l := f_l(\bsX_{l})\}_{l=1}^L$ be a sequence of approximations to $Y = f(\bsX)$, where $\bsX_l \sim \mathcal{U}[0,1)^{d_l}$.
+Here we assume that $d_1 < \cdots < d_L$, and that the cost of evaluating $f_l(\bsx_{l})$ is $\mathcal{O}(d_l)$.
+Also, $d_L = d$, $Y_{d_L} = Y$, and $f_L = f$.
 Then
@@
     \mu & := \mathbb{E}(Y) = \mathbb{E}\bigl[f(\bsX)\bigr]
-    = \mathbb{E}(Y_1) + \mathbb{E}(Y_2 - Y_1) + \cdots + \mathbb{E}(Y_{d_{L-1}} - Y_d) \\
+    = \mathbb{E}(Y_1) + \mathbb{E}(Y_2 - Y_1) + \cdots + \mathbb{E}(Y_{d_L} - Y_{d_{L-1}}) \\
     & = \int_{[0,1)^{d_1}} f_1(\bsx^{(1)}) \, \mathrm{d} \bsx^{(1)} +  \int_{[0,1)^{d_2}} \bigl[f_2(\bsx^{(2)}) - f_1(\bsx^{(1)})\bigr] \, \mathrm{d} \bsx^{(2)} + \cdots \\
-    & \qquad \qquad + \int_{[0,1)^{d_L}} \bigl[f_L(\bsx^{(L)}) - f_{L-1}(\bsx^{(L-1)}) \bigr] \, \mathrm{d} \bsx^{(L)}
-    \intertext{where}
-    \bsx^{(1)} & = (x_1, \dots, x_{d_1}), \quad \bsx^{(2)} = (x_1, \dots, x_{d_2}), \ldots, \bsx^{(L)} = (x_1, \dots, x_{d_L}).
+    & \qquad \qquad + \int_{[0,1)^{d_L}} \bigl[f_L(\bsx^{(L)}) - f_{L-1}(\bsx^{(L-1)}) \bigr] \, \mathrm{d} \bsx^{(L)},
 \end{align*}
+where $\bsx^{(1)}  = (x_1, \dots, x_{d_1}), \ldots, \bsx^{(L)} = (x_1, \dots, x_{d_L})$.
 This series of integrals may be approximated by a series of sample means, namely
@@
 \end{multline}
-The computational cost of this approximation is $\mathcal{O}(d_1n_1 + \cdots + d_Ln_L)$.
+If $C_1$ is the computational cost of $f^{(1)}(\bsx^{(1)})$ and $C_l$ is the computational cost of $f^{(l)}(\bsx^{(1)})$ of $f_L(\bsx_i^{(l)}) - f_{l-1}(\bsx_i^{(l-1)})$ for $l = 2, \ldots, L$, then the total computational cost of
+\eqref{eq:MLmuhat} is 
+\begin{equation} \label{eq:MLcostdef}
+    \text{cost}(\bsn,\bsd) = n_1 C_1 + \cdots + n_L C_L.
+\end{equation}
 
@@
     V_1/n_1 + V_2/n_2 + \cdots + V_L/n_L,  \\
-    \intertext{where} 
+    \intertext{where the $V_l$ are defined such that we expect them to have little dependence on $n_l$ since we are using quasi-Monte Carlo nodes:} 
     \nonumber
@@
 and $K^{(l)}$ is the reproducing kernel corresponding to the relevant  Hilbert space of functions of $d_l$ variables.
-The $V_l$ are defined such that we expect them to have little dependence on $n_l$ since we are using quasi-Monte Carlo nodes.
-
-If $C_1$ is the computational cost of $f^{(1)}(\bsx^{(1)})$ and $C_l$ is the computational cost of $f^{(l)}(\bsx^{(1)})$ of $f_L(\bsx_i^{(l)}) - f_{l-1}(\bsx_i^{(l-1)})$ for $l = 2, \ldots, L$, then the total computational cost of
-\eqref{eq:MLmuhat} is 
-\begin{equation}
-    \text{cost}(\bsn,\bsd) = n_1 C_1 + \cdots + n_L C_L.
-\end{equation}
+
 Optimizing $\bsn$ to minimize the error bound in \eqref{eq:MLerrbd} for a fixed cost and assuming constant $\bsC$ and $\bsV$, yields
@@
 
-To understand this result, suppose that the cost per function value at each level, $l$, is proportional to the dimension, $d_l$, and that the dimensions increase exponentially, i.e.,  $C_l \le \alpha d_1^l$.
-Moreover, suppose that the $V_l$, which are roughly the variation, decrease exponentially, i.e., $V_l \le \beta r^{-l} $.
+We see that the cost of approximating $\mu$ within an absolute error tolerance of $\varepsilon$ is $\mathcal{O}(\varepsilon^{-1})$, as expected, but the dependence of $\text{cost}(\bsn,\bsd)$ on the dimension of the problem depends critically on how quickly the variation at each level decays.
+Suppose that
+\begin{itemize}
+    \item The cost of a function value at each level, $l$, is proportional to the dimension, $d_l$, and that the dimensions increase exponentially, i.e.,  $d_l = d_1^{l}$ and $C_l \le \alpha d_1^l$. 
+    \item The $V_l$, which are roughly the variation, decrease exponentially, i.e., $V_l \le \beta r^{-l} $.
+\end{itemize}   
 Then the computational cost in \eqref{eq:MLcost} can be bounded above by
 \begin{align*}
-     \text{cost}(\bsn,\bsd) & \le \frac{\alpha \beta \bigl((d_1/r)^{-1/2} + \ldots + (d_1/r)^{-L/2}\bigr)^2}{\varepsilon} \\
+     \text{cost}(\bsn,\bsd) & \le \frac{\alpha \beta \bigl((d_1/r)^{1/2} + \ldots + (d_1/r)^{L/2}\bigr)^2}{\varepsilon} \\
      & = \frac{\alpha \beta \bigl((d_1/r)^{1/2} - (d_1/r)^{(L+1)/2}\bigr)^2}{\varepsilon \bigl( 1 - (d_1/r)^{1/2}\bigr)^2}.
 \end{align*}
+
 If the $V_l$ decay quickly enough, which means that the corrections $f_l - f_{l-1}$ depending on many variables decay quickly enough, in particular, if $r > d_1$, then 
@@
 \end{equation}
-
-\section{Ongoing Research} \label{sec:ongoing}
-\section{Conclusion} \label{sec:conclusion}
+In this case $\text{cost}(\bsn,\bsd)$ for this high dimensional problem is ont much worse than the cost of ensuring that just error of the rough $d_1$ approximation is smaller than $\varepsilon$, which would be $V_1C_1/\varepsilon \le \alpha \beta d_1/(r\varepsilon)$.
+
+However, if $r < d_1$, then $\text{cost}(\bsn,\bsd) \ge V_LC_L/\varepsilon$ which may be as large as $\alpha \beta (d_1/r)^{-L}{\varepsilon}$, which grows exponentially as $d = d_L = d_1^L$ increases via increasing $L$.
+Granted, this may not be as high a cost as the standard single-level approach where the cost would be roughly $V_1C_L$ if the $V_l$ are decaying.
+
+The ability to write one's problem to have the favorable case of high, or even infinite dimension with a computational cost comparable to the low dimensional case is somewhat an art.
+However, the multilevel method takes advantage of the principle that we pointed out in Section \ref{sec:coordwts}, namely that qMC thrives when problem have a low essential dimension.
+
+\section{Beyond Computing Expectations \textcolor{blue}{Nathan}}
+
+Quasi-Monte Carlo methods have found many applications in other areas of the computational sciences.
+We review several popular applications of qMC methods to highlight their practical impact and to motivate the further development of QMC theory and methodology.
+
+Besides computing expectations, it is often useful to know the distribution, $F$, or the density, $\varrho$, of a variable $Y$, i.e., 
+\[
+F(y) \coloneqq \mathbb{P}(Y\leq y)=\mathbb{P}(f(\mathbf{X})\leq y)=\int_{[0,1]^d} \mathbf{1}_{(-\infty,y]}(f(\mathbf{x}))d\mathbf{x}, \quad \varrho(y)\coloneqq F'(y).
+\]
+Several works in this direction use qMC \cite{LEcPuc20a,LEcuyer2022b,GilKuoSlo23a} and randomized qMC \cite{AbdEtal21a}.
+
+Sorokin and Rao \cite{SorRao23a} illustrates that integration of QMC methods with Gaussian process models enhance the reliability of estimating failure probabilities.
+
+Novel approaches are proposed for supervised deep learning algorithms for training Machine learning hyperparameter tuning \cite{longo21,mishra21}
+
+Keller \cite{kell95} was the first to adopt qMC for computer graphics.  \cite{kellprem12,shired08}
+
+Path planning is a critical aspect of efficient robotic operation, defining of a route from an initial position to a target destination while considering an optimality criterion.
+Most existing approaches utilize LD nodes, such as the Halton sequence \cite{veldel14,zhong24} with more recent efforts \cite{chahine24} demonstrating that employing state-of-the-art LD nodes from \cite{ruschkirk24} can significantly enhance performance compared to traditional implementations.
+
+PDEs - Frances and others
+
+%\section{Ongoing Research} \label{sec:ongoing}
+\section{Conclusion \textcolor{blue}{Fred}} \label{sec:conclusion}
 

### 2025-01-23 HickernellMCQMC2024_QMCTutorial.tex

@@
 
-\keywords{}
+\keywords{low discrepancy, randomization, sample mean, simulation}
 \end{abstract}
@@
 \setcounter{tocdepth}{2}
-\tableofcontents
+%\tableofcontents
 
@@
 In mathematical terms, the random outcome is $Y := f(\bsX)$, where $\bsX$ is a vector random input.
-Given an input, $\bsx$, the corresponding outcome  $y = f(\bsx)$ can be computed by an algorithm, whose complexity may might make $f$ a \emph{black box}, or at least a gray box.
-The user selects a sequence of input values, also known as data sites or \emph{nodes}, $\bsx_0, \bsx_1, \ldots$, which result in observed outcomes $y_0 = f(\bsx_0), y_1 = f(\bsx_1), \ldots$.
+Given an input, $\bsx$, the corresponding outcome  $y = f(\bsx)$ can be computed by an algorithm, whose complexity makes $f$ a \emph{black box}, or at least a gray box.
+The user selects a sequence of input values, also known as data sites or \emph{nodes}, $\bsx_0, \bsx_1, \ldots$, which lead to observed outcomes $y_0 = f(\bsx_0), y_1 = f(\bsx_1), \ldots$.
 The $y_i$ may be used to approximate $\bbE(Y)$, quartiles of $Y$, the probability density of $Y$, and other quantities of interest.
 
-Simple  Monte Carlo chooses the nodes, $\{\bsx_0, \bsx_1, \ldots\}$, to be independent and identically distributed (IID).
+Simple  Monte Carlo chooses the sequence of nodes, $\{\bsx_i\}_{i=0}^\infty$, to be independent and identically distributed (IID).
 Quasi-Monte Carlo (qMC)  chooses the nodes differently, so that their empirical distribution approximates well the true probability distribution of $\bsX$.
@@
 This tutorial describes what qMC is, why we would want to use qMC, and how qMC can be implemented well.
-The next session illustrates by example the advantages of qMC.
-This is followed by a description of how deterministic LD sequences are constructed and randomizations that preserve their low discrepancy.
-We describe various measures of discrepancy.
-We explain how to decide what sample size, $n$, is sufficient to meet the user's error criterion.
-We then discuss how to rewrite the problem of interest in a qMC-friendly way.
+The next section illustrates by example the advantages of qMC.
+This is followed by a description of how deterministic LD sequences are constructed (Section \ref{sec:construct}) and randomizations that preserve their low discrepancy (Section \ref{sec:random}).
+We describe various measures of discrepancy (Section \ref{sec:discrepancy}).
+We explain how to decide what sample size, $n$, is sufficient to meet the user's error requirements (Section \ref{sec:stop}).
+We then discuss how to rewrite the problem of interest in a qMC-friendly way (Section \ref{sec:reformulate}).
+We briefly discuss applications of qMC beyond computing the mean (Section \ref{sec:beyond}), and finish with a brief conclusion (Section \ref{sec:conclusion}).
 
@@
 Grids have the largest error and LD nodes have the smallest error.
-The numerical computations behind this figure and the others in the is chapter may be reproduced using the Jupyter notebook \url{??}.
+The numerical computations used to generate this figure and the others in the is chapter may be reproduced using the Jupyter notebook \url{??}.
 \label{fig:keister-err}}
@@
 
-	\item \emph{IID.}  Simple MC (orange diamonds) is a substantial improvement over grid nodes. As a reminder, for IID nodes the root mean squared error is
+	\item \emph{IID.}  Simple MC is a substantial improvement over grid nodes. As a reminder, for IID nodes the root mean squared error is
 	\begin{equation}\label{eq:IIDerror}
@@
 
-	\item \emph{LD.}  For qMC methods the error decays nearly like $\mathcal{O}(n^{-1})$, which for this example correponds to a reduction in error of several orders of magnitude compared to simple MC for large enough $n$.
+	\item \emph{LD.}  The error of qMC methods decays nearly like $\mathcal{O}(n^{-1})$, which for this example correponds to a reduction in error of several orders of magnitude compared to simple MC for large enough $n$.
 Typically, LD sequences have preferred sample sizes.
@@
 What sets these apart from IID sequences is that the nodes are deliberately chosen and highly correlated, whether they be deterministic or randomized.
-\texttt{qmcpy} \cite{QMCPy2020a} is a Python library that contains many of the constructions discussed here, as well as the stopping criteria discussed in Section \ref{sec:stop}.
+The Python library \texttt{qmcpy} \cite{QMCPy2020a} contains many of the constructions discussed here, as well as the stopping criteria discussed in Section \ref{sec:stop}.
 
@@
 However, the set of nodes defined by even $i$ do a reasonable job.
-This suggests a method for defining extensible lattice sequences that was proposed independently by \cite{Mai81a} and \cite{HicEtal00}.
-
-The one-dimensional extensible lattice is defined by the van der Corput sequence, $\{\phi_b(0), \phi_b(1), \ldots\}$.
+This suggests a method for defining extensible lattice sequences that was proposed independently in \cite{Mai81a} and \cite{HicEtal00}.
+
+The one-dimensional extensible lattice sequence is defined by the van der Corput sequence, $\{\phi_b(0), \phi_b(1), \ldots\}$.
 This involves reflecting the digits of the integer $i$ in base $b$ about the decimal point.
@@
 Figure \ref{fig:extensiblelatticeconstruct} shows the first $4$, $8$, and $16$ nodes of an extensible lattice in base $2$ with the same generator as in Figure \ref{fig:latticeconstruct}.
-The blue nodes are a copy of the nodes in the plot to the left, and the the orange nodes correspond to a shifted copy of the blue nodes modulo $\bsone$.
+The blue dots are a copy of the nodes in the plot to the left, and the the orange diamonds correspond to a shifted copy of the blue dots modulo $\bsone$.
 For the left plot, the shift is $\phi_2(2)\bsh = (1,11)/4$ or equivalently $(0.25,0.75)$.
@@
 The construction of great generating vectors for lattices has attracted a great deal of interest.
-The component-by-component constructions pioneered by Frances Kuo, Josef Dick, Ian Sloan, Dirk Nuyens and others are surveyed in \cite[Chapter ?]{DicEtal22a}.
+The component-by-component constructions pioneered by Frances Kuo, Josef Dick, Ian Sloan, Dirk Nuyens and others are explained in \cite[Chapters 3, 4 \& 6]{DicEtal22a}.
 LatNet builder \cite{LEcEtal22a,LatNet} is a software package by Pierre L'Ecuyer and collaborators, which constructs generators for lattices and digital sequences.
@@
 \begin{equation} \label{eq:digital}
-	\bsx_i^{\text{dig}} := i_0 \bsx^{\text{dig}} _1 \oplus i_1 \bsx^{\text{dig}} _2 \oplus i_2 \bsx^{\text{dig}} _4 \oplus \cdots \in [0,1)^d \quad \text{for }
+	\bsx_i^{\text{dig}} := i_0 \bsx^{\text{dig}} _1 \oplus i_1 \bsx^{\text{dig}}_2 \oplus i_2 \bsx^{\text{dig}} _4 \oplus \cdots \in [0,1)^d \quad \text{for }
 	i = i_0 + i_12 + i_2 2^2 + \cdots,
 \end{equation}
-where $\bsx_1, \bsx_2, \ldots$ are carefully chosen.
+where $\bsx^{\text{dig}}_1, \bsx^{\text{dig}}_2, \ldots$ are carefully chosen.
 (This is the van  der Corput sequence in base $2$ for $d=1$ and $x_\ell = 2^{-\ell}$.) The node set $\{ \bsx_0, \ldots, \bsx_{2^m-1}\}$  for non-negative integer $m$ is called a digital net and is closed under $\oplus$.
 
-If $x_{ij\ell}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ coordinate of $\bsx_i$, then the literature often defines generating matrices, $\mathsf{C}_j = (c_{ij\ell})_{\ell,i = 1}^{M,N}$ for $j = 1, \ldots, d$, where $M$ is the maximum number of bits in the expression for $\bsx_i$, say $52$, and $2^N$ is the maximum number of nodes that is intended to be generated.
+If $x_{ij\ell}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ coordinate of $\bsx_i$, then the literature often defines generating matrices, $\mathsf{C}_j = (x_{ij\ell})_{\ell,i = 1}^{M,N}$ for $j = 1, \ldots, d$, where $M$ is the maximum number of bits in the expression for $\bsx_i$, say $52$, and $2^N$ is the maximum number of nodes that is intended to be generated.
 The digital sequence can then be defined equivalently as
@@
 	\text{the first $k_2$ rows of the first $m$ columns of $\mathsf{C}_2$ plus} \\
-	 \cdots \\
+	 \vdots \\
 	\text{the first $k_d$ rows of the first $m$ columns of $\mathsf{C}_d$} \\
@@
 \subsection{Variable Transformations \textcolor{blue}{Aleksei}}
+
+Often (Quasi-)Monte Carlo problems may be written as $\mu = \bbE[g(\bsT)]$ where $\bsT$ is a $d$-dimensional random variable with has density $\lambda$.
+Since QMC points have low discrepancy with the standard uniform distribution $\bsX \sim \calU[0,1]^d$, we perform an invertible variable transformation $\bsPsi: [0,1]^d \to \calM$ satisfying $\mathrm{range}(\bsT) \subseteq \calM$.
+This implies
+$\mu = \bbE[f(\bsX)]$ where  
+$$f(\bsX) = g(\bsPsi(\bsX)) \frac{\lambda (\bsPsi(\bsX))}{\nu(\bsPsi(\bsX))}$$
+and $\nu$ is the density of $\bsPsi(\bsX)$.
+We may write $\nu(\bsPsi(\bsX))= \lvert \bsJ_\bsPsi(\bsX) \rvert^{-1}$ where $\lvert \bsJ_\bsPsi(\bsX) \rvert$ is the determinant of the Jacobian of $\bsPsi$.
+
+The above formulation is equivalent to importance sampling by a distribution with density $\nu$.
+Optimally, we would choose $\nu \propto g\lambda$ so that $f$ is constant, but this is infeasible in practice.
+Choosing $\bsPsi$ so that $\bsT = \bsPsi(\bsX)$ gives the simplification $f(\bsX) = g(\bsPsi(\bsX))$.
+An example of this was given in Section \ref{sec:practice} where $\bsT \sim \calN(\boldsymbol{0},\mathsf{I}/2)$ and $\bsPsi(\bsX) = \Phi^{-1}(\bsX)/\sqrt{2}$ where $\Phi^{-1}$ is the inverse CDF of the standard Gaussian distribution applied elementwise.
+
 \subsection{Variation Reduction \textcolor{blue}{Fred}} 
@@
 
-\section{Beyond Computing Expectations \textcolor{blue}{Nathan}}
-
-Quasi-Monte Carlo methods have found many applications in other areas of the computational sciences.
-We review several popular applications of qMC methods to highlight their practical impact and to motivate the further development of QMC theory and methodology.
-
-Besides computing expectations, it is often useful to know the distribution, $F$, or the density, $\varrho$, of a variable $Y$, i.e., 
+\section{Beyond Computing Expectations \textcolor{blue}{Nathan}} \label{sec:beyond}
+
+Expectations of the form $\mu \coloneqq \mathbb{E}(f(\mathbf{X}))$ appear in several prominent areas of mathematics and computational science, and qMC methods have long been applied in these contexts.
+Namely, in mathematical finance many problems can be formulated as high-dimensional integrals, where the large number of dimensions arises from small time steps in time discretization and/or a large number of variables.
+In this setting, Monte Carlo methods are applied to randomly simulate the evolution of stock prices over time, and the expected value is then estimated by averaging over all sample paths.
+The qMC approach replaces the randomly generated paths by carefully constructed paths in an effort to reduce the computational burden on the required number of paths.
+Excellent survey works for qMC and randomized qMC applied to finance include \cite{CafMor96,LEc09,Lem04a,Jae02,wangsloan05}.
+Another fruitful domain for qMC methods is in partial differential equations (PDEs), used for computing the expectations of nonlinear functionals of solutions of certain classes of PDEs.
+This includes problems in areas like fluid dynamics or material science, where random coefficients in the PDE model uncertain properties; see \cite{KuoNuy16a,KuoEtal12a,GraEtal11a} and references therein for qMC applied to PDEs.
+
+Besides computing expectations, qMC methods have found many applications in other areas of the computational sciences.
+We lightly review these applications to highlight their wider practical impact and to motivate the further development of QMC theory and methodology.
+
+For example, when considering a variable $Y$, it is often useful to determine its distribution, $F$, or the density, $\varrho$, i.e., 
 \[
@@
 \]
-Several works in this direction use qMC \cite{LEcPuc20a,LEcuyer2022b,GilKuoSlo23a} and randomized qMC \cite{AbdEtal21a}.
-
-Sorokin and Rao \cite{SorRao23a} illustrates that integration of QMC methods with Gaussian process models enhance the reliability of estimating failure probabilities.
-
-Novel approaches are proposed for supervised deep learning algorithms for training Machine learning hyperparameter tuning \cite{longo21,mishra21}
-
-Keller \cite{kell95} was the first to adopt qMC for computer graphics.  \cite{kellprem12,shired08}
+Here, $F(y)$ gives the cumulative probability that $Y$ does not exceed $y$, while $\varrho(y)$ represents the corresponding density function.
+Several recent works for density estimation use qMC \cite{LEcPuc20a,LEcuyer2022b,GilKuoSlo23a} alongside randomized qMC \cite{AbdEtal21a} and often incorporate kernel-based methods or rely on conditioning strategies (sometimes called preintegration).
+
+Taking a broader view, Sorokin and Rao \cite{SorRao23a} illustrates that integration of qMC methods with Gaussian process models enhance the reliability of estimating failure probabilities in the machine learning domain.
+Furthermore, novel approaches for training deep neural networks utilizing LD nodes have been proposed in recent works \cite{longo21,mishra21}.
+
+Keller \cite{kell95} was the first to adopt qMC for rendering photorealistic images in computer graphics where the problem of simulating light transport can be mathematically represented as an integral over all possible light paths from source to camera.
+Many algorithms involving sampling with LD nodes have since been developed much further in \cite{kellprem12,shired08,Keller2013a,kolkel00,kell96,kell95}.
+(Remarkably, Monte Carlo methods have even been awarded with an Academy Award for visual effects in movie making.)
 
 Path planning is a critical aspect of efficient robotic operation, defining of a route from an initial position to a target destination while considering an optimality criterion.
-Most existing approaches utilize LD nodes, such as the Halton sequence \cite{veldel14,zhong24} with more recent efforts \cite{chahine24} demonstrating that employing state-of-the-art LD nodes from \cite{ruschkirk24} can significantly enhance performance compared to traditional implementations.
-
-PDEs - Frances and others
+Most existing approaches utilize LD nodes, such as the Halton sequence \cite{veldel14,zhong24} with more recent efforts \cite{chahine24} demonstrating that employing state-of-the-art LD nodes from \cite{ruschkirk24} can significantly enhance the efficiency and flexibility of a robotic arm, demonstrated in both toy and real-world tasks.
+
+In summary, qMC methods have demonstrated their applicability and impact well beyond computing expectations, finding fruitful applications across fields such as computer graphics, robotics, and machine learning.
 

### 2025-01-24 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

@@
 %%Fred's temporary macros
+\newcommand{\FJHNote}[1]{{\textcolor{blue}{Fred: #1}}}
+\newcommand{\NKNote}[1]{{\textcolor{blue}{Nathan: #1}}}
 \newcommand{\figpath}{Figures}
@@
 	\item The porosity of rock, which affects the extraction of oil or gas, or
-	\item Elemetary particle interactions in a high energy physics experiment, which affect what is observed by detectors.
+	\item Elementary particle interactions in a high energy physics experiment, which affect what is observed by detectors.
 \end{itemize}
 In such situations, the unknown inputs are often modeled using random vector variables or stochastic processes.
-Computations are performed by generating a multidude of possible outcomes informed by the assumed probability distribution of the input.
+Computations are performed by generating a multitude of possible outcomes informed by the assumed probability distribution of the input.
 These are used to estimate the mean, quantiles, and/or probability distribution of the outcome.
@@
 	\centering
-	\includegraphics[width =0.7\textwidth]{\figpath/n_is_16777216_d_is_6_n_rep_is_50_KeisterErrors.pdf}
+	\includegraphics[width =0.5\textwidth]{\figpath/n_is_16777216_d_is_6_n_rep_is_50_KeisterErrors.pdf}
 	\caption{The relative error of approximating $\mu$ defined in \eqref{eq:keisterC} by the sample mean, $\hat{\mu}_n$, defined in \eqref{eq:sample_mean} for various choices of nodes, $\bsx_0, \bsx_1, \ldots$.
@@
 
-The one-dimensional extensible lattice sequence is defined by the van der Corput sequence, $\{\phi_b(0), \phi_b(1), \ldots\}$.
-This involves reflecting the digits of the integer $i$ in base $b$ about the decimal point.
+The extensible lattice sequence is defined by the one-dimensional van der Corput sequence, $\{\phi_b(0), \phi_b(1), \ldots\}$.
+This involves the so-called radical inverse function and essentially reflects the digits of the integer $i$ in base $b$ about the decimal point.
 For example, $\phi_2(6) = \phi_2(110_{2}) = {}_20.011 = 3/8$.
@@
 In practice, one searches computationally for a $\bsh$ that produces good set of lattice nodes of size $n = b^m$ for a range of $m$ \cite{HicEtal00}.
-
-The construction of great generating vectors for lattices has attracted a great deal of interest.
-The component-by-component constructions pioneered by Frances Kuo, Josef Dick, Ian Sloan, Dirk Nuyens and others are explained in \cite[Chapters 3, 4 \& 6]{DicEtal22a}.
+The construction of great generating vectors for lattices has attracted a great deal of interest and is discussed further in Section \ref{sec:optLD}.
+
+%The component-by-component constructions pioneered by Frances Kuo, Josef Dick, Ian Sloan, Dirk Nuyens and others are explained in \cite[Chapters 3, 4 \& 6]{DicEtal22a}.
 LatNet builder \cite{LEcEtal22a,LatNet} is a software package by Pierre L'Ecuyer and collaborators, which constructs generators for lattices and digital sequences.
@@
 
-To illustrate the $t$-value, consider the following digital net with $2^3$ (eight) nodes, i.e., $m = 3$:
+To illustrate the $t$-value, consider the following three dimensional digital net with $2^3$ (eight) nodes, i.e., $m = 3$:
 \begin{multline} \label{eq:smallnet}
@@
 
-Digital sequence generators can be found via number theory or numerical search \cite{}.  The earliest instance is due to Sobol' \cite{}.
+Digital sequence generators can be found via number theory or numerical search \cite{KuoJoe08a}.
+The earliest instance is due to Sobol' \cite{Sob67}.\NKNote{I added these two references which were previously left blank.
+Were you thinking of other papers to cite here?}
 
@@
 \end{equation}
-where $b_1, \ldots, b_d$ is a choice of $d$ distinct prime numbers.  Often they are chosen as the first $d$ prime numbers.  Figure \ref{fig:Halton} shows two dimensional projections of Halton nodes
+where $b_1, \ldots, b_d$ is a choice of $d$ distinct prime numbers.
+Often they are chosen as the first $d$ prime numbers.
+Figure \ref{fig:Halton} shows two dimensional projections of Halton nodes in six dimensions.
+\NKNote{Suggestion: The Halton construction is an extensible sequence of LD nodes.
+However, should one know $n \in \mathbb{N}$ in advance, a finite LD node set called the Hammersley point set can be defined similarly as
+\begin{equation}\label{eq:hammersley}
+    \bsx_i^{\text{Ham}} = \left( \frac{i}{n}, \phi_{b_1}(i), \ldots, \phi_{b_d}(i) \right) \quad i = 0, 1, \ldots, .
+\end{equation}
+}
 
@@
 When optimizing LD nodes, one requires an objective function which in some sense must assesses the quality of the LD nodes directly, or the expected performance when the node set is implemented in a qMC estimate \eqref{eq:sample_mean}.
-We will refer to such an objective function as a figure-of-merit (FOM) and popular choices include the discrepancy (see Section \ref{sec:discrepancy}) or worst-case error metrics when restricted a specific space of functions; see \cite{Hic00a,LatNet}.
+We will refer to such an objective function as a figure-of-merit (FOM) and popular choices include the discrepancy (see Section \ref{sec:discrepancy}) or worst-case error metrics when restricted to a specific space of functions; see \cite{Hic00a,LatNet}.
 
@@
 For certain choices of function spaces, there exist fast component-by-component construction implementations using fast Fourier transforms (FFTs) as the primary tool.
-We refer to \cite{DicKuo04a,JoeKuo03,KuoJoe02b} and references therein for works on the CBC construction for lattice rules and digital sequences, \cite{NuyCoo06a,NuyCoo06b} for their fast implementations and \cite{LatNet} for a QMC software tool called LatNet Builder by L'Ecuyer to generate LD nodes which relies heavily upon the CBC construction method.
-
-It is commonplace in qMC constructions that equidistribution is required not just in the (possibly high dimensional) $d$-dimensional space, but also for all or specific low-dimensional projections.
-MatBuilder \cite{paulin2022} is a software tool developed precisely for this task; to generate digital sequence LD nodes when low dimensional uniformity is critical.
-Uniformity constraints related to digital sequence stratification across boxes of the form \eqref{eq:element_box} are expressed as integer linear programming (ILP) problems and subsequently solved to optimize an objective function relating to the ILP reformulation ensuring good quality low dimensional projections.
-The solutions determine then build the columns of the generator matrices in a greedy manner for the LD digital sequence.
-
-As discussed in the previous section, the Halton sequence has the benefit of not having a preferred number of points for its construction.
-It does however possess the drawback of exhibiting poor uniformity in the lower dimensional projections when $d$ is moderate to high due to intrinsic number theoretic correlation between the coordinates \cite{kirklem24}.
+We refer to \cite{DicKuo04a,JoeKuo03,KuoJoe02b,DicEtal22a} and references therein for works on the CBC construction for lattice rules and digital sequences, \cite{NuyCoo06a,NuyCoo06b} for their fast implementations and \cite{LEcEtal22a,LatNet} for a QMC software tool \texttt{LatNet Builder} by L'Ecuyer to generate LD nodes which relies heavily on the CBC construction method.
+
+When constructing LD nodes, it is essential to ensure not only a well-distributed set of nodes in the $d$-dimensional space but also in its lower-dimensional projections (recall the Cartesian grid from \ref{sec:intro}, Figure \ref{fig:grid}).
+\texttt{MatBuilder} \cite{paulin2022} is a software tool developed precisely for this task; to generate generating matrices for digital sequence LD nodes possessing excellent low dimensional uniformity.
+In this software, the solutions to integer linear programming problems are used to build the columns of generator matrices in a greedy manner for the LD digital sequence.
+
+%Uniformity constraints related to digital sequence stratification across boxes of the form \eqref{eq:element_box} are expressed as integer linear programming (ILP) problems and subsequently solved to optimize an objective function relating to the ILP reformulation ensuring good quality low dimensional projections.
+
+In the previous section we discussed how the Halton sequence has the benefit of not having a preferred number of points for its construction.
+It does however possess the drawback of exhibiting poor uniformity in the lower dimensional projections when $d$ is moderate to high due to number theoretic correlation between the coordinates \cite{kirklem24}.
 One popular strategy to circumvent this poor behavior, dating back to Faure \cite{fau78} and Braaten and Weller \cite{brawell79}, is to randomly or deterministically permute the digits of the base $b_j$ expansions $j = 1,2\ldots,d$ to disrupt correlations between the components.
@@
 $$\bsx_i^{\text{G-Hal}} = \left( \phi^{\Sigma_1}_{b_1}(i), \ldots, \phi^{\Sigma_d}_{b_d}(i)\right), \quad i= 0,1,\ldots,$$
-There exist many proposed algorithms to optimize the permutations defining the generalized sequence, most of which involve so-called linear digit scramblings and usually involve directly optimizing a discrepancy FOM.
+There exist many proposed algorithms to optimize the permutations defining the generalized sequence, most of which involve so-called linear digit scramblings and involve directly optimizing a discrepancy FOM.
 We point the reader towards \cite{MasChiWar05,tuffin98,Atan04,vancools06} and the excellent survey of the work in this direction (until 2008) by Faure and Lemieux \cite{faulem09}.
 
-So far, we have discussed the attempts of researchers to optimize specific classes of LD constructions; lattices and digital sequences, and Halton sequences.
-There also exist several efforts, some very recently, to generate novel LD constructions which don't follow any number theoretic or any other underlying structure, but none-the-less, possess a high level of uniformity.
-These methods are often global optimization methods, which is often challenging due to the non-convex nature of the FOM as a function of the nodes.
-Traditional global optimization methods, such as gradient descent or its variants, struggle in this setting becoming trapped in suboptimal local minima and failing to identify the global optimum configuration of nodes.
+We have outlined various efforts by qMC researchers to optimize specific classes of LD constructions, including lattices, digital sequences, and Halton sequences.
+There also exist several efforts to generate novel LD constructions which don't adhere to any underlying number theoretic structure, but none-the-less, possess a high level of uniformity.
+These methods are often global optimization methods of sample point distributions, which is often challenging due to the non-convex nature of the FOM as a function of the nodes.
+%Traditional global optimization methods, such as gradient descent or its variants, struggle in this setting becoming trapped in suboptimal local minima and failing to identify the global optimum configuration of nodes.
 One of the first works to globally optimize LD nodes, albeit a heuristic approach, was by Winker and Fang \cite{WinFan97b} employing the threshold accepting optimization algorithm to directly optimize the uniformity of the LD nodes with respect to a discrepancy FOM.
 More recently, Rusch, Kirk, Bronstein, Lemieux and Rus \cite{ruschkirk24} developed the first machine-learning assisted optimization tool for LD node optimization in the Message-Passing Monte Carlo (MPMC) method.
-MPMC implements tools from geometric deep learning imposing a computational graph onto an initialized IID node set to transform the nodes to LD node set via a learned mapping, again guided by the minimization of a discrepancy-based FOM.
-We also highlight the method proposed by Hinrichs and Oettershagen \cite{hinoet16}, which generates optimal LD nodes for a specific class of functions in two-dimensions.
+MPMC implements tools from geometric deep learning, imposing a computational graph onto an initialized IID node set to transform the nodes to LD node set via a learned mapping.
+The optimization is gradient-based and is guided by the direct minimization of a discrepancy-based FOM.
+We also highlight the method proposed by Hinrichs and Oettershagen \cite{hinoet16} generating optimal LD nodes for a specific class of functions in two-dimensions.
 To address the non-convexity challenge, the authors decompose the global optimization of a specific FOM into a large number of smaller, convex sub-problems which are solved individually.
-There are several other recent attempts to construct optimal LD nodes, many of which are well summarized in the thesis of Cl\'{e}ment \cite{clethesis24}.
+There are several other recent attempts to construct optimal LD nodes, many of which are well summarized in the PhD thesis of Cl\'{e}ment \cite{clethesis24}.
 We highlight the subset selection method proposed in \cite{cle22} to choose from an $n$-element initialization set, the $k<n$ nodes which minimize a discrepancy FOM.
@@
 \section{Randomization} \label{sec:random}
-The LD constructions described in the previous section have so far been \emph{deterministic}. The sample mean, $\hat{\mu}_n$, does not change each time it is computed like it would for IID nodes.
-
+The LD constructions described in the previous section have so far been \emph{deterministic}.
+The sample mean, $\hat{\mu}_n$, does not change each time it is computed like it would for IID nodes.
 While determinism has some advantages, randomization has substantial advantages as well.
 \begin{itemize}
-	\item Randomization done right removes bias in the estimator.
+	\item Randomization done correctly removes bias in the estimator.
 	\item Replications of random estimators can facilitate error estimates for the sample mean.
 	\item Often the application of interest requires transforming the LD nodes to mimic other distributions.
-This is the case in the Keister example of Section \ref{sec:intro}, as seen in \eqref{eq:sample_mean_Keister}.
-There the nodes, $\{\bsx_i\}_{i=0}^{n-1}$, are transformed to mimic a Gaussian distribution.
+This is the case in the Keister example of Section \ref{sec:intro}, as seen in \eqref{eq:sample_mean_Keister} where the nodes, $\{\bsx_i\}_{i=0}^{n-1}$, are transformed to mimic a Gaussian distribution.
 
@@
 \end{equation}
-Since the first $2^m$ elements of the lattice sequence form a group under modulo $\bsone$ addition, the coresponding shifted node set is a coset.
+Since the first $2^m$ elements of the lattice sequence form a group under modulo $\bsone$ addition, the corresponding shifted node set is a coset.
 Figure \ref{fig:shift_lat} illustrates three shifts of the lattice plotted in Figure \ref{fig:latticeconstruct}.
@@
 
-As Halton nodes can be viewed as digital nets with different bases for each dimension, the randomization techniques used for digital nets can be applied to Halton pointsets.
-Random digital shifts for Halton set $x_{ij\ell} \mapsto (x_{ij\ell} + \Delta_{j\ell}) \mod b_j$ where the shift $\Delta_{j\ell} \in \{0,\dots,b_j-1\}$.
+As Halton nodes can be viewed as digital nets with different bases for each dimension, the randomization techniques used for digital nets can be applied to Halton point sets\NKNote{we should clarify here whether we mean truncated Halton, or Hammersley if it is introduced above} Random digital shifts for Halton set $x_{ij\ell} \mapsto (x_{ij\ell} + \Delta_{j\ell}) \mod b_j$ where the shift $\Delta_{j\ell} \in \{0,\dots,b_j-1\}$.
 More generally, a permutation scramble for digital nets sets $x_{ij\ell} \mapsto \Pi_{j\ell}(x_{ij\ell})$ where $\Pi_{j\ell}$ are independent random permutations of $\{0,\dots,b_j-1\}$.
-Random nested uniform scrambling and linear matrix scrambling have also been considered for Halton pointsets \cite{owen2024gain}.
+Randomization of Halton relating to permutations of digits was discussed briefly in Section \ref{sec:optLD}.
+Random nested uniform scrambling and linear matrix scrambling have also been considered for Halton point sets \cite{owen2024gain}.
 
@@
 \end{equation*}
-From this expression for $\eta_L(\bsx)$ one may then calculate the squared norm of the linear functional $L$ as the squared norm of its representer:
+From this expression for $\eta_L(\bsx)$ one may then calculate the squared norm of the linear functional $L$ as the squared norm of its representer
 \begin{equation} \label{eq:sqnormL}
@@
 First we write the dependence of $\mu - \hat{\mu}_n$ on  $f$ explicitly as $\mu(f) - \hat{\mu}_n(f)$ and note that the error functional, $\mu(\cdot) - \hat{\mu}_n(\cdot)$, is  linear.
-The Riesz Representation Theorem and the Cauchy-Schwarz inequality imply a tight error bound:
+The Riesz Representation Theorem and the Cauchy-Schwarz inequality imply a tight error bound
 \begin{align}
@@
 
-Error bound \eqref{eq:cuberrbd} has two factors:
+The error bound \eqref{eq:cuberrbd} has two factors:
 \begin{itemize}
@@
 	\item The reproducing kernel $K$ defines both the discrepancy and the variation.
-This means that differnt choices of $K$ will lead to different error bounds for the same $\hat{\mu}_n$, even though the error is unchanged.
+This means that different choices of $K$ will lead to different error bounds for the same $\hat{\mu}_n$, even though the error is unchanged.
 	\item The Hilbert space $\calH_K$ contains ones assumptions about $f$, such as smoothness or periodicity.
@@
 Conversely, having an explicit formula for  $\langle \cdot, \cdot \rangle_{\calH_K}$ does not automatically lead to an explicit formula for the reproducing kernel, $K$.
-In both cases educated guesswork is involved.
+In both cases, educated guesswork is involved.
 	\item Choosing a more restrictive $\calH_K$ may lead to a sharper error bound, however, it is often infeasible in practice to check whether the ``black box'' $f$ lies in $\calH_K$.
@@
 Once $K$ has been fixed, the discrepancy may be used as a comparison measure among different LD sequences, and its rate of decay---which sometimes can be computed theoretically---indicates how quickly $\hat{\mu}_n$ approaches $\mu$.
-	\item When an explicit formula for the inner product of the associated Hilbert space, $\langle \cdot, \cdot \rangle_{\calH_K}$, can be deduced, it suggests how to formulate $\mu$ to reduce variation$(f)$.
+	\item When an explicit formula for the inner product of the associated Hilbert space, $\langle \cdot, \cdot \rangle_{\calH_K}$, can be deduced, it suggests how to formulate $\mu$ to reduce variation$(f,K)$.
 \end{itemize}
@@
 
-We show that $K$ defined in \eqref{eq:ctrkernel} is the reproducing kernel for the $\calH_K$ with the above inner product for $d=1$ and assuming $0 \le x \le 1/2$:
+We show that $K$ defined in \eqref{eq:ctrkernel} is the reproducing kernel for the $\calH_K$ with the above inner product for $d=1$ and assuming $0 \le x \le 1/2$
 \begin{align*}
@@
 
-Figure  \ref{fig:ctrKer} (right)  displays the discrepancy for a Sobol' LD digital net for several modest values of $d$ and for $n =  1, 2, 4, \ldots$.
+Figure  \ref{fig:ctrKer} (right)  displays the discrepancy for a Sobol' LD digital net for several modest values of $d$ and for $n =  1, 2, 4, 8, 16, \ldots$.
 The decay rate of nearly $\calO(n^{-1})$ is observed for small $d$ and can be anticipated for a bit larger $d$.
-Thus, the decay of the discrepancy mirrors the convergenc rate for LD sequences for the Kiester example in Section \ref{sec:intro} and Figure \ref{fig:keister-err}.
+Thus, the decay of the discrepancy mirrors the convergence rate for LD sequences for the Keister example in Section \ref{sec:intro} and Figure \ref{fig:keister-err}.
 
@@
 \subsection{Coordinate weights} \label{sec:coordwts}
-When the dimension, $d$, is increased, the discrepancy increases and the rate of decay with increasing sample size deteriorates, as can be seen in Figure ??.
+When the dimension, $d$, is increased, the discrepancy increases and the rate of decay with increasing sample size deteriorates, as can be seen in Figure \ref{fig:ctrKer}.
 This raises the question of whether the LD sequences lose their effectiveness for higher dimensions or whether the theory described in the pervious section does not well describe what may occur in practice.
@@
     \centering
-    \includegraphics[width=\textwidth]{\figpath/IIDSobolScaledWtCtrDisc_n_8192_d64.pdf}
+    \includegraphics[width=0.75\textwidth]{\figpath/IIDSobolScaledWtCtrDisc_n_8192_d64.pdf}
     \caption{The scaled centered discrepancy with coordinate weights for a Sobol' LD sequence for various $d$ and the root mean squared scaled centered discrepancy with coordinate weights for IID nodes.
@@
 
-\section{Stopping Criteria \textcolor{blue}{Aleksei}} \label{sec:stop}
-
-Monte Carlo approximates the true mean $\mu = \bbE[f(\bsX)]$, $X \sim \mathcal{U}[0,1]^d$ by the sample mean $\hat{\mu} = n^{-1} \sum_{i=0}^{n-1} f(\bsx_i)$ \textcolor{green}{maybe these formulas are already in the paper.}.
-Stopping criterion determine how large $n$ should be so that the error $\lvert \mu - \hat{\mu} \rvert$ is below an error tolerance $\epsilon$ either certainly or with sufficiently low uncertainty $\alpha$.
+\section{Stopping Criteria} \label{sec:stop}
+
+%Monte Carlo approximates the true mean $\mu = \bbE[f(\bsX)]$, $\bsX \sim \mathcal{U}[0,1]^d$ by the sample mean $\hat{\mu} = n^{-1} \sum_{i=0}^{n-1} f(\bsx_i)$ \textcolor{green}{maybe these formulas are already in the paper.}.
+
+When approximating the true $\mu$ with $\hat{\mu}_n$, stopping criterion determine how large $n$ should be so that the error $\lvert \mu - \hat{\mu}_n \rvert$ is below an error tolerance $\epsilon$ either certainly or with sufficiently low uncertainty $\alpha$.
 The error tolerance and probability threshold are problem dependent.
-For example, in option pricing one may set the error tolerance to a penny, $\varepsilon=0.01$, and wish to find error bounds which hold with 95\% confidence i.e.
-$\alpha=.05$.
-
-For classic Monte Carlo with IID points, the Central Limit Theorem (CLT) may be used to get an approximate confidence interval $(\mu^-,\mu^+)$ where $\mu^\pm = \hat{\mu} \pm Z^{1-\alpha/2} \sigma/\sqrt{n}$ where $\sigma = \mathrm{Std}[f(X)]$ and $Z^{1-\alpha/2}$ is the $1-\alpha/2$ quantile of the standard Gaussian distribution.
+For example, in option pricing one may set the error tolerance to a penny, $\varepsilon=0.01$, and wish to find error bounds which hold with 95\% confidence, $\alpha=0.05$.
+
+For classic MC with IID points, the Central Limit Theorem (CLT) may be used to get an approximate confidence interval $(\mu^-,\mu^+)$ where $\mu^\pm = \hat{\mu} \pm Z^{1-\alpha/2} \sigma/\sqrt{n}$, $\sigma = \mathrm{Std}[f(X)]$ and $Z^{1-\alpha/2}$ is the $1-\alpha/2$ quantile of the standard Gaussian distribution.
 If
@@
 \end{equation} then $\mu \in (\mu^-,\mu^+)$ with probability $1-\alpha$.
-A simple IID Monte Carlo stopping criterion uses an initial sample of $n_0$ points to approximate $\sigma^2 \approx (n_0-1)^{-1} \sum_{i=0}^{n_0-1}(f(\bsx_i)-\hat{\mu})^2$, and then plugs this into \eqref{eq:sc_clt_iid_n} to determine $n$.
-As the CLT only holds as $n$ goes to infinity, this is only a approximate stopping criterion.
+A simple IID stopping criterion uses an initial sample of $n_0$ points to approximate $\sigma^2 \approx (n_0-1)^{-1} \sum_{i=0}^{n_0-1}(f(\bsx_i)-\hat{\mu})^2$, and then plugs this into \eqref{eq:sc_clt_iid_n} to determine $n$.
+As the CLT holds only as $n$ goes to infinity, this is an approximate stopping criterion.
 For integrands $f$ with bounded kurtosis, a guaranteed version of this two-stage-sampling idea has been developed to account for the finite sample size using Berry-Esseen inequalities \cite{JiaHic16a}.
 
-A straightforward Quasi-Monte Carlo stopping criterion may also be derived from the CLT using a few randomizations of a low-discrepancy pointset.
-Let $\bsx_{ir}$ be the $i^\mathrm{th}$ point in the $r^\mathrm{th}$ randomization of a low discrepancy pointset with $0 \leq i < n$ and $1 \leq r \leq R$.
-For example, we may use $R$ independent shifts of a lattice pointset or independent digital shifts of a digital net.
-Crucially, the randomized pointsets presented in Section \ref{sec:random} are also low discrepancy.
+A straightforward qMC stopping criterion may also be derived from the CLT using a few randomizations of a LD point set.
+Let $\bsx_{ir}$ be the $i^\mathrm{th}$ point in the $r^\mathrm{th}$ randomization of a LD point set with $0 \leq i < n$ and $1 \leq r \leq R$.
+For example, we may use $R$ independent shifts of a lattice point set or independent digital shifts of a digital net.
+Crucially, the randomized point sets presented in Section \ref{sec:random} are also LD.
 Let $\hat{\mu}_r = n^{-1} \sum_{i=0}^{n-1} f(\bsx_{ir})$ be approximate means for each randomization and set $\hat{\mu} = R^{-1} \sum_{r=1}^R \hat{\mu}_r$ to be the aggregate approximation.
@@
 Here we have used the Student $t$ distribution since $R$ is typically small, for example $R=15$.
-In \cite{l2023confidence}, comprehensive numerical experiments show that this straightforward QMC stopping criterion achieve better coverage than more involved bootstrap $t$ methods for a variety of test functions.
 To use this confidence interval for a stopping criterion, one may iteratively double the sample size $n$ until the interval width is sufficiently low.
-
-One of the drawbacks of this QMC-CLT stopping criterion is that, at least intuitively, independent randomizations of a low discrepancy pointset are less efficient than simply using more low discrepancy points from a single sequence.
-Concretely, we would expect an approximation with $n=2^8$ and $R=2^3$ to perform worse than simply taking $n=2^{11}$ points from a single randomized low discrepancy sequence ($R=1$).
+In \cite{l2023confidence}, comprehensive numerical experiments show that this straightforward qMC stopping criterion achieve better coverage than more involved bootstrap $t$ methods for a variety of test functions.
+
+One of the drawbacks of this qMC-CLT stopping criterion is that, at least intuitively, independent randomizations of LD nodes are less efficient than simply using more LD points from a single sequence.
+For instance, we would expect an approximation with $n=2^8$ and $R=2^3$ to perform worse than simply taking $n=2^{11}$ points from a single randomized LD sequence ($R=1$).
 One idea in this direction is to approximate the decay rate of Fourier coefficients of $f$.
 Bounds of absolute certainty have be derived for cones of functions with sufficient decay \cite{HicEtal17a}.
-As in the QMC CLT stopping criterion, one may iteratively double $n$ until the bounds are sufficiently tight.
-
-One cone tracks the decay of approximate complex-exponential Fourier coefficients \cite{JimHic16a}.
+As in the qMC-CLT stopping criterion, one may iteratively double $n$ until the bounds are sufficiently tight.
+
+One such cone tracks the decay of approximate complex-exponential Fourier coefficients \cite{JimHic16a}.
 When paired with a randomly shifted lattice, these approximate coefficients may be computed at only $\calO(n \log n)$ cost using the fast Fourier transform.
@@
 Conditioned on observations $f(\bsx_i)$, a closed form expression exists for the $1-\alpha$ credible intervals for $\mu$.
-While also applicable when using IID points, using low discrepancy sequences paired with special kernels enables these credible intervals to be computed at only $\calO(n \log n)$ cost \cite{rathinavel2019fast}.
+While also applicable when using IID points, using LD sequences paired with special kernels enables these credible intervals to be computed at only $\calO(n \log n)$ cost \cite{rathinavel2019fast}.
 Again we may iteratively double the sample size $n$ until the credible intervals are sufficiently tight.
@@
 Extensions of these stopping criteria to functions of multiple expectations was explored in \cite{sorokin2022bounding}.
-Such problems include computing the posterior mean in a Bayesian context or the sensitivity indices for quantifying feature importance to a functions output variability.
+Such problems include computing the posterior mean in a Bayesian context or computing  sensitivity indices which  quantifying feature importance to variability in a simulation.
 This work also considered sharing samples across approximations and generalized error tolerances such as relative error tolerance.
@@
 \section{Reformulating Our Problem} \label{sec:reformulate}
-\subsection{Variable Transformations \textcolor{blue}{Aleksei}}
-
-Often (Quasi-)Monte Carlo problems may be written as $\mu = \bbE[g(\bsT)]$ where $\bsT$ is a $d$-dimensional random variable with has density $\lambda$.
-Since QMC points have low discrepancy with the standard uniform distribution $\bsX \sim \calU[0,1]^d$, we perform an invertible variable transformation $\bsPsi: [0,1]^d \to \calM$ satisfying $\mathrm{range}(\bsT) \subseteq \calM$.
+\subsection{Variable Transformations}
+
+Often (quasi-)Monte Carlo problems may be written as $\mu = \bbE[g(\bsT)]$ where $\bsT$ is a $d$-dimensional random variable with density $\lambda$.
+Since qMC points have LD with the standard uniform distribution $\bsX \sim \calU[0,1]^d$, we perform an invertible variable transformation $\bsPsi: [0,1]^d \to \calM$ satisfying $\mathrm{range}(\bsT) \subseteq \calM$.
 This implies
@@
 and $\nu$ is the density of $\bsPsi(\bsX)$.
-We may write $\nu(\bsPsi(\bsX))= \lvert \bsJ_\bsPsi(\bsX) \rvert^{-1}$ where $\lvert \bsJ_\bsPsi(\bsX) \rvert$ is the determinant of the Jacobian of $\bsPsi$.
+Specifically, $\nu(\bsPsi(\bsX))= \lvert \bsJ_\bsPsi(\bsX) \rvert^{-1}$ where $\lvert \bsJ_\bsPsi(\bsX) \rvert$ is the determinant of the Jacobian of $\bsPsi$.
 
 The above formulation is equivalent to importance sampling by a distribution with density $\nu$.
-Optimally, we would choose $\nu \propto g\lambda$ so that $f$ is constant, but this is infeasible in practice.
+Optimally, we would choose $\nu \propto g\lambda$ so that $f$ is constant, but this is infeasible in practice due to our limited knowledge of $g$.
 Choosing $\bsPsi$ so that $\bsT = \bsPsi(\bsX)$ gives the simplification $f(\bsX) = g(\bsPsi(\bsX))$.
@@
 \subsection{Variation Reduction \textcolor{blue}{Fred}} 
-\subsection{Multilevel Methods \textcolor{blue}{Fred}} \label{sec:multi}
+\subsection{Multilevel Methods} \label{sec:multi}
 The cost of computing the sample mean defined in \eqref{eq:sample_mean} is typically $\mathcal{O}(dn)$ as the dimension, $d$, and the sample size, $n$, tend to infinity.
@@
 The ability to write one's problem to have the favorable case of high, or even infinite dimension with a computational cost comparable to the low dimensional case is somewhat an art.
-However, the multilevel method takes advantage of the principle that we pointed out in Section \ref{sec:coordwts}, namely that qMC thrives when problem have a low essential dimension.
-
-\section{Beyond Computing Expectations \textcolor{blue}{Nathan}} \label{sec:beyond}
-
-Expectations of the form $\mu \coloneqq \mathbb{E}(f(\mathbf{X}))$ appear in several prominent areas of mathematics and computational science, and qMC methods have long been applied in these contexts.
+However, the multilevel method takes advantage of the principle that we pointed out in Section \ref{sec:coordwts}, namely that qMC thrives when problem have a low essential\NKNote{effective?} dimension.
+
+\section{Beyond Computing Expectations} \label{sec:beyond}
+
+Expectations of the form $\mu \coloneqq \mathbb{E}(f(\mathbf{X}))$ appear in several prominent areas of mathematics and qMC methods have long been applied in these contexts.
 Namely, in mathematical finance many problems can be formulated as high-dimensional integrals, where the large number of dimensions arises from small time steps in time discretization and/or a large number of variables.

### 2025-01-24 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

@@
 We demonstrate the performance gains possible with quasi-Monte Carlo methods.
-Moreover, we describe how to formulate problems to realize the greatest performance gains using quasi-Monte Carlo methods.
+Moreover, we describe how to formulate problems to realize the greatest performance gains using quasi-Monte Carlo.
 We also briefly describe the use of quasi-Monte Carlo methods for problems beyond computing the mean, $\mu$.
@@
 \section{Introduction} \label{sec:intro}
-There are may settings where key underlying quantities that affect the outcome are unknown, e.g.,
+There are many settings where key underlying quantities that affect the outcome are unknown, e.g.,
 \begin{itemize}
 	\item Future market forces, which affect financial risk,
-	\item The porosity of rock, which affects the extraction of oil or gas, or
+	\item The porosity field of rock, which affects the extraction of oil or gas, or
 	\item Elementary particle interactions in a high energy physics experiment, which affect what is observed by detectors.
@@
 The resulting one-dimensional integral is amenable to quadrature methods.
-This non-trivial test case with a true value that can be easily calculated allows us to compute the numerical errors of various cubature schemes.
-
-For the purpose of this illustration, we work with $\mu$ in its original form, \eqref{eq:keisterA} and approximate it by a sample mean,
+This non-trivial test case with a true value that can be easily calculated allows us to compute the numerical errors of various cubature schemes and contrast their performance.
+
+For the purpose of this illustration, we work with $\mu$ in its original form, \eqref{eq:keisterA} rather than \eqref{eq:keisterExact}, and approximate it by a sample mean,
 \begin{equation} \label{eq:sample_mean}
@@
 \end{equation}
-This does require some preparation, which is discussed in generality in Section \ref{sec:reformulate}.
-
-This $\mu$ is the expectation of $Y := g(\bsT) := \pi^{d/2} \cos(\lVert \bsT \rVert)$, where $T_1, \ldots, T_d$ are IID Gaussian (normal) random variables with zero mean and variance $1/2$, i.e., $\bsT :=(T_1, \ldots, T_d) \sim \calN(\bszero,\mathsf{I}/2)$:
+This does require some preparation to identify a suitable $f$, which is discussed in generality in Section \ref{sec:reformulate}.
+
+This $\mu$ in \eqref{eq:keisterA} is the expectation of $Y := g(\bsT) := \pi^{d/2} \cos(\lVert \bsT \rVert)$, where $T_1, \ldots, T_d$ are IID Gaussian (normal) random variables with zero mean and variance $1/2$, i.e., $\bsT :=(T_1, \ldots, T_d) \sim \calN(\bszero,\mathsf{I}/2)$:
 \begin{equation}\label{eq:keisterB}
-	\mu = \int_{\mathbb{R}^d} \underbrace{\pi^{d/2} \cos(\lVert \bst \rVert)}_{=:g(\bst)} \cdot \underbrace{\frac{\exp(-\lVert \bst \rVert^2)}{\pi^{d/2}}}_{\text{density of } \calN(\bszero,\mathsf{I}/2)} \, \rmd \bst = \bbE(Y) = \bbE[g(\bsT)].
+	\mu = \int_{\mathbb{R}^d} \underbrace{\pi^{d/2} \cos(\lVert \bst \rVert)}_{=:g(\bst)}  \underbrace{\frac{\exp(-\lVert \bst \rVert^2)}{\pi^{d/2}}}_{\text{density of } \calN(\bszero,\mathsf{I}/2)} \, \rmd \bst = \bbE(Y) = \bbE[g(\bsT)].
 \end{equation}
@@
 Grids have the largest error and LD nodes have the smallest error.
-The numerical computations used to generate this figure and the others in the is chapter may be reproduced using the Jupyter notebook \url{??}.
+The numerical computations used to generate this figure and the others in this chapter may be reproduced using the Jupyter notebook \url{??}.
 \label{fig:keister-err}}
@@
 The error decay of $\mathcal{O}(n^{-1/5})$ rather than $\mathcal{O}(n^{-2/6})$ for this example may be due to the lack of smoothness of $f$ near the boundaries of the unit cube.
-
-	\item \emph{IID.}  Simple MC is a substantial improvement over grid nodes. As a reminder, for IID nodes the root mean squared error is
-	\begin{equation}\label{eq:IIDerror}
-		\sqrt{\mathbb{E}[(\mu - \hat{\mu})^2]} = \sqrt{\var(\hat{\mu})} = \sqrt{\frac{\var(f(\bsX))}{n}} = \frac{\mathrm{Std}(f(\bsX))}{n^{1/2}},
-	\end{equation}
-	where $\var$ denotes the variance and $\mathrm{Std}$ the standard deviation.
-This $\mathcal{O}(n^{-1/2})$ decay is observed in Figure \ref{fig:keister-err}.
-For simple MC the sample size, $n$, can be any positive integer without affecting the rate of decay of the error.
-
-	Whereas grid points collapse on top of one another when viewed in low dimensional projections (Figure \ref{fig:grid}), all IID points may be seen when viewed in any lower dimensional projection, as seen in Figure \ref{fig:iid}.
-The disadvantage of IID points is that they form clusters and leave gaps.
-This is because the position of any one node is independent of the position of the others.
-
-	\item \emph{LD.}  The error of qMC methods decays nearly like $\mathcal{O}(n^{-1})$, which for this example correponds to a reduction in error of several orders of magnitude compared to simple MC for large enough $n$.
-Typically, LD sequences have preferred sample sizes.
-For this Sobol' sequence, $n$ is chosen to be non-negative integer powers of $2$, which tends to give better accuracy than arbitrary $n$.
-
-	Figure \ref{fig:ld} shows typical two-dimensional projections of a size $64$ LD node set.
-Visually, these nodes fill the unit cube better than IID nodes.
-A quantitative measure of this, the discrepancy, is defined in Section \ref{sec:discrepancy}.
-Constructions of LD node sequences are explained in the next section.
-
-\end{itemize}
-
 \begin{figure}
@@
 
+	\item \emph{IID.}  Simple MC is a substantial improvement over grid nodes. For IID nodes the root mean squared error is
+	\begin{equation}\label{eq:IIDerror}
+		\sqrt{\mathbb{E}[(\mu - \hat{\mu})^2]} = \sqrt{\var(\hat{\mu})} = \sqrt{\frac{\var(f(\bsX))}{n}} = \frac{\mathrm{Std}(f(\bsX))}{n^{1/2}},
+	\end{equation}
+	where $\var$ denotes the variance and $\mathrm{Std}$ the standard deviation.
+This $\mathcal{O}(n^{-1/2})$ decay is observed in Figure \ref{fig:keister-err}.
+For simple MC the sample size, $n$, can be any positive integer without affecting the rate of decay of the error.
+
+	Whereas grid points collapse on top of one another when viewed in low dimensional projections (Figure \ref{fig:grid}), all IID points may be seen when viewed in any lower dimensional projection, as seen in Figure \ref{fig:iid}.
+The disadvantage of IID points is that they form clusters and leave gaps.
+This is because the position of any one node is independent of the position of the others.
+
 \begin{figure}
@@
 \end{figure}
+
+	\item \emph{LD.}  The error of qMC methods decays nearly like $\mathcal{O}(n^{-1})$, which for this example correponds to a reduction in error of several orders of magnitude compared to simple MC for large enough $n$.
+Typically, LD sequences have preferred sample sizes.
+For this Sobol' sequence, $n$ is chosen to be non-negative integer powers of $2$, which tends to give better accuracy than arbitrary $n$.
+
+	Figure \ref{fig:ld} shows typical two-dimensional projections of a size $64$ LD node set.
+Visually, these nodes fill the unit cube better than IID nodes.
+A quantitative measure of this, the discrepancy, is defined in Section \ref{sec:discrepancy}.
+Constructions of LD node sequences are explained in the next section.
+
+\end{itemize}
 
@@
 One disadvantage of this construction is that it is not extensible.
-For the example in Figure \ref{fig:latticeconstruct} the first $8$ nodes do not fill the unit square well.
+For the example in Figure \ref{fig:latticeconstruct}, the first $8$ nodes do not fill the unit square well.
 However, the set of nodes defined by even $i$ do a reasonable job.
@@
 
-The extensible lattice sequence is defined by the one-dimensional van der Corput sequence, $\{\phi_b(0), \phi_b(1), \ldots\}$.
+An extensible lattice sequence in one dimension is defined by the van der Corput sequence in base $b$, $\{\phi_b(0), \phi_b(1), \ldots\}$.
 This involves the so-called radical inverse function and essentially reflects the digits of the integer $i$ in base $b$ about the decimal point.
@@
 This reordering of nodes from the original construction allows us to preserve the lattice structure for the first $b^m$ nodes for any non-negative integer $m$.
-That is, $\{\bsx_0, \ldots, \bsx_{b^m-1}\}$ is a closed under addition modulo $\bsone$.
+That is, $\{\bsx^{\text{lat}}_0, \ldots, \bsx^{\text{lat}}_{b^m-1}\}$ is a closed under addition modulo $\bsone$.
 
@@
 The blue dots are a copy of the nodes in the plot to the left, and the the orange diamonds correspond to a shifted copy of the blue dots modulo $\bsone$.
-For the left plot, the shift is $\phi_2(2)\bsh = (1,11)/4$ or equivalently $(0.25,0.75)$.
-For the middle plot, the shift is $\phi_2(4)\bsh = (1,11)/8$ or equivalently $(0.125,0.375)$.
-For the right plot, the shift is $\phi_2(8)\bsh = (1,11)/16$  or equivalently $(0.0625,0.6875)$.
+For the left plot, the shift is $\phi_2(2)\bsh = (1,11)/4=(0.25,0.75)$.
+For the middle plot, the shift is $\phi_2(4)\bsh = (1,11)/8=(0.125,0.375)$.
+For the right plot, the shift is $\phi_2(8)\bsh = (1,11)/16=(0.0625,0.6875)$.
 
@@
 	\caption{An extensible lattice corresponding to  Figure \ref{fig:latticeconstruct}  with the nodes reordered using the van der Corput sequence in base $2$.
-For each plot the blue nodes are a copy of the nodes to the left and the orange nodes are a shifted copy (modulo $\bsone$) of the blue nodes.
+For each plot the blue dots are a copy of the nodes to the left and the orange diamonds are a shifted copy (modulo $\bsone$) of the blue dots.
 \label{fig:extensiblelatticeconstruct}}
@@
 For the example in Figure \ref{fig:extensiblelatticeconstruct}, increasing $n$ beyond $16$ repeats the original nodes because the generating vector, $\bsh$, contains integers all less than $16$.
-Obtaining a truly extensible lattice sequence requires that $\bsh$ be a vector of generalized integers---essentially integers with infinite numbers of nonzero digits---as explained in \cite{HicNie03a}, where the existence of good generating vectors is also proved.
+Obtaining a truly extensible lattice sequence requires that $\bsh$ be a vector of generalized integers---essentially integers with infinite numbers of nonzero digits---as explained in \cite{HicNie03a}, where the existence of good generating vectors for infinite lattice sequences is also proved.
 In practice, one searches computationally for a $\bsh$ that produces good set of lattice nodes of size $n = b^m$ for a range of $m$ \cite{HicEtal00}.
@@
 where $\bsx^{\text{dig}}_1, \bsx^{\text{dig}}_2, \ldots$ are carefully chosen.
-(This is the van  der Corput sequence in base $2$ for $d=1$ and $x_\ell = 2^{-\ell}$.) The node set $\{ \bsx_0, \ldots, \bsx_{2^m-1}\}$  for non-negative integer $m$ is called a digital net and is closed under $\oplus$.
-
-If $x_{ij\ell}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ coordinate of $\bsx_i$, then the literature often defines generating matrices, $\mathsf{C}_j = (x_{ij\ell})_{\ell,i = 1}^{M,N}$ for $j = 1, \ldots, d$, where $M$ is the maximum number of bits in the expression for $\bsx_i$, say $52$, and $2^N$ is the maximum number of nodes that is intended to be generated.
+(For $d=1$ and $x^{\text{dig}}_{2^m} = 2^{-m-1}$, this is the van  der Corput sequence in base $2$.) The node set $\{ \bsx^{\text{dig}}_0, \ldots, \bsx^{\text{dig}}_{2^m-1}\}$  for non-negative integer $m$ is called a digital net and is closed under $\oplus$.
+
+If $x^{\text{dig}}_{ij\ell}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ coordinate of $\bsx^{\text{dig}}_i$, then the literature often defines generating matrices, $\mathsf{C}_j = \bigl(x^{\text{dig}}_{ij\ell}\bigr)_{\ell,i = 1}^{M,N}$ for $j = 1, \ldots, d$, where $M$ is the maximum number of bits in the expression for $\bsx^{\text{dig}}_i$, say $52$, and $2^N$ is the maximum number of nodes that is intended to be generated.
 The digital sequence can then be defined equivalently as
@@
 
-To illustrate the $t$-value, consider the following three dimensional digital net with $2^3$ (eight) nodes, i.e., $m = 3$:
+To illustrate the $t$-value, consider the following $d=3$ dimensional digital net with $2^3$ (eight) nodes, i.e., $m = 3$:
 \begin{multline} \label{eq:smallnet}
@@
 
-Digital sequence generators can be found via number theory or numerical search \cite{KuoJoe08a}.
-The earliest instance is due to Sobol' \cite{Sob67}.\NKNote{I added these two references which were previously left blank.
+Digital sequence generators can be found via number theory \cite[Chapter 8]{DicPil10a} or numerical search (see \cite{KuoJoe08a} and \cite[Chapter 10]{DicPil10a}).
+The earliest instance is due to Sobol' \cite{Sob67}.
+%\NKNote{I added these two references which were previously left blank.
 Were you thinking of other papers to cite here?}
@@
 Figure \ref{fig:Halton} shows two dimensional projections of Halton nodes in six dimensions.
-\NKNote{Suggestion: The Halton construction is an extensible sequence of LD nodes.
-However, should one know $n \in \mathbb{N}$ in advance, a finite LD node set called the Hammersley point set can be defined similarly as
+%\NKNote{Suggestion:
+The Halton construction is an extensible LD sequence.
+However, should one know $n \in \mathbb{N}$ in advance, a finite $d+1$-dimensional LD node set called the Hammersley point set can be defined similarly as
 \begin{equation}\label{eq:hammersley}
@@
 \end{equation}
-}
 
@@
 Most quasi-Monte Carlo constructions are studied based on their asymptotic behavior as \(n \to \infty\), with numerous results on asymptotic rates \cite{NIED1978QMCBook,NovWoz10a}.
-However, few methods address finding optimal node configurations for specific \(n\) and \(d\) which can be critical in practical applications where each function evaluation is costly.
+However, few methods address finding optimal node configurations for specific \(n\) and \(d\) which can be critical in practical applications where each function evaluation is quite costly.
 Of course, one possible strategy is to truncate asymptotically optimal sequences to the desired \(n\), however, as discussed above many of the standard LD constructions have a preferred number of nodes with truncations leading to imbalances in equidistribution.
@@
 
-When constructing LD nodes, it is essential to ensure not only a well-distributed set of nodes in the $d$-dimensional space but also in its lower-dimensional projections (recall the Cartesian grid from \ref{sec:intro}, Figure \ref{fig:grid}).
+When constructing LD nodes, it is essential to ensure not only a well-distributed set of nodes in the $d$-dimensional space but also in its lower-dimensional projections (recall the Cartesian grid from Section \ref{sec:intro}, Figure \ref{fig:grid}).
 \texttt{MatBuilder} \cite{paulin2022} is a software tool developed precisely for this task; to generate generating matrices for digital sequence LD nodes possessing excellent low dimensional uniformity.
@@
 The sample mean, $\hat{\mu}_n$, does not change each time it is computed like it would for IID nodes.
-While determinism has some advantages, randomization has substantial advantages as well.
+While determinism has advantages, randomization has substantial advantages as well.
 \begin{itemize}
-	\item Randomization done correctly removes bias in the estimator.
+	\item Randomization done correctly removes bias in the estimator, $\hat{\mu}_n$.
 	\item Replications of random estimators can facilitate error estimates for the sample mean.
@@
 A simpler version called linear matrix scrambling was proposed by Matou\v{s}ek \cite{Mat98} and is described here.
-Starting from the formulation of digital sequences involving generator matrices in \ref{eq:digitalB}, we multiply each generator matrix on the left by a lower triangular matrix with ones along the diagonals and elements below the diagonal that are randomly chosen to be $0$ or $1$ with equal probability:
+Starting from the formulation of digital sequences involving generator matrices in \eqref{eq:digitalB}, we multiply each generator matrix on the left by a lower triangular matrix with ones along the diagonals and elements below the diagonal that are randomly chosen to be $0$ or $1$ with equal probability:
 \begin{subequations} \label{eq:scr_lattice}
@@
 
-[make a scramble of an image]
-
 \subsection{Other Randomizations for Halton}
@@
 Instead of explicitly searching for good generating vectors or matrices, one can instead sample a few random generating vectors or generating matrices and then prune out bad choices \cite{}.
-Random generating vectors $\bsh$ take $h_j \sim \calU\{0,\dots,n-1\}$.
-Random generating matrices for base $2$ digital nets set $c_{ij\ell} \sim \calU\{0,1\}$ for $i<\ell$ while enforcing $c_{iji}=1$ and $c_{ij\ell}=0$ for $i>\ell$.
+Random generating vectors for lattices take $\bsh \overset{\text{IID}}{\sim} \calU\{1,\dots,n-1\}^d$.
+Random generating matrices for base $2$ digital nets---as described in \eqref{eq:digitalB}---set $x^{\text{dig}}_{ij\ell} \overset{\text{IID}}{\sim} \calU\{0,1\}$ for $i<\ell$ while enforcing $x_{iji}=1$ and $x_{ij\ell}=0$ for $i>\ell$.
 
@@
 
-If $\mu(\cdot)$ is a bounded linear functional on $\calH_K$, we may use the argument above to derive our error bound.
+If $\mu(\cdot):\calH_{\calK} \to \RR$ is a bounded linear functional on $\calH_K$, we may use the argument above to derive our error bound.
 First we write the dependence of $\mu - \hat{\mu}_n$ on  $f$ explicitly as $\mu(f) - \hat{\mu}_n(f)$ and note that the error functional, $\mu(\cdot) - \hat{\mu}_n(\cdot)$, is  linear.
@@
 This means that different choices of $K$ will lead to different error bounds for the same $\hat{\mu}_n$, even though the error is unchanged.
-	\item The Hilbert space $\calH_K$ contains ones assumptions about $f$, such as smoothness or periodicity.
+	\item The Hilbert space $\calH_K$ contains one's assumptions about $f$, such as smoothness or periodicity.
 Knowing $K$, however, does not automatically lead to an explicit formula for the inner product of the associated Hilbert space, $\langle \cdot, \cdot \rangle_{\calH_K}$.
@@
 \begin{itemize}
-	\item There are a family of commonly used reproducing kernels \cite{BerT-A04,Hic97a,Hic98b,Hic99b}.
-Once $K$ has been fixed, the discrepancy may be used as a comparison measure among different LD sequences, and its rate of decay---which sometimes can be computed theoretically---indicates how quickly $\hat{\mu}_n$ approaches $\mu$.
+	\item There are families of commonly used reproducing kernels \cite{BerT-A04,Hic97a,Hic98b,Hic99b}.
+Once $K$ has been fixed, the discrepancy may be used to compare different LD sequences.
+The discrepancy's rate of decay---which sometimes can be computed theoretically---indicates how quickly $\hat{\mu}_n$ approaches $\mu$.
 	\item When an explicit formula for the inner product of the associated Hilbert space, $\langle \cdot, \cdot \rangle_{\calH_K}$, can be deduced, it suggests how to formulate $\mu$ to reduce variation$(f,K)$.
@@
 \end{equation}
-One can easily see that this $K$ satisfies the symmetry condition \eqref{eq:KcondSym}.  It also is symmetric about the middle of the unit cube, i.e.,
+One can verify that this $K$ satisfies the symmetry condition \eqref{eq:KcondSym}.  It also is symmetric about the middle of the unit cube, i.e.,
 \begin{equation*}
@@
 \end{equation*}
-which is a desirable property for problems where there is no preferred corner.
+which is a desirable property for problems where there is no preferred corner of the unit cube.
 
@@
 \end{align}
-where $\langle \cdot,\cdot \rangle_2$ denotes the $L^2$ inner product,
+where $\langle \cdot,\cdot \rangle_2$ denotes the $L^2$ inner product, the operator $D^{\bsell}$ is defined as
 \begin{multline*}
@@
 
-We show that $K$ defined in \eqref{eq:ctrkernel} is the reproducing kernel for the $\calH_K$ with the above inner product for $d=1$ and assuming $0 \le x \le 1/2$
+We show that $K$ defined in \eqref{eq:ctrkernel} is the reproducing kernel for the $\calH_K$ with the above inner product for $d=1$ and assuming $0 \le x \le 1/2$:
 \begin{align*}
@@
 \end{align*}
-The extensions to general $1/2 < x < 1$ and to $d > 1$ are straightforward but tedious \cite{Hic97a}.
+The extensions to general $1/2 < x < 1$ and to $d > 1$ are straightforward \cite{Hic97a}.
 
@@
 Figure  \ref{fig:ctrKer} (right)  displays the discrepancy for a Sobol' LD digital net for several modest values of $d$ and for $n =  1, 2, 4, 8, 16, \ldots$.
-The decay rate of nearly $\calO(n^{-1})$ is observed for small $d$ and can be anticipated for a bit larger $d$.
+The asymptotic decay rate of nearly $\calO(n^{-1})$ is observed for small $d$ and can be anticipated for a bit larger $d$.
 Thus, the decay of the discrepancy mirrors the convergence rate for LD sequences for the Keister example in Section \ref{sec:intro} and Figure \ref{fig:keister-err}.
 
-The centered discrepancy also has a geometric interpretation, as is described in \cite{Hic97a}. Other choices of kernels and their geometric interpretations are explained there as well.
+The centered discrepancy also has a geometric interpretation, as is described in \cite{Hic97a}.
+Other choices of kernels and their geometric interpretations are explained there as well.
+Moreover, it is possible to define discrepancies for Banach spaces of integrands, however, the formulas for these discrepancies are not as amenable to calculation.
 

### 2025-01-25 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

@@
 \newcommand{\NKNote}[1]{{\textcolor{blue}{Nathan: #1}}}
+\newcommand{\AGSNote}[1]{{\textcolor{blue}{Aleksei: #1}}}
+%\usepackage[normalem]{ulem}
+
 \newcommand{\figpath}{Figures}
@@
 \begin{equation} \label{eq:sample_mean}
-	\hat{\mu}_n := \frac 1n \sum_{i=1}^n f(\bsx_i).
+	\hat{\mu}_n := \frac 1n \sum_{i=0}^{n-1} f(\bsx_i).
 \end{equation}
@@
 \begin{equation} \label{eq:sample_mean_Keister}
-	\hat{\mu}_n = \frac 1n \sum_{i=0}^{n-1} f (\bsx) = \frac 1n \sum_{i=1}^n \pi^{d/2} \cos\Bigl (\bigl \lVert \bigl( \Phi^{-1}(x_{i1}), \ldots, \Phi^{-1}(x_{id})\bigr) \bigr\rVert/\sqrt{2}  \Bigr).
+	\hat{\mu}_n = \frac 1n \sum_{i=0}^{n-1} f (\bsx) = \frac 1n \sum_{i=0}^{n-1} \pi^{d/2} \cos\Bigl (\bigl \lVert \bigl( \Phi^{-1}(x_{i1}), \ldots, \Phi^{-1}(x_{id})\bigr) \bigr\rVert/\sqrt{2}  \Bigr).
 \end{equation}
@@
 Grids have the largest error and LD nodes have the smallest error.
-The numerical computations used to generate this figure and the others in this chapter may be reproduced using the Jupyter notebook \url{??}.
 \label{fig:keister-err}}
@@
 
+As an aside, the numerical computations used to generate the figures in this chapter may be reproduced using the Jupyter notebook \url{??}.
+The Python library \texttt{qmcpy} \cite{QMCPy2020a} is used to generate the LD node sequences.
+
 \section{LD Sequence Constructions} \label{sec:construct}
@@
 	\begin{pmatrix} x^{\text{dig}}_{ij1} \\ x^{\text{dig}}_{ij2} \\ \vdots \end{pmatrix}
-	= \mathsf{C}_j \begin{pmatrix} i_0 \\ i_1 \\ \vdots \end{pmatrix} \quad \pmod{2}, \qquad j = 1,\ldots, d, \ i = 0, 1, \ldots.
+	= \mathsf{C}_j \begin{pmatrix} i_0 \\ i_1 \\ \vdots \end{pmatrix}
+    \quad \pmod{2}, \qquad j = 1,\ldots, d, \ i = 0, 1, \ldots.
 \end{equation}
@@
 
-When optimizing LD nodes, one requires an objective function which in some sense must assesses the quality of the LD nodes directly, or the expected performance when the node set is implemented in a qMC estimate \eqref{eq:sample_mean}.
+When optimizing LD nodes, one requires an objective function which in some sense must assess the quality of the LD nodes directly, or the expected performance when the node set is implemented in a qMC estimate \eqref{eq:sample_mean}.
 We will refer to such an objective function as a figure-of-merit (FOM) and popular choices include the discrepancy (see Section \ref{sec:discrepancy}) or worst-case error metrics when restricted to a specific space of functions; see \cite{Hic00a,LatNet}.
@@
 For certain choices of function spaces, there exist fast component-by-component construction implementations using fast Fourier transforms (FFTs) as the primary tool.
-We refer to \cite{DicKuo04a,JoeKuo03,KuoJoe02b,DicEtal22a} and references therein for works on the CBC construction for lattice rules and digital sequences, \cite{NuyCoo06a,NuyCoo06b} for their fast implementations and \cite{LEcEtal22a,LatNet} for a QMC software tool \texttt{LatNet Builder} by L'Ecuyer to generate LD nodes which relies heavily on the CBC construction method.
+We refer to \cite{DicKuo04a,JoeKuo03,KuoJoe02b,DicEtal22a} and references therein for works on the CBC construction for lattice rules and digital sequences, \cite{NuyCoo06a,NuyCoo06b} for their fast implementations and \cite{LEcEtal22a,LatNet} for a qMC software tool \texttt{LatNet Builder} by L'Ecuyer to generate LD nodes which relies heavily on the CBC construction method.
 
@@
 We have outlined various efforts by qMC researchers to optimize specific classes of LD constructions, including lattices, digital sequences, and Halton sequences.
-There also exist several efforts to generate novel LD constructions which don't adhere to any underlying number theoretic structure, but none-the-less, possess a high level of uniformity.
+There also exist several efforts to generate novel LD constructions which do not adhere to any underlying number theoretic structure, but none-the-less, possess a high level of uniformity.
 These methods are often global optimization methods of sample point distributions, which is often challenging due to the non-convex nature of the FOM as a function of the nodes.
@@
 We highlight the subset selection method proposed in \cite{cle22} to choose from an $n$-element initialization set, the $k<n$ nodes which minimize a discrepancy FOM.
-A heuristic approach to this problem was later shown in \cite{cle24}.
+A heuristic approach to this problem was later shown in \cite{cle24_heuristic}.
 Furthermore, a method to generate optimal star-discrepancy node sets for fixed $n$ and $d$ based on a non-linear programming approach was suggested in \cite{cle24}, however is computationally restricted to very small numbers of points in dimensions two and three only.
@@
 More generally, a permutation scramble for digital nets sets $x_{ij\ell} \mapsto \Pi_{j\ell}(x_{ij\ell})$ where $\Pi_{j\ell}$ are independent random permutations of $\{0,\dots,b_j-1\}$.
-Randomization of Halton relating to permutations of digits was discussed briefly in Section \ref{sec:optLD}.
+\AGSNote{Ignore this permutation definition as it is the same as \eqref{eq:gen_vdc}?} Randomization of Halton relating to permutations of digits was discussed briefly in Section \ref{sec:optLD}.
 Random nested uniform scrambling and linear matrix scrambling have also been considered for Halton point sets \cite{owen2024gain}.
@@
 
-Instead of explicitly searching for good generating vectors or matrices, one can instead sample a few random generating vectors or generating matrices and then prune out bad choices \cite{}.
+Instead of explicitly searching for good generating vectors or matrices, one can instead sample a few random generating vectors or generating matrices and then prune out bad choices \AGSNote{citation?}.
 Random generating vectors for lattices take $\bsh \overset{\text{IID}}{\sim} \calU\{1,\dots,n-1\}^d$.
@@
 \subsection{Coordinate weights} \label{sec:coordwts}
-When the dimension, $d$, is increased, the discrepancy increases and the rate of decay with increasing sample size deteriorates, as can be seen in Figure \ref{fig:ctrKer}.
-This raises the question of whether the LD sequences lose their effectiveness for higher dimensions or whether the theory described in the pervious section does not well describe what may occur in practice.
-
-One thing to note is that the discrepancy of the null set, which is equivalent to the norm of the linear functional $\mu(\cdot)$, is
+When the dimension, $d$, is increased, the discrepancy increases and the rate of decay with increasing sample size deteriorates, as can be seen in Figure \ref{fig:scaledweightedcentered} (left).
+This raises the question of whether the LD sequences lose their effectiveness for higher dimensions or whether the theory of the pervious section does not coincide well with practice.
+
+Note that the discrepancy of the null set, which is equivalent to the norm of the linear functional $\mu(\cdot)$, is
 \begin{equation}
@@
 
-A better approach is to introduce coordinate weights, $\gamma_1, \ldots, \gamma_d$ into the reproducing kernel.  For the kernel defining the centered discrepancy, this means
+A good approach is to introduce coordinate weights, $\gamma_1, \ldots, \gamma_d$ into the reproducing kernel.  For the kernel defining the centered discrepancy, this means
 \begin{align} \label{eq:wtctrkernel}
 	K(\bst,\bsx) &= \prod_{\ell = 1}^d \biggl [ 1 + \frac {\gamma_\ell^2}2 \bigl ( \lvert t_\ell - 1/2 \rvert + \lvert x_\ell - 1/2 \rvert - \lvert t_\ell - x_\ell \rvert \bigr ) \biggr] , \\
-	\langle f, g \rangle_{\calH_K}  &= \sum_{\bsell \subseteq \{1, \ldots, d\}} \prod_k \gamma_{\ell_k}^2 \langle D^{\bsell}f, D^{\bsell}g \rangle_2, \\
-	\text{variation}(f)  &= \sqrt{\sum_{\bsell \subseteq \{1, \ldots, d\}} \prod_k \gamma_{\ell_k}^2 \lVert D^{\bsell}f \rVert_2^2}, \\
+	\langle f, g \rangle_{\calH_K}  &= \sum_{\bsell \subseteq \{1, \ldots, d\}} \biggl(\prod_k \gamma_{\ell_k}^2 \biggr) \langle D^{\bsell}f, D^{\bsell}g \rangle_2, \\
+    \label{eq:wtctrvariation}
+	\text{variation}(f)  &= \sqrt{\sum_{\bsell \subseteq \{1, \ldots, d\}} \biggl( \prod_k \gamma_{\ell_k}^{-2} \biggr) \lVert D^{\bsell}f \rVert_2^2}, \\
 	\label{eq:wt-ctrdisc}
@@
 
-Figure \ref{fig:scaledweightedcentered} plots the scaled centered discrepancy with coordinate weights $\bsgamma = (1/\ell)_{\ell =1}^d$,
+Figure \ref{fig:scaledweightedcentered} (right) plots the scaled centered discrepancy with coordinate weights $\bsgamma = (1/\ell)_{\ell =1}^d$, i.e., 
 \begin{equation*}
@@
     \centering
-    \includegraphics[width=0.75\textwidth]{\figpath/IIDSobolScaledWtCtrDisc_n_8192_d64.pdf}
-    \caption{The scaled centered discrepancy with coordinate weights for a Sobol' LD sequence for various $d$ and the root mean squared scaled centered discrepancy with coordinate weights for IID nodes.
+     \includegraphics[width=0.45\textwidth]{\figpath/SobolCtrDisc_n_8192_d64.pdf}
+    \includegraphics[width=0.45\textwidth]{\figpath/IIDSobolScaledWtCtrDisc_n_8192_d64.pdf}
+    \caption{Left: The centered discrepancy for a Sobol' LD sequence for a range of $d$.
+For larger $d$ only a decay rate of $\calO(n^{-1/2})$, is achieved for the range of $n$ shown.
+Moreover, the discrepancy for larger $d$ is huge.
+    Right:
+    The scaled centered discrepancy with coordinate weights for a Sobol' LD sequence for various $d$ and the root mean squared scaled centered discrepancy with coordinate weights for IID nodes.
 LD nodes have a much smaller discrepancy than IID nodes.}
@@
 \end{multline}
-This is also plotted in Figure \ref{fig:scaledweightedcentered}. which highlights the superiority of LD nodes.
+This is also plotted in Figure \ref{fig:scaledweightedcentered} (right). which highlights the superiority of LD nodes.
+
+The meaning of the coordinate weights is that those coordinates with a small weight should not contribute much to the variation.
+In \eqref{eq:wtctrvariation}, if the $\gamma_{\ell_1}, \ldots, \gamma_{\ell_s}$ are small, then $\prod_k \gamma_{\ell_k}^{-2}$ will be large.
+For $f$ to have modest variation requires $\lVert D^{\bsell}f \rVert_2$ to be small.
+Even though a function may have a large nominal dimension, $d$, its effective dimension, i.e., the number of variables that substantially contribute to he variation, must be relatively few, so that the variation under coordinate weights is modest.
 
@@
 Many useful constructions of LD sequences are constructed numerically by minimizing the discrepancy, however, as noted earlier, computing the discrepancy requires $\mathcal{O}(dn^2)$ operations, which can be expensive.
-However, there is a way to reduce this computational cost for lattices and digital nets when using friendly kernels.
+However, this computational cost can be reduced for lattices and digital nets when using friendly kernels.
 
@@
 \nonumber
-\MoveEqLeft \mathbb{E}\bigl[\text{discrepancy}^2 \bigl (\{ \bsx_{i} + \bsDelta \bmod \bsone \}_{i=0}^{n-1} \bigr) \bigr] \\
+\MoveEqLeft \mathbb{E}\bigl[\text{discrepancy}^2 \bigl (\{ \bsx_{i} + \bsDelta \bmod \bsone \}_{i=0}^{n-1},K \bigr) \bigr] \\
 \nonumber
@@
     & = \frac{1}{n^2} \sum_{i,j=0}^{n-1}  \widetilde{K}(\bsx_i - \bsx_j \bmod \bsone ) - \int_{[0,1)^d} \widetilde{K}(\bsx) \, \mathrm{d} \bsx, \\
-    \intertext{where the filtered version of the reproducing kernel is defined as}
+    \intertext{and where the filtered version of the reproducing kernel is defined as}
     \widetilde{K}(\bsx) &: = \int_{[0,1)^d} K(\bsx + \bst \bmod \bsone, \bst) \, \mathrm{d} \bst.
@@
 
-For arbitrary node sets, this mean squared discrepancy is of similar computational cost as the original, but if the unshifted node set $\{\bsx_i\}_{i=0}^{n-1}$ is a lattice as defined in \eqref{eq:latticepts}, then for any $j$, the set $\{\bsx_i - \bsx_j \bmod \bsone \}_{i=0}^{n-1}$ has the same values.
+Note that 
+\begin{equation} \label{eq:shinvK}
+    K^{\text{shift}} (\bst, \bsx) := \widetilde{K}(\bst - \bsx \bmod \bsone)
+\end{equation}
+is a reproducing kernel.
+We call it a shift-invariant kernel because $K^{\text{shift}} (\bst + \bsDelta \bmod \bsone, \bsx + \bsDelta \bmod1) = \widetilde{K}(\bst - \bsx \bmod \bsone) =  K^{\text{shift}} (\bst, \bsx) $ for all $\bsDelta \in [0,1)^d$.
+
+For arbitrary node sets, this mean squared discrepancy is of similar computational cost as the original, but if the unshifted node set $\{\bsx_i\}_{i=0}^{n-1}$ is a lattice, $\{\bsx^{\text{lat}}_i\}_{i=0}^{n-1}$, as defined in \eqref{eq:latticepts}, then for any $j$, the set $\{\bsx^{\text{lat}}_i - \bsx^{\text{lat}}_j \bmod \bsone \}_{i=0}^{n-1}$ has the same values for all $j = 0, \ldots, n-1$.
 Thus the mean squared discrepancy for a randomly shifted lattice can be computed in only $\mathcal{O}(dn)$ operations, assuming that $K(\bsx)$ requires $d$ operations to compute:
-\begin{equation}
-\mathbb{E}\bigl[\text{discrepancy}^2 \bigl (\{ \bsx^{\bsDelta-\text{lat}}_{i} \}_{i=0}^{n-1} \bigr) \bigr] = \frac{1}{n} \sum_{i=0}^{n-1}  \widetilde{K}(\bsx_i) - \int_{[0,1)^d} \widetilde{K}(\bsx) \, \mathrm{d} \bsx.
-\end{equation}
+\begin{multline}
+\mathbb{E}\bigl[\text{discrepancy}^2 \bigl (\{ \bsx^{\bsDelta-\text{lat}}_{i} \}_{i=0}^{n-1}, K \bigr) \bigr] = \frac{1}{n} \sum_{i=0}^{n-1}  \widetilde{K}(\bsx_i) - \int_{[0,1)^d} \widetilde{K}(\bsx) \, \mathrm{d} \bsx \\
+= \text{discrepancy}^2 \bigl (\{ \bsx^{\text{lat}}_{i} \}_{i=0}^{n-1}, K^{\text{shift}} \bigr).
+\end{multline}
 For the kernel defining the centered discrepancy with coordinate weights,
@@
 
-\subsection{Randomized Error and Bayesian Error}
-The error analysis leading to error bound \eqref{eq:cuberrbd} is a worst-case error analysis, meaning that the error in approximating $\mu(f)$ by $\hat{\mu}_n(f)$ is definitely smaller than the discrepancy times the variation, provided that $f$ lies inside the Hilbert space with reproducing kernel $K$.
-There are other error measures that we mention briefly here.
+For digital sequences, there is an analogous development of filtering the original kernel by digitally shifts and/or scrambling.
+This provides a root mean square discrepancy that is again requires only $\mathcal{O}(dn)$ operations to compute, not $\mathcal{O}(dn^2)$ operations.
+
+\subsection{Randomized Error and Bayesian Error} \label{sec:randBayes}
+The error analysis leading to error bound \eqref{eq:cuberrbd} is a worst-case error analysis, meaning that the error in approximating $\mu(f)$ by $\hat{\mu}_n(f)$ is definitely no larger than the discrepancy times the variation, provided that $f$ lies inside the Hilbert space with reproducing kernel $K$.
+We briefly mention other error measures here.
 More detail is given in \cite{HicEtal17a}.
@@
 \end{align}
-This quality measure $\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ is typically more optimistic than the root mean squared of  $\text{discrepancy}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ because the latter is finding a worst case function for each random instance of nodes, while the former is finding one worst function.
-
-Unfortunately, $\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ is normally not as easy to compute as the discrepancy.
+This quality measure $\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ is typically more optimisti, i.e., smaller, than the root mean squared of  $\text{discrepancy}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ because the latter is finding a worst case function for each random instance of nodes, while the former is finding one worst function.
+
+Unfortunately, $\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ is typically not as easy to compute as the discrepancy.
 If $\hat{\mu}_n$ is unbiased, then $\sqrt{\mathbb{E}[(\mu(f) - \hat{\mu}_n(f))^2]} = \text{std}(\hat{\mu}_n(f))$.
@@
 
-For randomly shifted lattices $\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ tends to have the same rate of decay.
-However, for randomly scrambled and shifted digital nets, there can be an improvement
-
-The symmetric, positive definite reproducing kernel, $K$, which appears in the worst-case error analysis in Section \ref{sec:kerdisc} may instead take the role of the covariance kernel of a Gaussian process with zero mean \cite{}.
-Bayesian
+For randomly shifted lattices $\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ tends to have the same rate of decay as does the root mean squared discrepancy.
+However, for randomly scrambled and shifted digital nets, there can be an improvement in rate of decay assuming a bit more smoothness \cite{Owe97,HeiHicYue02a,HicYue00}.
+
+The symmetric, positive definite reproducing kernel, $K$, which appears in the worst-case error analysis in Section \ref{sec:kerdisc} may instead take the role of the covariance kernel of a Gaussian process with zero mean.
+Average-case error analysis \cite{Rit00a} and probabilistic numerics \cite{BriEtal18a} have pursued this approach.
 
@@
 
-Another idea which avoids multiple randomizations is to treat $f$ as a Gaussian process.
+Another idea which avoids multiple randomizations is to treat $f$ as a Gaussian process, as mentioned in Section \ref{sec:randBayes}.
 Conditioned on observations $f(\bsx_i)$, a closed form expression exists for the $1-\alpha$ credible intervals for $\mu$.
@@
 
-Specifically, pairing shifted lattices with shift-invariant kernels yields circulant Gram matrices for which computations can be done quickly using the fast Fourier transform \cite{RatHic19a}.
+Specifically, pairing shifted lattices with shift-invariant kernels as defined in \eqref{eq:shinvK} yields circulant Gram matrices for which computations can be done quickly using the fast Fourier transform \cite{RatHic19a}.
 Similarly, pairing digitally shifted digital nets with digitally-shift-invariant kernels yields recursive-block Toeplitz Gram matrices for which computations can be done quickly using fast Walsh-Hadamard transform \cite{JagHic22a}.
@@
 \section{Reformulating Our Problem} \label{sec:reformulate}
+
+There are two themes of this section.
+One is that our original problem may not look like $\mu = f(\bsX)$ for $\bsX \sim \mathcal{U}[0,1)^d$, and thus it may need to be rewritten.
+The second is that qMC does best when most of the variation of $f$ is concentrated in the lower coordinates, and we should design $f$ so that this is so, if possible.
+
 \subsection{Variable Transformations}
@@
 The above formulation is equivalent to importance sampling by a distribution with density $\nu$.
-Optimally, we would choose $\nu \propto g\lambda$ so that $f$ is constant, but this is infeasible in practice due to our limited knowledge of $g$.
+Optimally, we would want to choose $\nu \propto g\lambda$ so that $f$ is nearly constant, but this is typically infeasible in practice due to our limited knowledge of $g$.
 Choosing $\bsPsi$ so that $\bsT = \bsPsi(\bsX)$ gives the simplification $f(\bsX) = g(\bsPsi(\bsX))$.
@@
 
-\subsection{Variation Reduction \textcolor{blue}{Fred}} 
+As observed in Section \ref{sec:coordwts}, qMC is more effective if the variation of $f$ is concentrated in the lower indexed coordinates.
+For finance examples where the outcome depends more on the gross behavior of the underlying stochastic process, rather than the high frequency behavior, it makes sense to choose the transformation accordingly.
+For example, if $\bsT$ represents a Brownian motion, then a principal component analysis coupled with the inverse Gaussian distribution shows greater convergence gains.
+
+\subsection{Control Variates}
+
+Since the upper bound of the qMC error includes the variation of $f$ as a factor, one can often improve the efficiency of qMC by reformulating the problem to reduce this variation.
+Choosing a good variable transformation, which is equivalent to importance sampling, to minimize $\text{variation}(f,K)$ is one approach.
+
+Another approach is control variates.
+If $f_{\text{ctrl}} : [0,1)^d \to \RR$ is a function whose integral is zero, then $\mu(f) = \mu(f - f_{\text{ctrl}})$, and one may approximate $\mu(f)$ by $\hat{\mu}_n(f - f_{\text{ctrl}})$.
+If the variation of  $f - f_{\text{ctrl}}$ is smaller than that of $f$, then it is likely that $\hat{\mu}_n(f - f_{\text{ctrl}})$ will have smaller error than $\hat{\mu}_n(f)$.
+
+An example of this in quantitative finance is where one may use the payoff of a European option, whose expectation is known, as a control variate for more exotic options.
+As noted in \cite{HicEtal03}, reducing the variation of $f$ via control variates may or may not bear resemblance to reducing the variance of $f$.
+Control variates have been implemented in the stopping criteria based on Fourier coefficients described in the previous section \cite{HicEtal17a}.
+
 \subsection{Multilevel Methods} \label{sec:multi}
@@
 The ability to write one's problem to have the favorable case of high, or even infinite dimension with a computational cost comparable to the low dimensional case is somewhat an art.
-However, the multilevel method takes advantage of the principle that we pointed out in Section \ref{sec:coordwts}, namely that qMC thrives when problem have a low essential\NKNote{effective?} dimension.
+However, the multilevel method takes advantage of the principle pointed out in Section \ref{sec:coordwts}, namely that qMC thrives when problem have a low effective dimension.
 
@@
 Besides computing expectations, qMC methods have found many applications in other areas of the computational sciences.
-We lightly review these applications to highlight their wider practical impact and to motivate the further development of QMC theory and methodology.
+We lightly review these applications to highlight their wider practical impact and to motivate the further development of qMC theory and methodology.
 
@@
 %\section{Ongoing Research} \label{sec:ongoing}
-\section{Conclusion \textcolor{blue}{Fred}} \label{sec:conclusion}
+\section{Conclusion} \label{sec:conclusion}
+Quasi-Monte Carlo methods, based on low discrepancy sequences, can approximate to the mean of a function, $f$, of $d$ variables to within an error tolerance of $\varepsilon$ at a cost of roughly $\mathcal{O}(\varepsilon^{-1})$ function evaluations, provided that the problem is formulated so that $f$ depends primarily on a modest number of variables.
+This means that qMC outperforms alternative methods such as simple Monte Carlo and sampling on grids.
+Using the multilevel qMC, the cost to reach the error tolerance is often essentially the same for $d=1$ and $d$ unbounded.
+Software libraries are available to facilitate qMC computations.
 

### 2025-01-26 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

@@
 For $f$ to have modest variation requires $\lVert D^{\bsell}f \rVert_2$ to be small.
-Even though a function may have a large nominal dimension, $d$, its effective dimension, i.e., the number of variables that substantially contribute to he variation, must be relatively few, so that the variation under coordinate weights is modest.
+Even though a function may have a large nominal dimension, $d$, its effective dimension, i.e., the number of variables that substantially contribute to the variation, must be relatively few, so that the variation under coordinate weights is modest.
 
@@
 \end{align}
-This quality measure $\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ is typically more optimisti, i.e., smaller, than the root mean squared of  $\text{discrepancy}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ because the latter is finding a worst case function for each random instance of nodes, while the former is finding one worst function.
+This quality measure $\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ is typically more optimistic, i.e., smaller, than the root mean squared of  $\text{discrepancy}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ because the latter is finding a worst case function for each random instance of nodes, while the former is finding one worst function.
 
@@
 For example, in option pricing one may set the error tolerance to a penny, $\varepsilon=0.01$, and wish to find error bounds which hold with 95\% confidence, $\alpha=0.05$.
-
-For classic MC with IID points, the Central Limit Theorem (CLT) may be used to get an approximate confidence interval $(\mu^-,\mu^+)$ where $\mu^\pm = \hat{\mu} \pm Z^{1-\alpha/2} \sigma/\sqrt{n}$, $\sigma = \mathrm{Std}[f(X)]$ and $Z^{1-\alpha/2}$ is the $1-\alpha/2$ quantile of the standard Gaussian distribution.
+The stopping criteria presented here are implemented in QMCPy \cite{QMCPy2020a}.
+
+For simple Monte Carlo with IID points, the Central Limit Theorem (CLT) may be used to get an approximate confidence interval $(\mu^-,\mu^+)$ where $\mu^\pm = \hat{\mu} \pm Z^{1-\alpha/2} \sigma/\sqrt{n}$, $\sigma = \mathrm{Std}[f(X)]$ and $Z^{1-\alpha/2}$ is the $1-\alpha/2$ quantile of the standard Gaussian distribution.
 If
@@
 As the CLT holds only as $n$ goes to infinity, this is an approximate stopping criterion.
-For integrands $f$ with bounded kurtosis, a guaranteed version of this two-stage-sampling idea has been developed to account for the finite sample size using Berry-Esseen inequalities \cite{JiaHic16a}.
+For integrands $f$ with bounded kurtosis, a guaranteed version of this two-stage-sampling idea has been developed to account for the finite sample size using  Berry-Esseen inequalities\cite{JiaHic16a}.
 
@@
 To use this confidence interval for a stopping criterion, one may iteratively double the sample size $n$ until the interval width is sufficiently low.
-In \cite{l2023confidence}, comprehensive numerical experiments show that this straightforward qMC stopping criterion achieve better coverage than more involved bootstrap $t$ methods for a variety of test functions.
+In \cite{l2023confidence}, comprehensive numerical experiments show that this straightforward qMC stopping criterion achieve better coverage than more involved bootstrap $t$ methods across a variety of test functions.
 
@@
 While also applicable when using IID points, using LD sequences paired with special kernels enables these credible intervals to be computed at only $\calO(n \log n)$ cost \cite{rathinavel2019fast}.
-Again we may iteratively double the sample size $n$ until the credible intervals are sufficiently tight.
+Again one may iteratively double the sample size $n$ until the credible intervals are sufficiently tight.
 
 Specifically, pairing shifted lattices with shift-invariant kernels as defined in \eqref{eq:shinvK} yields circulant Gram matrices for which computations can be done quickly using the fast Fourier transform \cite{RatHic19a}.
-Similarly, pairing digitally shifted digital nets with digitally-shift-invariant kernels yields recursive-block Toeplitz Gram matrices for which computations can be done quickly using fast Walsh-Hadamard transform \cite{JagHic22a}.
+Similarly, pairing digitally shifted digital nets with digitally-shift-invariant kernels yields recursive-block Toeplitz Gram matrices \AGSNote{Different name?} for which computations can be done quickly using the fast Walsh-Hadamard transform \cite{JagHic22a}.
 
@@
 
-Often (quasi-)Monte Carlo problems may be written as $\mu = \bbE[g(\bsT)]$ where $\bsT$ is a $d$-dimensional random variable with density $\lambda$.
+Often qMC problems may be written as $\mu = \bbE[g(\bsT)]$ where $\bsT$ is a $d$-dimensional random variable with density $\lambda$.
 Since qMC points have LD with the standard uniform distribution $\bsX \sim \calU[0,1]^d$, we perform an invertible variable transformation $\bsPsi: [0,1]^d \to \calM$ satisfying $\mathrm{range}(\bsT) \subseteq \calM$.
@@
 
-The above formulation is equivalent to importance sampling by a distribution with density $\nu$.
-Optimally, we would want to choose $\nu \propto g\lambda$ so that $f$ is nearly constant, but this is typically infeasible in practice due to our limited knowledge of $g$.
+The above formulation is equivalent to importance sampling \AGSNote{reference} by a distribution with density $\nu$.
+Optimally, one would choose $\nu \propto g\lambda$ so that $f$ is constant, but this is typically infeasible in practice due to our limited knowledge of $g$.
 Choosing $\bsPsi$ so that $\bsT = \bsPsi(\bsX)$ gives the simplification $f(\bsX) = g(\bsPsi(\bsX))$.
@@
 For finance examples where the outcome depends more on the gross behavior of the underlying stochastic process, rather than the high frequency behavior, it makes sense to choose the transformation accordingly.
-For example, if $\bsT$ represents a Brownian motion, then a principal component analysis coupled with the inverse Gaussian distribution shows greater convergence gains.
+For example, if $\bsT$ represents a Brownian motion, then a principal component analysis coupled with the inverse Gaussian distribution shows greater convergence gains \AGSNote{compared to the Cholesky decomposition or bridge construction?}.
 
@@
 \end{multline}
-If $C_1$ is the computational cost of $f^{(1)}(\bsx^{(1)})$ and $C_l$ is the computational cost of $f^{(l)}(\bsx^{(1)})$ of $f_L(\bsx_i^{(l)}) - f_{l-1}(\bsx_i^{(l-1)})$ for $l = 2, \ldots, L$, then the total computational cost of
+If $C_1$ is the computational cost of $f^{(1)}(\bsx^{(1)})$ and $C_l$ is the computational cost of \AGSNote{$f^{(l)}(\bsx^{(1)})$ of $f_L(\bsx_i^{(l)}) - f_{l-1}(\bsx_i^{(l-1)})$} for $l = 2, \ldots, L$, then the total computational cost of
 \eqref{eq:MLmuhat} is 
@@
 \end{equation}
-In this case $\text{cost}(\bsn,\bsd)$ for this high dimensional problem is ont much worse than the cost of ensuring that just error of the rough $d_1$ approximation is smaller than $\varepsilon$, which would be $V_1C_1/\varepsilon \le \alpha \beta d_1/(r\varepsilon)$.
+In this case $\text{cost}(\bsn,\bsd)$ for this high dimensional problem is not much worse than the cost of ensuring that just error of the rough $d_1$ approximation is smaller than $\varepsilon$, which would be $V_1C_1/\varepsilon \le \alpha \beta d_1/(r\varepsilon)$.
 
@@
 \section{Conclusion} \label{sec:conclusion}
-Quasi-Monte Carlo methods, based on low discrepancy sequences, can approximate to the mean of a function, $f$, of $d$ variables to within an error tolerance of $\varepsilon$ at a cost of roughly $\mathcal{O}(\varepsilon^{-1})$ function evaluations, provided that the problem is formulated so that $f$ depends primarily on a modest number of variables.
+Quasi-Monte Carlo methods, based on low discrepancy sequences, can approximate the mean of a function, $f$, of $d$ variables to within an error tolerance of $\varepsilon$ at a cost of roughly $\mathcal{O}(\varepsilon^{-1})$ function evaluations, provided that the problem is formulated so that $f$ depends primarily on a modest number of variables.
 This means that qMC outperforms alternative methods such as simple Monte Carlo and sampling on grids.

### 2025-01-27 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

@@
 In mathematical terms, the random outcome is $Y := f(\bsX)$, where $\bsX$ is a vector random input.
-Given an input, $\bsx$, the corresponding outcome  $y = f(\bsx)$ can be computed by an algorithm, whose complexity makes $f$ a \emph{black box}, or at least a gray box.
+Given an input, $\bsx$, the corresponding outcome  $y = f(\bsx)$ can be computed by an algorithm, whose complexity makes $f$ a \emph{black box}, or at least a \emph{gray box}.
 The user selects a sequence of input values, also known as data sites or \emph{nodes}, $\bsx_0, \bsx_1, \ldots$, which lead to observed outcomes $y_0 = f(\bsx_0), y_1 = f(\bsx_1), \ldots$.
@@
 We then discuss how to rewrite the problem of interest in a qMC-friendly way (Section \ref{sec:reformulate}).
-We briefly discuss applications of qMC beyond computing the mean (Section \ref{sec:beyond}), and finish with a brief conclusion (Section \ref{sec:conclusion}).
+We briefly discuss applications of qMC beyond computing the mean (Section \ref{sec:beyond}), and finish with a short conclusion (Section \ref{sec:conclusion}).
 
@@
 \end{equation}
-Virtually all LD sequences, which underlie qMC, are defined to approximate the standard uniform distribution, $\calU[0,1]^d$.
+Virtually all LD sequences underlying qMC are defined to approximate the standard uniform distribution, $\calU[0,1]^d$.
 Thus, we perform a variable transformation $\bst = \bigl (\Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d) \bigr )/\sqrt{2}$,
@@
 	\includegraphics[width =0.5\textwidth]{\figpath/n_is_16777216_d_is_6_n_rep_is_50_KeisterErrors.pdf}
-	\caption{The relative error of approximating $\mu$ defined in \eqref{eq:keisterC} by the sample mean, $\hat{\mu}_n$, defined in \eqref{eq:sample_mean} for various choices of nodes, $\bsx_0, \bsx_1, \ldots$.
+	\caption{The relative error of approximating $\mu$ defined in \eqref{eq:keisterC} by the sample mean, $\hat{\mu}_n$, defined in \eqref{eq:sample_mean_Keister} for various choices of nodes, $\bsx_0, \bsx_1, \ldots$.
 Grids have the largest error and LD nodes have the smallest error.
@@
 For $d = 6$, the possible sample sizes, $n$, are quite sparse, as shown in Figure \ref{fig:keister-err}.
+
+    Moreover, grids are not naturally extensible.  To move from $n = m^6$ nodes to $n = (m+1)^6$ nodes, one must discard the original nodes.
 
@@
 	\includegraphics[width =  0.3\textwidth]{\figpath/64gridpts_d6.pdf}
-	\caption{Although a two-dimensional grid covers the unit square rather well, a $d$-dimensional grid for modest $d$ does \emph{not} cover the unit cube well.
-For example, one can only see four distinct points in a two dimensional projection of a six-dimensional grid with $64$ points.
+	\caption{\FJHNote{Is the font size large enough?}Although a two-dimensional grid covers the unit square rather well, a $d$-dimensional grid for modest $d$ does \emph{not} cover the unit cube well.
+For example, one can only see four distinct nodes in a two dimensional projection of a six-dimensional grid with $64$ nodes.
 \label{fig:grid}}
@@
 
-	Whereas grid points collapse on top of one another when viewed in low dimensional projections (Figure \ref{fig:grid}), all IID points may be seen when viewed in any lower dimensional projection, as seen in Figure \ref{fig:iid}.
-The disadvantage of IID points is that they form clusters and leave gaps.
+	Whereas grid points collapse on top of one another when viewed in low dimensional projections (Figure \ref{fig:grid}), all IID nodes may be seen when viewed in any lower dimensional projection, as seen in Figure \ref{fig:iid}.
+The disadvantage of IID nodes is that they form clusters and leave gaps.
 This is because the position of any one node is independent of the position of the others.
@@
 	\includegraphics[width =\textwidth]{\figpath/64iidpts_d6.pdf}
-	\caption{IID points cover the unit cube better than grid points, although one does observe clusters and gaps.
-In any  projection there is  a similar looking distribution of all $64$ points \label{fig:iid}}
+	\caption{IID nodes cover the unit cube better than grid nodes, although one does observe clusters and gaps.
+In any  projection there is  a similar looking distribution of all $64$ nodes \label{fig:iid}}
 \end{figure}
@@
 Typically, LD sequences have preferred sample sizes.
-For this Sobol' sequence, $n$ is chosen to be non-negative integer powers of $2$, which tends to give better accuracy than arbitrary $n$.
+This particular LD sequence is a Sobol' sequence, where the preferred sample sizes are non-negative integer powers of $2$.
 
@@
 	\includegraphics[width = \textwidth]{\figpath/64sobolpts_d6.pdf}
-	\caption{LD points cover the unit cube even better than IID points or grids.  In any  projection there is  a similar looking distribution of all $64$ points. \label{fig:ld}}
+	\caption{LD nodes cover the unit cube even better than IID nodes or grids.  In any  projection there is  a similar looking distribution of all $64$ nodes. \label{fig:ld}}
 \end{figure}
 
-As an aside, the numerical computations used to generate the figures in this chapter may be reproduced using the Jupyter notebook \url{??}.
+The numerical computations used to generate the figures in this chapter may be reproduced using the Jupyter notebook \url{??}.
 The Python library \texttt{qmcpy} \cite{QMCPy2020a} is used to generate the LD node sequences.
@@
 	\includegraphics[width = \textwidth]{\figpath/16_lattice_construct_d2.pdf}
-	\caption{The construction of a good lattice point set in two dimensions with $16$ nodes is obtained by moving $\bsh/n$ beyond the present point and wrapping around the boundaries of the square (hypercube in $d$ dimensions) until one returns to the origin.
+	\caption{The construction of a good lattice node set in two dimensions with $16$ nodes is obtained by moving $\bsh/n$ beyond the present node and wrapping around the boundaries of the square (hypercube in $d$ dimensions) until one returns to the origin.
 \label{fig:latticeconstruct}}
@@
 
-Another family of LD sequences built upon the van der Corput sequence \eqref{eq:vdc} is digital sequences \cite{DicPil10a,Nie92}.
+Another family of LD sequences that is a generalization of the van der Corput sequence \eqref{eq:vdc} is digital sequences \cite{DicPil10a,Nie92}.
 For simplicity, we restrict ourselves to $b = 2$ and let $\oplus$ denote binary digitwise addition modulo $2$ (also known as digitwise exclusive or), e.g., $3/8 \oplus 3/4 = {}_20.011 \oplus {}_20.110 = {}_20.101 = 5/8$.
@@
 
-To understand how these nodes might be evenly distributed over $[0,1)^d$, imagine a box of the form
+To understand how well these nodes are evenly distributed over $[0,1)^d$, imagine a box of the form
 \begin{multline} \label{eq:element_box}
@@
 	\text{the first $k_d$ rows of the first $m$ columns of $\mathsf{C}_d$} \\
-	\text{are linearly independent over the finite field with the elements $\{0,1\}$, }
+	\text{are linearly independent over addition and multiplication modulo $2$, }
 \end{multline}
@@
 	\includegraphics[width = \textwidth]{\figpath/8_sobol_sequence_elementary_intervals_d3_coord_(1,3).pdf}
-	\caption{Each box in the two dimensional projections of the node set \eqref{eq:smallnet} plotted above show one node in each box except for the second row, third plot from the left.
+	\caption{Each box in the two dimensional projections of the node set \eqref{eq:smallnet} plotted above contain one node except for the second row, third plot from the left.
 Thus, this node set is not a $(0,3,3)$-net.
@@
 \subsection{Halton sequences} \label{sec:Halton}  While lattice sequences and digital sequences have preferred sample sizes, $n$, Halton sequences have no preferred sample size.
-The Halton sequence is defined in terms of the van der Corput sequence
+The Halton sequence is defined in terms of the van der Corput sequences for different bases:
 \begin{equation}\label{eq:Halton}
@@
 %\NKNote{Suggestion:
+
 The Halton construction is an extensible LD sequence.
-However, should one know $n \in \mathbb{N}$ in advance, a finite $d+1$-dimensional LD node set called the Hammersley point set can be defined similarly as
+However, should one know $n \in \mathbb{N}$ in advance, a finite $d+1$-dimensional LD node set called the Hammersley node set can be defined similarly as
 \begin{equation}\label{eq:hammersley}
@@
 
-In the previous section we discussed how the Halton sequence has the benefit of not having a preferred number of points for its construction.
+\FJHNote{Should this paragraph go here or in the next section?}
+In the previous section we discussed how the Halton sequence has the benefit of not having a preferred number of nodes for its construction.
 It does however possess the drawback of exhibiting poor uniformity in the lower dimensional projections when $d$ is moderate to high due to number theoretic correlation between the coordinates \cite{kirklem24}.
@@
 A heuristic approach to this problem was later shown in \cite{cle24_heuristic}.
-Furthermore, a method to generate optimal star-discrepancy node sets for fixed $n$ and $d$ based on a non-linear programming approach was suggested in \cite{cle24}, however is computationally restricted to very small numbers of points in dimensions two and three only.
+Furthermore, a method to generate optimal star-discrepancy node sets for fixed $n$ and $d$ based on a non-linear programming approach was suggested in \cite{cle24}, however is computationally restricted to very small numbers of nodes in dimensions two and three only.
 
@@
 That is, $\bsx_i^{\bsDelta} : = \bsx_i + \bsDelta \pmod \bsone$ or  $\bsx_i^{\bsDelta} : = \bsx_i \oplus \bsDelta$ and $\bsDelta \sim \calU[0,1)^d$.
-This then implies that each $\bsx_i^{\bsDelta} \sim \calU[0,1)^d$ and thus $\hat{\mu}_n$ is an unbiased estimator of $\mu$, which was mentioned at the beginning of this section.
+This then implies that each $\bsx_i^{\bsDelta} \sim \calU[0,1)^d$ and thus $\hat{\mu}_n$ is an unbiased estimator of $\mu$, which was mentioned as an advantage at the beginning of this section.
 
@@
 \begin{gather} \label{eq:digitalscr}
-	\begin{pmatrix} x^{\text{scr}}_{ij1} \\ x^{\text{scr}}_{ij2} \\ \vdots \end{pmatrix}
+	\begin{pmatrix} x^{\text{scr-net}}_{ij1} \\ x^{\text{scr-net}}_{ij2} \\ \vdots \end{pmatrix}
 	= \mathsf{L_j} \mathsf{C}_j \begin{pmatrix} i_0 \\ i_1 \\ \vdots \end{pmatrix} + \begin{pmatrix} \Delta_{1j} \\ \Delta_{2j} \\ \vdots \end{pmatrix} \quad \pmod{2}, \qquad j = 1,\ldots, d, \ i = 0, 1, \ldots, \\
@@
 	\includegraphics[width = \textwidth]{\figpath/n_8_n_scr_3_scrambled_sobol_sequence_d2.pdf}
-	\caption{Three linear scrambles of a digital $(0,3,2)$-net with nodes denoted by three different colors.
+	\caption{\FJHNote{Something went awry here.}Three linear scrambles of a digital $(0,3,2)$-net with nodes denoted by three different colors.
 There is one node in each smaller box of the form \eqref{eq:element_box} for $\lVert \bsk \rVert = 3$, for each digitally shifted net.
@@
 
-As Halton nodes can be viewed as digital nets with different bases for each dimension, the randomization techniques used for digital nets can be applied to Halton point sets\NKNote{we should clarify here whether we mean truncated Halton, or Hammersley if it is introduced above} Random digital shifts for Halton set $x_{ij\ell} \mapsto (x_{ij\ell} + \Delta_{j\ell}) \mod b_j$ where the shift $\Delta_{j\ell} \in \{0,\dots,b_j-1\}$.
+As Halton nodes can be viewed as digital nets with different bases for each dimension, the randomization techniques used for digital nets can be applied to Halton node sets\NKNote{we should clarify here whether we mean truncated Halton, or Hammersley if it is introduced above} Random digital shifts for Halton set $x_{ij\ell} \mapsto (x_{ij\ell} + \Delta_{j\ell}) \mod b_j$ where the shift $\Delta_{j\ell} \in \{0,\dots,b_j-1\}$.
 More generally, a permutation scramble for digital nets sets $x_{ij\ell} \mapsto \Pi_{j\ell}(x_{ij\ell})$ where $\Pi_{j\ell}$ are independent random permutations of $\{0,\dots,b_j-1\}$.
 \AGSNote{Ignore this permutation definition as it is the same as \eqref{eq:gen_vdc}?} Randomization of Halton relating to permutations of digits was discussed briefly in Section \ref{sec:optLD}.
-Random nested uniform scrambling and linear matrix scrambling have also been considered for Halton point sets \cite{owen2024gain}.
+Random nested uniform scrambling and linear matrix scrambling have also been considered for Halton node sets \cite{owen2024gain}.
 
@@
 	\intertext{since $\hat{\mu}_n$ is exact for constants.
-(It is assumed that $\calH$ contains constant functions so that $f \in \calH_K$ implies that $f - c \in \calH_K$.)  The squared norm of the error funtional can be expressed explicitly via \eqref{eq:sqnormL} as}
+(It is assumed that $\calH_K$ contains constant functions so that $f \in \calH_K$ implies that $f - c \in \calH_K$.)  The squared norm of the error funtional can be expressed explicitly via \eqref{eq:sqnormL} as}
 	\nonumber
@@
 \begin{itemize}
-	\item the discrepancy, which is the norm of the error functional, is a measure of deficiency of the node set, $\{\bsx\}_{i=0}^{n-1}$, and
-	\item the variation is a measure of the roughness of the function defining our random variable whose expectation we wish to compute.
+	\item The discrepancy, which is the norm of the error functional, depends only on the node set, $\{\bsx\}_{i=0}^{n-1}$, and \emph{not} on $f$.
+The discrepancy measures the deficiency of that node set.
+	\item The variation depends only on $f$, the function defining our random variable $Y$ whose expectation we wish to compute,  and \emph{not} on the node set, $\{\bsx\}_{i=0}^{n-1}$.
+The variation is a measure of the roughness of $f$.
 \end{itemize}
@@
 In both cases, educated guesswork is involved.
-	\item Choosing a more restrictive $\calH_K$ may lead to a sharper error bound, however, it is often infeasible in practice to check whether the ``black box'' $f$ lies in $\calH_K$.
-	\item   Moreover, even if one is confident that $f \in \calH_K$, it is usually impractical to compute variation$(f,K)$, as we shall see below.
+	\item Choosing a more restrictive $\calH_K$ may lead to a sharper error bound, however, it is often infeasible in practice to check whether the $f$ lies in $\calH_K$.
+This is what we mean in the introduction by referring to $f$ as a ``black box''.
+	\item   Moreover, even if one is confident that $f \in \calH_K$ ($f$ is a ``gray box''), it is usually impractical to compute variation$(f,K)$, as we shall see in the discrepancy example below.
 Error bound \eqref{eq:cuberrbd} cannot be used as a stopping criterion to determine the $n$ needed to satisfy an error tolerance.
@@
 \begin{align}
-	\langle f, g \rangle_{\calH_K} &= f(\boldsymbol{1/2}) g(\boldsymbol{1/2}) \nonumber
+	\langle f, g \rangle_{\calH_K} &= f(1/2, \ldots, 1/2) g(1/2, \ldots, 1/2) \nonumber
 	\\
@@
 	[D^{\{\ell_1, \ell_2, \ldots, \ell_s\} }f](x_1, x_2, \ldots, x_s) \\
-	: = \frac{\partial^s f(1/2,\ldots, 1/2, x_{\ell_1}, 1/2, \ldots, 1/2, x_{\ell_2}, 1/2, \ldots, 1/2, x_{\ell_s}, 1/2, \ldots, 1/2)}{\partial x_{\ell_1} \partial x_{\ell_2} \cdots \partial x_s},
+	: = \frac{\partial^s f(1/2,\ldots, 1/2, x_{\ell_1}, 1/2, \ldots, 1/2, x_{\ell_2}, 1/2, \ldots, 1/2, x_{\ell_s}, 1/2, \ldots, 1/2)}{\partial x_{\ell_1} \partial x_{\ell_2} \cdots \partial x_{\ell_s}},
 \end{multline*}
@@
 When the dimension, $d$, is increased, the discrepancy increases and the rate of decay with increasing sample size deteriorates, as can be seen in Figure \ref{fig:scaledweightedcentered} (left).
-This raises the question of whether the LD sequences lose their effectiveness for higher dimensions or whether the theory of the pervious section does not coincide well with practice.
+This raises the question of whether the LD sequences lose their effectiveness for higher dimensions or whether the theory of the previous subsection does not coincide with practice.
 
@@
 
-A good approach is to introduce coordinate weights, $\gamma_1, \ldots, \gamma_d$ into the reproducing kernel.  For the kernel defining the centered discrepancy, this means
+A good approach is to introduce coordinate weights, $\gamma_1, \ldots, \gamma_d$ into the reproducing kernel.
+These were introduced in \cite{Hic95} but begin to be fully exploited in \cite{SloWoz98} and the articles that followed.
+For the kernel defining the centered discrepancy, this means
 \begin{align} \label{eq:wtctrkernel}
-	K(\bst,\bsx) &= \prod_{\ell = 1}^d \biggl [ 1 + \frac {\gamma_\ell^2}2 \bigl ( \lvert t_\ell - 1/2 \rvert + \lvert x_\ell - 1/2 \rvert - \lvert t_\ell - x_\ell \rvert \bigr ) \biggr] , \\
-	\langle f, g \rangle_{\calH_K}  &= \sum_{\bsell \subseteq \{1, \ldots, d\}} \biggl(\prod_k \gamma_{\ell_k}^2 \biggr) \langle D^{\bsell}f, D^{\bsell}g \rangle_2, \\
+	K_{\bsgamma}(\bst,\bsx) &= \prod_{\ell = 1}^d \biggl [ 1 + \frac {\gamma_\ell^2}2 \bigl ( \lvert t_\ell - 1/2 \rvert + \lvert x_\ell - 1/2 \rvert - \lvert t_\ell - x_\ell \rvert \bigr ) \biggr] , \\
+	\langle f, g \rangle_{\calH_{K_{\bsgamma}}}  &= \sum_{\bsell \subseteq \{1, \ldots, d\}} \biggl(\prod_k \gamma_{\ell_k}^2 \biggr) \langle D^{\bsell}f, D^{\bsell}g \rangle_2, \\
     \label{eq:wtctrvariation}
-	\text{variation}(f)  &= \sqrt{\sum_{\bsell \subseteq \{1, \ldots, d\}} \biggl( \prod_k \gamma_{\ell_k}^{-2} \biggr) \lVert D^{\bsell}f \rVert_2^2}, \\
+	\text{variation}(f,K_{\bsgamma})  &= \sqrt{\sum_{\bsell \subseteq \{1, \ldots, d\}} \biggl( \prod_k \gamma_{\ell_k}^{-2} \biggr) \lVert D^{\bsell}f \rVert_2^2}, \\
 	\label{eq:wt-ctrdisc}
 	\nonumber
-	\MoveEqLeft[4]\text{wt-ctr-discrepancy}^2\bigl(\{ \bsx_i\}_{i=0}^{n-1} \bigr) \\
+	\MoveEqLeft[4]\text{wt-ctr-discrepancy}^2\bigl(\{ \bsx_i\}_{i=0}^{n-1} , \bsgamma\bigr) \\
 	\nonumber
@@
 \begin{equation*}
-    \text{wt-ctr-discrepancy}\bigl(\{ \bsx_i\}_{i=0}^{n-1} \bigr)/\text{wt-ctr-discrepancy}(\emptyset ),
+    \text{wt-ctr-discrepancy}\bigl(\{ \bsx_i\}_{i=0}^{n-1},\bsgamma \bigr)/\text{wt-ctr-discrepancy}(\emptyset,\bsgamma ),
 \end{equation*}
@@
 Even though a function may have a large nominal dimension, $d$, its effective dimension, i.e., the number of variables that substantially contribute to the variation, must be relatively few, so that the variation under coordinate weights is modest.
+Under these conditions, the corresponding discrepancy may retain the nearly $\mathcal{O}(n^{-1})$ decay.
 
@@
 \nonumber
-\MoveEqLeft \mathbb{E}\bigl[\text{discrepancy}^2 \bigl (\{ \bsx_{i} + \bsDelta \bmod \bsone \}_{i=0}^{n-1},K \bigr) \bigr] \\
+\MoveEqLeft \mathbb{E}\bigl[\text{discrepancy}^2 \bigl (\{ \bsx^{\bsDelta}_{i} \}_{i=0}^{n-1},K \bigr) \bigr] \\
+\nonumber
+& = \mathbb{E}\bigl[\text{discrepancy}^2 \bigl (\{ \bsx_{i} + \bsDelta \bmod \bsone \}_{i=0}^{n-1},K \bigr) \bigr] \\
 \nonumber
@@
 \end{align}
-This quality measure $\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ is typically more optimistic, i.e., smaller, than the root mean squared of  $\text{discrepancy}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ because the latter is finding a worst case function for each random instance of nodes, while the former is finding one worst function.
+This quality measure $\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ is typically more optimistic (smaller), than the root mean squared of  $\text{discrepancy}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ because the latter finds a worst case function for each random instance of nodes, while the former is finds just one worst function.
 
@@
 The symmetric, positive definite reproducing kernel, $K$, which appears in the worst-case error analysis in Section \ref{sec:kerdisc} may instead take the role of the covariance kernel of a Gaussian process with zero mean.
-Average-case error analysis \cite{Rit00a} and probabilistic numerics \cite{BriEtal18a} have pursued this approach.
+Average-case error analysis \cite{Rit00a} and probabilistic numerics \cite{BriEtal18a} pursue this approach.
 
@@
 
-When approximating the true $\mu$ with $\hat{\mu}_n$, stopping criterion determine how large $n$ should be so that the error $\lvert \mu - \hat{\mu}_n \rvert$ is below an error tolerance $\epsilon$ either certainly or with sufficiently low uncertainty $\alpha$.
+When approximating the true $\mu$ with $\hat{\mu}_n$, stopping criterion determine how large $n$ should be so that the error $\lvert \mu - \hat{\mu}_n \rvert$ is below an error tolerance $\varepsilon$ either certainly or with sufficiently low uncertainty $\alpha$.
 The error tolerance and probability threshold are problem dependent.
@@
 
-For simple Monte Carlo with IID points, the Central Limit Theorem (CLT) may be used to get an approximate confidence interval $(\mu^-,\mu^+)$ where $\mu^\pm = \hat{\mu} \pm Z^{1-\alpha/2} \sigma/\sqrt{n}$, $\sigma = \mathrm{Std}[f(X)]$ and $Z^{1-\alpha/2}$ is the $1-\alpha/2$ quantile of the standard Gaussian distribution.
+For simple Monte Carlo with IID nodes, the Central Limit Theorem (CLT) may be used to get an approximate confidence interval $[\mu^-,\mu^+]$ where $\mu^\pm = \hat{\mu} \pm Z^{1-\alpha/2} \sigma/\sqrt{n}$, $\sigma = \mathrm{Std}[f(\bsX)]$ and $Z^{1-\alpha/2}$ is the $1-\alpha/2$ quantile of the standard Gaussian distribution.
 If
@@
     \label{eq:sc_clt_iid_n}
-\end{equation} then $\mu \in (\mu^-,\mu^+)$ with probability $1-\alpha$.
-A simple IID stopping criterion uses an initial sample of $n_0$ points to approximate $\sigma^2 \approx (n_0-1)^{-1} \sum_{i=0}^{n_0-1}(f(\bsx_i)-\hat{\mu})^2$, and then plugs this into \eqref{eq:sc_clt_iid_n} to determine $n$.
+\end{equation} then $\mu \in [(]\mu^-,\mu^+]$ with probability $1-\alpha$.
+A simple IID stopping criterion uses an initial sample of $n_0$ nodes to approximate $\sigma^2 \approx S^2 := (n_0-1)^{-1} \sum_{i=0}^{n_0-1}(f(\bsx_i)-\hat{\mu})^2$, and then plugs this into \eqref{eq:sc_clt_iid_n} to determine $n$.
+(Sometimes the sample standard deviation, $S$, is multiplied by an inflation factor to account for the fact that it is an unbiased estimate and not an upper bound.)
+
 As the CLT holds only as $n$ goes to infinity, this is an approximate stopping criterion.
-For integrands $f$ with bounded kurtosis, a guaranteed version of this two-stage-sampling idea has been developed to account for the finite sample size using  Berry-Esseen inequalities\cite{JiaHic16a}.
-
-A straightforward qMC stopping criterion may also be derived from the CLT using a few randomizations of a LD point set.
-Let $\bsx_{ir}$ be the $i^\mathrm{th}$ point in the $r^\mathrm{th}$ randomization of a LD point set with $0 \leq i < n$ and $1 \leq r \leq R$.
-For example, we may use $R$ independent shifts of a lattice point set or independent digital shifts of a digital net.
-Crucially, the randomized point sets presented in Section \ref{sec:random} are also LD.
-Let $\hat{\mu}_r = n^{-1} \sum_{i=0}^{n-1} f(\bsx_{ir})$ be approximate means for each randomization and set $\hat{\mu} = R^{-1} \sum_{r=1}^R \hat{\mu}_r$ to be the aggregate approximation.
-In the spirit of CLT, a standard approximate $1-\alpha$ confidence interval $(\mu^-,\mu^+)$ has $\mu^\pm = \hat{\mu} \pm t_{R-1}^{1-\alpha/2} S/\sqrt{R}$ where $S^2 = (R-1)^{-1} \sum_{r=1}^R (\hat{\mu}_r - \hat{\mu})^2$ and $t_{R-1}^{1-\alpha/2}$ is the $1-\alpha/2$ quantile of the Student $t$ distribution with $R-1$ degrees of freedom.
+For integrands $f$ with bounded kurtosis, a guaranteed version of this two-stage-sampling idea has been developed to account for the finite sample size using  Berry-Esseen inequalities\cite{HicEtal14a}.
+
+A straightforward qMC stopping criterion may also be derived from the CLT using several randomizations of a LD node set.
+Let $\bsx_{ir}$ be the $i^\mathrm{th}$ node in the $r^\mathrm{th}$ randomization of a LD node set with $0 \leq i < n$ and $1 \leq r \leq R$.
+For example, we may use $R$ independent shifts of a lattice node set or independent digital shifts of a digital net.
+Crucially, the randomized node sets presented in Section \ref{sec:random} are also LD.
+\FJHNote{Tried to make the notation more consistent.}Let $\hat{\mu}^{(r)} = n^{-1} \sum_{i=0}^{n-1} f(\bsx_{ir})$ be sample means for each randomization and set $\hat{\mu}_{nR} = R^{-1} \sum_{r=1}^R \hat{\mu}^{(r)}$ to be the aggregate approximation.
+In the spirit of CLT, a standard approximate $1-\alpha$ confidence interval $(\mu^-,\mu^+)$ has $\mu^\pm = \hat{\mu}_{nR} \pm t_{R-1}^{1-\alpha/2} S/\sqrt{R}$ where $S^2 = (R-1)^{-1} \sum_{r=1}^R (\hat{\mu}^{(r)} - \hat{\mu}_{nR})^2$ and $t_{R-1}^{1-\alpha/2}$ is the $1-\alpha/2$ quantile of the Student $t$ distribution with $R-1$ degrees of freedom.
 Here we have used the Student $t$ distribution since $R$ is typically small, for example $R=15$.
@@
 
-One of the drawbacks of this qMC-CLT stopping criterion is that, at least intuitively, independent randomizations of LD nodes are less efficient than simply using more LD points from a single sequence.
-For instance, we would expect an approximation with $n=2^8$ and $R=2^3$ to perform worse than simply taking $n=2^{11}$ points from a single randomized LD sequence ($R=1$).
-One idea in this direction is to approximate the decay rate of Fourier coefficients of $f$.
-Bounds of absolute certainty have be derived for cones of functions with sufficient decay \cite{HicEtal17a}.
+One of the drawbacks of this qMC-CLT stopping criterion is that, at least intuitively, independent randomizations of LD nodes are less efficient than simply using more LD nodes from a single sequence.
+For instance, we would expect an approximation with $n=2^8$ and $R=2^3$ to perform worse than simply taking $n=2^{11}$ nodes from a single randomized LD sequence ($R=1$), but we need $R$ large enough to believe the CLT stopping criterion.
+
+One idea for constructing confidence intervals without replications is to approximate the decay rate of Fourier coefficients of $f$.
+Bounds of absolute certainty have be derived for cones of functions with reasonable decay of their Fourier coefficients \cite{HicEtal17a}.
 As in the qMC-CLT stopping criterion, one may iteratively double $n$ until the bounds are sufficiently tight.
 
-One such cone tracks the decay of approximate complex-exponential Fourier coefficients \cite{JimHic16a}.
-When paired with a randomly shifted lattice, these approximate coefficients may be computed at only $\calO(n \log n)$ cost using the fast Fourier transform.
-Another cone tracks the decay of approximate Walsh coefficients  \cite{HicJim16a}.
-When paired with a digitally shifted digital net, these approximate coefficients may be computed at only $\calO(n \log n)$ using the fast Walsh-Hadamard transform.
+For lattices one tracks the decay of the complex-exponential Fourier coefficients \cite{JimHic16a}.
+For digital sequences one tracks the decay of  Fourier-Walsh coefficients  \cite{HicJim16a}.
+In both cases, the definitions of the node sets paired with the appropriate Fourier basis allows these approximate Fourier coefficients to be computed from the data, $\{y_i = f(\bsx_i)\}_{i=0}^{n -1} $ at a cost of $\calO(n \log n)$ for $n$ a power of $2$ using fast  transforms.
 
@@
 Conditioned on observations $f(\bsx_i)$, a closed form expression exists for the $1-\alpha$ credible intervals for $\mu$.
-While also applicable when using IID points, using LD sequences paired with special kernels enables these credible intervals to be computed at only $\calO(n \log n)$ cost \cite{rathinavel2019fast}.
+While also applicable when using IID nodes, using LD sequences paired with special kernels enables these credible intervals to be computed at only $\calO(n \log n)$ cost \cite{Jag19a,JagHic22a,RatHic19a}.
 Again one may iteratively double the sample size $n$ until the credible intervals are sufficiently tight.
 
-Specifically, pairing shifted lattices with shift-invariant kernels as defined in \eqref{eq:shinvK} yields circulant Gram matrices for which computations can be done quickly using the fast Fourier transform \cite{RatHic19a}.
+For this Bayesian approach, one must find the eigenvalues and eigenvectors of the Gram matrix $\mathsf{K} : = \bigl(K(\bsx_i,\bsx_j) \bigr)_{i,j=0}^{n-1}$.
+In general this requires $\mathcal{O}(n^3)$ operations.
+However, pairing shifted lattices with shift-invariant kernels as defined in \eqref{eq:shinvK} yields circulant Gram matrices for which computations can be done quickly using the fast Fourier transform \cite{RatHic19a}.
 Similarly, pairing digitally shifted digital nets with digitally-shift-invariant kernels yields recursive-block Toeplitz Gram matrices \AGSNote{Different name?} for which computations can be done quickly using the fast Walsh-Hadamard transform \cite{JagHic22a}.
@@
 Such problems include computing the posterior mean in a Bayesian context or computing  sensitivity indices which  quantifying feature importance to variability in a simulation.
-This work also considered sharing samples across approximations and generalized error tolerances such as relative error tolerance.
+This work also considered sharing samples across approximations and generalized error criteria, such as a relative error tolerance.
 
@@
 One is that our original problem may not look like $\mu = f(\bsX)$ for $\bsX \sim \mathcal{U}[0,1)^d$, and thus it may need to be rewritten.
-The second is that qMC does best when most of the variation of $f$ is concentrated in the lower coordinates, and we should design $f$ so that this is so, if possible.
+The second is that qMC does best when most of the variation of $f$ is concentrated in the lower numbered coordinates, and, if possible, we should design $f$ so that this is so.
 
@@
 Often qMC problems may be written as $\mu = \bbE[g(\bsT)]$ where $\bsT$ is a $d$-dimensional random variable with density $\lambda$.
-Since qMC points have LD with the standard uniform distribution $\bsX \sim \calU[0,1]^d$, we perform an invertible variable transformation $\bsPsi: [0,1]^d \to \calM$ satisfying $\mathrm{range}(\bsT) \subseteq \calM$.
+Since qMC nodes have LD with the standard uniform distribution $\bsX \sim \calU[0,1]^d$, we perform an invertible variable transformation $\bsPsi: [0,1]^d \to \calM$ satisfying $\mathrm{range}(\bsT) \subseteq \calM$.
 This implies
 $\mu = \bbE[f(\bsX)]$ where  
-$$f(\bsX) = g(\bsPsi(\bsX)) \frac{\lambda (\bsPsi(\bsX))}{\nu(\bsPsi(\bsX))}$$
+$$
+f(\bsX) = g(\bsPsi(\bsX)) \frac{\lambda (\bsPsi(\bsX))}{\nu(\bsPsi(\bsX))},
+$$
 and $\nu$ is the density of $\bsPsi(\bsX)$.
@@
 For finance examples where the outcome depends more on the gross behavior of the underlying stochastic process, rather than the high frequency behavior, it makes sense to choose the transformation accordingly.
-For example, if $\bsT$ represents a Brownian motion, then a principal component analysis coupled with the inverse Gaussian distribution shows greater convergence gains \AGSNote{compared to the Cholesky decomposition or bridge construction?}.
+For example, if $\bsT$ represents a Brownian motion, then a principal component analysis coupled with the inverse Gaussian distribution shows greater convergence gains compared to the Cholesky decomposition.
 
@@
 
-Since the upper bound of the qMC error includes the variation of $f$ as a factor, one can often improve the efficiency of qMC by reformulating the problem to reduce this variation.
-Choosing a good variable transformation, which is equivalent to importance sampling, to minimize $\text{variation}(f,K)$ is one approach.
+Since the upper bound of the qMC error includes the variation of $f$ as a factor, one may improve the efficiency of qMC if one can reformulate the problem to reduce this variation.
+Choosing a good variable transformation, as mentioned in the previous subsection, is one approach.
 
@@
 An example of this in quantitative finance is where one may use the payoff of a European option, whose expectation is known, as a control variate for more exotic options.
-As noted in \cite{HicEtal03}, reducing the variation of $f$ via control variates may or may not bear resemblance to reducing the variance of $f$.
+As noted in \cite{HicEtal03}, reducing the variation of $f$ via control variates may or may not resemble reducing the variance of $f$.
 Control variates have been implemented in the stopping criteria based on Fourier coefficients described in the previous section \cite{HicEtal17a}.
@@
 The cost of computing the sample mean defined in \eqref{eq:sample_mean} is typically $\mathcal{O}(dn)$ as the dimension, $d$, and the sample size, $n$, tend to infinity.
-Some applications have $d$ that is either huge or infinite.
+Some applications have a $d$ that is either huge or infinite.
 On example is in quantitative finance, where $d$ denotes the number of times that the asset price is monitored over the life of a financial derivative times the number of assets involved.
 
-However, there is a way to reduce the cost of approximating $\mu$ by employing multilevel methods so that the computational cost may be comparable to solving a low dimensional approximation \cite{Gil15a}.
-We give an introduction to these methods here.
-
-Let $\{Y_l := f_l(\bsX_{l})\}_{l=1}^L$ be a sequence of approximations to $Y = f(\bsX)$, where $\bsX_l \sim \mathcal{U}[0,1)^{d_l}$.
-Here we assume that $d_1 < \cdots < d_L$, and that the cost of evaluating $f_l(\bsx_{l})$ is $\mathcal{O}(d_l)$.
+However, the $\mathcal{O}(dn)$ cost of approximating $\mu$ may often be reduced by employing multilevel methods.
+In fact, these methods may reduce the cost to the level comparable to solving a low dimensional approximation \cite{Gil15a}.
+We introduce these methods here.
+
+Let $\{Y_l := f_l(\bsX^{(l)})\}_{l=1}^L$ be a sequence of approximations to $Y = f(\bsX)$, where $\bsX^{(l)} = (X_1, \ldots, X_l) \sim \mathcal{U}[0,1)^{d_l}$.
+Here we assume that $d_1 < \cdots < d_L$, and that the cost of evaluating $f_l(\bsx^{(l)})$ is $\mathcal{O}(d_l)$, where $\bsx^{(l)} = (x_1, \ldots, x_l)$.
 Also, $d_L = d$, $Y_{d_L} = Y$, and $f_L = f$.
-Then
+Then the population mean may written as a telescoping sum:
 \begin{align*}
@@
     & = \int_{[0,1)^{d_1}} f_1(\bsx^{(1)}) \, \mathrm{d} \bsx^{(1)} +  \int_{[0,1)^{d_2}} \bigl[f_2(\bsx^{(2)}) - f_1(\bsx^{(1)})\bigr] \, \mathrm{d} \bsx^{(2)} + \cdots \\
-    & \qquad \qquad + \int_{[0,1)^{d_L}} \bigl[f_L(\bsx^{(L)}) - f_{L-1}(\bsx^{(L-1)}) \bigr] \, \mathrm{d} \bsx^{(L)},
+    & \qquad \qquad + \int_{[0,1)^{d_L}} \bigl[f_L(\bsx^{(L)}) - f_{L-1}(\bsx^{(L-1)}) \bigr] \, \mathrm{d} \bsx^{(L)}.
 \end{align*}
-where $\bsx^{(1)}  = (x_1, \dots, x_{d_1}), \ldots, \bsx^{(L)} = (x_1, \dots, x_{d_L})$.
 This series of integrals may be approximated by a series of sample means, namely
@@
 \begin{itemize}
-    \item The cost of a function value at each level, $l$, is proportional to the dimension, $d_l$, and that the dimensions increase exponentially, i.e.,  $d_l = d_1^{l}$ and $C_l \le \alpha d_1^l$. 
-    \item The $V_l$, which are roughly the variation, decrease exponentially, i.e., $V_l \le \beta r^{-l} $.
+    \item The cost of a function value at each level, $l$, is proportional to the dimension, $d_l$, and that the dimensions \emph{increase} exponentially, i.e.,  $d_l = d_1^{l}$ and $C_l \le \alpha d_1^l$.
+    \item The $V_l$, which are roughly the variation, \emph{decrease} exponentially, i.e., $V_l \le \beta r^{-l} $.
 \end{itemize}   
@@
 
-The ability to write one's problem to have the favorable case of high, or even infinite dimension with a computational cost comparable to the low dimensional case is somewhat an art.
-However, the multilevel method takes advantage of the principle pointed out in Section \ref{sec:coordwts}, namely that qMC thrives when problem have a low effective dimension.
+It is an art to write one's problem to have the favorable case of high or even infinite nominal dimension, $d$, be solvable with a computational cost comparable to the low dimensional case.
+However, the multilevel method takes advantage of the principle noted in Section \ref{sec:coordwts}, namely that qMC thrives when problem have a low effective dimension, even if the nominal dimension is large.
 
@@
 Excellent survey works for qMC and randomized qMC applied to finance include \cite{CafMor96,LEc09,Lem04a,Jae02,wangsloan05}.
+
 Another fruitful domain for qMC methods is in partial differential equations (PDEs), used for computing the expectations of nonlinear functionals of solutions of certain classes of PDEs.
@@
 
-Taking a broader view, Sorokin and Rao \cite{SorRao23a} illustrates that integration of qMC methods with Gaussian process models enhance the reliability of estimating failure probabilities in the machine learning domain.
+Taking a broader view, Sorokin and Rao \cite{SorRao23a} illustrate that integration of qMC methods with Gaussian process models enhance the reliability of estimating failure probabilities in the machine learning domain.
 Furthermore, novel approaches for training deep neural networks utilizing LD nodes have been proposed in recent works \cite{longo21,mishra21}.
+\FJHNote{Should include the water paper of Sorokin.  Also, these two sentences do not fit together.}
 
@@
 Many algorithms involving sampling with LD nodes have since been developed much further in \cite{kellprem12,shired08,Keller2013a,kolkel00,kell96,kell95}.
-(Remarkably, Monte Carlo methods have even been awarded with an Academy Award for visual effects in movie making.)
+(Remarkably, Monte Carlo methods have even been awarded with an Academy Award for visual effects in movie making \FJHNote{citation?}.)
 
@@
 
+\FJHNote{Our citations need to have Monte Carlo, Halton, etc. capitalized.}
 \bibliographystyle{spmpsci}

### 2025-01-27 HickernellMCQMC2024_QMCTutorial.tex
aleksei updates

@@
 	\includegraphics[width =  0.3\textwidth]{\figpath/64gridpts_d6.pdf}
-	\caption{\FJHNote{Is the font size large enough?}Although a two-dimensional grid covers the unit square rather well, a $d$-dimensional grid for modest $d$ does \emph{not} cover the unit cube well.
+	\caption{Although a two-dimensional grid covers the unit square rather well, a $d$-dimensional grid for modest $d$ does \emph{not} cover the unit cube well.
 For example, one can only see four distinct nodes in a two dimensional projection of a six-dimensional grid with $64$ nodes.
@@
 	\caption{IID nodes cover the unit cube better than grid nodes, although one does observe clusters and gaps.
-In any  projection there is  a similar looking distribution of all $64$ nodes \label{fig:iid}}
+In any  projection there is  a similar looking distribution of all $64$ nodes.
+\label{fig:iid}}
 \end{figure}
@@
 \begin{equation}\label{eq:hammersley}
-    \bsx_i^{\text{Ham}} = \left( \frac{i}{n}, \phi_{b_1}(i), \ldots, \phi_{b_d}(i) \right) \quad i = 0, 1, \ldots, .
+    \bsx_i^{\text{Ham}} = \left( \frac{i}{n}, \phi_{b_1}(i), \ldots, \phi_{b_d}(i) \right) \quad i = 0, \ldots, n-1.
 \end{equation}
@@
 
-Most quasi-Monte Carlo constructions are studied based on their asymptotic behavior as \(n \to \infty\), with numerous results on asymptotic rates \cite{NIED1978QMCBook,NovWoz10a}.
+Most quasi-Monte Carlo constructions are studied based on their asymptotic behavior as \(n \to \infty\), with numerous results on asymptotic rates \cite{Nie72,Nie92,NovWoz10a}.
 However, few methods address finding optimal node configurations for specific \(n\) and \(d\) which can be critical in practical applications where each function evaluation is quite costly.
@@
 	\includegraphics[width = \textwidth]{\figpath/n_8_n_scr_3_scrambled_sobol_sequence_d2.pdf}
-	\caption{\FJHNote{Something went awry here.}Three linear scrambles of a digital $(0,3,2)$-net with nodes denoted by three different colors.
+	\caption{Three linear scrambles of a digital $(0,3,2)$-net with nodes denoted by three different colors.
 There is one node in each smaller box of the form \eqref{eq:element_box} for $\lVert \bsk \rVert = 3$, for each digitally shifted net.
@@
 
-\subsection{Other Randomizations for Halton}
+\subsection{Other Randomizations for Halton \FJHNote{Nathan will tackle this subsection and next}}
 
@@
 
-Instead of explicitly searching for good generating vectors or matrices, one can instead sample a few random generating vectors or generating matrices and then prune out bad choices \AGSNote{citation?}.
+Instead of explicitly searching for good generating vectors or matrices, one can instead sample a few random generating vectors or generating matrices and then prune out bad choices \cite{PanOwe23a}.
 Random generating vectors for lattices take $\bsh \overset{\text{IID}}{\sim} \calU\{1,\dots,n-1\}^d$.
@@
 Crucially, the randomized node sets presented in Section \ref{sec:random} are also LD.
-\FJHNote{Tried to make the notation more consistent.}Let $\hat{\mu}^{(r)} = n^{-1} \sum_{i=0}^{n-1} f(\bsx_{ir})$ be sample means for each randomization and set $\hat{\mu}_{nR} = R^{-1} \sum_{r=1}^R \hat{\mu}^{(r)}$ to be the aggregate approximation.
+Let $\hat{\mu}^{(r)} = n^{-1} \sum_{i=0}^{n-1} f(\bsx_{ir})$ be sample means for each randomization and set $\hat{\mu}_{nR} = R^{-1} \sum_{r=1}^R \hat{\mu}^{(r)}$ to be the aggregate approximation.
 In the spirit of CLT, a standard approximate $1-\alpha$ confidence interval $(\mu^-,\mu^+)$ has $\mu^\pm = \hat{\mu}_{nR} \pm t_{R-1}^{1-\alpha/2} S/\sqrt{R}$ where $S^2 = (R-1)^{-1} \sum_{r=1}^R (\hat{\mu}^{(r)} - \hat{\mu}_{nR})^2$ and $t_{R-1}^{1-\alpha/2}$ is the $1-\alpha/2$ quantile of the Student $t$ distribution with $R-1$ degrees of freedom.
@@
 However, pairing shifted lattices with shift-invariant kernels as defined in \eqref{eq:shinvK} yields circulant Gram matrices for which computations can be done quickly using the fast Fourier transform \cite{RatHic19a}.
-Similarly, pairing digitally shifted digital nets with digitally-shift-invariant kernels yields recursive-block Toeplitz Gram matrices \AGSNote{Different name?} for which computations can be done quickly using the fast Walsh-Hadamard transform \cite{JagHic22a}.
+Similarly, pairing digitally shifted digital nets with digitally-shift-invariant kernels yields recursive-block Toeplitz Gram matrices for which computations can be done quickly using the fast Walsh-Hadamard transform \cite{JagHic22a}.
 
@@
 
-The above formulation is equivalent to importance sampling \AGSNote{reference} by a distribution with density $\nu$.
+The above formulation is equivalent to importance sampling by a distribution with density $\nu$.
 Optimally, one would choose $\nu \propto g\lambda$ so that $f$ is constant, but this is typically infeasible in practice due to our limited knowledge of $g$.
@@
 \end{multline}
-If $C_1$ is the computational cost of $f^{(1)}(\bsx^{(1)})$ and $C_l$ is the computational cost of \AGSNote{$f^{(l)}(\bsx^{(1)})$ of $f_L(\bsx_i^{(l)}) - f_{l-1}(\bsx_i^{(l-1)})$} for $l = 2, \ldots, L$, then the total computational cost of
+If $C_1$ is the computational cost of $f^{(1)}(\bsx^{(1)})$ and $C_l$ is the computational cost of $f^{(l)}(\bsx^{(1)}) - f_{l-1}(\bsx_i^{(l-1)})$ for $l = 2, \ldots, L$, then the total computational cost of
 \eqref{eq:MLmuhat} is 
@@
 
+QMC methods have also found recent applications in scientific machine learning of PDE solutions.
+It has been shown that LD nodes are more efficient than IID nodes for training neural networks modeling PDE solution \cite{longo21,mishra21}.
+LD nodes have also been used to more efficiently fit Gaussian process regression models for operator learning the solution process of PDEs with uncertain coefficients  \cite{sorokin2024computationally}.
+There the computational speedup follows from pairing LD nodes with special kernels to induce nicely structured gram matrices, as used in the fast Bayesian cubature stopping criteria discussed in Section \ref{sec:stop}.
+
 Besides computing expectations, qMC methods have found many applications in other areas of the computational sciences.
@@
 
-Taking a broader view, Sorokin and Rao \cite{SorRao23a} illustrate that integration of qMC methods with Gaussian process models enhance the reliability of estimating failure probabilities in the machine learning domain.
-Furthermore, novel approaches for training deep neural networks utilizing LD nodes have been proposed in recent works \cite{longo21,mishra21}.
-\FJHNote{Should include the water paper of Sorokin.  Also, these two sentences do not fit together.}
-
 Keller \cite{kell95} was the first to adopt qMC for rendering photorealistic images in computer graphics where the problem of simulating light transport can be mathematically represented as an integral over all possible light paths from source to camera.
@@
 This means that qMC outperforms alternative methods such as simple Monte Carlo and sampling on grids.
-Using the multilevel qMC, the cost to reach the error tolerance is often essentially the same for $d=1$ and $d$ unbounded.
+QMC is particularly powerful when the effective dimension of the problem is small, even though the nominal dimension, $d$, may be large.
 Software libraries are available to facilitate qMC computations.
@@
 \bibliographystyle{spmpsci}
-\bibliography{FJH23,FJHown23,main}
+\bibliography{FJH25,FJHown23,main}
 

### 2025-01-28 HickernellMCQMC2024_QMCTutorial.tex
Updates from Overleaf

@@
 \begin{abstract}
-Many problems in  quantitative finance, uncertainty quantification, and other disiplines are answered by computing the population mean, $\mu := \mathbb{E}(Y)$, where instances of $Y:=f(\boldsymbol{X})$ may be generated by numerical simulation and $\bsX$ has a simple probability  distribution.
+Many questions in  quantitative finance, uncertainty quantification, and other disciplines are answered by computing the population mean, $\mu := \mathbb{E}(Y)$, where instances of $Y:=f(\boldsymbol{X})$ may be generated by numerical simulation and $\bsX$ has a simple probability  distribution.
 The population mean can be approximated by the sample mean, $\hat{\mu}_n := n^{-1} \sum_{i=0}^{n-1} f(\bsx_i)$ for a well chosen sequence of nodes, $\{\bsx_0, \bsx_1, \ldots\}$ and a sufficiently large sample size, $n$.
@@
 \begin{equation} \label{eq:sample_mean_Keister}
-	\hat{\mu}_n = \frac 1n \sum_{i=0}^{n-1} f (\bsx) = \frac 1n \sum_{i=0}^{n-1} \pi^{d/2} \cos\Bigl (\bigl \lVert \bigl( \Phi^{-1}(x_{i1}), \ldots, \Phi^{-1}(x_{id})\bigr) \bigr\rVert/\sqrt{2}  \Bigr).
+	\hat{\mu}_n = \frac 1n \sum_{i=0}^{n-1} f (\bsx_i) = \frac 1n \sum_{i=0}^{n-1} \pi^{d/2} \cos\Bigl (\bigl \lVert \bigl( \Phi^{-1}(x_{i1}), \ldots, \Phi^{-1}(x_{id})\bigr) \bigr\rVert/\sqrt{2}  \Bigr).
 \end{equation}
@@
 
-The numerical computations used to generate the figures in this chapter may be reproduced using the Jupyter notebook \url{??}.
+The numerical computations used to generate the figures in this chapter may be reproduced using the Jupyter notebook in \cite{MCQMC2024TutorialNotebook}.
 The Python library \texttt{qmcpy} \cite{QMCPy2020a} is used to generate the LD node sequences.
@@
 Figure \ref{fig:latticeconstruct} illustrates this construction for $n = 16$ and $\bsh = (1,11)$.
-Note that the set $\{\bsx_0, \ldots, \bsx_{n-1}\}$ defined in \eqref{eq:latticepts} is closed under addition modulo $\bsone$ so that it forms a group.
+Note that the set $\{\bsx^{\text{lat}}_0, \ldots, \bsx^{\text{lat}}_{n-1}\}$ defined in \eqref{eq:latticepts} is closed under addition modulo $\bsone$ so that it forms a group.
 
@@
 
-Most quasi-Monte Carlo constructions are studied based on their asymptotic behavior as \(n \to \infty\), with numerous results on asymptotic rates \cite{Nie72,Nie92,NovWoz10a}.
+Most quasi-Monte Carlo constructions are studied based on their asymptotic behavior as \(n \to \infty\), with numerous results on asymptotic rates \cite{Nie92,NovWoz10a}.
 However, few methods address finding optimal node configurations for specific \(n\) and \(d\) which can be critical in practical applications where each function evaluation is quite costly.
@@
 
-\FJHNote{Should this paragraph go here or in the next section?}
-In the previous section we discussed how the Halton sequence has the benefit of not having a preferred number of nodes for its construction.
-It does however possess the drawback of exhibiting poor uniformity in the lower dimensional projections when $d$ is moderate to high due to number theoretic correlation between the coordinates \cite{kirklem24}.
-One popular strategy to circumvent this poor behavior, dating back to Faure \cite{fau78} and Braaten and Weller \cite{brawell79}, is to randomly or deterministically permute the digits of the base $b_j$ expansions $j = 1,2\ldots,d$ to disrupt correlations between the components.
-To do so, we first define the generalized van der Corput sequence by obtaining a sequence $\Sigma = (\sigma_r)_{r\geq0}$ of permutations of $\{0,\ldots,b-1\}$ where the $i^{th}$ term of the one-dimensional sequence is defined as
-\begin{multline}\label{eq:gen_vdc}
-    \phi_b^{\Sigma}(i_0 + i_1b + i_2 b^2 + \cdots ) := \sigma_0(i_0) b^{-1} + \sigma_1(i_1) b^{-2} + \sigma_2(i_2) b^{-3} + \cdots \in [0,1)
-	\\
-	 \text{where } i_0, \sigma_0(i_0), i_1, \sigma_1(i_1), \ldots \in \{0,\ldots b-1\}.
-\end{multline}
-The generalized Halton sequence is subsequently defined by choosing $d$ sequences of permutations $\Sigma_j = (\sigma_{j,r})_{r\geq0}, j = 1, 2,\ldots,d$ and setting
-$$\bsx_i^{\text{G-Hal}} = \left( \phi^{\Sigma_1}_{b_1}(i), \ldots, \phi^{\Sigma_d}_{b_d}(i)\right), \quad i= 0,1,\ldots,$$
-There exist many proposed algorithms to optimize the permutations defining the generalized sequence, most of which involve so-called linear digit scramblings and involve directly optimizing a discrepancy FOM.
-We point the reader towards \cite{MasChiWar05,tuffin98,Atan04,vancools06} and the excellent survey of the work in this direction (until 2008) by Faure and Lemieux \cite{faulem09}.
-
-We have outlined various efforts by qMC researchers to optimize specific classes of LD constructions, including lattices, digital sequences, and Halton sequences.
+We have outlined various efforts by qMC researchers to optimize specific classes of LD constructions, including lattices and digital sequences.
+Significant effort has even been taken to optimize the uniformity of so-called generalized Halton sequences due to the poor lower dimensional projections exhibited by even a moderate dimensional Halton sequence \cite{kirklem24}.
+These methods are discussed briefly in Section \ref{sec:randHalton}.
 There also exist several efforts to generate novel LD constructions which do not adhere to any underlying number theoretic structure, but none-the-less, possess a high level of uniformity.
@@
 
-\subsection{Other Randomizations for Halton \FJHNote{Nathan will tackle this subsection and next}}
-
-As Halton nodes can be viewed as digital nets with different bases for each dimension, the randomization techniques used for digital nets can be applied to Halton node sets\NKNote{we should clarify here whether we mean truncated Halton, or Hammersley if it is introduced above} Random digital shifts for Halton set $x_{ij\ell} \mapsto (x_{ij\ell} + \Delta_{j\ell}) \mod b_j$ where the shift $\Delta_{j\ell} \in \{0,\dots,b_j-1\}$.
+\subsection{Other Randomizations for Halton}\label{sec:randHalton}
+
+%As Halton nodes can be viewed as digital nets with different bases for each dimension, the randomization techniques used for digital nets can be applied to Halton node sets.
+Random digital shifts for Halton set $x_{ij\ell} \mapsto (x_{ij\ell} + \Delta_{j\ell}) \mod b_j$ where the shift $\Delta_{j\ell} \in \{0,\dots,b_j-1\}$.
 More generally, a permutation scramble for digital nets sets $x_{ij\ell} \mapsto \Pi_{j\ell}(x_{ij\ell})$ where $\Pi_{j\ell}$ are independent random permutations of $\{0,\dots,b_j-1\}$.
 \AGSNote{Ignore this permutation definition as it is the same as \eqref{eq:gen_vdc}?} Randomization of Halton relating to permutations of digits was discussed briefly in Section \ref{sec:optLD}.
+
+As Halton nodes can be viewed as digital nets with different bases for each dimension, the randomization techniques used for digital nets can be applied to Halton node sets.
 Random nested uniform scrambling and linear matrix scrambling have also been considered for Halton node sets \cite{owen2024gain}.
 
+We detail one possible randomization of the Halton sequence which is equivalent to applying a random digital shift \eqref{eq:sh_digital} in the respective base of each dimension of the Halton sequence.
+We first define the generalized van der Corput sequence in base $b$ by obtaining a sequence $\Sigma = (\sigma_r)_{r\geq0}$ of permutations of $\{0,\ldots,b-1\}$ where the $i^{th}$ term of the one-dimensional sequence is defined as
+\begin{multline}\label{eq:gen_vdc}
+    \phi_b^{\Sigma}(i_0 + i_1b + i_2 b^2 + \cdots ) := \sigma_0(i_0) b^{-1} + \sigma_1(i_1) b^{-2} + \sigma_2(i_2) b^{-3} + \cdots \in [0,1)
+	\\
+	 \text{where } i_0, \sigma_0(i_0), i_1, \sigma_1(i_1), \ldots \in \{0,\ldots b-1\}.
+\end{multline}
+The generalized Halton sequence is subsequently defined by choosing $d$ sequences of permutations $\Sigma_j = (\sigma_{j,r})_{r\geq0}$ of $\{0,\ldots,b_j-1\}$ for $j = 1, 2,\ldots,d$ and setting
+$$\bsx_i^{\text{G-Hal}} = \left( \phi^{\Sigma_1}_{b_1}(i), \ldots, \phi^{\Sigma_d}_{b_d}(i)\right), \quad i= 0,1,\ldots,$$ A randomized qMC estimate $\hat{\mu}_n$ is then an unbiased estimator of $\mu$ if the permutations $\Sigma_j$ are sampled independently and uniformly, i.e $\mu$ if $\Sigma_j \overset{\text{IID}}{\sim} \calU\{0,\dots,b_j-1\}^d$ since this ensures $\bsx_i^{\text{G-Hal}} \sim \mathcal{U}[0,1)^d$.
+
+We note in passing that many works \cite{MasChiWar05,tuffin98,Atan04,vancools06} have sought to optimize the permutations $\Sigma_j$ deterministically to enhance uniformity.
+For a comprehensive survey of the Halton sequence (up to 2008), we refer readers to the excellent work by Faure and Lemieux \cite{faulem09}.
+
 \subsection{Randomized Generators}
 
-Instead of explicitly searching for good generating vectors or matrices, one can instead sample a few random generating vectors or generating matrices and then prune out bad choices \cite{PanOwe23a}.
-Random generating vectors for lattices take $\bsh \overset{\text{IID}}{\sim} \calU\{1,\dots,n-1\}^d$.
-Random generating matrices for base $2$ digital nets---as described in \eqref{eq:digitalB}---set $x^{\text{dig}}_{ij\ell} \overset{\text{IID}}{\sim} \calU\{0,1\}$ for $i<\ell$ while enforcing $x_{iji}=1$ and $x_{ij\ell}=0$ for $i>\ell$.
+Instead of explicitly searching for good generating vectors or matrices, one can instead sample a few random generating vectors or generating matrices and then prune out bad choices.
+One recent method has explored selecting the median of $r$ randomized qMC estimates, rather than the mean \cite{PanOwe23a}.
+Random generating vectors for lattices take $\bsh \overset{\text{IID}}{\sim} \calU\{1,\dots,n-1\}^d$ while random generating matrices for base $2$ digital nets---as described in \eqref{eq:digitalB}---set $x^{\text{dig}}_{ij\ell} \overset{\text{IID}}{\sim} \calU\{0,1\}$ for $i<\ell$ while enforcing $x_{iji}=1$ and $x_{ij\ell}=0$ for $i>\ell$.
 
@@
 Many algorithms involving sampling with LD nodes have since been developed much further in \cite{kellprem12,shired08,Keller2013a,kolkel00,kell96,kell95}.
-(Remarkably, Monte Carlo methods have even been awarded with an Academy Award for visual effects in movie making \FJHNote{citation?}.)
+(Remarkably, Monte Carlo methods have even been awarded with an Academy Award for visual effects in movie making \cite{veachthesis,oscars}.)
 
@@
 
-\FJHNote{Our citations need to have Monte Carlo, Halton, etc. capitalized.}
 \bibliographystyle{spmpsci}

### 2025-06-04 HickernellMCQMC2024_QMCTutorial rev 1.tex
Creating revision 1 copy

--- 2025-01-28 HickernellMCQMC2024_QMCTutorial.tex
@@
 \end{equation*}
-For $k_1=2$, $k_2 = 0$, and $k_3 = 1$ condition \eqref{eq:tcond} is not satisfied, and so $t$ cannot be $0$ for the node set  \eqref{eq:smallnet}.
+For $\bsk = (2, 0, 1)$ and $\lVert \bsk \rVert_1 = 3$, condition \eqref{eq:tcond} is not satisfied, and so $t$ cannot be $0$ for the node set  \eqref{eq:smallnet}.
 However, condition \eqref{eq:tcond} is satisfied for all $\bsk$ with $\lVert \bsk \rVert_1 = 2$, again confirming that we have a $(1,3,3)$-net.
@@
 %While good generating vectors are known for \(d \leq 2\) \cite{DicEtal22a,SloJoe94}, the same does not hold for \(d \geq 3\),
-one often resorts to finding good $\boldsymbol{h}$ via computer search to minimize a chosen FOM.
-An exhaustive search is infeasible due to the exponential growth of the search space with \(d\).
+one often resorts to finding good $\boldsymbol{h}$ via computer search to minimize a chosen FOM. An exhaustive search is infeasible due to the exponential growth of the search space with \(d\).
+
 Thus, the component-by-component (CBC) construction, introduced by Korobov \cite{kor63} and later revisited by Sloan and collaborators \cite{Slo02}, reduces the search space to size \(dn\) by constructing \(\boldsymbol{h}\) one component at a time while keeping the previous components fixed.
@@
 One of the first works to globally optimize LD nodes, albeit a heuristic approach, was by Winker and Fang \cite{WinFan97b} employing the threshold accepting optimization algorithm to directly optimize the uniformity of the LD nodes with respect to a discrepancy FOM.
+
 More recently, Rusch, Kirk, Bronstein, Lemieux and Rus \cite{ruschkirk24} developed the first machine-learning assisted optimization tool for LD node optimization in the Message-Passing Monte Carlo (MPMC) method.
@@
 The optimization is gradient-based and is guided by the direct minimization of a discrepancy-based FOM.
+
 We also highlight the method proposed by Hinrichs and Oettershagen \cite{hinoet16} generating optimal LD nodes for a specific class of functions in two-dimensions.
 To address the non-convexity challenge, the authors decompose the global optimization of a specific FOM into a large number of smaller, convex sub-problems which are solved individually.
+
 There are several other recent attempts to construct optimal LD nodes, many of which are well summarized in the PhD thesis of Cl\'{e}ment \cite{clethesis24}.
@@
 \begin{multline} \label{eq:sh_digital}
-	\bsx_i^{\bsDelta\text{-dig}} := \bsx_i^{\text{dig}} \oplus \bsDelta = i_0 \bsx^{\text{dig}}_1 \oplus i_1 \bsx^{\text{dig}}_2 \oplus i_2 \bsx^{\text{dig}}_4 \oplus \cdots \oplus \bsDelta \in [0,1)^d \\
+	\bsx_i^{\bsDelta\text{-dig}} := \bsx_i^{\text{dig}} \oplus \bsDelta = i_0 \bsx^{\text{dig}}_1 \oplus i_1 \bsx^{\text{dig}}_2 \oplus i_2 \bsx^{\text{dig}}_4 \oplus \cdots \oplus \bsDelta \\
     \text{for }
@@
 	\includegraphics[width = \textwidth]{\figpath/n_8_n_sh_3_shifted_sobol_sequence_d2.pdf}
-	\caption{Three digital shifts of a digital $(0,3,2)$-net with nodes denoted by three different colors.
+	\caption{Three digital shifts of a digital $(0,3,2)$-net with nodes denoted by three different colors/shapes.
 There is one node in each smaller box of the form \eqref{eq:element_box} for $\lVert \bsk \rVert = 3$, for each digitally shifted net.
@@
 A random digital shift, $\bsDelta$ is also added, where $\Delta_{\ell j}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ dimension of the shift.
+Digital scrambling of a $(t,m,d)$-net does not alter its $t$-value.
 
@@
 	\includegraphics[width = \textwidth]{\figpath/n_8_n_scr_3_scrambled_sobol_sequence_d2.pdf}
-	\caption{Three linear scrambles of a digital $(0,3,2)$-net with nodes denoted by three different colors.
+	\caption{Three linear scrambles of a digital $(0,3,2)$-net with nodes denoted by three different colors/shapes.
 There is one node in each smaller box of the form \eqref{eq:element_box} for $\lVert \bsk \rVert = 3$, for each digitally shifted net.
@@
 
-\subsection{Other Randomizations for Halton}\label{sec:randHalton}
+\subsection{Randomizations for Halton}\label{sec:randHalton}
 

### 2025-06-10 response_letter.tex
Updates from Overleaf

\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{amsmath}
\usepackage{lipsum}

% ---------- Custom Commands and Environments ----------

% Define Response environment

\newenvironment{response}{%
  \vspace{1em}
  \begin{quote}\itshape
}{%
  \end{quote}
  \noindent\textbf{Reply:}~
}

% Optional: define formatting for section titles
\titleformat{\section}{\normalfont\Large\bfseries}{\thesection.}{1em}{}
\titleformat{\subsection}{\normalfont\large\bfseries}{\thesubsection.}{1em}{}

% ---------- Document Starts ----------
\title{Author Response for “Quasi-Monte Carlo Methods: What, Why, and How?”}
\author{Fred J. Hickernell, Nathan Kirk and Aleksei G. Sorokin}
\date{}

\begin{document}

\maketitle

We thank the reviewers for their careful reading of the manuscript and for their thoughtful suggestions, which have helped improve both the content and presentation of the work. All grammatical errors and typographical mistakes have been corrected. Below, we provide responses to all comments; substantial changes are highlighted in blue in the revised manuscript

\section*{Response to Reviewer 1}

\begin{response}{1.}
It is unavoidable that surveys of this type are not comprehensive and will choose to focus
on some areas and ignore others. It would be useful to mention something along those lines
somewhere in the introduction.
\end{response}
Good point. Sentence added at the end of the introduction.

\begin{response}{2.}
...This is an unusual way to define a digital
sequence....
\end{response}
...

\begin{response}{3.}
...the definition of a digital sequence also requires segments between powers of $2$ to be nets with parameter (at most) $t$. This also brings a minor point: especially when defining a $(t, d)$-sequence, one does not usually assume that the t parameter applies as the exact $t$ for all segments...
\end{response}
....

\begin{response}{4.}
Halton sequences do exhibit specific equidistribution properties when $n$ is a product of powers of the $d$ bases used. So it is not completely accurate to say that Halton sequences do not have preferred sample sizes
\end{response}
Good point. We have fixed this inaccuracy in Section 3.3.

\begin{response}{5.}
    It would be important to cite other works than [49] when discussing the Halton sequences. For example [18] which includes many other important references
\end{response}
Good point. Lemieux and Faure has been added.

\begin{response}{6.}
“One example” rather than “On example”. Also in such cases $d$ may be larger than the number of times the asset is monitored, e.g., when dealing with stochastic volatility models or multiple assets. So this should say something like “where $d$ could denote...”
\end{response}
Good point. Fixed and added.

\subsection*{Minor Comments}

The authors agree with all minor comments  flagged by the reviewer and thus have been rectified.

\newpage

\section*{Response to Reviewer 2}

\begin{response}{1.}
    Add a simple but practical application, like option pricing under Black-Scholes model, as an illustrative example. This reviewer finds the illustrative example in Section 2 lacks practical relevance
\end{response}
....

\begin{response}{2.}
    Section 7 mentions variance reduction techniques such as importance sampling (Section 7.1) and control variate (Section 7.2). This reviewer suggests elaborating how variance reduction techniques, which were originally proposed for Monte Carlo methods, can be applied to qMC methods and whether these techniques and effective and in what ways they are effective.
\end{response}
.....

\begin{response}{3.}
    Section 7.3 seems to be a bit disconnected with the rest of the paper, with some undefined notations like $\boldsymbol{n}$ and $\boldsymbol{d}$. Are these the vector versions of n and d? This reviewer suggests adding more details and discussions in this section.
\end{response}
....

\subsubsection*{Minor Issues}

\begin{response}{1.}
    One line below Equation (6) on page 3: Is there a specific reason to show the value of $\mu$ in such a high accuracy? If there is a reason, maybe it should be clearly explained
\end{response}
There is no specific reason. We have kept the same accuracy.

\begin{response}{2.}
    The authors mention the van der Corput sequence a few times (e.g., 4 lines above Equation (9), first line on page 7, and other occurrences). For readers who don’t know what this sequence is, such as this reviewer, it would be helpful to provide some explanations and references
\end{response}
The authors have slightly rewritten the paragraph above equation (9) to more clearly define the van der Corput sequence the first time it is mentioned.

\begin{response}{3.}
    Is Equation (14) supposed to be an equation? It looks very strange
\end{response}
.....

\begin{response}{4.}
    The last line on page 9 is just a reference...
\end{response}
With some rewriting, this is no longer an issue.

\begin{response}{5.}
    First line in Section 3.4: The acronym qMC is defined and used many times, why not use it here?
\end{response}
    Thank you. Acronym now used here.

\begin{response}{6.}
    Figure 9 on page 13. The $x_0 = \Delta$ equation in the figures overlaps with some markers, which makes it hard to see. Please reposition it.
\end{response}
.....

\begin{response}{7.}
    The legends in Figure 12 and Figure 13 are too small
\end{response}
....

\begin{response}{8.}
    First line after Equation (43): “[(]” seems to be a typo
\end{response}
Thank you. Fixed.

\begin{response}{9.}
    Third line in Section 7: On example $\rightarrow$ One example
\end{response}
Thank you. Fixed.

\end{document}

### 2025-06-10 HickernellMCQMC2024_QMCTutorial_rev_1.tex
Updates from Overleaf

--- 2025-06-04 HickernellMCQMC2024_QMCTutorial rev 1.tex
@@
 %
-\author{Fred J. Hickernell\inst{1} \and Nathan Kirk\inst{2} \and Aleksei G. Sorokin\inst{3}}
+\author{Fred J. Hickernell\inst{1,2} \and Nathan Kirk\inst{1,3} \and Aleksei G. Sorokin\inst{1,4}}
 %
@@
 Department of Applied Mathematics, Illinois Institute of Technology, \\ Chicago, IL, 60616, USA\\
+\and
 \email{hickernell@iit.edu}, \
-\texttt{http://www.iit.edu/~hickernell}
+%\texttt{http://www.iit.edu/\textasciitilde{}hickernell}
+\url{http://www.iit.edu/~hickernell}
 \and
@@
 
-Quasi-Monte Carlo methods replace independent and identically distributed  sequences of random vector nodes, $\{\bsx_i \}_{i = 0}^{\infty}$, by low discrepancy sequences.
+Quasi-Monte Carlo methods replace independent and identically distributed  sequences of random vector nodes, $\{\bsx_i \}_{i = 0}^{\infty}$, by low-discrepancy sequences.
 This accelerates the convergence of $\hat{\mu}_n$ to $\mu$ as $n \to \infty$.
 
-This tutorial describes  low discrepancy sequences  and their quality measures.
+This tutorial describes  low-discrepancy sequences  and their quality measures.
 We demonstrate the performance gains possible with quasi-Monte Carlo methods.
@@
 
-\keywords{low discrepancy, randomization, sample mean, simulation}
+\keywords{low-discrepancy, randomization, sample mean, simulation}
 \end{abstract}
@@
 Quasi-Monte Carlo (qMC)  chooses the nodes differently, so that their empirical distribution approximates well the true probability distribution of $\bsX$.
-The difference between  the true and empirical distributions is called a \emph{discrepancy}, and the node sequences used in qMC are called low discrepancy (LD) sequences.
+The difference between  the true and empirical distributions is called a \emph{discrepancy}, and the node sequences used in qMC are called low-discrepancy (LD) sequences.
 
@@
 The next section illustrates by example the advantages of qMC.
-This is followed by a description of how deterministic LD sequences are constructed (Section \ref{sec:construct}) and randomizations that preserve their low discrepancy (Section \ref{sec:random}).
+This is followed by a description of how deterministic LD sequences are constructed (Section \ref{sec:construct}) and randomizations that preserve their low-discrepancy (Section \ref{sec:random}).
 We describe various measures of discrepancy (Section \ref{sec:discrepancy}).
@@
 We briefly discuss applications of qMC beyond computing the mean (Section \ref{sec:beyond}), and finish with a short conclusion (Section \ref{sec:conclusion}).
+\NKNote{Inevitably, this tutorial is not comprehensive of the full breadth of over a century of development in qMC and uniform distribution theory, dating back to Hermann Weyl's seminal 1916 paper \cite{Wey16}.
+Instead, our goal is to provide a solid foundation in the core concepts along with practical guidance, and to spark further interest and study in the field.}
 
@@
 Thus, we perform a variable transformation $\bst = \bigl (\Phi^{-1}(x_1), \ldots, \Phi^{-1}(x_d) \bigr )/\sqrt{2}$,
-where is $\Phi$ is the cumulative distribution function of the standard Gaussian random variable.
+where $\Phi$ is the cumulative distribution function of the standard Gaussian random variable.
 This reimagines the integral $\mu$ as the expectation of a function, $f$, of a standard uniform random variable:
@@
 
-	\item \emph{Grids.} Although they may be attractive for low dimensional problems, e.g., $d = 1$, $2$, or $3$, grids do poorly for this modest dimension, $d=6$.
+	\item \emph{Grids.} Although they may be attractive for low-dimensional problems, e.g., $d = 1$, $2$, or $3$, grids do poorly for this modest dimension, $d=6$.
 Figure \ref{fig:grid} compares a $64$ point grid with $d = 2$ and $6$.
@@
 
-	Whereas grid points collapse on top of one another when viewed in low dimensional projections (Figure \ref{fig:grid}), all IID nodes may be seen when viewed in any lower dimensional projection, as seen in Figure \ref{fig:iid}.
+	Whereas grid points collapse on top of one another when viewed in low-dimensional projections (Figure \ref{fig:grid}), all IID nodes may be seen when viewed in any lower dimensional projection, as seen in Figure \ref{fig:iid}.
 The disadvantage of IID nodes is that they form clusters and leave gaps.
@@
 
-An extensible lattice sequence in one dimension is defined by the van der Corput sequence in base $b$, $\{\phi_b(0), \phi_b(1), \ldots\}$.
-This involves the so-called radical inverse function and essentially reflects the digits of the integer $i$ in base $b$ about the decimal point.
+An extensible lattice sequence in one dimension is defined by the van der Corput sequence in base $b$, $\{\phi_b(0), \phi_b(1), \ldots\}$, where $\phi_b$ is the radical inverse function and essentially reflects the digits of the integer $i$ in base $b$ about the decimal point.
 For example, $\phi_2(6) = \phi_2(110_{2}) = {}_20.011 = 3/8$.
@@
 
-\subsection{Halton sequences} \label{sec:Halton}  While lattice sequences and digital sequences have preferred sample sizes, $n$, Halton sequences have no preferred sample size.
+\subsection{Halton sequences} \label{sec:Halton} 
+%While lattice sequences and digital sequences have preferred sample sizes, $n$, Halton sequences have no preferred sample size.  
 The Halton sequence is defined in terms of the van der Corput sequences for different bases:
@@
 \end{equation}
-where $b_1, \ldots, b_d$ is a choice of $d$ distinct prime numbers.
+where $b_1, \ldots, b_d$ is a choice of $d$ distinct co-prime numbers.
 Often they are chosen as the first $d$ prime numbers.
-Figure \ref{fig:Halton} shows two dimensional projections of Halton nodes in six dimensions.
-%\NKNote{Suggestion:
+\NKNote{Halton sequences can be used with arbitrary sample sizes \( n \), although its equidistribution properties are best when \( n \) is a product of powers of the \( d \) bases used.} Figure \ref{fig:Halton} shows two dimensional projections of Halton nodes in six dimensions.
 
@@
 	\includegraphics[width=\textwidth]{\figpath/64Halton_pts_d6.pdf}
-	\caption{Halton nodes as defined in \eqref{eq:Halton} have low discrepancy but no preferred sample size. \label{fig:Halton}}
+	\caption{Halton nodes as defined in \eqref{eq:Halton} have low-discrepancy. 
+    %but no preferred sample size. 
+    \label{fig:Halton}}
 \end{figure}
@@
 
-Most quasi-Monte Carlo constructions are studied based on their asymptotic behavior as \(n \to \infty\), with numerous results on asymptotic rates \cite{Nie92,NovWoz10a}.
+Most qMC constructions are studied based on their asymptotic behavior as \(n \to \infty\), with numerous results on asymptotic rates \cite{Nie92,NovWoz10a}.
 However, few methods address finding optimal node configurations for specific \(n\) and \(d\) which can be critical in practical applications where each function evaluation is quite costly.
@@
 When constructing LD nodes, it is essential to ensure not only a well-distributed set of nodes in the $d$-dimensional space but also in its lower-dimensional projections (recall the Cartesian grid from Section \ref{sec:intro}, Figure \ref{fig:grid}).
-\texttt{MatBuilder} \cite{paulin2022} is a software tool developed precisely for this task; to generate generating matrices for digital sequence LD nodes possessing excellent low dimensional uniformity.
+\texttt{MatBuilder} \cite{paulin2022} is a software tool developed precisely for this task; to generate generating matrices for digital sequence LD nodes possessing excellent low-dimensional uniformity.
 In this software, the solutions to integer linear programming problems are used to build the columns of generator matrices in a greedy manner for the LD digital sequence.
 
-%Uniformity constraints related to digital sequence stratification across boxes of the form \eqref{eq:element_box} are expressed as integer linear programming (ILP) problems and subsequently solved to optimize an objective function relating to the ILP reformulation ensuring good quality low dimensional projections.
+%Uniformity constraints related to digital sequence stratification across boxes of the form \eqref{eq:element_box} are expressed as integer linear programming (ILP) problems and subsequently solved to optimize an objective function relating to the ILP reformulation ensuring good quality low-dimensional projections.
 
 We have outlined various efforts by qMC researchers to optimize specific classes of LD constructions, including lattices and digital sequences.
-Significant effort has even been taken to optimize the uniformity of so-called generalized Halton sequences due to the poor lower dimensional projections exhibited by even a moderate dimensional Halton sequence \cite{kirklem24}.
+Significant effort has even been taken to optimize the uniformity of so-called generalized Halton sequences due to the poor lower dimensional projections exhibited by even a moderate dimensional Halton sequence \cite{faulem09,kirklem24}.
 These methods are discussed briefly in Section \ref{sec:randHalton}.
@@
 \begin{equation} \label{eq:sqnormL}
-	\lVert \eta_L \rVert_{\calH_K}^2 = \langle\eta_L, \eta_L \rangle_{\calH_K} = L(\eta_L) = L^{\cdot\cdot} \Bigl(L^{\cdot}\bigl( K(\cdot,\cdot \cdot) \bigr) \Bigr).
-\end{equation}
+	\lVert \eta_L \rVert_{\calH_K}^2 = \langle\eta_L, \eta_L \rangle_{\calH_K} = L(\eta_L) = L^{\cdot\cdot} \Bigl(L^{\cdot}\bigl( K(\cdot,\cdot \cdot) \bigr) \Bigr)
+\end{equation}
+
+\NKNote{where $L^\cdot$ denotes applying linear functional $L$ to the first argument of the kernel $K$, and $L^{\cdot\cdot}$ indicates applying $L$ to the resulting function of the second variable only.}
 
@@
 \end{align}
+
+\NKNote{where again the $(\cdot)$ and $(\cdot\cdot)$ notations represent the functional acting on the first and second arguments respectively.}
 
@@
 	&= \left ( \frac{13}{12} \right )^d
-	- \frac 2n \sum_{i=0}^{n-1} \prod_{\ell=1}^d \left [ 1 + \frac 12 \bigl ( \lvert x_{i\ell} - 1/2 \rvert - \lvert x_{i\ell} - 1/2 \rvert^d \bigr )\right ] \\
+	- \frac 2n \sum_{i=0}^{n-1} \prod_{\ell=1}^d \left [ 1 + \frac 12 \bigl ( \lvert x_{i\ell} - 1/2 \rvert - \lvert x_{i\ell} - 1/2 \rvert^2 \bigr )\right ] \\
 	& \qquad \qquad \frac{1}{n^2} \sum_{i,j=0}^{n-1} \prod_{\ell = 1}^d \left[ 1 + \frac 12 \bigl ( \lvert x_{i\ell} - 1/2 \rvert + \lvert x_{j\ell} - 1/2 \rvert - \lvert x_{i\ell} - x_{j\ell} \rvert \bigr ) \right].
@@
 \end{align}
-This quality measure $\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ is typically more optimistic (smaller), than the root mean squared of  $\text{discrepancy}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ because the latter finds a worst case function for each random instance of nodes, while the former is finds just one worst function.
+This quality measure $\text{RERR}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ is typically more optimistic (smaller), than the root mean squared of  $\text{discrepancy}\bigl(\{\bsx_i\}_{i=0}^{n-1},K\bigr)$ because the latter finds a worst case function for each random instance of nodes, while the former finds just one worst function.
 
@@
     \label{eq:sc_clt_iid_n}
-\end{equation} then $\mu \in [(]\mu^-,\mu^+]$ with probability $1-\alpha$.
+\end{equation} then $\mu \in (\mu^-,\mu^+]$ with probability $1-\alpha$.
 A simple IID stopping criterion uses an initial sample of $n_0$ nodes to approximate $\sigma^2 \approx S^2 := (n_0-1)^{-1} \sum_{i=0}^{n_0-1}(f(\bsx_i)-\hat{\mu})^2$, and then plugs this into \eqref{eq:sc_clt_iid_n} to determine $n$.
@@
 
-Often qMC problems may be written as $\mu = \bbE[g(\bsT)]$ where $\bsT$ is a $d$-dimensional random variable with density $\lambda$.
+Often qMC problems may be written as $\mu = \bbE[g(\bsT)]$ where $\bsT$ is a $d$-dimensional random variable with joint density $\lambda$.
 Since qMC nodes have LD with the standard uniform distribution $\bsX \sim \calU[0,1]^d$, we perform an invertible variable transformation $\bsPsi: [0,1]^d \to \calM$ satisfying $\mathrm{range}(\bsT) \subseteq \calM$.
@@
 For finance examples where the outcome depends more on the gross behavior of the underlying stochastic process, rather than the high frequency behavior, it makes sense to choose the transformation accordingly.
-For example, if $\bsT$ represents a Brownian motion, then a principal component analysis coupled with the inverse Gaussian distribution shows greater convergence gains compared to the Cholesky decomposition.
+For example, if $\bsT$ represents a Brownian motion, then a principal component analysis coupled with the inverse Gaussian distribution shows greater convergence gains compared to the Cholesky decomposition \cite{AcwBroGla97}.
 
@@
 Some applications have a $d$ that is either huge or infinite.
-On example is in quantitative finance, where $d$ denotes the number of times that the asset price is monitored over the life of a financial derivative times the number of assets involved.
+One example is in quantitative finance, where $d$ could denote the number of times that the asset price is monitored over the life of a financial derivative times the number of assets involved.
 
 However, the $\mathcal{O}(dn)$ cost of approximating $\mu$ may often be reduced by employing multilevel methods.
-In fact, these methods may reduce the cost to the level comparable to solving a low dimensional approximation \cite{Gil15a}.
+In fact, these methods may reduce the cost to the level comparable to solving a low-dimensional approximation \cite{Gil15a}.
 We introduce these methods here.
@@
     \nonumber
-    V_1 &= n_1 \, \text{discrepancy}\bigl(\{\bsx_i^{(1)}\}_{i=0}^{n_2-1}, K^{(1)} \bigr) \, \text{variation}(f^{(1)}, K^{(1)})\\
-    \nonumber
-    V_l & = n_l \,\text{discrepancy}\bigl(\{\bsx_i^{(l)}\}_{i=0}^{n_l-1}, K^{(l)} \bigr) \, \text{variation}(f^{(l)} - f^{(l-1)}, K^{(l)}), \\
+    V_1 &= n_1 \, \textcolor{blue}{\times} \, \text{discrepancy}\bigl(\{\bsx_i^{(1)}\}_{i=0}^{n_2-1}, K^{(1)} \bigr) \, \textcolor{blue}{\times} \, \text{variation}(f^{(1)}, K^{(1)})\\
+    \nonumber
+    V_l & = n_l \, \textcolor{blue}{\times} \, \text{discrepancy}\bigl(\{\bsx_i^{(l)}\}_{i=0}^{n_l-1}, K^{(l)} \bigr) \, \textcolor{blue}{\times} \, \text{variation}(f^{(l)} - f^{(l-1)}, K^{(l)}), \\
     \nonumber
@@
 
-It is an art to write one's problem to have the favorable case of high or even infinite nominal dimension, $d$, be solvable with a computational cost comparable to the low dimensional case.
+It is an art to write one's problem to have the favorable case of high or even infinite nominal dimension, $d$, be solvable with a computational cost comparable to the low-dimensional case.
 However, the multilevel method takes advantage of the principle noted in Section \ref{sec:coordwts}, namely that qMC thrives when problem have a low effective dimension, even if the nominal dimension is large.
@@
 Many algorithms involving sampling with LD nodes have since been developed much further in \cite{kellprem12,shired08,Keller2013a,kolkel00,kell96,kell95}.
-(Remarkably, Monte Carlo methods have even been awarded with an Academy Award for visual effects in movie making \cite{veachthesis,oscars}.)
+(Remarkably, MC and qMC methods have even been awarded with an Academy Award for visual effects in movie making \cite{veachthesis,oscars}.)
 
@@
 \section{Conclusion} \label{sec:conclusion}
-Quasi-Monte Carlo methods, based on low discrepancy sequences, can approximate the mean of a function, $f$, of $d$ variables to within an error tolerance of $\varepsilon$ at a cost of roughly $\mathcal{O}(\varepsilon^{-1})$ function evaluations, provided that the problem is formulated so that $f$ depends primarily on a modest number of variables.
+Quasi-Monte Carlo methods, based on low-discrepancy sequences, can approximate the mean of a function, $f$, of $d$ variables to within an error tolerance of $\varepsilon$ at a cost of roughly $\mathcal{O}(\varepsilon^{-1})$ function evaluations, provided that the problem is formulated so that $f$ depends primarily on a modest number of variables.
 This means that qMC outperforms alternative methods such as simple Monte Carlo and sampling on grids.

### 2025-06-10 response_letter.tex
Updates from Overleaf

@@
 \usepackage{lipsum}
+
+\input{macros.tex}
 
@@
 \end{response}
-...
+We choose this ``unusual'' way because we want to show the connection to the van der Corput sequence.
+Lattice sequences, digital sequences, and Halton sequences all can be considered extensions of van der Corput.
+
+By stating that $\bsx^{\text{dig}}_1, \bsx^{\text{dig}}_2, \ldots$ are carefully chosen, we were unintentionally vague, and have now stated that $\bsx^{\text{dig}}_1, \bsx^{\text{dig}}_2, \bsx^{\text{dig}}_4, \bsx^{\text{dig}}_8, \ldots$ are carefully chosen.
+
+Thanks to the reviewer for pointing out the error in our correspondence between generator matrices and the $\boldsymbol{x}_{2^m}$ for $m =  0, 1, \ldots$.
+That has been corrected.
+These paragraphs have also been reworded.
 
@@
 \end{response}
-....
+This has been corrected.  Thank you.
 

### 2025-06-10 HickernellMCQMC2024_QMCTutorial_rev_1.tex
Updates from Overleaf

@@
 %%Fred's temporary macros
-\newcommand{\FJHNote}[1]{{\textcolor{blue}{Fred: #1}}}
+\usepackage[dvipsnames]{xcolor}
+\newcommand{\FJHNote}[1]{{\textcolor{ForestGreen}{#1}}}
 \newcommand{\NKNote}[1]{{\textcolor{blue}{Nathan: #1}}}
@@
 \NKNote{Inevitably, this tutorial is not comprehensive of the full breadth of over a century of development in qMC and uniform distribution theory, dating back to Hermann Weyl's seminal 1916 paper \cite{Wey16}.
-Instead, our goal is to provide a solid foundation in the core concepts along with practical guidance, and to spark further interest and study in the field.}
+Instead, our goal is to provide a solid foundation in the core concepts along with practical guidance\FJHNote{, especially for the newcomer}, and to spark further interest and study in the field.}
 
@@
 
-Another family of LD sequences that is a generalization of the van der Corput sequence \eqref{eq:vdc} is digital sequences \cite{DicPil10a,Nie92}.
+\FJHNote{Another family of LD sequences that is a generalization of the van der Corput sequence \eqref{eq:vdc} is digital sequences \cite{DicPil10a,Nie92}.
 For simplicity, we restrict ourselves to $b = 2$ and let $\oplus$ denote binary digitwise addition modulo $2$ (also known as digitwise exclusive or), e.g., $3/8 \oplus 3/4 = {}_20.011 \oplus {}_20.110 = {}_20.101 = 5/8$.
@@
 \end{equation}
-where $\bsx^{\text{dig}}_1, \bsx^{\text{dig}}_2, \ldots$ are carefully chosen.
-(For $d=1$ and $x^{\text{dig}}_{2^m} = 2^{-m-1}$, this is the van  der Corput sequence in base $2$.) The node set $\{ \bsx^{\text{dig}}_0, \ldots, \bsx^{\text{dig}}_{2^m-1}\}$  for non-negative integer $m$ is called a digital net and is closed under $\oplus$.
-
-If $x^{\text{dig}}_{ij\ell}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ coordinate of $\bsx^{\text{dig}}_i$, then the literature often defines generating matrices, $\mathsf{C}_j = \bigl(x^{\text{dig}}_{ij\ell}\bigr)_{\ell,i = 1}^{M,N}$ for $j = 1, \ldots, d$, where $M$ is the maximum number of bits in the expression for $\bsx^{\text{dig}}_i$, say $52$, and $2^N$ is the maximum number of nodes that is intended to be generated.
-The digital sequence can then be defined equivalently as
+where $\bsx^{\text{dig}}_1, \bsx^{\text{dig}}_2, \bsx^{\text{dig}}_4, \bsx^{\text{dig}}_8, \ldots$ are carefully chosen to ensure low discrepancy.
+(For $d=1$ and $x^{\text{dig}}_{2^m} = 2^{-m-1}$, this is the van  der Corput sequence in base $2$.) The node set $\{ \bsx^{\text{dig}}_0, \ldots, \bsx^{\text{dig}}_{2^m-1}\}$  for non-negative integer $m$ is called a digital net and is closed under $\oplus$.}
+
+\FJHNote{Traditionally, digital nets are defined in terms of $N \times M$ generating matrices, $\mathsf{C}_j$, as follows:
 \begin{equation} \label{eq:digitalB}
 \renewcommand{\arraystretch}{1.5}
-	\begin{pmatrix} x^{\text{dig}}_{ij1} \\ x^{\text{dig}}_{ij2} \\ \vdots \end{pmatrix}
+	\begin{pmatrix} x^{\text{dig}}_{i,j,1} \\ x^{\text{dig}}_{i,j,2} \\ \vdots \end{pmatrix}
 	= \mathsf{C}_j \begin{pmatrix} i_0 \\ i_1 \\ \vdots \end{pmatrix}
-    \quad \pmod{2}, \qquad j = 1,\ldots, d, \ i = 0, 1, \ldots.
-\end{equation}
+    \quad \pmod{2}, \qquad j = 1,\ldots, d, \ i = 0, 1, \ldots,
+\end{equation}
+where $x^{\text{dig}}_{i,j,\ell}$ denotes the $\ell^{\text{th}}$ digit of the $j^{\text{th}}$ component of the $i^{\text{th}}$ node.
+Here $N$ is the maximum number of bits in the expression for $\bsx^{\text{dig}}_i$, say $52$, and $2^M$ is the maximum number of nodes that is intended to be generated.
+It can be seen that definitions  \eqref{eq:digital} and \eqref{eq:digitalB} are equivalent by noting that $\mathsf{C}_j = \bigl(x^{\text{dig}}_{2^m,j,\ell}\bigr)_{\ell,m = 0}^{N,M-1}$.
+}
 
@@
 regardless of the choice of $\bsa = (a_1, \ldots, a_d)$.
-The $t$-value of a digital net with $2^m$ nodes is defined as the smallest $t$ for which condition \eqref{eq:tcond} holds for all $\bsk$ with $\lVert \bsk \rVert_1 \le m-t$.
-Equivalently, this $t$ is the smallest value for which every box of the form \eqref{eq:element_box} with volume $2^{\lVert \bsk \rVert_1 - m}$ contains its fair share of $2^t$ nodes.
+The $t$-value of a digital net with $2^m$ nodes is defined as \FJHNote{any} $t$ for which condition \eqref{eq:tcond} holds for all $\bsk$ with $\lVert \bsk \rVert_1 \le m-t$.
+Equivalently, \FJHNote{for} this $t$, every box of the form \eqref{eq:element_box} with volume $2^{\lVert \bsk \rVert_1 - m}$ contains its fair share of $2^t$ nodes.
 Such a digital net is then called a $(t,m,d)$-net.
 An infinite sequence of the form \eqref{eq:digitalB} for which this condition holds for all non-negative $m$ is called a  $(t,d)$-sequence.
+\FJHNote{Ideally, one would like to construct $(t,m,d)$-nets and $(t,d)$-sequences with $t$ as small as possible.}
 

### 2025-06-12 response_letter.tex
Fred correcting some typos

@@
 \end{response}
-....
+.... [Aleksei]
 
@@
 \end{response}
-.....
+Thanks.  Text has been added to this effect.
 
@@
 \end{response}
-....
+We are constrained by page limitations, but more definitions were added to clarify the meaning of symbols.
 
@@
 \end{response}
-There is no specific reason. We have kept the same accuracy.
+Because we show a plot where some cubature errors are $10^{-8}$ and smaller, we want to show that the answer can be known to high accuracy.
 
@@
 The authors have slightly rewritten the paragraph above equation (9) to more clearly define the van der Corput sequence the first time it is mentioned.
+We have added a reference to the digital nets book by Dick \& Pillichshammer.
 
@@
 \end{response}
-.....
+It is a condition and it is numbered so that we can refer to it.  It has been rewritten.
 
@@
 \end{response}
-.....
+Done.
 
@@
 \end{response}
-....
+We have increase the font size of the legends for these figures.
 

### 2025-06-12 HickernellMCQMC2024_QMCTutorial_rev_1.tex
Fred correcting some typos

@@
 \newcommand{\FJHNote}[1]{{\textcolor{ForestGreen}{#1}}}
-\newcommand{\NKNote}[1]{{\textcolor{blue}{Nathan: #1}}}
-\newcommand{\AGSNote}[1]{{\textcolor{blue}{Aleksei: #1}}}
+\newcommand{\NKNote}[1]{{\textcolor{blue}{ #1}}}
+\newcommand{\AGSNote}[1]{{\textcolor{magenta}{#1}}}
 %\usepackage[normalem]{ulem}
@@
 Many questions in  quantitative finance, uncertainty quantification, and other disciplines are answered by computing the population mean, $\mu := \mathbb{E}(Y)$, where instances of $Y:=f(\boldsymbol{X})$ may be generated by numerical simulation and $\bsX$ has a simple probability  distribution.
-The population mean can be approximated by the sample mean, $\hat{\mu}_n := n^{-1} \sum_{i=0}^{n-1} f(\bsx_i)$ for a well chosen sequence of nodes, $\{\bsx_0, \bsx_1, \ldots\}$ and a sufficiently large sample size, $n$.
+The population mean can be approximated by the sample mean, $\hat{\mu}_n := n^{-1} \sum_{i=0}^{n-1} f(\bsx_i)$, for a well chosen sequence of nodes, $\{\bsx_0, \bsx_1, \ldots\}$, and a sufficiently large sample size, $n$.
 Computing $\mu$ is equivalent to computing a $d$-dimensional integral, $\int f(\bsx) \varrho(\bsx) \, \mathrm{d} \bsx$, where $\varrho$ is the probability density for $\bsX$.
@@
 \NKNote{Inevitably, this tutorial is not comprehensive of the full breadth of over a century of development in qMC and uniform distribution theory, dating back to Hermann Weyl's seminal 1916 paper \cite{Wey16}.
-Instead, our goal is to provide a solid foundation in the core concepts along with practical guidance\FJHNote{, especially for the newcomer}, and to spark further interest and study in the field.}
+Instead, our goal is to provide a solid foundation in the core concepts along with practical guidance\FJHNote{---especially for the newcomer---}and to spark further interest and study in the field.}
 
@@
 
-Now we can apply IID MC, qMC, and other computational methods to approximate $\mu$ by the sample mean
+Now we can apply IID MC, qMC, and other computational methods to approximate $\mu$ by the sample mean,
 \begin{equation} \label{eq:sample_mean_Keister}
@@
 \end{equation}
-Consider the specific case of $d=6$ for which $\mu \approx -2.327303729298$.
+Consider the specific case of $d=6$ for which $\mu = -2.327303729298\ldots$ (by quadrature).
 We approximate $\mu$ by $\hat{\mu}_n$ for different, $\{\bsx_i\}_{i=0}^{n-1}$, for various sample sizes, $n$, and plot the relative errors, $\lvert (\mu - 	\hat{\mu}_n)/\mu\rvert$ in Figure \ref{fig:keister-err}.
@@
 
-An extensible lattice sequence in one dimension is defined by the van der Corput sequence in base $b$, $\{\phi_b(0), \phi_b(1), \ldots\}$, where $\phi_b$ is the radical inverse function and essentially reflects the digits of the integer $i$ in base $b$ about the decimal point.
+An extensible lattice sequence in one dimension is defined by the van der Corput sequence in base $b$, $\{\phi_b(0), \phi_b(1), \ldots\}$, where $\phi_b$ is the radical inverse function and essentially reflects the digits of the integer $i$ in base $b$ about the decimal point \cite[Chapter 3]{DicPil10a}.
 For example, $\phi_2(6) = \phi_2(110_{2}) = {}_20.011 = 3/8$.
@@
 \begin{equation} \label{eq:extensiblelattice}
-	\bsx_i^{\text{lat}} =\phi_b( i)\bsh \pmod \bsone, \qquad i = 0, 1 , \ldots.
+	\bsx_i^{\text{lat}} =\phi_b( i)\bsh \pmod \bsone, \qquad i = 0, 1 , \ldots .
 \end{equation}
@@
 Here $N$ is the maximum number of bits in the expression for $\bsx^{\text{dig}}_i$, say $52$, and $2^M$ is the maximum number of nodes that is intended to be generated.
-It can be seen that definitions  \eqref{eq:digital} and \eqref{eq:digitalB} are equivalent by noting that $\mathsf{C}_j = \bigl(x^{\text{dig}}_{2^m,j,\ell}\bigr)_{\ell,m = 0}^{N,M-1}$.
+It can be seen that definitions  \eqref{eq:digital} and \eqref{eq:digitalB} are equivalent by noting that $\mathsf{C}_j = \bigl(x^{\text{dig}}_{2^m,j,\ell}\bigr)_{\ell,m = 0}^{N-1,M-1}$.
 }
@@
 This will occur if and only if
-\begin{multline} \label{eq:tcond}
-	\text{the first $k_1$ rows of the first $m$ columns of $\mathsf{C}_1$ plus} \\
-	\text{the first $k_2$ rows of the first $m$ columns of $\mathsf{C}_2$ plus} \\
-	 \vdots \\
-	\text{the first $k_d$ rows of the first $m$ columns of $\mathsf{C}_d$} \\
+\FJHNote{
+\begin{equation} \label{eq:tcond}
+\begin{aligned}
+	\text{the first $k_1$ rows of the first $m$ columns of $\mathsf{C}_1$} & \text{ plus} \\
+	\text{the first $k_2$ rows of the first $m$ columns of $\mathsf{C}_2$} & \text{ plus} \\
+	 \vdots \hspace{10ex} & \\
+	\text{the first $k_d$ rows of the first $m$ columns of $\mathsf{C}_d$} &\\
 	\text{are linearly independent over addition and multiplication modulo $2$, }
-\end{multline}
+\end{aligned}
+\end{equation}}
 regardless of the choice of $\bsa = (a_1, \ldots, a_d)$.
@@
 \end{equation}
-
 \NKNote{where $L^\cdot$ denotes applying linear functional $L$ to the first argument of the kernel $K$, and $L^{\cdot\cdot}$ indicates applying $L$ to the resulting function of the second variable only.}
@@
 \subsection{Coordinate weights} \label{sec:coordwts}
-When the dimension, $d$, is increased, the discrepancy increases and the rate of decay with increasing sample size deteriorates, as can be seen in Figure \ref{fig:scaledweightedcentered} (left).
+When the dimension, $d$, is increased, the discrepancy increases and the rate of decay with increasing sample size deteriorates, as can be seen in Figure \ref{fig:scaledweightedcentered} (top).
 This raises the question of whether the LD sequences lose their effectiveness for higher dimensions or whether the theory of the previous subsection does not coincide with practice.
@@
 
-Figure \ref{fig:scaledweightedcentered} (right) plots the scaled centered discrepancy with coordinate weights $\bsgamma = (1/\ell)_{\ell =1}^d$, i.e., 
+Figure \ref{fig:scaledweightedcentered} (bottom) plots the scaled centered discrepancy with coordinate weights $\bsgamma = (1/\ell)_{\ell =1}^d$, i.e., 
 \begin{equation*}
@@
 
-\begin{figure}
-    \centering
-     \includegraphics[width=0.45\textwidth]{\figpath/SobolCtrDisc_n_8192_d64.pdf}
-    \includegraphics[width=0.45\textwidth]{\figpath/IIDSobolScaledWtCtrDisc_n_8192_d64.pdf}
-    \caption{Left: The centered discrepancy for a Sobol' LD sequence for a range of $d$.
+\begin{figure}[!ht]
+    \centering
+    \includegraphics[width=.8\textwidth]{\figpath/SobolCtrDisc_n_8192_d64.pdf} \\
+    \includegraphics[width=.8\textwidth]{\figpath/IIDSobolScaledWtCtrDisc_n_8192_d64.pdf}
+    \caption{Top: The centered discrepancy for a Sobol' LD sequence for a range of $d$.
 For larger $d$ only a decay rate of $\calO(n^{-1/2})$, is achieved for the range of $n$ shown.
 Moreover, the discrepancy for larger $d$ is huge.
-    Right:
+    Bottom:
     The scaled centered discrepancy with coordinate weights for a Sobol' LD sequence for various $d$ and the root mean squared scaled centered discrepancy with coordinate weights for IID nodes.
@@
 Conditioned on observations $f(\bsx_i)$, a closed form expression exists for the $1-\alpha$ credible intervals for $\mu$.
-While also applicable when using IID nodes, using LD sequences paired with special kernels enables these credible intervals to be computed at only $\calO(n \log n)$ cost \cite{Jag19a,JagHic22a,RatHic19a}.
+While also applicable when using IID nodes, using LD sequences paired with special kernels enables these credible intervals to be computed at only $\calO(n \log n)$ cost \cite{Jag19a,RatHic19a,JagHic22a}.
 Again one may iteratively double the sample size $n$ until the credible intervals are sufficiently tight.
@@
 The second is that qMC does best when most of the variation of $f$ is concentrated in the lower numbered coordinates, and, if possible, we should design $f$ so that this is so.
+
+\FJHNote{The methods described in this section for increasing the efficiency of qMC have their parallels in variance reduction methods for IID MC.
+The difference is that rather than decreasing the variance of $f(\bsX)$, we aim to decrease  $\text{variation}(f,K)$, which appears in error bound \eqref{eq:cuberrbd}.
+Sometimes the same mathematical constructs work for both IID MC and qMC, but not always.}
+
+\FJHNote{Not all variance reduction methods for IID MC can be easily transferred to qMC.
+For example, acceptance-rejection sampling requires an undetermined number of IID samples before one is accepted.
+This is not a qMC friendly method.}
 
@@
 For finance examples where the outcome depends more on the gross behavior of the underlying stochastic process, rather than the high frequency behavior, it makes sense to choose the transformation accordingly.
-For example, if $\bsT$ represents a Brownian motion, then a principal component analysis coupled with the inverse Gaussian distribution shows greater convergence gains compared to the Cholesky decomposition \cite{AcwBroGla97}.
+\AGSNote{For example, in Figure \ref{fig:asian_option_pda_vs_chol} we see that if  $\bsT$ represents a Brownian motion, then a principal component analysis coupled with the inverse Gaussian distribution shows greater convergence gains compared to the Cholesky decomposition \cite{AcwBroGla97}.}
+
+\begin{figure}
+	\centering
+    \includegraphics[width = 0.48\textwidth]{\figpath//NumSamples_Asian_Option_tol_0.001_d52.pdf}
+	\includegraphics[width = 0.48\textwidth]{\figpath/Times_Asian_Option_tol_0.001_d52.pdf} \quad 
+	\caption{\AGSNote{Comparison of the required number of samples (left) and execution time (right) versus the requested error tolerance for (Q)MC algorithms.
+We compare decompositions of the Brownian motion using principal component analysis (PCA) against the Cholesky decomposition for IID and Sobol' points.
+\label{fig:asian_option_pda_vs_chol}}}
+\end{figure}
 
@@
     \hat{\mu}^{\text{ML}}_{\bsn,\bsd}  := \frac{1}{n_1}  \sum_{i=0}^{n_1-1} f_1(\bsx^{(1)}_i)  +  \frac{1}{n_2}  \sum_{i=0}^{n_2-1}  \bigl[f_2(\bsx^{(2)}_i) - f_1(\bsx^{(1)}_i)\bigr] + \cdots \\
-     + \frac{1}{n_L}  \sum_{i=0}^{n_L-1}  \bigl[f_L(\bsx_i^{(L)}) - f_{L-1}(\bsx_i^{(L-1)}) \bigr].
+     + \frac{1}{n_L}  \sum_{i=0}^{n_L-1}  \bigl[f_L(\bsx_i^{(L)}) - f_{L-1}(\bsx_i^{(L-1)}) \bigr],
 \end{multline}
+\FJHNote{where $\bsn = (n_1, \ldots, n_L)$ and $\bsd = (d_1, \ldots, d_L$.} 
 If $C_1$ is the computational cost of $f^{(1)}(\bsx^{(1)})$ and $C_l$ is the computational cost of $f^{(l)}(\bsx^{(1)}) - f_{l-1}(\bsx_i^{(l-1)})$ for $l = 2, \ldots, L$, then the total computational cost of
@@
 
-Optimizing $\bsn$ to minimize the error bound in \eqref{eq:MLerrbd} for a fixed cost and assuming constant $\bsC$ and $\bsV$, yields
+Optimizing \FJHNote{vector of sample sizes} $\bsn$ to minimize the error bound in \eqref{eq:MLerrbd} for a fixed cost and assuming constant \FJHNote{$\bsC = (C_1, \ldots, C_L)$ and $\bsV = (V_1, \ldots, V_L)$}, yields \FJHNote{$\bigl \lvert \mu - \hat{\mu}^{\text{ML}}_{\bsn,\bsd} \bigr \rvert \le \varepsilon$ for}
 \begin{align}
-    \bigl \lvert \mu - \hat{\mu}^{\text{ML}}_{\bsn,\bsd} \bigr \rvert &\le \varepsilon \\
-\intertext{for}
-    n_l & = \frac{\sqrt{V_l} \bigl(\sqrt{V_1 C_1} + \ldots + \sqrt{V_LC_L}\bigr)}{\varepsilon \sqrt{C_l} }, \qquad l = 1, \ldots, d,\\
-    \text{cost}(\bsn,\bsd) & = \frac{\bigl(\sqrt{V_1 C_1} + \ldots + \sqrt{V_LC_L}\bigr)^2}{\varepsilon}. \label{eq:MLcost}
+\nonumber
+    n_l & = \frac{\sqrt{V_l} \bigl(\sqrt{V_1 C_1} + \cdots + \sqrt{V_LC_L}\bigr)}{\varepsilon \sqrt{C_l} }, \qquad l = 1, \ldots, d,\\
+    \text{cost}(\bsn,\bsd) & = \frac{\bigl(\sqrt{V_1 C_1} + \cdots + \sqrt{V_LC_L}\bigr)^2}{\varepsilon}. \label{eq:MLcost}
 \end{align}
@@
 \bibliographystyle{spmpsci}
-\bibliography{FJH25,FJHown23,main}
+\bibliography{FJH25,FJHown25,main}
 

### 2025-06-15 HickernellMCQMC2024_QMCTutorial_rev_1.tex
aleksei: updates to quantiles in convergence plots + asian option geometric mean CV

@@
 	\centering
-	\includegraphics[width =0.5\textwidth]{\figpath/n_is_16777216_d_is_6_n_rep_is_50_KeisterErrors.pdf}
+	\includegraphics[width =0.75\textwidth]{\figpath/n_is_16777216_d_is_6_n_rep_is_50_KeisterErrors.pdf}
 	\caption{The relative error of approximating $\mu$ defined in \eqref{eq:keisterC} by the sample mean, $\hat{\mu}_n$, defined in \eqref{eq:sample_mean_Keister} for various choices of nodes, $\bsx_0, \bsx_1, \ldots$.
@@
 
-\NKNote{where again the $(\cdot)$ and $(\cdot\cdot)$ notations represent the functional acting on the first and second arguments respectively.}
+\NKNote{where again the $(\cdot)$ and $(\cdot\cdot)$ notations represent the functional acting on the first and second arguments, respectively.}
 
@@
 	\centering
-    \includegraphics[width = 0.48\textwidth]{\figpath//NumSamples_Asian_Option_tol_0.001_d52.pdf}
-	\includegraphics[width = 0.48\textwidth]{\figpath/Times_Asian_Option_tol_0.001_d52.pdf} \quad 
-	\caption{\AGSNote{Comparison of the required number of samples (left) and execution time (right) versus the requested error tolerance for (Q)MC algorithms.
-We compare decompositions of the Brownian motion using principal component analysis (PCA) against the Cholesky decomposition for IID and Sobol' points.
+ %    \includegraphics[width = 0.48\textwidth]{\figpath//NumSamples_Asian_Option_tol_0.001_d52.pdf}
+	% \includegraphics[width = 0.48\textwidth]{\figpath/Times_Asian_Option_tol_0.001_d52.pdf} \quad 
+    \includegraphics[width = 1\textwidth]{\figpath//Times_NumSamples_Asian_Option_tol_0.001_d52.pdf}
+	\caption{\AGSNote{Comparison of Monte Carlo and Quasi-Monte Carlo stopping criteria for estimating the fair price of an Asian option with arithmetic mean.
+For QMC with Sobol' points, using a principal component analysis (PCA) decomposition of the Brownian motion provides substantial savings compared to the Cholesky decomposition.
+Using an Asian option with geometric mean control variate (CV) provides additional savings.
 \label{fig:asian_option_pda_vs_chol}}}
@@
 Control variates have been implemented in the stopping criteria based on Fourier coefficients described in the previous section \cite{HicEtal17a}.
+\AGSNote{In Figure \ref{fig:asian_option_pda_vs_chol} we show that, for pricing an Asian option with arithmetic mean, using an Asian option with geometric mean control variate can provide improved convergence.}
 
@@
 \end{multline}
-\FJHNote{where $\bsn = (n_1, \ldots, n_L)$ and $\bsd = (d_1, \ldots, d_L$.} 
+\FJHNote{where $\bsn = (n_1, \ldots, n_L)$ and $\bsd = (d_1, \ldots, d_L)$.} 
 If $C_1$ is the computational cost of $f^{(1)}(\bsx^{(1)})$ and $C_l$ is the computational cost of $f^{(l)}(\bsx^{(1)}) - f_{l-1}(\bsx_i^{(l-1)})$ for $l = 2, \ldots, L$, then the total computational cost of

### 2025-06-24 response_letter.tex
Updates from Overleaf

@@
 \end{response}
-.... [Aleksei]
+We thank the reviewer for the suggestion to include a more practical application.
+We have added an illustrative example on the fair price of an Asian option using MC and QMC methods to tie in with the discussions of variable transforms and control variates (see below).
 
@@
 \end{response}
-We have increase the font size of the legends for these figures.
+We have increased the font size of the legends for these figures.
 

### 2025-06-24 HickernellMCQMC2024_QMCTutorial_rev_1.tex
Updates from Overleaf

@@
 \usepackage[dvipsnames]{xcolor}
-\newcommand{\FJHNote}[1]{{\textcolor{ForestGreen}{#1}}}
+\newcommand{\FJHNote}[1]{{\textcolor{blue}{#1}}}
 \newcommand{\NKNote}[1]{{\textcolor{blue}{ #1}}}
-\newcommand{\AGSNote}[1]{{\textcolor{magenta}{#1}}}
+\newcommand{\AGSNote}[1]{{\textcolor{blue}{#1}}}
 %\usepackage[normalem]{ulem}
@@
 \and
-\email{hickernell@iit.edu}, \
+\email{hickernell@illinoistech.edu}, \
 %\texttt{http://www.iit.edu/\textasciitilde{}hickernell}
-\url{http://www.iit.edu/~hickernell}
+\href{https://www.iit.edu/directory/people/fred-hickernell}{http://www.iit.edu/hickernell}
 \and
-\email{nkirk@iit.edu}
+\email{nkirk@illinoistech.edu}
 \and
-\email{asorokin@hawk.iit.edu}
+\email{asorokin@hawk.illinoistech.edu}
 }
@@
 	 \vdots \hspace{10ex} & \\
-	\text{the first $k_d$ rows of the first $m$ columns of $\mathsf{C}_d$} &\\
+	\text{ the first $k_d$ rows of the first $m$ columns of $\mathsf{C}_d$} &\\
 	\text{are linearly independent over addition and multiplication modulo $2$, }
@@
 For example, in option pricing one may set the error tolerance to a penny, $\varepsilon=0.01$, and wish to find error bounds which hold with 95\% confidence, $\alpha=0.05$.
-The stopping criteria presented here are implemented in QMCPy \cite{QMCPy2020a}.
+The stopping criteria presented here are implemented in \texttt{qmcpy} \cite{QMCPy2020a}.
 
@@
 As in the qMC-CLT stopping criterion, one may iteratively double $n$ until the bounds are sufficiently tight.
+\AGSNote{Figure \ref{fig:asian_option_pda_vs_chol} compares these decay-tracking qMC algorithms against IID Monte Carlo stopping criterion across a variety of tolerances for pricing an Asian option from finance \cite[p.\ 8]{Gla03}.}
 
@@
     \includegraphics[width = 1\textwidth]{\figpath//Times_NumSamples_Asian_Option_tol_0.001_d52.pdf}
-	\caption{\AGSNote{Comparison of Monte Carlo and Quasi-Monte Carlo stopping criteria for estimating the fair price of an Asian option with arithmetic mean.
+	\caption{\AGSNote{Comparison of IID Monte Carlo and qMC stopping criteria for estimating the fair price of an Asian option with arithmetic mean.
 For QMC with Sobol' points, using a principal component analysis (PCA) decomposition of the Brownian motion provides substantial savings compared to the Cholesky decomposition.
@@
 
-An example of this in quantitative finance is where one may use the payoff of a European option, whose expectation is known, as a control variate for more exotic options.
+An example of this in quantitative finance is where one may use the payoff of a \AGSNote{similar} option, whose expectation is known, as a control variate for more exotic options.
 As noted in \cite{HicEtal03}, reducing the variation of $f$ via control variates may or may not resemble reducing the variance of $f$.
@@
 Software libraries are available to facilitate qMC computations.
+
+\subsubsection*{Acknowledgments.} This work is supported by National Science Foundation (DMS Grant No.
+2316011).
+The third author would like to acknowledge support from the DOE Office of Science Graduate Student Research (SCGSR) Program.
 

### 2025-08-04 strip_notes.py
Strip out notes

@@ -1,0 +1,102 @@
+#!/usr/bin/env python3
+# strip_notes.py
+#
+# Description:
+#   Recursively strips LaTeX macros like \FJHNote{...}, \NKNote{...}, \AGSNote{...}
+#   from a .tex file, preserving their content.
+#
+# Usage:
+#   ./strip_notes.py myfile.tex
+#
+#   or
+#
+#   python strip_notes.py <input_file.tex>
+#
+#   Example:
+#     python strip_notes.py myfile.tex
+#     → creates myfile_cleaned.tex
+#
+# Customization:
+#   You can edit the `macros` list inside the script to remove different macros.
+#
+# Author: FJH and ChatGPT
+# Date: 2025-08-04
+
+import re
+import sys
+import os
+
+def strip_macros_recursive(content, macros_to_strip):
+    pattern = re.compile(
+        r'\\(' + '|'.join(macros_to_strip) + r')\s*\{',
+        re.DOTALL
+    )
+
+    def find_macro_at(text, start):
+        match = pattern.match(text, start)
+        if not match:
+            return None
+        i = match.end()
+        depth = 1
+        brace_content = []
+        while i < len(text) and depth > 0:
+            if text[i] == '{':
+                depth += 1
+            elif text[i] == '}':
+                depth -= 1
+            if depth > 0:
+                brace_content.append(text[i])
+            i += 1
+        return (match.start(), i, ''.join(brace_content))
+
+    while True:
+        i = 0
+        found = False
+        new_text = []
+        while i < len(content):
+            result = find_macro_at(content, i)
+            if result:
+                found = True
+                start, end, inner = result
+                new_text.append(content[i:start])
+                new_text.append(inner)
+                i = end
+            else:
+                new_text.append(content[i])
+                i += 1
+        updated_content = ''.join(new_text)
+        if not found:
+            break
+        content = updated_content
+
+    return content
+
+def generate_output_filename(input_filename):
+    base, ext = os.path.splitext(input_filename)
+    return f"{base}_cleaned{ext}"
+
+# === Entry Point ===
+
+if __name__ == "__main__":
+    if len(sys.argv) < 2:
+        print("Usage: python strip_notes.py <input_file.tex>")
+        sys.exit(1)
+
+    input_file = sys.argv[1]
+    output_file = generate_output_filename(input_file)
+
+    macros = ['FJHNote', 'NKNote', 'AGSNote']  # Customize as needed
+
+    try:
[…]

### 2025-08-04 HickernellMCQMC2024_QMCTutorial_rev_3.tex
fred updates before aleksei takes a look

--- 2025-06-24 HickernellMCQMC2024_QMCTutorial_rev_1.tex
@@
 %\texttt{http://www.iit.edu/\textasciitilde{}hickernell}
-\href{https://www.iit.edu/directory/people/fred-hickernell}{http://www.iit.edu/hickernell}
 \and
@@
 We briefly discuss applications of qMC beyond computing the mean (Section \ref{sec:beyond}), and finish with a short conclusion (Section \ref{sec:conclusion}).
-\NKNote{Inevitably, this tutorial is not comprehensive of the full breadth of over a century of development in qMC and uniform distribution theory, dating back to Hermann Weyl's seminal 1916 paper \cite{Wey16}.
-Instead, our goal is to provide a solid foundation in the core concepts along with practical guidance\FJHNote{---especially for the newcomer---}and to spark further interest and study in the field.}
+Inevitably, this tutorial is not comprehensive of the full breadth of over a century of development in qMC and uniform distribution theory, dating back to Hermann Weyl's seminal 1916 paper \cite{Wey16}.
+Instead, our goal is to provide a solid foundation in the core concepts along with practical guidance---especially for the newcomer---and to spark further interest and study in the field.
 
@@
 
-\FJHNote{Another family of LD sequences that is a generalization of the van der Corput sequence \eqref{eq:vdc} is digital sequences \cite{DicPil10a,Nie92}.
+Another family of LD sequences that is a generalization of the van der Corput sequence \eqref{eq:vdc} is digital sequences \cite{DicPil10a,Nie92}.
 For simplicity, we restrict ourselves to $b = 2$ and let $\oplus$ denote binary digitwise addition modulo $2$ (also known as digitwise exclusive or), e.g., $3/8 \oplus 3/4 = {}_20.011 \oplus {}_20.110 = {}_20.101 = 5/8$.
@@
 where $\bsx^{\text{dig}}_1, \bsx^{\text{dig}}_2, \bsx^{\text{dig}}_4, \bsx^{\text{dig}}_8, \ldots$ are carefully chosen to ensure low discrepancy.
-(For $d=1$ and $x^{\text{dig}}_{2^m} = 2^{-m-1}$, this is the van  der Corput sequence in base $2$.) The node set $\{ \bsx^{\text{dig}}_0, \ldots, \bsx^{\text{dig}}_{2^m-1}\}$  for non-negative integer $m$ is called a digital net and is closed under $\oplus$.}
-
-\FJHNote{Traditionally, digital nets are defined in terms of $N \times M$ generating matrices, $\mathsf{C}_j$, as follows:
+(For $d=1$ and $x^{\text{dig}}_{2^m} = 2^{-m-1}$, this is the van  der Corput sequence in base $2$.) The node set $\{ \bsx^{\text{dig}}_0, \ldots, \bsx^{\text{dig}}_{2^m-1}\}$  for non-negative integer $m$ is called a digital net and is closed under $\oplus$.
+
+Traditionally, digital nets are defined in terms of $N \times M$ generating matrices, $\mathsf{C}_j$, as follows:
 \begin{equation} \label{eq:digitalB}
@@
 It can be seen that definitions  \eqref{eq:digital} and \eqref{eq:digitalB} are equivalent by noting that $\mathsf{C}_j = \bigl(x^{\text{dig}}_{2^m,j,\ell}\bigr)_{\ell,m = 0}^{N-1,M-1}$.
-}
 
@@
 This will occur if and only if
-\FJHNote{
+
 \begin{equation} \label{eq:tcond}
@@
 \end{aligned}
-\end{equation}}
+\end{equation}
 regardless of the choice of $\bsa = (a_1, \ldots, a_d)$.
-The $t$-value of a digital net with $2^m$ nodes is defined as \FJHNote{any} $t$ for which condition \eqref{eq:tcond} holds for all $\bsk$ with $\lVert \bsk \rVert_1 \le m-t$.
-Equivalently, \FJHNote{for} this $t$, every box of the form \eqref{eq:element_box} with volume $2^{\lVert \bsk \rVert_1 - m}$ contains its fair share of $2^t$ nodes.
+The $t$-value of a digital net with $2^m$ nodes is defined as any $t$ for which condition \eqref{eq:tcond} holds for all $\bsk$ with $\lVert \bsk \rVert_1 \le m-t$.
+Equivalently, for this $t$, every box of the form \eqref{eq:element_box} with volume $2^{\lVert \bsk \rVert_1 - m}$ contains its fair share of $2^t$ nodes.
 Such a digital net is then called a $(t,m,d)$-net.
-An infinite sequence of the form \eqref{eq:digitalB} for which this condition holds for all non-negative $m$ is called a  $(t,d)$-sequence.
-\FJHNote{Ideally, one would like to construct $(t,m,d)$-nets and $(t,d)$-sequences with $t$ as small as possible.}
+\FJHNote{An infinite sequence defined by \eqref{eq:digitalB} where every box of the form \eqref{eq:element_box} contains its fair share of elements of the subsequence $\{\bsx^{\text{dig}}_{p 2^m}, \ldots \bsx^{\text{dig}}_{(p+1) 2^m-1} \}$ for all non-negative $m$ and $p$ is called a digital $(t,d)$-sequence.} Ideally, one would like to construct $(t,m,d)$-nets and $(t,d)$-sequences with $t$ as small as possible.
 
@@
 The earliest instance is due to Sobol' \cite{Sob67}.
-%\NKNote{I added these two references which were previously left blank.
-Were you thinking of other papers to cite here?}
+%I added these two references which were previously left blank.
+Were you thinking of other papers to cite here?
 
@@
 Often they are chosen as the first $d$ prime numbers.
-\NKNote{Halton sequences can be used with arbitrary sample sizes \( n \), although its equidistribution properties are best when \( n \) is a product of powers of the \( d \) bases used.} Figure \ref{fig:Halton} shows two dimensional projections of Halton nodes in six dimensions.
+Halton sequences can be used with arbitrary sample sizes \( n \), although its equidistribution properties are best when \( n \) is a product of powers of the \( d \) bases used.
+Figure \ref{fig:Halton} shows two dimensional projections of Halton nodes in six dimensions.
 
@@
 More generally, a permutation scramble for digital nets sets $x_{ij\ell} \mapsto \Pi_{j\ell}(x_{ij\ell})$ where $\Pi_{j\ell}$ are independent random permutations of $\{0,\dots,b_j-1\}$.
-\AGSNote{Ignore this permutation definition as it is the same as \eqref{eq:gen_vdc}?} Randomization of Halton relating to permutations of digits was discussed briefly in Section \ref{sec:optLD}.
+Ignore this permutation definition as it is the same as \eqref{eq:gen_vdc}?
+Randomization of Halton relating to permutations of digits was discussed briefly in Section \ref{sec:optLD}.
 
@@
 The generalized Halton sequence is subsequently defined by choosing $d$ sequences of permutations $\Sigma_j = (\sigma_{j,r})_{r\geq0}$ of $\{0,\ldots,b_j-1\}$ for $j = 1, 2,\ldots,d$ and setting
-$$\bsx_i^{\text{G-Hal}} = \left( \phi^{\Sigma_1}_{b_1}(i), \ldots, \phi^{\Sigma_d}_{b_d}(i)\right), \quad i= 0,1,\ldots,$$ A randomized qMC estimate $\hat{\mu}_n$ is then an unbiased estimator of $\mu$ if the permutations $\Sigma_j$ are sampled independently and uniformly, i.e $\mu$ if $\Sigma_j \overset{\text{IID}}{\sim} \calU\{0,\dots,b_j-1\}^d$ since this ensures $\bsx_i^{\text{G-Hal}} \sim \mathcal{U}[0,1)^d$.
-
-We note in passing that many works \cite{MasChiWar05,tuffin98,Atan04,vancools06} have sought to optimize the permutations $\Sigma_j$ deterministically to enhance uniformity.
+$$
+\bsx_i^{\text{G-Hal}} = \left( \phi^{\Sigma_1}_{b_1}(i), \ldots, \phi^{\Sigma_d}_{b_d}(i)\right), \quad i= 0,1,\ldots.
+$$ 
+A randomized qMC estimate $\hat{\mu}_n$ is then an unbiased estimator of $\mu$ if the permutations $\Sigma_j$ are sampled independently and uniformly, i.e $\mu$ if $\Sigma_j \overset{\text{IID}}{\sim} \calU\{0,\dots,b_j-1\}^d$ since this ensures $\bsx_i^{\text{G-Hal}} \sim \mathcal{U}[0,1)^d$.
+
+We note in passing that many works \cite{Atan04,MasChiWar05,tuffin98,vancools06} have sought to optimize the permutations $\Sigma_j$ deterministically to enhance uniformity.
 For a comprehensive survey of the Halton sequence (up to 2008), we refer readers to the excellent work by Faure and Lemieux \cite{faulem09}.
@@
 		\label{eq:KcondSym}
-		\text{Symmetry:} \quad & K(\bst,\bsx) = K(\bsx,\bst) \qquad \forall \bst,\bsx \in [0,1)^d \\
+		\text{Symmetry:} \quad & K(\bst,\bsx) = K(\bsx,\bst) \qquad \forall \bst,\bsx \in [0,1)^d, \\
 		\nonumber
@@
 	\begin{align}
-	\text{Belonging:} \quad & K(\cdot,\bsx) \in \calH_K \qquad \forall \bsx \in [0,1)^d \\
+	\text{Belonging:} \quad & K(\cdot,\bsx) \in \calH_K \qquad \forall \bsx \in [0,1)^d, \\
 	\text{Reproducing:} \quad & f(\bsx) = \langle K(\cdot,\bsx),f \rangle \qquad  \forall \bsx \in [0,1)^d, \ f \in \calH_K.
@@
 \end{equation}
-\NKNote{where $L^\cdot$ denotes applying linear functional $L$ to the first argument of the kernel $K$, and $L^{\cdot\cdot}$ indicates applying $L$ to the resulting function of the second variable only.}
+where $L^\cdot$ denotes applying linear functional $L$ to the first argument of the kernel $K$, and $L^{\cdot\cdot}$ indicates applying $L$ to the resulting function of the second variable only.
 
@@
 	& = \int_{[0,1)^d \times [0,1)^d} K(\bst,\bsx) \, \mathrm{d} \bst \, \mathrm{d} \bsx  -\frac 2n  \sum_{i=0}^{n-1} \int_{[0,1)^d} K(\bst,\bsx_i) \, \mathrm{d} \bst \\
-	& \qquad \qquad + \frac{1}{n^2} \sum_{i,j=0}^{n-1}  K(\bsx_i,\bsx_j). \label{eq:sqdisc}
+	& \qquad \qquad + \frac{1}{n^2} \sum_{i,j=0}^{n-1}  K(\bsx_i,\bsx_j), \label{eq:sqdisc}
 \end{align}
-
-\NKNote{where again the $(\cdot)$ and $(\cdot\cdot)$ notations represent the functional acting on the first and second arguments, respectively.}
+where again the $(\cdot)$ and $(\cdot\cdot)$ notations represent the functional acting on the first and second arguments, respectively.
 
@@
 As in the qMC-CLT stopping criterion, one may iteratively double $n$ until the bounds are sufficiently tight.
-\AGSNote{Figure \ref{fig:asian_option_pda_vs_chol} compares these decay-tracking qMC algorithms against IID Monte Carlo stopping criterion across a variety of tolerances for pricing an Asian option from finance \cite[p.\ 8]{Gla03}.}
+Figure \ref{fig:asian_option_pda_vs_chol} compares these decay-tracking qMC algorithms against IID Monte Carlo stopping criterion across a variety of tolerances for pricing an Asian option from finance \cite[p.\ 8]{Gla03}.
 
@@
 
-\FJHNote{The methods described in this section for increasing the efficiency of qMC have their parallels in variance reduction methods for IID MC.
+The methods described in this section for increasing the efficiency of qMC have their parallels in variance reduction methods for IID MC.
 The difference is that rather than decreasing the variance of $f(\bsX)$, we aim to decrease  $\text{variation}(f,K)$, which appears in error bound \eqref{eq:cuberrbd}.
-Sometimes the same mathematical constructs work for both IID MC and qMC, but not always.}
-
-\FJHNote{Not all variance reduction methods for IID MC can be easily transferred to qMC.
+Sometimes the same mathematical constructs work for both IID MC and qMC, but not always.
+
+%Not all variance reduction methods for IID MC can be easily transferred to qMC.
 For example, acceptance-rejection sampling requires an undetermined number of IID samples before one is accepted.
-This is not a qMC friendly method.}
+This is not a qMC friendly method.
 
@@
 For finance examples where the outcome depends more on the gross behavior of the underlying stochastic process, rather than the high frequency behavior, it makes sense to choose the transformation accordingly.
-\AGSNote{For example, in Figure \ref{fig:asian_option_pda_vs_chol} we see that if  $\bsT$ represents a Brownian motion, then a principal component analysis coupled with the inverse Gaussian distribution shows greater convergence gains compared to the Cholesky decomposition \cite{AcwBroGla97}.}
+For example, in Figure \ref{fig:asian_option_pda_vs_chol} we see that if  $\bsT$ \FJHNote{contains observations from a Brownian motion}, then a principal component analysis coupled with the inverse Gaussian distribution shows greater convergence gains compared to the Cholesky decomposition \cite{AcwBroGla97}.
 
@@
     \includegraphics[width = 1\textwidth]{\figpath//Times_NumSamples_Asian_Option_tol_0.001_d52.pdf}
-	\caption{\AGSNote{Comparison of IID Monte Carlo and qMC stopping criteria for estimating the fair price of an Asian option with arithmetic mean.
+	\caption{Comparison of IID Monte Carlo and qMC stopping criteria for estimating the fair price of an Asian option with arithmetic mean.
 For QMC with Sobol' points, using a principal component analysis (PCA) decomposition of the Brownian motion provides substantial savings compared to the Cholesky decomposition.
 Using an Asian option with geometric mean control variate (CV) provides additional savings.
-\label{fig:asian_option_pda_vs_chol}}}
+\label{fig:asian_option_pda_vs_chol}}
 \end{figure}
@@
 
-An example of this in quantitative finance is where one may use the payoff of a \AGSNote{similar} option, whose expectation is known, as a control variate for more exotic options.
+An example of this in quantitative finance is where one may use the payoff of a similar option, whose expectation is known, as a control variate for more exotic options.
 As noted in \cite{HicEtal03}, reducing the variation of $f$ via control variates may or may not resemble reducing the variance of $f$.
 Control variates have been implemented in the stopping criteria based on Fourier coefficients described in the previous section \cite{HicEtal17a}.
-\AGSNote{In Figure \ref{fig:asian_option_pda_vs_chol} we show that, for pricing an Asian option with arithmetic mean, using an Asian option with geometric mean control variate can provide improved convergence.}
+In Figure \ref{fig:asian_option_pda_vs_chol} we show that, for pricing an Asian option with arithmetic mean, using an Asian option with geometric mean control variate can provide improved convergence.
 
@@
 \end{multline}
-\FJHNote{where $\bsn = (n_1, \ldots, n_L)$ and $\bsd = (d_1, \ldots, d_L)$.} 
+where $\bsn = (n_1, \ldots, n_L)$ and $\bsd = (d_1, \ldots, d_L)$. 
 If $C_1$ is the computational cost of $f^{(1)}(\bsx^{(1)})$ and $C_l$ is the computational cost of $f^{(l)}(\bsx^{(1)}) - f_{l-1}(\bsx_i^{(l-1)})$ for $l = 2, \ldots, L$, then the total computational cost of
@@
     \nonumber
-    V_1 &= n_1 \, \textcolor{blue}{\times} \, \text{discrepancy}\bigl(\{\bsx_i^{(1)}\}_{i=0}^{n_2-1}, K^{(1)} \bigr) \, \textcolor{blue}{\times} \, \text{variation}(f^{(1)}, K^{(1)})\\
-    \nonumber
-    V_l & = n_l \, \textcolor{blue}{\times} \, \text{discrepancy}\bigl(\{\bsx_i^{(l)}\}_{i=0}^{n_l-1}, K^{(l)} \bigr) \, \textcolor{blue}{\times} \, \text{variation}(f^{(l)} - f^{(l-1)}, K^{(l)}), \\
+    V_1 &= n_1 \, \times \, \text{discrepancy}\bigl(\{\bsx_i^{(1)}\}_{i=0}^{n_2-1}, K^{(1)} \bigr) \, \times \, \text{variation}(f^{(1)}, K^{(1)})\\
+    \nonumber
+    V_l & = n_l \, \times \, \text{discrepancy}\bigl(\{\bsx_i^{(l)}\}_{i=0}^{n_l-1}, K^{(l)} \bigr) \, \times \, \text{variation}(f^{(l)} - f^{(l-1)}, K^{(l)}), \\
     \nonumber
@@
 
-Optimizing \FJHNote{vector of sample sizes} $\bsn$ to minimize the error bound in \eqref{eq:MLerrbd} for a fixed cost and assuming constant \FJHNote{$\bsC = (C_1, \ldots, C_L)$ and $\bsV = (V_1, \ldots, V_L)$}, yields \FJHNote{$\bigl \lvert \mu - \hat{\mu}^{\text{ML}}_{\bsn,\bsd} \bigr \rvert \le \varepsilon$ for}
+Optimizing vector of sample sizes $\bsn$ to minimize the error bound in \eqref{eq:MLerrbd} for a fixed cost and assuming constant $\bsC = (C_1, \ldots, C_L)$ and $\bsV = (V_1, \ldots, V_L)$, yields $\bigl \lvert \mu - \hat{\mu}^{\text{ML}}_{\bsn,\bsd} \bigr \rvert \le \varepsilon$ for
 \begin{align}
@@
 
-\subsubsection*{Acknowledgments.} This work is supported by National Science Foundation (DMS Grant No.
-2316011).
+\subsubsection*{Acknowledgments.} This work is supported by National Science Foundation DMS Grant No.
+2316011.
 The third author would like to acknowledge support from the DOE Office of Science Graduate Student Research (SCGSR) Program.

### 2025-08-04 Older versions/HickernellMCQMC2024_QMCTutorial_rev_2_cleaned.tex
fred updates before aleksei takes a look

--- 2025-08-04 HickernellMCQMC2024_QMCTutorial_rev_3.tex
@@
 Such a digital net is then called a $(t,m,d)$-net.
-\FJHNote{An infinite sequence defined by \eqref{eq:digitalB} where every box of the form \eqref{eq:element_box} contains its fair share of elements of the subsequence $\{\bsx^{\text{dig}}_{p 2^m}, \ldots \bsx^{\text{dig}}_{(p+1) 2^m-1} \}$ for all non-negative $m$ and $p$ is called a digital $(t,d)$-sequence.} Ideally, one would like to construct $(t,m,d)$-nets and $(t,d)$-sequences with $t$ as small as possible.
+An infinite sequence of the form \eqref{eq:digitalB} for which this condition holds for all non-negative $m$ is called a  $(t,d)$-sequence.
+Ideally, one would like to construct $(t,m,d)$-nets and $(t,d)$-sequences with $t$ as small as possible.
 
@@
 The generalized Halton sequence is subsequently defined by choosing $d$ sequences of permutations $\Sigma_j = (\sigma_{j,r})_{r\geq0}$ of $\{0,\ldots,b_j-1\}$ for $j = 1, 2,\ldots,d$ and setting
-$$
-\bsx_i^{\text{G-Hal}} = \left( \phi^{\Sigma_1}_{b_1}(i), \ldots, \phi^{\Sigma_d}_{b_d}(i)\right), \quad i= 0,1,\ldots.
-$$ 
-A randomized qMC estimate $\hat{\mu}_n$ is then an unbiased estimator of $\mu$ if the permutations $\Sigma_j$ are sampled independently and uniformly, i.e $\mu$ if $\Sigma_j \overset{\text{IID}}{\sim} \calU\{0,\dots,b_j-1\}^d$ since this ensures $\bsx_i^{\text{G-Hal}} \sim \mathcal{U}[0,1)^d$.
-
-We note in passing that many works \cite{Atan04,MasChiWar05,tuffin98,vancools06} have sought to optimize the permutations $\Sigma_j$ deterministically to enhance uniformity.
+$$\bsx_i^{\text{G-Hal}} = \left( \phi^{\Sigma_1}_{b_1}(i), \ldots, \phi^{\Sigma_d}_{b_d}(i)\right), \quad i= 0,1,\ldots,$$ A randomized qMC estimate $\hat{\mu}_n$ is then an unbiased estimator of $\mu$ if the permutations $\Sigma_j$ are sampled independently and uniformly, i.e $\mu$ if $\Sigma_j \overset{\text{IID}}{\sim} \calU\{0,\dots,b_j-1\}^d$ since this ensures $\bsx_i^{\text{G-Hal}} \sim \mathcal{U}[0,1)^d$.
+
+We note in passing that many works \cite{MasChiWar05,tuffin98,Atan04,vancools06} have sought to optimize the permutations $\Sigma_j$ deterministically to enhance uniformity.
 For a comprehensive survey of the Halton sequence (up to 2008), we refer readers to the excellent work by Faure and Lemieux \cite{faulem09}.
@@
 		\label{eq:KcondSym}
-		\text{Symmetry:} \quad & K(\bst,\bsx) = K(\bsx,\bst) \qquad \forall \bst,\bsx \in [0,1)^d, \\
+		\text{Symmetry:} \quad & K(\bst,\bsx) = K(\bsx,\bst) \qquad \forall \bst,\bsx \in [0,1)^d \\
 		\nonumber
@@
 	\begin{align}
-	\text{Belonging:} \quad & K(\cdot,\bsx) \in \calH_K \qquad \forall \bsx \in [0,1)^d, \\
+	\text{Belonging:} \quad & K(\cdot,\bsx) \in \calH_K \qquad \forall \bsx \in [0,1)^d \\
 	\text{Reproducing:} \quad & f(\bsx) = \langle K(\cdot,\bsx),f \rangle \qquad  \forall \bsx \in [0,1)^d, \ f \in \calH_K.
@@
 	& = \int_{[0,1)^d \times [0,1)^d} K(\bst,\bsx) \, \mathrm{d} \bst \, \mathrm{d} \bsx  -\frac 2n  \sum_{i=0}^{n-1} \int_{[0,1)^d} K(\bst,\bsx_i) \, \mathrm{d} \bst \\
-	& \qquad \qquad + \frac{1}{n^2} \sum_{i,j=0}^{n-1}  K(\bsx_i,\bsx_j), \label{eq:sqdisc}
+	& \qquad \qquad + \frac{1}{n^2} \sum_{i,j=0}^{n-1}  K(\bsx_i,\bsx_j). \label{eq:sqdisc}
 \end{align}
+
 where again the $(\cdot)$ and $(\cdot\cdot)$ notations represent the functional acting on the first and second arguments, respectively.
@@
 
-%Not all variance reduction methods for IID MC can be easily transferred to qMC.
+Not all variance reduction methods for IID MC can be easily transferred to qMC.
 For example, acceptance-rejection sampling requires an undetermined number of IID samples before one is accepted.
@@
 For finance examples where the outcome depends more on the gross behavior of the underlying stochastic process, rather than the high frequency behavior, it makes sense to choose the transformation accordingly.
-For example, in Figure \ref{fig:asian_option_pda_vs_chol} we see that if  $\bsT$ \FJHNote{contains observations from a Brownian motion}, then a principal component analysis coupled with the inverse Gaussian distribution shows greater convergence gains compared to the Cholesky decomposition \cite{AcwBroGla97}.
+For example, in Figure \ref{fig:asian_option_pda_vs_chol} we see that if  $\bsT$ represents a Brownian motion, then a principal component analysis coupled with the inverse Gaussian distribution shows greater convergence gains compared to the Cholesky decomposition \cite{AcwBroGla97}.
 
@@
     \nonumber
-    V_1 &= n_1 \, \times \, \text{discrepancy}\bigl(\{\bsx_i^{(1)}\}_{i=0}^{n_2-1}, K^{(1)} \bigr) \, \times \, \text{variation}(f^{(1)}, K^{(1)})\\
-    \nonumber
-    V_l & = n_l \, \times \, \text{discrepancy}\bigl(\{\bsx_i^{(l)}\}_{i=0}^{n_l-1}, K^{(l)} \bigr) \, \times \, \text{variation}(f^{(l)} - f^{(l-1)}, K^{(l)}), \\
+    V_1 &= n_1 \, \textcolor{blue}{\times} \, \text{discrepancy}\bigl(\{\bsx_i^{(1)}\}_{i=0}^{n_2-1}, K^{(1)} \bigr) \, \textcolor{blue}{\times} \, \text{variation}(f^{(1)}, K^{(1)})\\
+    \nonumber
+    V_l & = n_l \, \textcolor{blue}{\times} \, \text{discrepancy}\bigl(\{\bsx_i^{(l)}\}_{i=0}^{n_l-1}, K^{(l)} \bigr) \, \textcolor{blue}{\times} \, \text{variation}(f^{(l)} - f^{(l-1)}, K^{(l)}), \\
     \nonumber

### 2025-08-05 HickernellMCQMC2024_QMCTutorial_rev_3.tex
Updates from Overleaf

@@
 As the CLT holds only as $n$ goes to infinity, this is an approximate stopping criterion.
-For integrands $f$ with bounded kurtosis, a guaranteed version of this two-stage-sampling idea has been developed to account for the finite sample size using  Berry-Esseen inequalities\cite{HicEtal14a}.
+For integrands $f$ with bounded kurtosis, a guaranteed version of this two-stage-sampling idea has been developed to account for the finite sample size using  Berry-Esseen inequalities \cite{HicEtal14a}.
 

### 2025-08-05 Older versions/HickernellMCQMC2024_QMCTutorial_rev_1.tex
Updates from Overleaf

--- 2025-06-24 HickernellMCQMC2024_QMCTutorial_rev_1.tex
@@
 \begin{itemize}
-	\item Future market forces, which affect financial risk,
-	\item The porosity field of rock, which affects the extraction of oil or gas, or
-	\item Elementary particle interactions in a high energy physics experiment, which affect what is observed by detectors.
+	\item future market forces, which affect financial risk,
+	\item the porosity field of rock, which affects the extraction of oil or gas, or
+	\item elementary particle interactions in a high energy physics experiment, which affect what is observed by detectors.
 \end{itemize}
@@
 
-\FJHNote{Another family of LD sequences that is a generalization of the van der Corput sequence \eqref{eq:vdc} is digital sequences \cite{DicPil10a,Nie92}.
+\FJHNote{Another family of LD sequences that generalize the van der Corput sequence \eqref{eq:vdc} are digital sequences \cite{DicPil10a,Nie92}.
 For simplicity, we restrict ourselves to $b = 2$ and let $\oplus$ denote binary digitwise addition modulo $2$ (also known as digitwise exclusive or), e.g., $3/8 \oplus 3/4 = {}_20.011 \oplus {}_20.110 = {}_20.101 = 5/8$.

### 2025-08-05 Older versions/HickernellMCQMC2024_QMCTutorial_rev_1.tex
small changes from Nathan and Aleksei

@@
 As in the qMC-CLT stopping criterion, one may iteratively double $n$ until the bounds are sufficiently tight.
-\AGSNote{Figure \ref{fig:asian_option_pda_vs_chol} compares these decay-tracking qMC algorithms against IID Monte Carlo stopping criterion across a variety of tolerances for pricing an Asian option from finance \cite[p.\ 8]{Gla03}.}
+\AGSNote{Figure \ref{fig:asian_option_pda_vs_chol} compares these decay-tracking qMC algorithms against IID Monte Carlo stopping criterion across a range of tolerances for pricing an Asian option from finance \cite[p.\ 8]{Gla03}.}
 
@@
 	\caption{\AGSNote{Comparison of IID Monte Carlo and qMC stopping criteria for estimating the fair price of an Asian option with arithmetic mean.
-For QMC with Sobol' points, using a principal component analysis (PCA) decomposition of the Brownian motion provides substantial savings compared to the Cholesky decomposition.
+For qMC with Sobol' points, using a principal component analysis (PCA) decomposition of the Brownian motion provides substantial savings compared to the Cholesky decomposition.
 Using an Asian option with geometric mean control variate (CV) provides additional savings.

### 2025-08-05 HickernellMCQMC2024_QMCTutorial_rev_3.tex
aleksei: put all updates in rev3

@@
 
-Another family of LD sequences that is a generalization of the van der Corput sequence \eqref{eq:vdc} is digital sequences \cite{DicPil10a,Nie92}.
+Another family of LD sequences that generalize the van der Corput sequence \eqref{eq:vdc} are digital sequences \cite{DicPil10a,Nie92}.
 For simplicity, we restrict ourselves to $b = 2$ and let $\oplus$ denote binary digitwise addition modulo $2$ (also known as digitwise exclusive or), e.g., $3/8 \oplus 3/4 = {}_20.011 \oplus {}_20.110 = {}_20.101 = 5/8$.
@@
 \end{equation*}
-against sample size for a variety of $d$.  Now, even for larger $d$ there is a reasonable decay that approximates $\mathcal{O}(n^{-1})$.
+against sample size for a range of $d$.  Now, even for larger $d$ there is a reasonable decay that approximates $\mathcal{O}(n^{-1})$.
 
@@
 As in the qMC-CLT stopping criterion, one may iteratively double $n$ until the bounds are sufficiently tight.
-Figure \ref{fig:asian_option_pda_vs_chol} compares these decay-tracking qMC algorithms against IID Monte Carlo stopping criterion across a variety of tolerances for pricing an Asian option from finance \cite[p.\ 8]{Gla03}.
+Figure \ref{fig:asian_option_pda_vs_chol} compares these decay-tracking qMC algorithms against IID Monte Carlo stopping criterion across a range of tolerances for pricing an Asian option from finance \cite[p.\ 8]{Gla03}.
 
@@
 	\caption{Comparison of IID Monte Carlo and qMC stopping criteria for estimating the fair price of an Asian option with arithmetic mean.
-For QMC with Sobol' points, using a principal component analysis (PCA) decomposition of the Brownian motion provides substantial savings compared to the Cholesky decomposition.
+For qMC with Sobol' points, using a principal component analysis (PCA) decomposition of the Brownian motion provides substantial savings compared to the Cholesky decomposition.
 Using an Asian option with geometric mean control variate (CV) provides additional savings.
